### 2023-09-20 corridor_docs/proposal/eec_corridor_proposal.Rmd
initial commit of sample docs and rmd of our corridor proposal

---
title: "Corridor Project Proposal"
author: "Eric Escobar-Chena and Emilio M. Bruna"
date: "updated: `r Sys.Date()`"
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: darkmidnightblue
urlcolor: darkmidnightblue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
# - \fancyfoot[R]{last update `r Sys.Date()`}
- \fancyfoot[C]{ }
- \fancyhead[R]{p. \thepage}
- \fancyhead[L]{last update `r Sys.Date()`}
- \usepackage[default]{sourcesanspro}
# - \usepackage{parskip}
# - \usepackage{xcolor}
- \AtBeginDocument{\let\maketitle\relax}
# - \usepackage{sectsty} \sectionfont{\centering}
# - \usepackage{titlesec}
# - \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
# - \titlespacing{\section}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
# - \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
# - \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

# Title

## Project Personnel

- List of people involved, including an estimate of the number of technicians (if any)

## PI Contact Information

- Eric Escobar-Chena (eescobarchena@ufl.edu)
- Emilio Bruna (embruna@ufl.edu)

## Research proposal (1500 words max):

(a) Description of research, including type of sampling and description of any type of manipulation (even if minor) 
(b) Number and location of study areas, including a map of proposed sites and their relation to ongoing studies
(c) Potential impacts on corridor plots and ongoing studies  
(d) Study duration 
(e) A plan for how the study will be cleaned-up
(f) Funding sources (if applicable)
(g) Plan for making storing data on a publicly accessible repository (e.g. Environemental Data Intiative (EDI), Dryad, etc.)

### 2023-09-20 corridor_docs/thesis_ideas.Rmd
adding thesis ideas doc

---
title: "Project Ideas and Notes"
author: "Eric Escobar-Chena and Emilio M. Bruna"
date: "updated: `r Sys.Date()`"
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: darkmidnightblue
urlcolor: darkmidnightblue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
# - \fancyfoot[R]{last update `r Sys.Date()`}
- \fancyfoot[C]{ }
- \fancyhead[R]{p. \thepage}
- \fancyhead[L]{last update `r Sys.Date()`}
- \usepackage[default]{sourcesanspro}
# - \usepackage{parskip}
# - \usepackage{xcolor}
- \AtBeginDocument{\let\maketitle\relax}
# - \usepackage{sectsty} \sectionfont{\centering}
# - \usepackage{titlesec}
# - \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
# - \titlespacing{\section}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
# - \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
# - \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
bibliography: "eric-thesis.bib"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(citr)
```

### _Notes from conversation with Julian:_

1. Low-hanging fruit: Effect of connectivity on (a) species and functional group diversity and abundance
1. Next level: Movement & Dispersal. There are large areas of forest where it is possible to capture beetles for Mark-Release-Recapture (MRR) experiments.
1. Higher-risk, Higher reward (even if they don't go in MS, can be set-up for potential PHD projects): Experiments on Ecosystem Services

- dung burial, decomposition, soil properties based on results of 1 & 2 (buckets with dung and beetle assemblages)
- gas (e.g., methane) emissions
- seed dispersal/burial/germination

### Example studies

1. Diversity and Abundance

+ @WOS:000178929600002: "56% of individuals were capturedin the continuous forest, 29% in the mosaic habitat and 15% in the forest fragments"

2. Movement & Dispersal

3. Ecosystem Services

# Other stuff to work on:

1. Species List & Keys for Dung Beetles of the Southeastern US
1. Any previous work done on dung beetles in Southeastern US
1. List of Equipment and Tools needed
+ Dung beetle traps
+ alcohol
+ bait

References

### 2023-10-18 corridor_docs/thesis_ideas.Rmd
adding questions
added prelim questions, lit search questions

@@
 
-+ @WOS:000178929600002: "56% of individuals were capturedin the continuous forest, 29% in the mosaic habitat and 15% in the forest fragments"
++ @WOS:000178929600002: "56% of individuals were captured in the continuous forest, 29% in the mosaic habitat and 15% in the forest fragments"
 
@@
 
+# Species
+
+\newpage
+
+INTRODUCTION
+
+<!-- Good stuff in here: -->
+<!-- https://link.springer.com/article/10.1007/s10531-017-1299-7 -->
+<!-- https://www.sciencedirect.com/science/article/pii/S1470160X19303486 -->
+
+1. Paragraph 1: Habitat Fragmentation is a common feature of most human dominated landscapes, and Corridors are thought to be an important means of connecting these isolated patches.
+1. Paragraph 2: It is predicted that Corridors will increase diversity of connected patches. Summary of findings. However, because of this gap in research, it remains unclear how_____.
+1.
+Paragraph 3: Part of the reason we havbe this gap is that few studis have assessed dispersal of animals through corridors from one patch to another.
+I know...crazy.
+Lots of studies have assessed movement, but few of these have actually done so in cases where one patch is conncected to another.
+1.
+Paragraph 4: Dung Beetles are a critical and divese functional group that is often used as an indicator species.
+I am proposing to use a large-scale experiement of connectd and unconnected patches to assess how corridors shape dung beetle diversity and movement.
+Specifically, i will address the following questions:
+
+\newpage
+
+## What is predicted about how corridors influence the diversity of insects communities (with an emphasis on Dung Beetles) in patches?
+
+Are corridors predicted to enhance diversity in connected patches?
+What is the mechanism>?
+A certain species groups are predicted to drive this change in diversity - rare species, larger species, specialists, particular function groups, better fliers, etc.?
+(_Note: "Corridors" = Riparian strips, living fences, linear fragments, etc._)
+
+- **Prediction 1:** Diversity higher in connected patches (citations in Damschen et al PNAS, others)
+
+- **Prediction 2:**
+
+### Empiricial Results
+
+### 1. Studies that have compared diversity of Dung Beetles in Corridors and Other Habitats (e.g., primary forest, pastures) have found that:
+
+- **Diversity increases:**
+
+- **Diversity decreases:**
+
+- **Similar diversity in both:**
+
+### 2. Studies that have actually compared the diversity of patches that are connected by a corridor vs. patches that are unconnected
+
+- **Diversity increases:**
+
+- **Diversity decreases:**
+
+- **Similar diversity in both:**
+
+## What is predicted or known about insect movement through corridors (with an emphasis on Dung Beetles)?
+
+Do Corridors facilitate movement between patches?
+Are particular species or groups more able to move through corridors than others - larger species, specialists on particular resources, particular functional groups, better fliers, etc.?
+
+### 1. Studies that have assessed movement of Dung Beetles have found the following can play an important role:
+
+- **Habitat Preference:** [@WOS:000360850000002, @WOS:000262759300008].
+
+- **Wing Loading:** [@WOS:000360850000002]
+
+- **Sex:**
+
+- **Foraging Mode:**
+
+- **Size:**
+
+- **Invasive vs. Native:**
+
+- **No inter-specific or inter-group differences in movement:**
+
+### 2. Studies that have estimated dispersal distance:
+
+\newpage
+
+# THE QUESTIONS
+
+Question 1: Is dung beetle abundance greater in connected patches than isolated ones?
+
+<!-- Review predictions (or results) from studies of habitat fragments, other corridor studies, etc to refine this prediction -->
+
+Question 2: Is dung beetle diversity greater in connected patches than isolated ones?
+
+> Prediction: Species diversity is greater in connected patches than isolated ones (due to the greater representation of rare species) 
+> OR
+> Prediction: Diversity of _______ (e.g., specialists, foraging strategy X, etc.) is greater in connected patches, but the Diversity of ________ (e.g., generalists, foraging strategy Y) will be the similar in both.
+> OR
+> The diversity is similar, but the community composition changes to one dominated by specialists instead of generalists
+
+2.  Do Dunge beetles move through corridors?
+
 References

### 2023-10-25 corridor_docs/thesis_ideas.Rmd
made a change to test committing
added citation

@@
 
-### _Notes from conversation with Julian:_
+### *Notes from conversation with Julian:*
 
 1.  Low-hanging fruit: Effect of connectivity on (a) species and functional group diversity and abundance
-1. Next level: Movement & Dispersal. There are large areas of forest where it is possible to capture beetles for Mark-Release-Recapture (MRR) experiments.
-1. Higher-risk, Higher reward (even if they don't go in MS, can be set-up for potential PHD projects): Experiments on Ecosystem Services
+2.  Next level: Movement & Dispersal. There are large areas of forest where it is possible to capture beetles for Mark-Release-Recapture (MRR) experiments.
+3.  Higher-risk, Higher reward (even if they don't go in MS, can be set-up for potential PHD projects): Experiments on Ecosystem Services
 
@@
 
-+ @WOS:000178929600002: "56% of individuals were captured in the continuous forest, 29% in the mosaic habitat and 15% in the forest fragments"
+-   @WOS:000178929600002: "56% of individuals were captured in the continuous forest, 29% in the mosaic habitat and 15% in the forest fragments"
 
@@
 1.  Species List & Keys for Dung Beetles of the Southeastern US
-1. Any previous work done on dung beetles in Southeastern US
-1. List of Equipment and Tools needed
-+ Dung beetle traps
-+ alcohol
-+ bait
+2.  Any previous work done on dung beetles in Southeastern US
+3.  List of Equipment and Tools needed
+
+-   Dung beetle traps
+-   alcohol
+-   bait
 
@@
 <!-- Good stuff in here: -->
+
 <!-- https://link.springer.com/article/10.1007/s10531-017-1299-7 -->
+
 <!-- https://www.sciencedirect.com/science/article/pii/S1470160X19303486 -->
@@
 1.  Paragraph 1: Habitat Fragmentation is a common feature of most human dominated landscapes, and Corridors are thought to be an important means of connecting these isolated patches.
-1. Paragraph 2: It is predicted that Corridors will increase diversity of connected patches. Summary of findings. However, because of this gap in research, it remains unclear how_____.
-1.
+2.  Paragraph 2: It is predicted that Corridors will increase diversity of connected patches. Summary of findings. However, because of this gap in research, it remains unclear how\_\_\_\_\_.
+3.
 Paragraph 3: Part of the reason we havbe this gap is that few studis have assessed dispersal of animals through corridors from one patch to another.
@@
 Lots of studies have assessed movement, but few of these have actually done so in cases where one patch is conncected to another.
-1.
+4.
 Paragraph 4: Dung Beetles are a critical and divese functional group that is often used as an indicator species.
@@
 Are corridors predicted to enhance diversity in connected patches?
-What is the mechanism>?
+What is the mechanism\>?
 A certain species groups are predicted to drive this change in diversity - rare species, larger species, specialists, particular function groups, better fliers, etc.?
-(_Note: "Corridors" = Riparian strips, living fences, linear fragments, etc._)
+(*Note: "Corridors" = Riparian strips, living fences, linear fragments, etc.*)
 
@@
 
-- **Diversity increases:**
+-   **Diversity increases:** [@WOS:000399235300014]
 
@@
 
-> Prediction: Species diversity is greater in connected patches than isolated ones (due to the greater representation of rare species) 
-> OR
-> Prediction: Diversity of _______ (e.g., specialists, foraging strategy X, etc.) is greater in connected patches, but the Diversity of ________ (e.g., generalists, foraging strategy Y) will be the similar in both.
-> OR
-> The diversity is similar, but the community composition changes to one dominated by specialists instead of generalists
+> Prediction: Species diversity is greater in connected patches than isolated ones (due to the greater representation of rare species) OR Prediction: Diversity of \_\_\_\_\_\_\_ (e.g., specialists, foraging strategy X, etc.) is greater in connected patches, but the Diversity of \_\_\_\_\_\_\_\_ (e.g., generalists, foraging strategy Y) will be the similar in both.
+OR The diversity is similar, but the community composition changes to one dominated by specialists instead of generalists
 
@@
 References
+
+#practice push

### 2023-11-13 corridor_docs/thesis_ideas.Rmd
adding paragraphs for proposal

@@
 
-1.  Paragraph 1: Habitat Fragmentation is a common feature of most human dominated landscapes, and Corridors are thought to be an important means of connecting these isolated patches.
-2.  Paragraph 2: It is predicted that Corridors will increase diversity of connected patches. Summary of findings. However, because of this gap in research, it remains unclear how\_\_\_\_\_.
+1.
+Paragraph 1: In an increasingly fragmented world, corridors are thought to be an important strategy of connecting isolated patches to promote population persistence and increase diversity in connected patches.
+This can ultimately have consequences for ecosystem processes in fragments.
+2.
+Paragraph 2: Despite this, direct evidence of movement between patches is limited.
+Plenty of studies have demonstrated that the diversity or abundance of organisms is similar in corridors and the primary habitat to which they are connected.
+However, studies demonstrating animals actually move from one patch through a corridor to another patch are surprisingly rare (but see examples).
+NB: might have to say "although there is ample evidence that vertbrates move through corridors, studies documenting the movements of invertebrates - the dominant species on earth - are limited".
 3.
-Paragraph 3: Part of the reason we havbe this gap is that few studis have assessed dispersal of animals through corridors from one patch to another.
-I know...crazy.
-Lots of studies have assessed movement, but few of these have actually done so in cases where one patch is conncected to another.
+Paragraph 3: Part of the challenge in assessing the  efficacy of corridors is the confounding effects of patch area and the length of habitat edges (see resasco 2014 etc); to address this it is necessary to compare connected and unconnected areas of equal area and shape.
 4.
-Paragraph 4: Dung Beetles are a critical and divese functional group that is often used as an indicator species.
-I am proposing to use a large-scale experiement of connectd and unconnected patches to assess how corridors shape dung beetle diversity and movement.
-Specifically, i will address the following questions:
+Paragraph 4: Dung beetles have emerged as a model system with which to test hypotheses on how anthropogenic landscape alterations influence biodiversity.
+They are locally species rich, exhibit variety in key funcitonal traits (size, foraging style,,....), and they provide critical ecosystem services.
+robust (but not *too* robust): studies have shown diversity of dung beetles can be lower in fragments, but that they can be found in corridors and are capable of dispersing for up to 1 km.
+5.
+Paragraph 5: I am proposing to use a landscape-scale fragmentation experiment to test the effects of corridors on the diversity and dispersal dung beetles and the ecosystem services they provide.
+Specifically, I will test the following predictions:
+
+1. Diversity will be higher in connected patches
+
+2. Dung beetles disperse through corridors, with larger beetles dispersing more quickly (size-dependent dispersal ability)
+
+3. Dung removal rates will be highest in connected patches.
+
+To test there predictions, I will use a combination of passive sampling and a MRR experiment and control for the confounding effects of edge and area.
 
@@
 
--   **Prediction 2:**
+-   **Prediction 2:** Diversity could be lower in patches (in specific case of corridor project, only getting open space species)
 
@@
 
--   **Diversity increases:** [@WOS:000399235300014]
+-   **Diversity increases:** [@WOS:000399235300014; @WOS:A1995TL67200023]
 
@@
 References
-
-#practice push

### 2024-01-23 corridor_docs/proposal/eec_corridor_proposal.Rmd
update for submition to review committee

@@
 
-# Title
+# The effect of corridors on dung beetle diversty, abundance, and dispersal
 
@@
 
-- List of people involved, including an estimate of the number of technicians (if any)
+- Eric Escobar-Chena (MS student, UF) 
+- Emilio Bruna (UF)
+- Julian Resasco (UC Boulder)
+- 1 field intern (TBD)
 
@@
 
+- Emilio Bruna (embruna@ufl.edu)
 - Eric Escobar-Chena (eescobarchena@ufl.edu)
-- Emilio Bruna (embruna@ufl.edu)
 
@@
 (a) Description of research, including type of sampling and description of any type of manipulation (even if minor)
+
+### Background
+
+In an increasingly fragmented world, corridors are considered a key strategy for promoting connectivity and facilitating the movement of organisms through the landscape.
+Such movement is hypothesized to help maintain species diversity in fragments; it could also have patch- and landscape-level consequences for the ecosystem processes in which influenced by these species.
+Although there is some evidence that animals disperse between patches via corridors, and that connected patches have higher species diversity than unconnected ones, little work to date has investigated the consequences of these corridor-driven patterns for ecosystem ecosystem.
+
+Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity.
+They are locally species rich, exhibit variety in key functional traits (e.g., size, foraging style, resource-use), are capable of long-distance dispersal (e.g., some species up to 1 km), and that they influence a number of critical ecosystem services (e.g., seed dispersal, nutrient cycling).
+Previous work has shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors.
+However, it remains unknown if corridors can reduce the loss of species from fragments, how such declines are influenced by inter-specific differences in dispersal capability, and what the consequences of these patterns are for the ecosystems sefvices provided by dung beetles.
+
+We propose using the SRS corridor experiment to test whether landscape connectivity via corridors can increase gene flow to and genetic diversity of isolated inbred populations via pollination using a model organism, the sunflower, Helianthus annus.
+Our experimental landscapes are ideal because they allow us to test for the effects of connectivity while controlling for typically confounding factors and to conduct our experiments at large spatial scales that approximate those of conservation management activities.
+Our design is also ideal because it allows us to separate the effects of genetic rescue from those of demographic rescue that are typically confounded due to their difficulty in teasing apart in other studies.
+
+We propose using the SRS Corridor Experiment to test for corridor use by dung beetles, whether corridor results in higher species diversity in connected than unconnected patches, and if differences in diversity influence nutrient cycling.
+Specifically, we will test the following predictions:
+
+1. Diversity will be higher in connected patches. To test this prediction, I will sample dung beetles with pitfall traps and use non-parametric estimators of species diversity
+
+2. Dung beetles disperse through corridors, with larger beetles dispersing more quickly (size-dependent dispersal ability)
+
+3. Dung removal rates will be highest in connected patches.
+
+To test there predictions, I will use a combination of passive sampling and a MRR experiment and control for the confounding effects of edge and area.
+
 (b) Number and location of study areas, including a map of proposed sites and their relation to ongoing studies
-(c) Potential impacts on corridor plots and ongoing studies  
+
+(c) Potential impacts on corridor plots and ongoing studies:
+
+CHnages in duyng beetle abundance due to collection (short term, unlikely to be high impact)
+Dung introduced into plots (contained in above ground plots, so also unlikely)
+note no holes being dug for pitfalls
+
 (d) Study duration
-(e) A plan for how the study will be cleaned-up
+
+Preliminary sampling during Spring 2023
+
+Summer and Fall Sampling
+
+(e) Plan for how the study will be cleaned-up
+
+removing all traps and pin flags after each sampling bout
+
 (f) Funding sources (if applicable)
-(g) Plan for making storing data on a publicly accessible repository (e.g. Environemental Data Intiative (EDI), Dryad, etc.)
+
+(g) Plan for making data publicly accessible
+
+> Data will be entered into spreadsheets and backed up by saving them to a repository on the Bruna Lab's Github site (https://github.com/BrunaLab)  along with a .txt file of metadata and all R scripts for data correction and analysis.
+When new data are added they will be automatically validated using  Github actions and the `pointblank` library (e.g., https://brunalab.github.io/HeliconiaSurveys/survey_validation/survey_validation.html).
+This approach allows us to share the all materials with collaborators and easily archive code and data at Zenodo and Dryad (respectively) upon the acceptance of a manuscript.
+
+> For an overview of our approach to data archiving and accessibility see https://github.com/BrunaLab/HeliconiaSurveys

### 2024-01-24 corridor_docs/proposal/eec_corridor_proposal.Rmd
uodated, added srs plot schematic

--- 2023-09-20 corridor_docs/proposal/eec_corridor_proposal.Rmd
@@
 author: "Eric Escobar-Chena and Emilio M. Bruna"
-date: "updated: `r Sys.Date()`"
-mainfont: SourceSansPro
+date: 'updated: `r Sys.Date()`'
+output:
+  pdf_document: default
+  html_document:
+    df_print: paged
 geometry: margin=1in
@@
 - \pagestyle{fancy}
-# - \fancyfoot[R]{last update `r Sys.Date()`}
 - \fancyfoot[C]{ }
@@
 - \usepackage[default]{sourcesanspro}
-# - \usepackage{parskip}
-# - \usepackage{xcolor}
 - \AtBeginDocument{\let\maketitle\relax}
-# - \usepackage{sectsty} \sectionfont{\centering}
-# - \usepackage{titlesec}
-# - \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
-# - \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
-# - \titlespacing{\section}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
-# - \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
-# - \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
-# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
 - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
-output: pdf_document
+mainfont: SourceSansPro
 ---
@@
 
-# Title
+## Effect of corridors on dung beetle diversity, dispersal, & ecosystem services
 
-## Project Personnel
+### Project Personnel
 
-- List of people involved, including an estimate of the number of technicians (if any)
+- Eric Escobar-Chena (MS student, UF) 
+- Emilio Bruna (UF)
+- Julian Resasco (UC Boulder)
+- 1 field intern (TBD)
 
-## PI Contact Information
+### PI Contact Information
 
+- Emilio Bruna (embruna@ufl.edu)
 - Eric Escobar-Chena (eescobarchena@ufl.edu)
-- Emilio Bruna (embruna@ufl.edu)
 
-## Research proposal (1500 words max):
+### Research proposal (1500 words max): 
+<!-- (a) Description of research, including type of sampling and description of any type of manipulation (even if minor)  -->
 
-(a) Description of research, including type of sampling and description of any type of manipulation (even if minor) 
-(b) Number and location of study areas, including a map of proposed sites and their relation to ongoing studies
-(c) Potential impacts on corridor plots and ongoing studies  
-(d) Study duration 
-(e) A plan for how the study will be cleaned-up
-(f) Funding sources (if applicable)
-(g) Plan for making storing data on a publicly accessible repository (e.g. Environemental Data Intiative (EDI), Dryad, etc.)
+### Background
+
+In an increasingly fragmented world, corridors are considered a key strategy for promoting connectivity and facilitating the movement of organisms through the landscape.
+Such movement is hypothesized to help maintain species diversity in fragments; it could also have patch- and landscape-level consequences for the ecosystem processes in which influenced by these species.
+Although there is some evidence that animals disperse between patches via corridors, and that connected patches have higher species diversity than unconnected ones, little work to date has investigated the consequences of these corridor-driven patterns for ecosystem ecosystem.
+
+Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity.
+They are locally species rich, exhibit variety in key functional traits (e.g., size, foraging style, resource-use), are capable of long-distance dispersal (e.g., some species up to 1 km), and that they influence a number of critical ecosystem services (e.g., seed dispersal, nutrient cycling).
+Previous work has shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors.
+However, it remains unknown if corridors can reduce the loss of species from fragments, how such declines are influenced by inter-specific differences in dispersal capability, and what the consequences of these patterns are for the ecosystems services provided by dung beetles.
+Though there are several reasons for this, one major factor is the challenge in finding locations where one can assess the role of corridors while also while controlling for confounding factors such as patch size, edge, and corridor length.
+
+**_We propose using the community of dung beetles at the SRS Corridor Experiment to test for_** **(1)** higher species richness, species diversity, and functional diversity in patches connected by corridors than in unconnected patches **(2)** interspecific differences in corridor use and inter-patch movement, and **(3)** higher rates of dung removal in connected than unconnected patches.
+The SRS Corridor Experiment is an ideal location in which to conduct our stud.
+The design overcome the primary technical impediments to isolating the effects of corridors _per se_, its spatial scale allows for drawing and realistic and relevant conclusions regarding dung beetle dispersal in fragmented landscapes, and the local dung beetle assemblage is a highly tractable one with which to address our questions (i.e., taxonomically and functionally diverse, keys for species identification are readily available, there are multiple abundant species from different functional groups with which to conduct experiments, and the experimental/survey methods are standardized and straightforward).
+
+**_Our sampling and experiments will test the following predictions:_**
+
+1. Prediction 1: Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches
+
+<!-- 2. Predictions 2: Corridors facilitate inter-patch movement of dung beetle species,  -->
+2.
+Prediction 2: Corridors are suitable habitat for all species.
+However, the speed at which in individuals move through corridors and the probability of successful inter-patch movement is size-dependent (i.e., larger beetle species move more quickly and are more likely to reach the connected patch).
+
+3. Dung removal rates will be highest in connected patches. This is due to the higher functional diversity of beetles in these locations.
+
+Below we describe the combinations of sampling and experiments with which we will test these predictions.
+
+### Methods
+
+<!-- (b) Number and location of study areas, including a map of proposed sites and their relation to ongoing studies -->
+
+_Study site_
+
+1. Design Prediction 1 (pitfall sampling)
+1. Design Prediction 2 (MRR)
+1. Design Prediction 3 (Dung Removal Experiment)
+
+![Sampling Design](./images/srs_wt.png){fig-align="center" width=95% height=95%}
+
+Each experimental landscape block consists of a 1-ha square center patch surrounded by four peripheral patches that are equal in area (~1.4 ha) but differ in the amount of habitat connectivity or edge (Fig.
+1).
+One peripheral patch is connected to the center patch by a 150 x 25-m corridor (hereafter “connected”).
+The isolated patches are also 150 m from the center patch and vary in their edge amount.
+The winged patch type (hereafter “isolated high edge”) is a 100 x 100-m square with two 75 x 25 -m projections on either side of the patch.
+The rectangular patch type (hereafter “isolated low edge”) consists of a 100 x 100-m square with an area equal in size to the corridor added to the part of the patch farthest from the center patch.
+Each block contains one of each patch type and a duplicate of one of the isolated patch types.
+
+<!-- Fig.
+1 – Schematic of experimental design showing 1 of the 7 blocks.
+Center patches (lower left) are planted by with 16 source populations of wild sunflowers from locations throughout the US (from US-NPGS; represented by rainbow colors) and will be randomly assigned to a grid pole location.
+Connected and one unconnected patch, C location (rectangular or winged) will each be planted with an inbred line (HA 467).
+-->
+
+_Focal Species Assemblage_
+
+1. How many species, functional groups, range in size, what is known about them
+1. Table of expected species
+
+_Potential impacts on corridor plots and ongoing studies_
+
+1. no holes are being dug for pitfalls: no disturbance of the seeds or soil
+1. changes in dung beetle abundance due to collection: short term, unlikely to be high impact
+1. Dung introduced into plots: contained in above ground traps, so also unlikely, standardized and from cows taking abntibiotics, so no disease introduced,.
+
+_Study duration_
+
+1. Preliminary sampling during Spring 2023
+1. Summer and Fall Sampling
+<!-- Plan for how the study will be cleaned-up -->
+1. All traps and pin flags will be removed after each sampling bout
+
+_Funding sources (if applicable)_
+
+_Plan for making data publicly accessible_
+
+Data will be entered into spreadsheets and backed up by saving them to a repository on the Bruna Lab's Github site (https://github.com/BrunaLab)  along with a .txt file of metadata and all R scripts for data correction and analysis.
+When new data are added they will be automatically validated using  Github actions and the `pointblank` library (e.g., https://brunalab.github.io/HeliconiaSurveys/survey_validation/survey_validation.html).
+This approach allows us to share the all materials with collaborators and easily archive code and data at Zenodo and Dryad (respectively) upon the acceptance of a manuscript.
+For an overview of our approach to data archiving and accessibility see https://github.com/BrunaLab/HeliconiaSurveys

### 2024-01-26 corridor_docs/proposal/eec_corridor_proposal.Rmd
adding plots for each experiment

@@
 
-1. Prediction 1: Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches
+1. **Prediction 1:** Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches
 
@@
 2.
-Prediction 2: Corridors are suitable habitat for all species.
+**Prediction 2:** Corridors are suitable habitat for all species.
 However, the speed at which in individuals move through corridors and the probability of successful inter-patch movement is size-dependent (i.e., larger beetle species move more quickly and are more likely to reach the connected patch).
 
-3. Dung removal rates will be highest in connected patches. This is due to the higher functional diversity of beetles in these locations.
+3. **Prediction 3:** Dung removal rates will be highest in connected patches. This is due to the higher functional diversity of beetles in these locations.
 
@@
 
-### Methods
+### Study Site and Methods
 
@@
 
+_Methods_
+**Methods - Prediction 1:**
+
+To calculate species richness, species diversity, and functional diversity we will sample the dung beetle community in each location with pitfall traps baited with cow dung (Spector 2006, Andrade et al.
+2011).
+In each site we will arrange traps in a grid, with traps separated from each other by 25 m.
+The traps will be 25 m from the edge of the patch to avoid attracting beetles from the matrix outside of the patch.
+Traps in each location will be set on the same day and left out for 24 hours.
+
+The traps, modified from the design of Wood and Kaufman (2008), are constructed from 6 inch sections of (4?) inch(cm) diameter PVC pipe fitted with a screen on the bottom and a -- cm diameter funnel on top with an extended --cm plastic platform used to suspend the trap over a hole in the ground.
+The traps will be set level to the ground by placing them into a semi-permanent hole that was maintained by a larger section of PVC while traps were not in use.
+Dung is wrapped in paper towel to exclude beetles and suspended over the trap on a wire cloth platform.
+Note that although the traps of Wood and Kaufman (2008) were designed as a live catch trap, ours will be converted to a wet trap by placing a --ml cup containing 95% ethanol inside the trap (the original design allows some beetles to escape and for predatory beetles to enter the traps).
+
+The dung used to bait traps will be collected from the University of Florida’s Beef Teaching Unit from cattle that were unsupplemented – (with what).
+It is  collected fresh in the morning, homogenized, and formed into approximately (_ml) balls and frozen at (_degrees Celsius) until deployment.
+While some dung beetles have a clear preference for a particular type of dung, most species are generalists that use a broad diversity of vertebrate excrement (cite - check Gordon 1983, and Mart´in Piera and Lobo 1996, and see what Woodruff has to say; see page 13 of Haffter and Edmonds).
+All beetles captured will stored in 95% ethanol and identified to species or morphospecies using Woodruff (1973).
+A collection of voucher specimens will be deposited at the FLMNH and ------.
+
 1. Design Prediction 1 (pitfall sampling)
-1. Design Prediction 2 (MRR)
-1. Design Prediction 3 (Dung Removal Experiment)
 
-![Sampling Design](./images/srs_wt.png){fig-align="center" width=95% height=95%}
+- 5 traps in a grid in the central 50 x 50 m of each patch + 1 sampling grid in the matrix. The 25 m buffer from the edge is to avoid beetles coming into the plot from the matrix.
+- will try to test buffer width with marked beetles released at different distances, including from outside matrix
+- 4 arrays per blocks = 36 traps per array, 7 blocks = 252 traps total
+- need to make sure - ebnough "coverage" without a buffer too small, meaning bugs from matrix come into patch
+
+![Sampling Design](./images/plots1.png){fig-align="center" width=50% height=50%}
+
+**Methods - Prediction 2:** 
+2. Design Prediction 2 (Dispersal)
+
+- using connected patches + unconected: 2 dung baits in center patch.
+one at ende closest to isolated patch, one at end of corridor,.
+release marked beetles at ende of isolated patch and entry of corridor.
+Check at intervals of X to see if they will make it.
+- to consider: passive traps to see if they are moving through? shouidl we start experiment with smaller distances into the orridor and then eventually the full length? 
+- tech to follow them?
+
+![Sampling Design](./images/plots2.png){fig-align="center" width=50% height=50%}
+
+**Field Methods - Prediction 3:** 
+3. Design Prediction 3 (Dung Removal Experiment)
+
+- field experiment: 300 g of dung on surface, 300 g on surface covered with mesh (controls for dessication), weigh after 48 hours
+-  to consider: gather and sieve soil from outside, place in shallow pot, and place dung on that. Allows breakdown of soil without getting into their soils, if this is a concern
+- repeat biweekly for X months 
+- lab experiment: meso-cosms in GNV based on results from diversity sampling 
+- need to make sure each point inside a plot is independent, and that each sampling date is independent
+
+![Sampling Design](./images/plots3.png){fig-align="center" width=50% height=50%}
 

### 2024-01-29 corridor_docs/proposal/eec_corridor_proposal.Rmd
update proposal text

--- 2024-01-24 corridor_docs/proposal/eec_corridor_proposal.Rmd
@@
 geometry: margin=1in
+mainfont: Times New Roman
+# mainfont: SourceSansPro
 fontsize: 12pt
@@
 - \usepackage{fancyhdr}
+- \renewcommand{\headrulewidth}{0pt}
 - \pagestyle{fancy}
+
+# - \fancyhead[R]{p. \thepage}
+# - \fancyhead[L]{ }
+- \fancyhead[L]{Corridor Proposal, E. Escobar-Chena}
+- \fancyhead[R]{p. \thepage}
 - \fancyfoot[C]{ }
-- \fancyhead[R]{p. \thepage}
-- \fancyhead[L]{last update `r Sys.Date()`}
-- \usepackage[default]{sourcesanspro}
+# - \fancyhead[L]{updated `r Sys.Date()`}
+- \fancyfoot[R]{updated `r Sys.Date()`}
+# - \usepackage[default]{sourcesanspro}
 - \AtBeginDocument{\let\maketitle\relax}
 - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
-mainfont: SourceSansPro
+- \raggedright
 ---
@@
 
-## Effect of corridors on dung beetle diversity, dispersal, & ecosystem services
+## Influence of corridors on dung beetle diversity, dispersal, & ecosystem services
 
@@
 - Eric Escobar-Chena (MS student, UF) 
-- Emilio Bruna (UF)
-- Julian Resasco (UC Boulder)
-- 1 field intern (TBD)
+- Emilio M. Bruna (MS Committee Chair, UF)
+- Julian Resasco (MS Committee Member, UC Boulder)
+- 1 undergraduate field intern (TBD)
 
@@
 
-### Research proposal (1500 words max): 
+### Research proposal 
 <!-- (a) Description of research, including type of sampling and description of any type of manipulation (even if minor)  -->
 
-### Background
-
-In an increasingly fragmented world, corridors are considered a key strategy for promoting connectivity and facilitating the movement of organisms through the landscape.
-Such movement is hypothesized to help maintain species diversity in fragments; it could also have patch- and landscape-level consequences for the ecosystem processes in which influenced by these species.
+**_Background_:**  Corridors are considered an important mechanism for promoting connectivity and facilitating the movement of organisms through fragmented landscapes.
+These movements are hypothesized to help maintain species diversity in fragments; it could also have patch- and landscape-level consequences for the ecosystem processes in which influenced by these species.
 Although there is some evidence that animals disperse between patches via corridors, and that connected patches have higher species diversity than unconnected ones, little work to date has investigated the consequences of these corridor-driven patterns for ecosystem ecosystem.
@@
 Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity.
-They are locally species rich, exhibit variety in key functional traits (e.g., size, foraging style, resource-use), are capable of long-distance dispersal (e.g., some species up to 1 km), and that they influence a number of critical ecosystem services (e.g., seed dispersal, nutrient cycling).
+They are locally species rich, exhibit variety in key functional traits (e.g., size, foraging style, resource-use), are capable of long-distance dispersal (e.g., up to 1 km for some tropical species), and that they influence a number of critical ecosystem services (e.g., seed dispersal, nutrient cycling).
 Previous work has shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors.
@@
 
-**_We propose using the community of dung beetles at the SRS Corridor Experiment to test for_** **(1)** higher species richness, species diversity, and functional diversity in patches connected by corridors than in unconnected patches **(2)** interspecific differences in corridor use and inter-patch movement, and **(3)** higher rates of dung removal in connected than unconnected patches.
-The SRS Corridor Experiment is an ideal location in which to conduct our stud.
-The design overcome the primary technical impediments to isolating the effects of corridors _per se_, its spatial scale allows for drawing and realistic and relevant conclusions regarding dung beetle dispersal in fragmented landscapes, and the local dung beetle assemblage is a highly tractable one with which to address our questions (i.e., taxonomically and functionally diverse, keys for species identification are readily available, there are multiple abundant species from different functional groups with which to conduct experiments, and the experimental/survey methods are standardized and straightforward).
+**We propose using the community of dung beetles at the SRS Corridor Experiment to test for** **(1)** higher species richness, species diversity, and functional diversity in patches connected by corridors than in unconnected patches **(2)** interspecific differences in corridor use and inter-patch movement, and **(3)** higher rates of dung removal in connected than unconnected patches.
 
-**_Our sampling and experiments will test the following predictions:_**
+The Corridor Experiment is an ideal location in which to conduct this project.
+The design overcomes the primary technical impediments to isolating the effects of corridors _per se_, its spatial scale allows for drawing and realistic and relevant conclusions regarding dung beetle dispersal in fragmented landscapes, and the local dung beetle assemblage is a highly tractable one with which to address our questions (i.e., taxonomically and functionally diverse, keys for species identification are readily available, there are multiple abundant species from different functional groups with which to conduct experiments, and the experimental/survey methods are standardized and straightforward).
+**_Our sampling and experiments will be used to test the following predictions:_**
 
-1. Prediction 1: Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches
+1. **Prediction 1:** Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches
 
@@
 2.
-Prediction 2: Corridors are suitable habitat for all species.
+**Prediction 2:** Corridors are suitable habitat for all species.
 However, the speed at which in individuals move through corridors and the probability of successful inter-patch movement is size-dependent (i.e., larger beetle species move more quickly and are more likely to reach the connected patch).
 
-3. Dung removal rates will be highest in connected patches. This is due to the higher functional diversity of beetles in these locations.
+3. **Prediction 3:** Dung removal rates will be highest in connected patches. This is due to the higher functional diversity of beetles in these locations.
 
-Below we describe the combinations of sampling and experiments with which we will test these predictions.
+Below we describe the location of sampling or experiments with which we will test these predictions.
 
-### Methods
+### Study Site and Methods
 
-<!-- (b) Number and location of study areas, including a map of proposed sites and their relation to ongoing studies -->
+We propose to conduct the field component of our study in seven of the Savannah River Site's experimental landscapes (i.e., 'blocks') designed to assess the ecological effects of corridors.
+Each experimental landscape block consists of a 1-ha square center patch surrounded by four peripheral patches that are equal in area (~1.4 ha) but differ in the amount of habitat connectivity or edge (Figure 1).
+One peripheral patch is connected to the center patch by a 150 $\times$ 25m corridor (hereafter “connected”).
+The isolated patches are also 150 m from the center patch and vary in their edge amount.
+The winged patch type (hereafter “isolated high edge”) is a 100m $\times$ 100m square with two 75 $\times$ 25 m projections on either side of the patch.
+The rectangular patch type (hereafter “isolated low edge”) consists of a 100m $\times$ 100m square with an area equal in size to the corridor added to the part of the patch farthest from the center patch.
+Each block contains one of each patch type and a duplicate of one of the isolated patch types.
 
-_Study site_
+**_Methods: Beetle Diversity_:** To test the effect of corridors on species richness, species diversity, and functional diversity we will sample the dung beetle community in each location with pitfall traps baited with cow dung.
+Sampling will be conducted in four locations within in each landscape: the "connected" patch, the "rectangular" patch, one of the "winged" patches, and an area of matrix habitat at least 100 m from any of the patches (Fig.
+1).
+In each location we will arrange N = 5 pitfall traps in a grid in the center of the patch; the grid will be 25 m from the edge of the patch to reduce the likelihood of attracting beetles from the matrix habitat into the patch.
+The trap will be baited with 300 mg of cow dung (collected at the University of Florida’s Beef Teaching Unit and frozen until needed) suspended over a cup with 96% ethanol.
+All traps in a landscape will be set on the same day and left open for 24 hours; sampling will be conducted monthly in each landscape for 6 months to capture potential monthly and seasonal (wet/dry) variation in species abundance.
+A collection of voucher specimens will be deposited at the Florida Museum of Natural History, with a reference collection left at the Corridor Project.
 
-1. Design Prediction 1 (pitfall sampling)
-1. Design Prediction 2 (MRR)
-1. Design Prediction 3 (Dung Removal Experiment)
+<!-- Pitfall traps are constructed from 15 cm sections of PVC pipe (diameter = 10 cm).
+A metal screen has been attached to the bottom of the PVC, while a 14 cm diameter funnel has been inserted to the top.
+The with an extended --cm plastic platform used to suspend the trap over a hole in the ground.
+The traps will be set level to the ground by placing them into a semi-permanent hole that was maintained by a larger section of PVC while traps were not in use.
+Dung is wrapped in paper towel to exclude beetles and suspended over the trap on a wire cloth platform.
+Note that although the traps of Wood and Kaufman (2008) were designed as a live catch trap, ours will be converted to a wet trap by placing a --ml cup containing 95% ethanol inside the trap (the original design allows some beetles to escape and for predatory beetles to enter the traps).
+-->
 
-![Sampling Design](./images/srs_wt.png){fig-align="center" width=95% height=95%}
+<!-- The dung used to bait traps will be collected from the University of Florida’s Beef Teaching Unit from cattle that were unsupplemented – (with what).
+It is  collected fresh in the morning, homogenized, and formed into approximately (_ml) balls and frozen at (_degrees Celsius) until deployment.
+While some dung beetles have a clear preference for a particular type of dung, most species are generalists that use a broad diversity of vertebrate excrement (cite - check Gordon 1983, and Mart´in Piera and Lobo 1996, and see what Woodruff has to say; see page 13 of Haffter and Edmonds).
+All beetles captured will stored in 95% ethanol and identified to species or morphospecies using Woodruff (1973).
+A collection of voucher specimens will be deposited at the FLMNH and ------.
+-->
 
-Each experimental landscape block consists of a 1-ha square center patch surrounded by four peripheral patches that are equal in area (~1.4 ha) but differ in the amount of habitat connectivity or edge (Fig.
-1).
-One peripheral patch is connected to the center patch by a 150 x 25-m corridor (hereafter “connected”).
-The isolated patches are also 150 m from the center patch and vary in their edge amount.
-The winged patch type (hereafter “isolated high edge”) is a 100 x 100-m square with two 75 x 25 -m projections on either side of the patch.
-The rectangular patch type (hereafter “isolated low edge”) consists of a 100 x 100-m square with an area equal in size to the corridor added to the part of the patch farthest from the center patch.
-Each block contains one of each patch type and a duplicate of one of the isolated patch types.
+<!-- - 5 traps in a grid in the central 50 $\times$ 50 m of each patch + 1 sampling grid in the matrix. The 25 m buffer from the edge is to avoid beetles coming into the plot from the matrix. -->
+<!-- - will try to test buffer width with marked beetles released at different distances, including from outside matrix -->
+<!-- - 4 arrays per blocks = 36 traps per array, 7 blocks = 252 traps total -->
+<!-- - need to make sure - ebnough "coverage" without a buffer too small, meaning bugs from matrix come into patch -->
+
+![Design for pitfall sampling in each of the seven blocks.](./images/plots1.png){fig-align="center" width=50% height=50%}
+
+**_Methods: Beetle dispersal_:** To determine if beetles use corridors to move between patches, and if patterns of movement differ between species, we will conduct a mark-release-recapture (i.e., MRR) experiment.
+A pitfall trap will be placed in the middle of the landscape's "center" patch and baited with 300 g of cow dung.
+We will then release marked beetles at two points equidistant from the bait: the "connected" patch and the edge of the "rectangular" patch(Figure 2); these locations are separated from the bait by either the corridor or matrix habitat, allowing us to determine if corridors allow for more rapid movement and how this might vary by beetle species or functional group.
+We will check the pitfall trap for marked beetles twice per day for three days following their release.
+
+We anticipate initially conducting this experiment with two species:LARGE AND SMALL, but may modifify rthethe choice of species based on our initial sampling.
+We will also conduct preliomiary expertiments to determine the beetle dispral, and may place passive (i.e., unbaited) traps in the corridor, matrix, and patches.
+
+<!-- - to consider: passive traps to see if they are moving through? should we start experiment with smaller distances into the orridor and then eventually the full length?  -->
+<!-- - is there any technology to track or follow them? RFID? etc? -->
+
+![Design for experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=50% height=50%}
+
+**_Field Methods: Dung removal_:** The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles.
+We will compare the efficacy of dung removal by beetle communities in connected, unconnected, and winged patches using both mesocosms (at the University of Florida) and a field experiment conducted in seven of the experimental blocks.
+In each of the patches and a equally sized plot in the matrix we will establish a grid of four points (Figure 3).
+Each point of the grid will be 25 m from the patch edge to avoid attracting beetles from the matrix surrounding the patch.
+At each point we will place two 'plant saucers' filled with 3 inches of homogenized local soil in which we place 300 g of cow dung; one of the saucers will be protected with mesh to prevent beetle access.
+After 48 hours the saucers will be removed and the remaining dung will be weighed (the weight of the protected dung is used to correct for weight loss due to desiccation).
+Experiments will be conducted monthly in each landscape for 6 months to capture potential monthly and seasonal (wet/dry) variation.
+
+<!-- -  to consider: gather and sieve soil from outside, place in shallow pot, and place dung on that. Allows breakdown of soil without getting into their soils, if this is a concern -->
+<!-- - repeat biweekly for X months  -->
+<!-- - lab experiment: meso-cosms in GNV based on results from diversity sampling  -->
+<!-- - need to make sure each point inside a plot is independent, and that each sampling date is independent -->
+<!-- plant saucers: http://tinyurl.com/3ph2wd6b -->
+
+![Arrangement of the dung removal experiment to be conducted in each of the seven experimental blocks.](./images/plots3.png){fig-align="center" width=50% height=50%}
 

### 2024-01-29 corridor_docs/proposal/eec_corridor_proposal.Rmd
design added

@@
 # - \fancyhead[R]{p. \thepage}
-# - \fancyhead[L]{ }
-- \fancyhead[L]{Corridor Proposal, E. Escobar-Chena}
-- \fancyhead[R]{p. \thepage}
+- \fancyhead[L]{ }
+# - \fancyhead[L]{Corridor Proposal, E. Escobar-Chena}
+- \fancyhead[R]{Escobar-Chena Corridor Proposal, p. \thepage}
 - \fancyfoot[C]{ }
@@
 - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
+# - \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
+# to center to sub-sub-headings (###) and change color of font
+- \usepackage{sectsty} \subsectionfont{\centering\color{darkmidnightblue}}
+# - \usepackage{sectsty} \subsubsectionfont{\color{darkmidnightblue}}
 - \raggedright
@@
 
-### Research proposal 
+## Research proposal 
 <!-- (a) Description of research, including type of sampling and description of any type of manipulation (even if minor)  -->
 
-**_Background_:**  Corridors are considered an important mechanism for promoting connectivity and facilitating the movement of organisms through fragmented landscapes.
+### Background
+
+Corridors are considered an important mechanism for promoting connectivity and facilitating the movement of organisms through fragmented landscapes.
 These movements are hypothesized to help maintain species diversity in fragments; it could also have patch- and landscape-level consequences for the ecosystem processes in which influenced by these species.
@@
 The Corridor Experiment is an ideal location in which to conduct this project.
-The design overcomes the primary technical impediments to isolating the effects of corridors _per se_, its spatial scale allows for drawing and realistic and relevant conclusions regarding dung beetle dispersal in fragmented landscapes, and the local dung beetle assemblage is a highly tractable one with which to address our questions (i.e., taxonomically and functionally diverse, keys for species identification are readily available, there are multiple abundant species from different functional groups with which to conduct experiments, and the experimental/survey methods are standardized and straightforward).
+The design overcomes the primary technical impediments to isolating the effects of corridors _per se_, its spatial scale allows for drawing and realistic and relevant conclusions regarding dung beetle dispersal in fragmented landscapes, and the local dung beetle assemblage is a highly tractable one with which to address our questions (i.e., taxonomically and functionally diverse, keys for species identification are readily available, there are multiple abundant species from different functional groups with which to conduct experiments, and the experimental/survey methods are straightforward).
 **_Our sampling and experiments will be used to test the following predictions:_**
@@
 
-<!-- 2. Predictions 2: Corridors facilitate inter-patch movement of dung beetle species,  -->
+3. **Prediction 2:** Dung removal rates will be highest in connected patches. This is due to the higher functional diversity of beetles in these locations.
+
 2.
-**Prediction 2:** Corridors are suitable habitat for all species.
+**Prediction 3:** Corridors facilitate the dispersal of all species.
 However, the speed at which in individuals move through corridors and the probability of successful inter-patch movement is size-dependent (i.e., larger beetle species move more quickly and are more likely to reach the connected patch).
 
-3. **Prediction 3:** Dung removal rates will be highest in connected patches. This is due to the higher functional diversity of beetles in these locations.
-
-Below we describe the location of sampling or experiments with which we will test these predictions.
+Below we describe the sampling and experiments with which we will test these predictions in June 2024-March 2025.
+The proposed designs are based on a year-long study of dung beetle community structure we conducted at the University of Florida's Ordway-Swisher Biological Station; any revisions made to the design following preliminary data collection in Spring 2024 will be submitted for review by the Steering Committee.
 
@@
 We propose to conduct the field component of our study in seven of the Savannah River Site's experimental landscapes (i.e., 'blocks') designed to assess the ecological effects of corridors.
-Each experimental landscape block consists of a 1-ha square center patch surrounded by four peripheral patches that are equal in area (~1.4 ha) but differ in the amount of habitat connectivity or edge (Figure 1).
-One peripheral patch is connected to the center patch by a 150 $\times$ 25m corridor (hereafter “connected”).
-The isolated patches are also 150 m from the center patch and vary in their edge amount.
-The winged patch type (hereafter “isolated high edge”) is a 100m $\times$ 100m square with two 75 $\times$ 25 m projections on either side of the patch.
-The rectangular patch type (hereafter “isolated low edge”) consists of a 100m $\times$ 100m square with an area equal in size to the corridor added to the part of the patch farthest from the center patch.
-Each block contains one of each patch type and a duplicate of one of the isolated patch types.
+Each experimental landscape (i.e., "block") consists of a 1-ha square "central patch" surrounded by four peripheral patches.
+All peripheral patches separated from the 150 m from the central patch and have the same area (~1.4 ha), However, the four peripheral patches differ in their level of connectivity to the central patch and edge:area ratio (Figure 1).
+<!-- The "connected patch" is 100m $\times$ 100m and connected to the center patch by a 150m $\times$ 25m corridor.
+The two "winged patches" are 100m $\times$ 100m; each of these patches has two linear projections of 75m $\times$ 25m extending from opposite sides of the patch.
+Finally, the "rectangular patch" is a 100m $\times$ 100m square to which an area equal to that of the corridor/wings has been attached to the edge furthest from the central patch  .
+-->
 
-**_Methods: Beetle Diversity_:** To test the effect of corridors on species richness, species diversity, and functional diversity we will sample the dung beetle community in each location with pitfall traps baited with cow dung.
-Sampling will be conducted in four locations within in each landscape: the "connected" patch, the "rectangular" patch, one of the "winged" patches, and an area of matrix habitat at least 100 m from any of the patches (Fig.
-1).
-In each location we will arrange N = 5 pitfall traps in a grid in the center of the patch; the grid will be 25 m from the edge of the patch to reduce the likelihood of attracting beetles from the matrix habitat into the patch.
-The trap will be baited with 300 mg of cow dung (collected at the University of Florida’s Beef Teaching Unit and frozen until needed) suspended over a cup with 96% ethanol.
-All traps in a landscape will be set on the same day and left open for 24 hours; sampling will be conducted monthly in each landscape for 6 months to capture potential monthly and seasonal (wet/dry) variation in species abundance.
-A collection of voucher specimens will be deposited at the Florida Museum of Natural History, with a reference collection left at the Corridor Project.
+**_Methods - Beetle Diversity_:** To test the effect of corridors on species richness, species diversity, and functional diversity we will sample the dung beetle community in each location with pitfall traps baited with cow dung.
+In each of the seven landscapes we will sample in four locations: the "connected" patch, the "rectangular" patch, one of the "winged" patches, and an area of matrix habitat at least 150 m from any of the patches (Figure 1).
+In each location we will arrange N = 5 traps located	$\geq$ 25m from the patch edge to reduce the likelihood of attracting beetles from the the matrix habitat.
+Each trap will be baited with 300 mg of cow dung (collected at the University of Florida’s Beef Teaching Unit and frozen until needed) suspended over a vial of 95% ethanol.
+The N = 20 traps in a landscape will be set on the same day and remain open for 24 hours; sampling will be conducted monthly for 8 months to capture temporal variation in community structure.
+We will establish a reference collection with identification keys for the Corridor Project and deposit voucher specimens at the Florida Museum of Natural History.
 
@@
 
-![Design for pitfall sampling in each of the seven blocks.](./images/plots1.png){fig-align="center" width=50% height=50%}
+<!-- ![Arrangement of pitfall traps used to sample beetles in experimental blocks.](./images/plots1.png){fig-align="center" width=40% height=40%} -->
 
-**_Methods: Beetle dispersal_:** To determine if beetles use corridors to move between patches, and if patterns of movement differ between species, we will conduct a mark-release-recapture (i.e., MRR) experiment.
-A pitfall trap will be placed in the middle of the landscape's "center" patch and baited with 300 g of cow dung.
-We will then release marked beetles at two points equidistant from the bait: the "connected" patch and the edge of the "rectangular" patch(Figure 2); these locations are separated from the bait by either the corridor or matrix habitat, allowing us to determine if corridors allow for more rapid movement and how this might vary by beetle species or functional group.
-We will check the pitfall trap for marked beetles twice per day for three days following their release.
-
-We anticipate initially conducting this experiment with two species:LARGE AND SMALL, but may modifify rthethe choice of species based on our initial sampling.
-We will also conduct preliomiary expertiments to determine the beetle dispral, and may place passive (i.e., unbaited) traps in the corridor, matrix, and patches.
-
-<!-- - to consider: passive traps to see if they are moving through? should we start experiment with smaller distances into the orridor and then eventually the full length?  -->
-<!-- - is there any technology to track or follow them? RFID? etc? -->
-
-![Design for experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=50% height=50%}
-
-**_Field Methods: Dung removal_:** The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles.
-We will compare the efficacy of dung removal by beetle communities in connected, unconnected, and winged patches using both mesocosms (at the University of Florida) and a field experiment conducted in seven of the experimental blocks.
-In each of the patches and a equally sized plot in the matrix we will establish a grid of four points (Figure 3).
-Each point of the grid will be 25 m from the patch edge to avoid attracting beetles from the matrix surrounding the patch.
-At each point we will place two 'plant saucers' filled with 3 inches of homogenized local soil in which we place 300 g of cow dung; one of the saucers will be protected with mesh to prevent beetle access.
+**_Field Methods - Dung removal_:** The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles.
+We will compare the efficacy of dung removal by beetle communities in connected, unconnected, and winged patches with a field experiment to be conducted in seven of the experimental blocks.
+We first will establish a grid of N = 4 points in each of the patches in which we previously sampled beetle diversity to test Prediction 1 (Figure 2).
+At each point in the grid we will place two 'plant saucers' filled with 3 inches of homogenized local soil in which we place 300 g of cow dung; one of the saucers will be protected with mesh to prevent beetle access.
 After 48 hours the saucers will be removed and the remaining dung will be weighed (the weight of the protected dung is used to correct for weight loss due to desiccation).
-Experiments will be conducted monthly in each landscape for 6 months to capture potential monthly and seasonal (wet/dry) variation.
+Experiments will be conducted monthly in each landscape (N = 6) months to to capture temporal variation in patterns of dung removal.
+This field experiment will be complemented by a mesocosm study with experimentally assembled communities (based on the results of surveys for testing Prediction 1) to be conducted at the University of Florida.
 
@@
 
-![Arrangement of the dung removal experiment to be conducted in each of the seven experimental blocks.](./images/plots3.png){fig-align="center" width=50% height=50%}
+<!-- ![Locations and treatments for the dung removal experiment.](./images/plots3.png){fig-align="center" width=40% height=40%} -->
+
+**_Methods - Beetle dispersal_:** To determine if beetles use corridors to move between patches, and if patterns of movement differ between species based on their size, we will conduct a mark-release-recapture (i.e., MRR) experiment.
+A pitfall trap will be placed in the middle of the landscape's "center" patch and baited with 300 g of cow dung.
+We will then release marked beetles at two points equidistant from the bait: the "connected" patch and the edge of the "rectangular" patch (Figure 3), and monitor the baited pitfall trap for marked beetles twice per day for N = 3 days following their release.
+This design will allow us to to determine (a) if beetles disperse between patches using corridors, (b) if beetles disperse through the matrix, and (c) if beetles dispersing via corridors move more quickly or have higher a higher probability of successful dispersal.
+
+We anticipate conducting this experiment with two locally common and abundant species that differ 70-fold in biomass: _Dochotomius carolinus_ (avg.
+dry biomass = 0.6339 g $\pm$ 0.2454 SD, N=10 individuals) and _Ateuchus lecontei_ (avg dry biomass  = 0.009 g $\pm$ 0.0019 SD, N = 22 individuals).
+We also anticipate conducting preliminary experiments with passive traps (i.e., no dung bait) in the corridor and matrix to determine the optimal number of beetles to release and if a single 'destination trap' will capture sufficient individuals with which to conduct robust analyses of dispersal.
+These will in turn determine the number of landscapes in which we can conduct the experiment and the number of times it can be repeated.
+In additional while we can conduct the experiment with individuals captured in the matrix, we hope to use individuals bred in lab colonies established with individuals captured at the Corridor Site.
+
+<!-- - to consider: passive traps to see if they are moving through? should we start experiment with smaller distances into the orridor and then eventually the full length?  -->
+<!-- - is there any technology to track or follow them? RFID? etc? -->
+
+<!-- ![Design for the experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=40% height=40%} -->
 
@@
 
-_Focal Species Assemblage_
+### Potential impacts on corridor plots and ongoing studies
 
-1. How many species, functional groups, range in size, what is known about them
-1. Table of expected species
+Each pitfall trap requires diffing a cylindrical hole ~10 cm wide $\times$ ~10 cm deep, but the trap's base can remain in place until all surveys are complete (with a cover when not in use).
+This, coupled with the number of traps we are deploying per patch, means that disturbance of plots will be minimal.
+The number of insects collected is also unlikely to have a large or long-term impact on their populations or ecosystem processes.
+The dung removal experiment was also designed to have a minimal impact.
+The amount of dung used in the plot is relatively small, and by conducting the trials with plant saucers placed on the soil surface we greatly reduce the possibility of dung beetle activity disturbing seed banks or altering soil properties (chemistry, structure) in experimental microsites.
+Finally, it is highly unlikely that the dung used in baits will expose local animals to novel disease.
+Any pathogens or parasites that persist in the guts of donor cows despite the efforts of UF's veterinarians, and then survive several months in a freezer, will only be in the field for 24-72 hours.
 
-_Potential impacts on corridor plots and ongoing studies_
+All traps and markings will be removed at the study's conclusion.
 
-1. no holes are being dug for pitfalls: no disturbance of the seeds or soil
-1. changes in dung beetle abundance due to collection: short term, unlikely to be high impact
-1. Dung introduced into plots: contained in above ground traps, so also unlikely, standardized and from cows taking abntibiotics, so no disease introduced,.
+### Study duration
 
-_Study duration_
+1. Preliminary sampling during March-May 2024
+1. Sampling and Dung Removal Experiments: June 2024-December 2024
+1. Dispersal Experiments: January-February 2025
 
-1. Preliminary sampling during Spring 2023
-1. Summer and Fall Sampling
-<!-- Plan for how the study will be cleaned-up -->
-1. All traps and pin flags will be removed after each sampling bout
+### Funding sources
 
-_Funding sources (if applicable)_
+We have much of the equipment necessary to complete the project, and a modest budget to buy new equipment or defer some transportation expenses.
+We are actively seeking funds to support travel and living expenses duering the summer.
 
-_Plan for making data publicly accessible_
+### Plan for making data publicly accessible
 
@@
 For an overview of our approach to data archiving and accessibility see https://github.com/BrunaLab/HeliconiaSurveys
+
+![Arrangement of pitfall traps used to sample beetles in experimental blocks.](./images/plots1.png){fig-align="center" width=40% height=40%}
+
+![Locations and treatments for the dung removal experiment.](./images/plots3.png){fig-align="center" width=40% height=40%}
+
+![Design for the experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=40% height=40%}

### 2024-01-30 corridor_docs/proposal/eec_corridor_proposal.Rmd
update draft

@@
 output:
-  pdf_document: default
+  word_document: default
   html_document:
     df_print: paged
+  pdf_document: default
 geometry: margin=1in
 mainfont: Times New Roman
-# mainfont: SourceSansPro
 fontsize: 12pt
@@
 urlcolor: darkmidnightblue
+indent: true
 header-includes:
@@
 - \pagestyle{fancy}
-
-# - \fancyhead[R]{p. \thepage}
 - \fancyhead[L]{ }
-# - \fancyhead[L]{Corridor Proposal, E. Escobar-Chena}
-- \fancyhead[R]{Escobar-Chena Corridor Proposal, p. \thepage}
+- "\\fancyhead[R]{Escobar-Chena Corridor Proposal, p. \\thepage}"
 - \fancyfoot[C]{ }
-# - \fancyhead[L]{updated `r Sys.Date()`}
 - \fancyfoot[R]{updated `r Sys.Date()`}
-# - \usepackage[default]{sourcesanspro}
 - \AtBeginDocument{\let\maketitle\relax}
 - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
-# - \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
-# to center to sub-sub-headings (###) and change color of font
-- \usepackage{sectsty} \subsectionfont{\centering\color{darkmidnightblue}}
-# - \usepackage{sectsty} \subsubsectionfont{\color{darkmidnightblue}}
+- \usepackage{sectsty} \subsectionfont{\centering}
 - \raggedright
+- \usepackage{indentfirst}
+- \setlength\parindent{24pt}
 ---
@@
 library(knitr)
+library(here)
 ```
@@
 
-Corridors are considered an important mechanism for promoting connectivity and facilitating the movement of organisms through fragmented landscapes.
-These movements are hypothesized to help maintain species diversity in fragments; it could also have patch- and landscape-level consequences for the ecosystem processes in which influenced by these species.
-Although there is some evidence that animals disperse between patches via corridors, and that connected patches have higher species diversity than unconnected ones, little work to date has investigated the consequences of these corridor-driven patterns for ecosystem ecosystem.
+Corridors have been hypothesized to be an important mechanism for facilitating the movement of organisms through fragmented landscapes.
+These movements are hypothesized to prevent species diversity from declining in fragments, as well help maintain the ecosystem services provided by these species (at both the patch- and landscape-level).
+Although there is some evidence that animals disperse between patches via corridors, and that connected patches have higher species diversity than unconnected ones, little work to date has investigated the consequences of these corridor-driven patterns for ecosystem services.
 
-Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity.
-They are locally species rich, exhibit variety in key functional traits (e.g., size, foraging style, resource-use), are capable of long-distance dispersal (e.g., up to 1 km for some tropical species), and that they influence a number of critical ecosystem services (e.g., seed dispersal, nutrient cycling).
-Previous work has shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors.
-However, it remains unknown if corridors can reduce the loss of species from fragments, how such declines are influenced by inter-specific differences in dispersal capability, and what the consequences of these patterns are for the ecosystems services provided by dung beetles.
-Though there are several reasons for this, one major factor is the challenge in finding locations where one can assess the role of corridors while also while controlling for confounding factors such as patch size, edge, and corridor length.
+Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity and their ecosystem services.
+Local assemblages can be species rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use).
+There is evidence they are capable of long-distance dispersal -- up to 1 km for some tropical species -- and they influence a number of critical ecosystem services ranging from seed dispersal to nutrient cycling.
+Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors.
+However, it remains unknown if corridors actually act to reduce the loss of dung beetle species from fragments, if such declines are influenced to inter-specific differences in dispersal capability, and what the consequences of these patterns are for the ecosystems services they provide.
+One major factor behind this lack of information is the challenge in finding locations where one can assess the role of corridors while also while controlling for confounding factors such as patch size, edge, and corridor length.
 
-**We propose using the community of dung beetles at the SRS Corridor Experiment to test for** **(1)** higher species richness, species diversity, and functional diversity in patches connected by corridors than in unconnected patches **(2)** interspecific differences in corridor use and inter-patch movement, and **(3)** higher rates of dung removal in connected than unconnected patches.
-
-The Corridor Experiment is an ideal location in which to conduct this project.
-The design overcomes the primary technical impediments to isolating the effects of corridors _per se_, its spatial scale allows for drawing and realistic and relevant conclusions regarding dung beetle dispersal in fragmented landscapes, and the local dung beetle assemblage is a highly tractable one with which to address our questions (i.e., taxonomically and functionally diverse, keys for species identification are readily available, there are multiple abundant species from different functional groups with which to conduct experiments, and the experimental/survey methods are straightforward).
-**_Our sampling and experiments will be used to test the following predictions:_**
+**We propose using the community of dung beetles at the SRS Corridor Experiment to test the following predictions:**
 
@@
 
-Below we describe the sampling and experiments with which we will test these predictions in June 2024-March 2025.
-The proposed designs are based on a year-long study of dung beetle community structure we conducted at the University of Florida's Ordway-Swisher Biological Station; any revisions made to the design following preliminary data collection in Spring 2024 will be submitted for review by the Steering Committee.
+**_The Corridor Experiment is an ideal location in which to conduct this project._** Its design overcomes the primary technical impediments to isolating the effects of corridors _per se_, and the spatial scale allows for drawing and realistic and relevant conclusions regarding dung beetle dispersal in fragmented landscapes.
+In addition the local dung beetle community is a highly tractable one with which to address our questions: the ~16 species span a broad range of abundance and functional traits, their taxonomy is well resolved, their identification is straightforward, and methods surveying their diversity and using them in experiments are well-established.
+Below we describe the sampling and experiments with which we will test our predictions.
+The proposed designs are informed by a year-long study of dung beetle community structure we conducted at the University of Florida's Ordway-Swisher Biological Station; any revisions made to the design following preliminary data collection in Spring 2024 will be submitted for review by the Steering Committee.
 
@@
 
-<!-- ![Arrangement of pitfall traps used to sample beetles in experimental blocks.](./images/plots1.png){fig-align="center" width=40% height=40%} -->
+<!-- ![Arrangement of pitfall traps used to sample beetles in experimental blocks.](./images/plots1.png){fig-align="center" width=45% height=45%} -->
+
+<!-- ```{r echo=FALSE, out.width = "45%", fig.align = "center"} -->
+
+<!-- knitr::include_graphics(here("corridor_docs","proposal","images", "plots1.png")) -->
+<!-- ``` -->
 
@@
 
-<!-- ![Locations and treatments for the dung removal experiment.](./images/plots3.png){fig-align="center" width=40% height=40%} -->
+<!-- ![Locations and treatments for the dung removal experiment.](./images/plots3.png){fig-align="center" width=45% height=45%,fig.pos="H"} -->
+
+<!-- ```{r fig1, echo=FALSE, fig.align="center", out.width = "100%",fig.cap='TEXT'} -->
+<!-- p1<-cowplot::ggdraw() +cowplot::draw_image("./corridor_docs/proposal/images/plots3.png") -->
+<!-- p1 -->
+<!-- ``` -->
 
@@
 
-<!-- ![Design for the experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=40% height=40%} -->
+<!-- ![Design for the experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=45% height=45%,fig.pos="H"} -->
 
@@
 We have much of the equipment necessary to complete the project, and a modest budget to buy new equipment or defer some transportation expenses.
-We are actively seeking funds to support travel and living expenses duering the summer.
+We are actively seeking funds to support travel and living expenses during the summer.
 
@@
 
-![Arrangement of pitfall traps used to sample beetles in experimental blocks.](./images/plots1.png){fig-align="center" width=40% height=40%}
+\newpage
 
-![Locations and treatments for the dung removal experiment.](./images/plots3.png){fig-align="center" width=40% height=40%}
+<!-- ## Figures -->
 
-![Design for the experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=40% height=40%}
+![Arrangement of pitfall traps used to sample beetles in experimental blocks.](./images/plots1.png){fig-align="center" width=45% height=45%}
+
+![Locations and treatments for the dung removal experiment.](./images/plots3.png){fig-align="center" width=45% height=45%}
+
+![Design for the experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=45% height=45%}

### 2024-02-08 corridor_docs/ideas/thesis_ideas.tex
repo reorg

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont[]{SourceSansPro}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyfoot[C]{ }
\fancyhead[R]{p. \thepage}
\fancyhead[L]{last update 2024-02-02}
\usepackage[default]{sourcesanspro}
\AtBeginDocument{\let\maketitle\relax}
\definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Project Ideas and Notes},
  pdfauthor={Eric Escobar-Chena and Emilio M. Bruna},
  colorlinks=true,
  linkcolor={darkmidnightblue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={darkmidnightblue},
  pdfcreator={LaTeX via pandoc}}

\title{Project Ideas and Notes}
\author{Eric Escobar-Chena and Emilio M. Bruna}
\date{updated: 2024-02-02}

\begin{document}
\maketitle

\hypertarget{notes-from-conversation-with-julian}{%
\subsubsection{\texorpdfstring{\emph{Notes from conversation with
Julian:}}{Notes from conversation with Julian:}}\label{notes-from-conversation-with-julian}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Low-hanging fruit: Effect of connectivity on (a) species and
  functional group diversity and abundance
\item
  Next level: Movement \& Dispersal. There are large areas of forest
  where it is possible to capture beetles for Mark-Release-Recapture
  (MRR) experiments.
\item
  Higher-risk, Higher reward (even if they don't go in MS, can be set-up
  for potential PHD projects): Experiments on Ecosystem Services
\end{enumerate}

\begin{itemize}
\tightlist
\item
  dung burial, decomposition, soil properties based on results of 1 \& 2
  (buckets with dung and beetle assemblages)
\item
  gas (e.g., methane) emissions
\item
  seed dispersal/burial/germination
\end{itemize}

\hypertarget{example-studies}{%
\subsubsection{Example studies}\label{example-studies}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Diversity and Abundance
\end{enumerate}

\begin{itemize}
\tightlist
\item
  Estrada and Coates-Estrada (2002): ``56\% of individuals were captured
  in the continuous forest, 29\% in the mosaic habitat and 15\% in the
  forest fragments''
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\item
  Movement \& Dispersal
\item
  Ecosystem Services
\end{enumerate}

\hypertarget{other-stuff-to-work-on}{%
\section{Other stuff to work on:}\label{other-stuff-to-work-on}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Species List \& Keys for Dung Beetles of the Southeastern US
\item
  Any previous work done on dung beetles in Southeastern US
\item
  List of Equipment and Tools needed
\end{enumerate}

\begin{itemize}
\tightlist
\item
  Dung beetle traps
\item
  alcohol
\item
  bait
\end{itemize}

\hypertarget{species}{%
\section{Species}\label{species}}

\newpage

\hypertarget{introduction}{%
\section{INTRODUCTION}\label{introduction}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Paragraph 1: In an increasingly fragmented world, corridors are
  thought to be an important strategy of connecting isolated patches to
  promote population persistence and increase diversity in connected
  patches. This can ultimately have consequences for ecosystem processes
  in fragments.
\item
  Paragraph 2: Despite this, direct evidence of movement between patches
  is limited. Plenty of studies have demonstrated that the diversity or
  abundance of organisms is similar in corridors and the primary habitat
  to which they are connected. However, studies demonstrating animals
  actually move from one patch through a corridor to another patch are
  surprisingly rare (but see examples). NB: might have to say ``although
  there is ample evidence that vertbrates move through corridors,
  studies documenting the movements of invertebrates - the dominant
  species on earth - are limited''.
\item
  Paragraph 3: Part of the challenge in assessing the efficacy of
  corridors is the confounding effects of patch area and the length of
  habitat edges (see resasco 2014 etc); to address this it is necessary
  to compare connected and unconnected areas of equal area and shape.
\item
  Paragraph 4: Dung beetles have emerged as a model system with which to
  test hypotheses on how anthropogenic landscape alterations influence
  biodiversity. They are locally species rich, exhibit variety in key
  funcitonal traits (size, foraging style,,\ldots.), and they provide
  critical ecosystem services. robust (but not \emph{too} robust):
  studies have shown diversity of dung beetles can be lower in
  fragments, but that they can be found in corridors and are capable of
  dispersing for up to 1 km.
\item
  Paragraph 5: I am proposing to use a landscape-scale fragmentation
  experiment to test the effects of corridors on the diversity and
  dispersal dung beetles and the ecosystem services they provide.
  Specifically, I will test the following predictions:
\item
  Diversity will be higher in connected patches. To test this
  prediction, I will sample dung beetles with pitfall traps and use
  non-parametric estimators of species diversity
\item
  Dung beetles disperse through corridors, with larger beetles
  dispersing more quickly (size-dependent dispersal ability)
\item
  Dung removal rates will be highest in connected patches.
\end{enumerate}

To test there predictions, I will use a combination of passive sampling
and a MRR experiment and control for the confounding effects of edge and
area.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Prediction 1: Pitfall trapping (arrangement, timing)

Prediction 2: MRR experiment, where will recapture traps be placed?

Prediction 3: experimental design poop disappearance

\newpage

\hypertarget{what-is-predicted-about-how-corridors-influence-the-diversity-of-insects-communities-with-an-emphasis-on-dung-beetles-in-patches}{%
\subsection{What is predicted about how corridors influence the
diversity of insects communities (with an emphasis on Dung Beetles) in
patches?}\label{what-is-predicted-about-how-corridors-influence-the-diversity-of-insects-communities-with-an-emphasis-on-dung-beetles-in-patches}}

Are corridors predicted to enhance diversity in connected patches? What
is the mechanism\textgreater? A certain species groups are predicted to
drive this change in diversity - rare species, larger species,
specialists, particular function groups, better fliers, etc.?
(\emph{Note: ``Corridors'' = Riparian strips, living fences, linear
fragments, etc.})

\begin{itemize}
\item
  \textbf{Prediction 1:} Diversity higher in connected patches
  (citations in Damschen et al PNAS, others)
\item
  \textbf{Prediction 2:} Diversity could be lower in patches (in
  specific case of corridor project, only getting open space species)
\end{itemize}

\hypertarget{empiricial-results}{%
\subsubsection{Empiricial Results}\label{empiricial-results}}

\hypertarget{studies-that-have-compared-diversity-of-dung-beetles-in-corridors-and-other-habitats-e.g.-primary-forest-pastures-have-found-that}{%
\subsubsection{1. Studies that have compared diversity of Dung Beetles
in Corridors and Other Habitats (e.g., primary forest, pastures) have
found
that:}\label{studies-that-have-compared-diversity-of-dung-beetles-in-corridors-and-other-habitats-e.g.-primary-forest-pastures-have-found-that}}

\begin{itemize}
\item
  \textbf{Diversity increases:} (Schalkwyk, Pryke, and Samways 2017;
  Hill 1995)
\item
  \textbf{Diversity decreases:}
\item
  \textbf{Similar diversity in both:}
\end{itemize}

\hypertarget{studies-that-have-actually-compared-the-diversity-of-patches-that-are-connected-by-a-corridor-vs.-patches-that-are-unconnected}{%
\subsubsection{2. Studies that have actually compared the diversity of
patches that are connected by a corridor vs.~patches that are
unconnected}\label{studies-that-have-actually-compared-the-diversity-of-patches-that-are-connected-by-a-corridor-vs.-patches-that-are-unconnected}}

\begin{itemize}
\item
  \textbf{Diversity increases:}
\item
  \textbf{Diversity decreases:}
\item
  \textbf{Similar diversity in both:}
\end{itemize}

\hypertarget{what-is-predicted-or-known-about-insect-movement-through-corridors-with-an-emphasis-on-dung-beetles}{%
\subsection{What is predicted or known about insect movement through
corridors (with an emphasis on Dung
Beetles)?}\label{what-is-predicted-or-known-about-insect-movement-through-corridors-with-an-emphasis-on-dung-beetles}}

Do Corridors facilitate movement between patches? Are particular species
or groups more able to move through corridors than others - larger
species, specialists on particular resources, particular functional
groups, better fliers, etc.?

\hypertarget{studies-that-have-assessed-movement-of-dung-beetles-have-found-the-following-can-play-an-important-role}{%
\subsubsection{1. Studies that have assessed movement of Dung Beetles
have found the following can play an important
role:}\label{studies-that-have-assessed-movement-of-dung-beetles-have-found-the-following-can-play-an-important-role}}

\begin{itemize}
\item
  \textbf{Habitat Preference:} Arellano, Leon-Cortes, and Halffter
  (2008).
\item
  \textbf{Wing Loading:} (Cultid-Medina et al. 2015)
\item
  \textbf{Sex:}
\item
  \textbf{Foraging Mode:}
\item
  \textbf{Size:}
\item
  \textbf{Invasive vs.~Native:}
\item
  \textbf{No inter-specific or inter-group differences in movement:}
\end{itemize}

\hypertarget{studies-that-have-estimated-dispersal-distance}{%
\subsubsection{2. Studies that have estimated dispersal
distance:}\label{studies-that-have-estimated-dispersal-distance}}

\newpage

\hypertarget{the-questions}{%
\section{THE QUESTIONS}\label{the-questions}}

Question 1: Is dung beetle abundance greater in connected patches than
isolated ones?

Question 2: Is dung beetle diversity greater in connected patches than
isolated ones?

\begin{quote}
Prediction: Species diversity is greater in connected patches than
isolated ones (due to the greater representation of rare species) OR
Prediction: Diversity of \_\_\_\_\_\_\_ (e.g., specialists, foraging
strategy X, etc.) is greater in connected patches, but the Diversity of
\_\_\_\_\_\_\_\_ (e.g., generalists, foraging strategy Y) will be the
similar in both. OR The diversity is similar, but the community
composition changes to one dominated by specialists instead of
generalists
\end{quote}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Do Dunge beetles move through corridors?
\end{enumerate}

References

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-WOS:000262759300008}{}}%
Arellano, Lucrecia, Jorge L. Leon-Cortes, and Gonzalo Halffter. 2008.
{``Response of Dung Beetle Assemblages to Landscape Structure in Remnant
Natural and Modified Habitats in Southern {Mexico}.''} \emph{INSECT
CONSERVATION AND DIVERSITY} 1 (4): 253--62.
\url{https://doi.org/10.1111/j.1752-4598.2008.00033.x}.

\leavevmode\vadjust pre{\hypertarget{ref-WOS:000360850000002}{}}%
Cultid-Medina, Carlos A., Bedir G. Martinez-Quintero, Federico Escobar,
and Patricia Chacon de Ulloa. 2015. {``Movement and Population Size of
Two Dung Beetle Species in an {Andean} Agricultural Landscape Dominated
by Sun-Grown Coffee.''} \emph{JOURNAL OF INSECT CONSERVATION} 19 (4):
617--26. \url{https://doi.org/10.1007/s10841-015-9784-3}.

\leavevmode\vadjust pre{\hypertarget{ref-WOS:000178929600002}{}}%
Estrada, A, and R Coates-Estrada. 2002. {``Dung Beetles in Continuous
Forest, Forest Fragments and in an Agricultural Mosaic Habitat Island at
{Los} {Tuxtlas}, {Mexico}.''} \emph{BIODIVERSITY AND CONSERVATION} 11
(11): 1903--18. \url{https://doi.org/10.1023/A:1020896928578}.

\leavevmode\vadjust pre{\hypertarget{ref-WOS:A1995TL67200023}{}}%
Hill, CJ. 1995. {``Linear Strips of Rain Forest Vegetation as Potential
Dispersal Corridors for Rain Forest Insects.''} \emph{CONSERVATION
BIOLOGY} 9 (6): 1559--66.
\url{https://doi.org/10.1046/j.1523-1739.1995.09061559.x}.

\leavevmode\vadjust pre{\hypertarget{ref-WOS:000399235300014}{}}%
Schalkwyk, J. van, J. S. Pryke, and M. J. Samways. 2017. {``Wide
Corridors with Much Environmental Heterogeneity Best Conserve High Dung
Beetle and Ant Diversity.''} \emph{BIODIVERSITY AND CONSERVATION} 26
(5): 1243--56. \url{https://doi.org/10.1007/s10531-017-1299-7}.

\end{CSLReferences}

\end{document}

### 2024-02-08 corridor_docs/site_steering_cmtee_proposal/eec_corridor_proposal.Rmd
repo reorg

--- 2024-01-30 corridor_docs/proposal/eec_corridor_proposal.Rmd
@@
 date: 'updated: `r Sys.Date()`'
-output:
-  word_document: default
-  html_document:
-    df_print: paged
-  pdf_document: default
+output: pdf_document
+  # pdf_document: default
+  # html_document:
+  #   df_print: paged
+  # word_document: default
 geometry: margin=1in
@@
 fontsize: 12pt
+indent: true
 linkcolor: darkmidnightblue
 urlcolor: darkmidnightblue
-indent: true
 header-includes:
@@
 - \pagestyle{fancy}
-- \fancyhead[L]{ }
-- "\\fancyhead[R]{Escobar-Chena Corridor Proposal, p. \\thepage}"
+- \fancyhead[L]{}
+- \fancyhead[R]{Escobar-Chena, p. \thepage}
 - \fancyfoot[C]{ }
@@
 ```
-
 ## Influence of corridors on dung beetle diversity, dispersal, & ecosystem services
@@
 
+- Eric Escobar-Chena (eescobarchena@ufl.edu)
 - Emilio Bruna (embruna@ufl.edu)
-- Eric Escobar-Chena (eescobarchena@ufl.edu)
 
@@
 
-Corridors have been hypothesized to be an important mechanism for facilitating the movement of organisms through fragmented landscapes.
-These movements are hypothesized to prevent species diversity from declining in fragments, as well help maintain the ecosystem services provided by these species (at both the patch- and landscape-level).
+Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes.
+These movements are hypothesized to prevent species diversity from declining in fragments, as well as help maintain the ecosystem services provided by these species (at both the patch- and landscape-level).
 Although there is some evidence that animals disperse between patches via corridors, and that connected patches have higher species diversity than unconnected ones, little work to date has investigated the consequences of these corridor-driven patterns for ecosystem services.
@@
 Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity and their ecosystem services.
-Local assemblages can be species rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use).
-There is evidence they are capable of long-distance dispersal -- up to 1 km for some tropical species -- and they influence a number of critical ecosystem services ranging from seed dispersal to nutrient cycling.
+The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles such as enhanced nutrient cycling and soil quality and reduction of parasites on methane emissions from dung.
+Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use).
 Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors.
-However, it remains unknown if corridors actually act to reduce the loss of dung beetle species from fragments, if such declines are influenced to inter-specific differences in dispersal capability, and what the consequences of these patterns are for the ecosystems services they provide.
+However, it remains unknown if corridors actually act to reduce the loss of dung beetle species from fragments, if such declines are influenced by inter-specific differences in dispersal capability, and what the consequences of these patterns are for the ecosystems services they provide.
 One major factor behind this lack of information is the challenge in finding locations where one can assess the role of corridors while also while controlling for confounding factors such as patch size, edge, and corridor length.
@@
 
-1. **Prediction 1:** Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches
+1. **Prediction 1:** Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches.
 
@@
 
-**_The Corridor Experiment is an ideal location in which to conduct this project._** Its design overcomes the primary technical impediments to isolating the effects of corridors _per se_, and the spatial scale allows for drawing and realistic and relevant conclusions regarding dung beetle dispersal in fragmented landscapes.
-In addition the local dung beetle community is a highly tractable one with which to address our questions: the ~16 species span a broad range of abundance and functional traits, their taxonomy is well resolved, their identification is straightforward, and methods surveying their diversity and using them in experiments are well-established.
+<!-- **_The Corridor Experiment is an ideal location in which to conduct this project._** Its design overcomes the primary technical impediments to isolating the effects of corridors _per se_, and the spatial scale allows for drawing and realistic and relevant conclusions regarding dung beetle dispersal in fragmented landscapes.
+In addition, the local dung beetle community is a highly tractable one with which to address our questions: the ~16 species span a broad range of abundance and functional traits, their taxonomy is well resolved, their identification is straightforward, and methods surveying their diversity and using them in experiments are well-established.
 Below we describe the sampling and experiments with which we will test our predictions.
 The proposed designs are informed by a year-long study of dung beetle community structure we conducted at the University of Florida's Ordway-Swisher Biological Station; any revisions made to the design following preliminary data collection in Spring 2024 will be submitted for review by the Steering Committee.
+-->
 
@@
 We propose to conduct the field component of our study in seven of the Savannah River Site's experimental landscapes (i.e., 'blocks') designed to assess the ecological effects of corridors.
-Each experimental landscape (i.e., "block") consists of a 1-ha square "central patch" surrounded by four peripheral patches.
+Each experimental landscape (i.e., "block") consists of a 1-ha square central patch surrounded by four peripheral patches.
 All peripheral patches separated from the 150 m from the central patch and have the same area (~1.4 ha), However, the four peripheral patches differ in their level of connectivity to the central patch and edge:area ratio (Figure 1).
@@
 
-**_Field Methods - Dung removal_:** The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles.
-We will compare the efficacy of dung removal by beetle communities in connected, unconnected, and winged patches with a field experiment to be conducted in seven of the experimental blocks.
+**_Field Methods - Dung removal_:** We will compare the efficacy of dung removal by beetle communities in connected, unconnected, and winged patches with a field experiment to be conducted in seven of the experimental blocks.
 We first will establish a grid of N = 4 points in each of the patches in which we previously sampled beetle diversity to test Prediction 1 (Figure 2).
 At each point in the grid we will place two 'plant saucers' filled with 3 inches of homogenized local soil in which we place 300 g of cow dung; one of the saucers will be protected with mesh to prevent beetle access.
-After 48 hours the saucers will be removed and the remaining dung will be weighed (the weight of the protected dung is used to correct for weight loss due to desiccation).
-Experiments will be conducted monthly in each landscape (N = 6) months to to capture temporal variation in patterns of dung removal.
+After 48 hours the saucers will be removed and the remaining dung will be weighed (the weight of the protected dung is used to correct for weight loss from desiccation).
+Experiments will be conducted monthly in each landscape (N = 6) months to capture temporal variation in patterns of dung removal.
 This field experiment will be complemented by a mesocosm study with experimentally assembled communities (based on the results of surveys for testing Prediction 1) to be conducted at the University of Florida.
@@
 A pitfall trap will be placed in the middle of the landscape's "center" patch and baited with 300 g of cow dung.
-We will then release marked beetles at two points equidistant from the bait: the "connected" patch and the edge of the "rectangular" patch (Figure 3), and monitor the baited pitfall trap for marked beetles twice per day for N = 3 days following their release.
-This design will allow us to to determine (a) if beetles disperse between patches using corridors, (b) if beetles disperse through the matrix, and (c) if beetles dispersing via corridors move more quickly or have higher a higher probability of successful dispersal.
+We will then release marked beetles at two points equidistant from the bait -- the "connected" patch and the edge of the "rectangular" patch (Figure 3) -- and monitor the baited pitfall trap for N = 3 days.
+This design will allow us to determine (a) if beetles disperse between patches using corridors, (b) if beetles disperse through the matrix, and (c) if beetles dispersing via corridors move more quickly or have higher a higher probability of successful dispersal.
 
 We anticipate conducting this experiment with two locally common and abundant species that differ 70-fold in biomass: _Dochotomius carolinus_ (avg.
-dry biomass = 0.6339 g $\pm$ 0.2454 SD, N=10 individuals) and _Ateuchus lecontei_ (avg dry biomass  = 0.009 g $\pm$ 0.0019 SD, N = 22 individuals).
-We also anticipate conducting preliminary experiments with passive traps (i.e., no dung bait) in the corridor and matrix to determine the optimal number of beetles to release and if a single 'destination trap' will capture sufficient individuals with which to conduct robust analyses of dispersal.
-These will in turn determine the number of landscapes in which we can conduct the experiment and the number of times it can be repeated.
-In additional while we can conduct the experiment with individuals captured in the matrix, we hope to use individuals bred in lab colonies established with individuals captured at the Corridor Site.
+dry biomass = 0.634 g $\pm$ 0.245 SD, N=10 individuals) and _Ateuchus lecontei_ (avg dry biomass  = 0.009 g $\pm$ 0.002 SD, N = 22 individuals).
+The individuals used in the experiment will be captured locally in matrix habitat; our prior sampling in Florida indicates these species are among the most abundant in Pine savanna.
+While we anticipate field collection will yield sufficient beetles to conduct a robust MRR analysis, we will also attempt to establish breeding colonies at UF.
 
+As part of our efforts to determine the necessary sample sizes of individuals to use in trials, we will conduct preliminary dispersal trials in both the corridor and matrix with grids of passive traps (i.e., no dung bait).
+This will also help us determine if a single 'destination trap' is sufficient.
+These preliminary results will in turn determine the number of landscapes in which we can conduct the experiment and how many times it will be repeated.
 <!-- - to consider: passive traps to see if they are moving through? should we start experiment with smaller distances into the orridor and then eventually the full length?  -->
@@
 
-Each pitfall trap requires diffing a cylindrical hole ~10 cm wide $\times$ ~10 cm deep, but the trap's base can remain in place until all surveys are complete (with a cover when not in use).
+Each pitfall trap requires digging a cylindrical hole ~10 cm wide $\times$ ~10 cm deep, but the trap's base can remain in place until all surveys are complete (with a cover when not in use).
 This, coupled with the number of traps we are deploying per patch, means that disturbance of plots will be minimal.
@@
 1. Preliminary sampling during March-May 2024
-1. Sampling and Dung Removal Experiments: June 2024-December 2024
-1. Dispersal Experiments: January-February 2025
+1. Sampling and Dung Removal Experiments: June 2024-March 2025
+1. Dispersal Experiments: August-September 2024
 
@@
 
-\newpage
+<!-- \newpage -->
 
-<!-- ## Figures -->
+## Figures
 

### 2024-02-08 corridor_docs/masters_cmtee_proposal/eec_ms_proposal.Rmd
link to .bib

--- 2024-02-08 corridor_docs/site_steering_cmtee_proposal/eec_corridor_proposal.Rmd
@@
 date: 'updated: `r Sys.Date()`'
+bibliography: "`r here::here('corridor_docs', 'eric-thesis.bib')`"
 output: pdf_document
@@
 - \setlength\parindent{24pt}
+
 ---
@@
 
-Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes.
+Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes .
 These movements are hypothesized to prevent species diversity from declining in fragments, as well as help maintain the ecosystem services provided by these species (at both the patch- and landscape-level).

### 2024-02-23 corridor_docs/masters_cmtee_proposal/eec_ms_proposal.Rmd
added some citations.

--- 2024-02-08 corridor_docs/site_steering_cmtee_proposal/eec_corridor_proposal.Rmd
@@
 date: 'updated: `r Sys.Date()`'
+bibliography: "eric-thesis.bib"
 output: pdf_document
@@
 library(here)
+# knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
 ```
@@
 
-Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes.
-These movements are hypothesized to prevent species diversity from declining in fragments, as well as help maintain the ecosystem services provided by these species (at both the patch- and landscape-level).
+Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes [@haddad_corridor_2003].
+These movements are hypothesized to prevent species diversity from declining in fragments, as well as help maintain the ecosystem services provided by these species (at both the patch- and landscape-level)[@burt_ants_2022].
 Although there is some evidence that animals disperse between patches via corridors, and that connected patches have higher species diversity than unconnected ones, little work to date has investigated the consequences of these corridor-driven patterns for ecosystem services.
 
-Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity and their ecosystem services.
-The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles such as enhanced nutrient cycling and soil quality and reduction of parasites on methane emissions from dung.
-Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use).
-Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors.
+Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity and their ecosystem services[@ros_how_2012; @roslin_dung_2000].
+The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles such as enhanced nutrient cycling and soil quality and reduction of parasites on methane emissions from dung[@slade_disentangling_2016; @WOS:000350444300013].
+Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use)[@WOS:000891747700001].
+Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors[@WOS:000707546700001].
 However, it remains unknown if corridors actually act to reduce the loss of dung beetle species from fragments, if such declines are influenced by inter-specific differences in dispersal capability, and what the consequences of these patterns are for the ecosystems services they provide.
-One major factor behind this lack of information is the challenge in finding locations where one can assess the role of corridors while also while controlling for confounding factors such as patch size, edge, and corridor length.
+One major factor behind this lack of information is the challenge in finding locations where one can assess the role of corridors while also while controlling for confounding factors such as patch size, edge, and corridor length[@haddad_habitat_2015].
 
@@
 
+[@WOS:000341406900011]
+
 ### Study duration
@@
 
-![Arrangement of pitfall traps used to sample beetles in experimental blocks.](./images/plots1.png){fig-align="center" width=45% height=45%}
+```{r fig1, echo=FALSE, fig.cap="Arrangement of pitfall traps used to sample beetles in experimental blocks.", fig.align="center", out.width='45%'}
 
-![Locations and treatments for the dung removal experiment.](./images/plots3.png){fig-align="center" width=45% height=45%}
+knitr::include_graphics("images/plots1.png")
 
-![Design for the experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=45% height=45%}
+```
+
+<!-- ![Arrangement of pitfall traps used to sample beetles in experimental blocks.](./images/plots1.png){fig-align="center" width=45% height=45%} -->
+
+<!-- ![Locations and treatments for the dung removal experiment.](./images/plots3.png){fig-align="center" width=45% height=45%} -->
+
+<!-- ![Design for the experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=45% height=45%} -->
+
+## References

### 2024-03-04 corridor_docs/masters_cmtee_proposal/eec_ms_proposal.Rmd
method citations added

@@
 
-We propose to conduct the field component of our study in seven of the Savannah River Site's experimental landscapes (i.e., 'blocks') designed to assess the ecological effects of corridors.
+We propose to conduct the field component of our study in seven of the Savannah River Site's experimental landscapes (i.e., 'blocks') designed to assess the ecological effects of corridors[@tewksbury_corridors_2002].
 Each experimental landscape (i.e., "block") consists of a 1-ha square central patch surrounded by four peripheral patches.
@@
 At each point in the grid we will place two 'plant saucers' filled with 3 inches of homogenized local soil in which we place 300 g of cow dung; one of the saucers will be protected with mesh to prevent beetle access.
-After 48 hours the saucers will be removed and the remaining dung will be weighed (the weight of the protected dung is used to correct for weight loss from desiccation).
+After 48 hours the saucers will be removed and the remaining dung will be weighed (the weight of the protected dung is used to correct for weight loss from desiccation)[@noriega_dung_2023].
 Experiments will be conducted monthly in each landscape (N = 6) months to capture temporal variation in patterns of dung removal.
-This field experiment will be complemented by a mesocosm study with experimentally assembled communities (based on the results of surveys for testing Prediction 1) to be conducted at the University of Florida.
+This field experiment will be complemented by a mesocosm study with experimentally assembled communities (based on the results of surveys for testing Prediction 1) to be conducted at the University of Florida[@WOS:000311855000018].
 

### 2024-03-28 corridor_docs/ideas/thesis_ideas.Rmd
repo reorg
added ideas from badging trip

--- 2023-11-13 corridor_docs/thesis_ideas.Rmd
@@
 - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
-bibliography: "eric-thesis.bib"
+bibliography: "`r here::here('corridor_docs', 'eric-thesis.bib')`"
 output: pdf_document
@@
 ```
-
+insert_citation
 ### *Notes from conversation with Julian:*
@@
 
-INTRODUCTION
+# INTRODUCTION
 
@@
 This can ultimately have consequences for ecosystem processes in fragments.
+
 2.
@@
 NB: might have to say "although there is ample evidence that vertbrates move through corridors, studies documenting the movements of invertebrates - the dominant species on earth - are limited".
+
 3.
 Paragraph 3: Part of the challenge in assessing the  efficacy of corridors is the confounding effects of patch area and the length of habitat edges (see resasco 2014 etc); to address this it is necessary to compare connected and unconnected areas of equal area and shape.
+
 4.
@@
 robust (but not *too* robust): studies have shown diversity of dung beetles can be lower in fragments, but that they can be found in corridors and are capable of dispersing for up to 1 km.
+
 5.
@@
 
-1. Diversity will be higher in connected patches
+1. Diversity will be higher in connected patches. To test this prediction, I will sample dung beetles with pitfall traps and use non-parametric estimators of species diversity
 
@@
 To test there predictions, I will use a combination of passive sampling and a MRR experiment and control for the confounding effects of edge and area.
+
+# Methods
+
+Prediction 1: Pitfall trapping (arrangement, timing)
+
+Prediction 2: MRR experiment, where will recapture traps be placed?
+
+Prediction 3: experimental design poop disappearance
 

### 2024-04-04 corridor_docs/masters_cmtee_proposal/eec_ms_proposal.Rmd

@@
 date: 'updated: `r Sys.Date()`'
-bibliography: "eric-thesis.bib"
+bibliography: "`r here::here('corridor_docs', 'eric-thesis.bib')`"
+
 output: pdf_document
@@
 
-Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity and their ecosystem services[@ros_how_2012; @roslin_dung_2000].
-The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles such as enhanced nutrient cycling and soil quality and reduction of parasites on methane emissions from dung[@slade_disentangling_2016; @WOS:000350444300013].
-Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use)[@WOS:000891747700001].
-Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors[@WOS:000707546700001].
+Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity and their ecosystem services [@ros_how_2012; @roslin_dung_2000].
+The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles such as enhanced nutrient cycling and soil quality and reduction of parasites on methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
+Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) [@WOS:000891747700001].
+Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].
 However, it remains unknown if corridors actually act to reduce the loss of dung beetle species from fragments, if such declines are influenced by inter-specific differences in dispersal capability, and what the consequences of these patterns are for the ecosystems services they provide.
-One major factor behind this lack of information is the challenge in finding locations where one can assess the role of corridors while also while controlling for confounding factors such as patch size, edge, and corridor length[@haddad_habitat_2015].
+One major factor behind this lack of information is the challenge in finding locations where one can assess the role of corridors while also while controlling for confounding factors such as patch size, edge, and corridor length [@haddad_habitat_2015].
 
@@
 
-```{r fig1, echo=FALSE, fig.cap="Arrangement of pitfall traps used to sample beetles in experimental blocks.", fig.align="center", out.width='45%'}
+<!-- ```{r fig1, echo=FALSE, fig.cap="Arrangement of pitfall traps used to sample beetles in experimental blocks.", fig.align="center", out.width='45%'} -->
 
-knitr::include_graphics("images/plots1.png")
+<!-- knitr::include_graphics("images/plots1.png") -->
 
-```
+<!-- ``` -->
 

### 2024-05-02 corridor_docs/masters_cmtee_proposal/eec_ms_proposal.Rmd
updating ms proposal
comments from corridor steering cmmmtee

@@
 bibliography: "`r here::here('corridor_docs', 'eric-thesis.bib')`"
-
+csl: "`r here::here('corridor_docs', 'ecology.csl')`"
 output: pdf_document
@@
 
-All traps and markings will be removed at the study's conclusion.
-
-[@WOS:000341406900011]
+All traps and markings will be removed at the study's conclusion. [@WOS:000341406900011]
 
@@
 
-<!-- \newpage -->
+\newpage
 
-## Figures
+# References
+
+&nbsp;
+
+::: {#refs custom-style="Bibliography"}
+:::
+
+\newpage
 
@@
 
-<!-- ![Arrangement of pitfall traps used to sample beetles in experimental blocks.](./images/plots1.png){fig-align="center" width=45% height=45%} -->
+![Arrangement of pitfall traps used to sample beetles in experimental blocks.](./images/plots1.png){fig-align="center" width=45% height=45%}
 
-<!-- ![Locations and treatments for the dung removal experiment.](./images/plots3.png){fig-align="center" width=45% height=45%} -->
+\newpage
 
-<!-- ![Design for the experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=45% height=45%} -->
+![Locations and treatments for the dung removal experiment.](./images/plots3.png){fig-align="center" width=45% height=45%}
 
-## References
+\newpage
+
+![Design for the experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=45% height=45%}
+
+\newpage
+
+## CORRIDOR STEERING COMMITTEE FEEDBACK
+
+### Sampling in the matrix
+
+**Doug:** What's the point of sampling in the matrix?  I think I know the answer.  I ask because it doesn't seem related to any of your predictions.
+
+**Christopher Warneke:** I do like the idea of sampling the matrix to get a sense of what is going on with dung beetles there, so I am glad to see that in the proposal.
+When Carolyn Graham and I looked at foliar mites in the Corridor patches, we did look at the mites in the matrix too and found that mites were in higher abundance in the matrix.
+For our story, it was useful information for our Discussion, but it didn't matter too much for our study questions.
+But I could see it mattering for dung beetles, particularly if they are viewing the matrix and patches as essentially the same habitat type (though I doubt that is the case).
+
+### Traps and Baits
+
+**Doug: **I've trapped dung beetles in FL and learned that dung dries quickly in the sun and loses its attractiveness in the process.
+My traps contained dung suspended from a modified coat-hanger in a bag made of panty hose.
+Briefly soaking the bag in water 1-2 times/day kept the dung attractive.
+I don't think I used 95% ethanol in the traps (maybe ethylene glycol?) because I was concerned about the odor of ethanol.
+Finally, think about the type of dung.
+I used cow dung but learned afterwards that it's not the best choice.
+Pig dung is far better and probably easily available at UF.
+There are papers on what type of dung works best (e.g., https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0073147
+
+**Christopher Warneke:** I remember Sean Griffen (who worked with a megachild bee and movement between patches) that there were some challenges with number of bees found in the peripheral patches after being released in the center patch?
+I just glanced at the paper again [link here](https://nsojournals.onlinelibrary.wiley.com/doi/pdf/10.1111/ecog.05299) and it looks like it worked out better than my memory, but I worry about question 3 of the proposal and having enough recaps to be able to discuss movement rates (as it looks like is a common theme in feedback about the proposal).
+I don't know whether it would be helpful to check with Sean about things, but his website [is here](https://srgriffin.com/) if you want to reach out to him.
+Question 3 would be awesome if you had enough recaptures though!
+
+**Lars: **Your preliminary surveys will help a lot, but what if you don’t catch much through your initial surveys.
+Would you look to ramp up the number of pitfalls you use in each patch (and would you have this capacity)?
+
+I think there could be something interesting with the temporal dynamics, but it was less clear to me how critical this was compared to simply comparing among the different patch types.
+Do you know if there is a predominant season that you expect to catch the most beetles/when they might be most active dung foragers?
+If so, it might be worth considering ramping up sampling intensity during this time (and maybe sacrificing some of the temporal extent?).
+Just a suggestion – definitely feel free to ignore this!
+
+### Dispersal
+
+**Thomas:**  I think it is a good idea to assess species use of matrix both in dispersal trials and the richness sampling so you could control for this in species responses to the corridor.
+You could also include habitat specialist versus generalist users in the dispersal trials and potentially conduct an "inverse" of the Baum et al.
+(2004) corridor experiment, where instead of manipulating matrix directly, you use species that have different permeabilities or affinities for the matrix.
+The last things to think about with the dispersal trails are both the effect of the corridors on winds patterns and the prevailing wind direction for each block.
+I imagine olfaction is the primary way dung beetles locate dung sources, so this might be an important avenue with which either the corridors promote movement or corridor use gets obscured.
+One way to address this could be to either release in multiple unconnected patches or to sufficiently randomize across the blocks.
+If releases are conducted at the edge of patches, it might not require that only rectangular patches be used.
+I imagine the feasibility of this would depend on the numbers needed per release.
+
+**Lars:**  Something to consider might also be to do multiple release distances from the bait, perhaps with different color markings (or something like that – I’m not sure how marking works), to work out beetle movement capacity.
+This could be preferable to having multiple bait distances, in case beetles are prone to stopping at the first bait they encounter.
+If you wanted to try out something like this, a recent large-ish clear cut could be good (bait in the middle, differing release distances radiating out from it) and I’m sure we could help locate one to work in.
+
+### Beetles/Bycatch
+
+**John Orrock: **In [this paper](https://link.springer.com/article/10.1007/s10980-011-9656-5) we report data from two years (6 total sampling sessions across 2001 and 2002) representing over 33,000 individuals from 58 genera.
+The list of arthropods we captured is at the end (some scarabs, but we stopped at genus).
+The pitfalls were relatively large (compared to centrifuge tubes and things we’ve used for other studies) so they collected a lot of things.
+This paper uses pitfalls near the center of patches, but I also sampled near patch edges, in wings, and in corridors in 2022.
+Those data were never published, but I think I just found a spreadsheet with them (over 13,000 individuals identified to genus).
+That, said, the core-pitfall dataset in the attached paper is the one that is the most comprehensive.
+
+### Assessing risk of plant introduction
+
+**Christopher Warneke:** From a plant perspective, I wonder if it would potentially be possible to maybe plant some of the cow dung (or pig or whatever you go with) in the boneyard at station on site?
+For dung from ag animals, I suspect there won't be many germinants that come out from dung that gets moved by beetles, but if some is able to be planted at station, then Thomas/Benjamin/Eva could check on them and if any plants come up, that will help with the plant surveys within the patches if any plants pop up in the patches from the dung.
+That would just help us sort out unknown plants easier if we know they are accidental germinants from the cow/pig dung.
+Though I don't think this will really be a big issue, so it is probably also fine to not worry about it.
+
+**Lars: **Christopher’s point about possibly introducing plants through dung is important (especially if this happens unequally across the patches, since one/block isn’t used).
+It’d be good to hear your thoughts on this possibility.

### 2024-05-02 corridor_docs/ideas/thesis_ideas.Rmd

@@
 insert_citation
+
+Post SRS Badging ideas
+1. Camera monitoring - Big mammals
+2. Matrix densities
+3. Go through John Orrock’s unpublished data
+Other stuff to figure out
+1. Housing
+2. Funding
+3. Steering committee questions
+
 ### *Notes from conversation with Julian:*

### 2025-01-24 corridor_docs/eric_ms_thesis/01_cleanup_sample_info.R
editing data import

@@ -4,6 +4,7 @@
 # load libraries ----------------------------------------------------------
 
 library(tidyverse)
+library(here)
 library(lubridate) # to deal with dates https://lubridate.tidyverse.org/
 # library(here)
 # load data ---------------------------------------------------------------
@@ -63,3 +64,98 @@ write_csv(btl,"./corridor_docs/eric_ms_thesis/ms_data/data_clean/clean_sample_in
 # 
 # btl<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","samples.csv"))
 #   
+
+# load beetle id data -----------------------------------------------------
+library(vegan)
+library(ggvegan)
+# vegan cheatsheet: https://www.rpubs.com/an-bui/vegan-cheat-sheet
+
+
+
+env <- read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","samples.csv")) %>% 
+  mutate(sample_id=as.factor(sample_id)) 
+
+btl_counts<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","beetle_counts.csv")) %>% 
+  mutate(sample_id=as.factor(sample_id)) %>% 
+  rowwise() %>%
+  mutate(sum = sum(c_across(pvin:ocon), na.rm = TRUE)) %>% 
+  mutate(sum=ifelse(sum==0,NA,sum)) %>% 
+  replace(is.na(.), 0) %>% 
+  left_join(env) %>% 
+  relocate(c(block,patch, point, month, day),.after = sample_id)
+
+btl_summary<-btl_counts %>% 
+  select(-c(sample_id,block,point,month,day)) %>% 
+  group_by(patch) %>% 
+  summarize(across(everything(), ~ sum(.x, na.rm = TRUE))) %>% 
+  filter(!is.na(patch)) %>% 
+  select(-sum) %>% 
+  column_to_rownames("patch")
+
+
+  
+  summarize(sum=sum(c_across(pvin:ocon), na.rm = TRUE)) %>% 
+  filter(!is.na(patch))
+btls<-btl_summary %>% 
+  
+
+
+
+
+btls<-btl_counts %>% 
+  select(-c(sum,sample_id,block,point, month, day)) %>% 
+  # column_to_rownames("sample_id")
+  column_to_rownames("patch")
+
+
+
+
+
+
+
+site_type <- env %>% 
+  select(block, patch)
+# How speciose are my communities?
+  
+# specnumber() will tell you the number of species within each sample. 
[…]

### 2025-01-24 corridor_docs/eric_ms_thesis/ms_thesis_uf/eric-ms-thesis.tex
thsis, latex

\documentclass[editMode]{ufdissertation}\sloppy
\usepackage[colorlinks=true]{hyperref}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                 User Package and Style File loading.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage{CustomMacros}%  This is a user macro/style file.

\usepackage{tikz}%       tikz is used by almost everyone, but certainly by me for this.
\usepackage{pgfplots}%   pgfplots is tikz but better.
%\usepackage{amsrefs}%   amsrefs contains the .bibtex style content for mathematician papers.
\usepackage{algpseudocode}

\usepackage{scrextend}
\deffootnote{1.5em}{0em}{\thefootnotemark\quad}
\usepackage[all]{nowidow} %avoids single lines of text at the top or bottom of page.  If any of your packages cause this package not to work, you will have to manually remove any orphans/widows or add a line penalty
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                     User Configuration commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Uncomment the relevant line below if you have tables or figures.
\haveTablestrue%        Uncomment this if you have tables in your thesis.
\haveFigurestrue%       Uncomment this if you have figures in your thesis.
\haveObjectstrue%       Uncomment this if you have Objects in your thesis. This is almost certainly not the case however.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Below are the commands to set the degree type, department, graduation time, and chair. 
%       Most of these are self explanatory. 
%       Note: The \chair command takes an optional argument for a cochair. 
%           So if John was your chair and Jacob was a cochair, you would use \chair[Jacob]{John}.
%           If John was your chair and you had no cochair, you can simply use \chair{John}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Dissertation and Thesis Example File}%  Put your title here.

\degreeType{Doctor of Philosophy}%   Official name of your degree; eg "Doctor of Philosophy".
\major{Mathematics}%                    Your official Department
\author{Jason Nowell}%                  Your Name
\thesisType{Dissertation}%              Dissertation (PhD) or Thesis (Masters)
\degreeYear{2024}%                      Intended graduation year (not the year you submit the thesis)
\degreeMonth{August}%                   Month of graduation should be May, August, or December.
\chair[Jack Griffin]{Claude Rains}%                   Chair and Cochair (see comment block above).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% For each of the following, type in the name of the file that contains each section. 
% They are assumed to be tex files, but if they aren't the command takes an optional argument for the extension.
%So, you could load dedication.tex as your dedication file using \setDedicationFile{dedication}
% You could load dedication.txt instead with \setDedicationFile[txt]{dedication}.

% NOTE: For some compilers they may or may not add a .tex to the end of the file automatically.
% If you get a "couldn't find dedication.tex.tex" type error, try the command with an empty optional argument,
%e.g. \setDedicationFile[]{dedication}
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% These are REQUIRED sections; easiest to do via these commands.

\setDedicationFile{dedicationFile}%                 Dedication Page
\setAcknowledgementsFile{acknowledgementsFile}%     Acknowledgements Page
\setAbstractFile{abstractFile}%                     Abstract Page (This should only include the abstract itself)
\setReferenceFile{referenceFile}{amsplain}%         References. First argument is your bibtex source file
%                                                       the second argument is your bibtex style file.

\setBiographicalFile{biographyFile}%                Biography file of the Author (you).

%%% These are NOT required, so only use them if you actually need/have them.

\setAbbreviationsFile{abbreviations}%           Abbreviations Page
\setAppendixFile{appendix}%                     Appendix Content; hyperlinking might be weird.
\multipleAppendixtrue%                          Uncomment this if you have more than one appendix, 
%                                                   comment it if you have only one appendix.

%%%%%%%                     End of File Assignment
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%% Here you just need to include/input your actual work. 
%       The above files (dedication, acknowledgement, titlepage, etc etc) will all be added for you 
%       using the files you assigned above. 
%       If you want to input the above files manually you can comment out the \setFILE command above 
%       and use \input or \include here. Generally you want to use \include to get your pagebreak.
%       NOTE: If you input manually you will have to do some/all the formatting manually.

\include{chapter1}% Modified from old template.
\include{chapter2}% Modified from old template.
\include{chapter3}% Modified from old template.

\chapter{EXAMPLES OF EDITOR/Author TOOLS}% Notice that we can use chapter/section etc breaks in the master file if we want, and then use \input instead of \include to avoid unneccessary page breaks.
\input{editorAndAuthorRemarks}%     Stuff about using editorRemark and authorRemark commands

\include{chapter5}% Modified from old template.

\end{document}

### 2025-01-24 corridor_docs/eric_ms_thesis/ms_thesis_uf/uf-latex/chapter1.tex
thsis, latex

\chapter{INTRODUCTION and opening remarks} {\renewcommand*{\thefootnote}{\fnsymbol{footnote}}\footnotetext{An un-numbered footnote - this is how you tell the readers that this chapter was previously published and then cite the Journal where it was published. It is typically formatted like "Reprinted with permission from...}\label{intro}
\counterwithin{algorithm}{chapter}
\renewcommand*{\thealgorithm}{\thechapter-\arabic{algorithm}} %Lines 2 and 3 are specifically for those that are using multiple aglorithms in multiple chapters. These must be added to all chapters with algorithms to make the numbering of the objects work correctly. You are welcome to remove if this does not apply to you.

We automatically capitalize all chapters, but if you need to suppress this you can use the class option ``overrideTitles" and/or ``overrideChapter" to allow you to use non-capitalized letters in the title and/or chapter names respectively. For more detailed information on the template's features and options, see the included file ``ufdissertation-Doc-and-Troubleshooting".

 We don't recommend that you change much of anything in the class file unless you're absolutely sure of what your are doing.\renewcommand*{\thefootnote}{\arabic{footnote}}\setcounter{footnote}{0}\footnote{and now we're back to normal footnote marking}

\section{First-Level Heading or Section Heading}

This is a first-level or Section heading. They should always be in Title-Case. Title case is where all principal words are capitalized except prepositions, articles, and conjunctions.  %\cite{green2008wrinkle}

\subsection{Second-Level or Subsection Heading}

This is a second-level or subsection heading. They will always be in title-case but are left-aligned.

\subsubsection{Third-Level or Subsubsections}
The third level subheadings are left-aligned but in sentence case. Only the first letter and any proper nouns are capitalized.

\subsubsection{If you divide a section, you must divide it into two, or more, parts}
What this means is that if you are going to break down text via subheading, you will want to make sure it is paired with at least one other subheading of the same "level."

{\bf Paragraph headings.} There is no official fourth level heading. Do not use the Paragraph heading feature in LaTeX, simply apply the bold characteristic to the first few words of a paragraph followed by a colon or period.

\subsection{Subsection}

\(\Omega\) Aliquam mi nisi, tristique at rhoncus quis, consectetur non mi. Phasellus blandit quam ligula, a viverra lacus commodo at. In iaculis nisl vel pretium sollicitudin. In efficitur massa vel elit sollicitudin, vel auctor sapien cursus. Proin feugiat sapien a mi tempus;

\begin{equation}
       X-X'=D+D' 
\end{equation}
Augue sapien mattis leo, nec accumsan turpis quam at neque. Ut pellentesque velit sed
placerat cursus. Integer congue urna non massa dictum, a pellentesque arcu accumsan. Nulla
posuere, elit accumsan eleifend elementum, ipsum massa tristique metus, in ornare neque nisl sed
odio. Nullam eget elementum nisi. Duis a consectetur erat, sit amet malesuada sapien. Aliquam
nec sapien et leo sagittis porttitor at ut lacus. Vivamus vulputate elit vitae libero condimentum
dictum. Nulla facilisi. Quisque non nibh et massa ullamcorper iaculis.
Augue sapien mattis leo, nec accumsan turpis quam at neque. Ut pellentesque velit sed
placerat cursus. Integer congue urna non massa dictum, a pellentesque arcu accumsan. Nulla
posuere, elit accumsan eleifend elementum, ipsum massa tristique metus, in ornare neque nisl sed
odio. Nullam eget elementum nisi. Duis a consectetur erat, sit amet malesuada sapien. Aliquam
nec sapien et leo sagittis porttitor at ut lacus. Vivamus vulputate elit vitae libero condimentum
dictum. Nulla facilisi. Quisque non nibh et massa ullamcorper iaculis.

\begin{quote}
    \begin{singlespace}
  This is an example of a block quote. Aliquam mi nisi, tristique at rhoncus quis, consectetur non mi. Phasellus blandit quam ligula, a viverra lacus commodo at. 
    \end{singlespace}

\end{quote}

\subsection{Subsection}

Aliquam mi nisi, tristique at rhoncus quis, consectetur non mi. Phasellus blandit quam ligula, a viverra lacus commodo at. In iaculis nisl vel pretium sollicitudin. In efficitur massa vel elit sollicitudin, vel auctor sapien cursus. Proin feugiat sapien a mi tempus;

\begin{equation}
    X-X'=D+D'
\end{equation}

\noindent in consequat augue cursus. Nulla sed sagittis purus. Nunc eu consequat orci, eu laoreet enim. Ut euismod tincidunt sem, eget lacinia dui luctus eu. Aliquam mi augue, faucibus id semper vitae, porta ac ligula. Morbi sed ultrices odio. Mauris id luctus ex. Nulla ac libero dictum, interdum turpis lacinia, scelerisque leo. Praesent varius orci ac eros varius pharetra.

\subsection{Subsection}
Donec convallis scelerisque ante, in sollicitudin orci laoreet eu. Nam arcu magna, semper vel lorem eu, venenatis ultrices est. Nam aliquet ut erat ac scelerisque. Maecenas ut molestie mi. Phasellus ipsum magna, sollicitudin eu ipsum quis, imperdiet cursus turpis. Etiam pretium enim a fermentum accumsan. Morbi vel vehicula enim.

\section{Objects}

\begin{algorithm}
\caption{An algorithm with caption. Example from Overleaf.}\label{alg:cap}
\begin{algorithmic}
\Require $n \geq 0$
\Ensure $y = x^n$
\State $y \gets 1$
\State $X \gets x$
\State $N \gets n$
\While{$N \neq 0$}
\If{$N$ is even}
    \State $X \gets X \times X$
    \State $N \gets \frac{N}{2}$  
    \State $y \gets y \times X$
    \State $N \gets N - 1$
\EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}

Please note, the 'Objects' section of this document is based off of the Algorithm environment. If you wish to use external links to a repository, UF recommends using Zenodo (https://guides.uflib.ufl.edu/etds/supplemental). Please reach out to our office at TandDSupport-hd@ufl.edu if you have any questions about adding it to your document, and the ETD team at the library if you have any questions about external repositories (IRManager@uflib.ufl.edu).

### 2025-01-24 corridor_docs/eric_ms_thesis/ms_thesis_uf/uf-latex/chapter2.tex
thsis, latex

\chapter{EXAMPLE TABLE FORMATTING} \label{lit}

\section{Table Examples}

 The UF Graduate Counsel is very specific in the Table Requirements.There should be no vertical lines in tables and only three horizontal lines. Tables should be left aligned or the full width of the margins. No bold text, etc., see the web site for the complete list of requirements. One simple improvement can be incorporated by using tabularx instead of the tabular environment. This allows a table to be stretched the full text width easily, which avoids the centered or left aligned issue. Table \ref{sttab} is an example of the tabularx code.

\begin{table}[htbp]% Fix the table captions to sit directly on the table, but figures do NOT sit directly on the figure.
    \captionof{table}{A sample Table using tabularx. All tables should be left aligned OR the full width of the margin, like this one. }\label{sample}
    \begin{tabularx}{6.5in}{XXX}
      \hline
      First & Second & Third \\
      \hline
      12 & 45 & 26 \\
      17 & 32 & 93 \\
      text & 51 & can be there too. \\	
      \hline
    \end{tabularx}
\end{table}

 Praesent fermentum felis nec massa interdum, vel dapibus mi luctus. Cras id fringilla mauris. Ut molestie eros mi, ut hendrerit nulla tempor et. Pellentesque tortor quam, mattis a scelerisque nec, euismod et odio. Mauris rhoncus metus sit amet risus mattis, eu mattis sem interdum.

 \begin{table}[t!]
    \caption{A sample Table using standard tablular}\label{tab}
    \begin{tabular}{c c c}
      \hline
      First & Second & Third \\
      \hline
      12 & 45 & 26 \\
      17 & 32 & 93 \\
      text & 51 & can be there too. \\	
      \hline
    \end{tabular}
\end{table}

\section{Additional Table Examples}
Augue sapien mattis leo, nec accumsan turpis quam at neque. Ut pellentesque velit sed placerat cursus. Integer congue urna non massa dictum, a pellentesque arcu accumsan. Nulla posuere, elit accumsan eleifend elementum, ipsum massa tristique metus, in ornare neque nisl sed odio. Nullam eget elementum nisi. Duis a consectetur erat, sit amet malesuada sapien. Aliquam nec sapien et leo sagittis porttitor at ut lacus. Vivamus vulputate elit vitae libero condimentum dictum. Nulla facilisi. Quisque non nibh et massa ullamcorper iaculis.

 \begin{table}[htbp]
    \caption{Another sample Table using standard tablular}\label{sttab}
    \begin{tabular}{c c c}
      \hline
      First & Second & Third \\
      \hline
      12 & 45 & 26 \\
      17 & 32 & 93 \\
      text & 51 & can be there too. \\	
      \hline
    \end{tabular}
\end{table}

\section{Additional Table Examples}
You may notice that some tables get moved outside of where you placed them. This is because \LaTeX{} is a little too helpful when it comes to placement of `float' types. You can get around this by using the ``H" parameter in the table environment, or the `multiFigure' environment described in the ``adding graphics section"; ie section \ref{Sec:addingGraphics}

\begin{table}[H]
\caption[Caption Location]{ Caption Location. You should also make a note that the caption command is placed after the table itself, which means the caption occurs after the table. The graduate school requires tables to have captions placed {before} the actual table data, so the caption command should be located before the table data. See the next table for an example.}
\begin{tabular}{llcr}
\hline
Some    & Data  & Goes  & Here\\
\hline
Some    & Data  & Goes  & Here\\
Some    & Data  & Goes  & Here\\
Some    & Data  & Goes  & Here\\
\hline
\end{tabular}

\end{table}

\begin{table}[h]
\caption[A proper table caption location]{Notice that this caption is included above the table data, as per the graduate school requirements. Also note that the caption itself has a short version in the ``List of Tables" which is achieved by using the optional argument of the caption command. See the file source code directly to see the example.Unfortunately, since we did not use the ``H" parameter in the table environment, this table was placed after the next section heading, which is almost certainly not where an author would have wanted it.}
\begin{tabularx}{\textwidth}{XXXX}\hline
Some    & Data  & Goes  & Here\\\hline
Some    & Data  & Goes  & Here\\
Some    & Data  & Goes  & Here\\
Some    & Data  & Goes  & Here\\\hline
\end{tabularx}
\end{table}

\section{Very Long Tables}

There are two approaches to inputting very long tables. You can do it manually, or you can do it using the longtables package. Here we include an example of both. Table \ref{tbl1} is done manually, whereas \ref{tbl2} is done using the longtables package. Whatever works for you is fine to use, and you can use both in one document.

Augue sapien mattis leo, nec accumsan turpis quam at neque. Ut pellentesque velit sed placerat cursus. Integer congue urna non massa dictum, a pellentesque arcu accumsan. Nulla posuere, elit accumsan eleifend elementum, ipsum massa tristique metus, in ornare neque nisl sed odio. Nullam eget elementum nisi. Duis a consectetur erat, sit amet malesuada sapien. Aliquam nec sapien et leo sagittis porttitor at ut lacus. Vivamus vulputate elit vitae libero condimentum dictum. Nulla facilisi. Quisque non nibh et massa ullamcorper iaculis.Augue sapien mattis leo, nec accumsan turpis quam at neque. Ut pellentesque velit sed placerat cursus. Integer congue urna non massa dictum, a pellentesque arcu accumsan. Nulla posuere, elit accumsan eleifend elementum, ipsum massa tristique metus, in ornare neque nisl sed odio. Nullam eget elementum nisi. Duis a consectetur erat, sit amet malesuada sapien. Aliquam nec sapien et leo sagittis porttitor at ut lacus. Vivamus vulputate elit vitae libero condimentum dictum. Nulla facilisi. Quisque non nibh et massa ullamcorper iaculis.

Augue sapien mattis leo, nec accumsan turpis quam at neque. Ut pellentesque velit sed placerat cursus. Integer congue urna non massa dictum, a pellentesque arcu accumsan. Nulla posuere, elit accumsan eleifend elementum, ipsum massa tristique metus, in ornare neque nisl sed odio. Nullam eget elementum nisi. Duis a consectetur erat, sit amet malesuada sapien. Aliquam nec sapien et leo sagittis porttitor at ut lacus. Vivamus vulputate elit vitae libero condimentum dictum. Nulla facilisi. Quisque non nibh et massa ullamcorper iaculis.Augue sapien mattis leo, nec accumsan turpis quam at neque. Ut pellentesque velit sed placerat cursus. Integer congue urna non massa dictum, a pellentesque arcu accumsan. Nulla posuere, elit accumsan eleifend elementum, ipsum massa tristique metus, in ornare neque nisl sed odio. Nullam eget elementum nisi. Duis a consectetur erat, sit amet malesuada sapien. Aliquam nec sapien et leo sagittis porttitor at ut lacus.

\begin{table}[H]
\caption{Feasible triples for highly variable Grid, MLMMH.} \label{tbl1}
\begin{tabularx}{6.5 in}{r l X}
\hline {{Time (s)}} & {{Triple chosen}} & {{Other feasible triples}} \\ \hline
0 & (1, 11, 13725) & (1, 12, 10980), (1, 13, 8235), (2, 2, 0), (3, 1, 0) \\
2745 & (1, 12, 10980) & (1, 13, 8235), (2, 2, 0), (2, 3, 0), (3, 1, 0) \\
5490 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
8235 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
10980 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
13725 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
16470 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
19215 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
21960 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
24705 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
27450 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
30195 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
32940 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
35685 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
38430 & (1, 13, 10980) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
41175 & (1, 12, 13725) & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
43920 & (1, 13, 10980) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
46665 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
49410 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
52155 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
54900 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
57645 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
60390 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
63135 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
65880 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\

\hline
\end{tabularx}
\end{table}
\clearpage

\begin{table}[h!t!]
\begin{tabularx}{6.5 in}{r l X}
\multicolumn{3}{l}{Table \ref{tbl1}. Continued}\\%
\hline {{Time (s)}} & {{Triple chosen}} & {{Other feasible triples}} \\ \hline
104310 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
107055 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
109800 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
112545 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
115290 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
118035 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
120780 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
123525 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
126270 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
129015 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
131760 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
134505 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
137250 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
139995 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
142740 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
145485 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0)\\%
148230 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
150975 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
153720 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
156465 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
159210 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
161955 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
164700 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
169210 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
171955 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
184700 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
253720 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
256465 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
259210 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
261955 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
262955 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
263555 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
271955 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\

\hline
\end{tabularx}
\end{table}

Alternatively, compared to the previous example where we used manual breaks to break the table, we can let LaTeX do this for us, as well as taking care of any recurrent headers and footers, utilizing the \verb|\longtable| command,\footnote{note that the longtable environment is not in a table environment; putting it inside a table environment will stop it from correctly page breaking as needed.} as follows:

\clearpage

\begin{longtable}[h!t!]{p{0.6in}p{1in}p{4.4in}}
    \caption{Duplicate of Previous table, using longtables environment.}\label{tbl2}\\% Default caption at top of table
    \hline {{Time (s)}} & {{Triple chosen}} & {{Other feasible triples}}\\ \hline \endfirsthead% The top row of the first page
    \hline\endfoot%         This line should always be included; it includes a line at the end of the table on every page.
    \caption*{Table 2-17. Continued} \\% This caption is added to every page after the first as per the \endhead next line.
    \hline{{Time (s)}} & {{Triple chosen}} & {{Other feasible triples}}\\ \hline \endhead%
% Everything between \endfoot and \endhead here is added at
%  the top of the table on every page except the first;
% The first page is an exception because we have defined a
% \endfirsthead row already which supersedes \endhead.
0 & (1, 11, 13725) & (1, 12, 10980), (1, 13, 8235), (2, 2, 0), (3, 1, 0)\\
2745 & (1, 12, 10980) & (1, 13, 8235), (2, 2, 0), (2, 3, 0), (3, 1, 0) \\
5490 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
8235 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
10980 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
13725 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
16470 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
19215 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
21960 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
24705 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
27450 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
30195 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
32940 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
35685 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
38430 & (1, 13, 10980) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
41175 & (1, 12, 13725) & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
43920 & (1, 13, 10980) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
46665 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
49410 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
52155 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
54900 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
57645 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
60390 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
63135 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
65880 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
68625 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
71370 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
74115 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
76860 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
79605 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
82350 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
85095 & (1, 12, 13725) & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
87840 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
90585 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
93330 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
96075 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
98820 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
101565 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
104310 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
107055 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
109800 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
112545 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
115290 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
118035 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
120780 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
123525 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
126270 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
129015 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
131760 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
134505 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
137250 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
139995 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
142740 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
145485 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0)\\%
148230 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
150975 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
153720 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
156465 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
159210 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
161955 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
164700 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
\end{longtable}

\section{Tables with Notes}

As there is no cell shading or other descriptive visuals in a table, you prefer to use tables notes to explain additional information. Below is an example of a table made with the 'threeparttable' environemnt.

\begin{table}[h]
\begin{threeparttable}
\caption{Here is my table that needs notes for further content information.}
\begin{tabular*}{\textwidth}{c @{\extracolsep{\fill}} ccccc}
\hline
Some    & Data  & Goes  & Here\\
\hline
Some    & Data  & Goes  & Here\\
Some    & Data  & Goes  & Here\\
Some    & Data  & Goes  & Here\\
\hline
\end{tabular*}

\begin{tablenotes}
\item[1] Here is my table note.
\end{tablenotes}

\end{threeparttable}
\end{table}

### 2025-01-24 corridor_docs/eric_ms_thesis/ms_thesis_uf/uf-latex/chapter3.tex
thsis, latex

\chapter{EXAMPLE FIGURE FORMATTING} \label{materials}

\counterwithin{algorithm}{chapter}
\renewcommand*{\thealgorithm}{\thechapter-\arabic{algorithm}} %Lines 2 and 3 are specifically for those that are using multiple aglorithms in multiple chapters. These must be added to all chapters with algorithms to make the numbering of the objects work correctly. You are welcome to remove if this does not apply to you.

\section{Section Heading}

 Fusce eget tempus lectus, non porttitor tellus. Aliquam molestie sed urna quis convallis. Aenean nibh eros, aliquam non eros in, tempus lacinia justo. \cite{2008arXiv0807.1715B} In magna sapien, blandit a faucibus ac, scelerisque nec purus. Praesent fermentum felis nec massa interdum, vel dapibus mi luctus. Cras id fringilla mauris. Ut molestie eros mi, ut hendrerit nulla tempor et. Pellentesque tortor quam, mattis a scelerisque nec, euismod et odio. Mauris rhoncus metus sit amet risus mattis, eu mattis sem interdum.

\subsection{Subsection Heading}
Fusce eget tempus lectus, non porttitor tellus. Aliquam molestie sed urna quis convallis. Aenean nibh eros, aliquam non eros in, tempus lacinia justo. In magna sapien, blandit a faucibus ac, scelerisque nec purus. Praesent fermentum felis nec massa interdum, vel dapibus mi luctus. Cras id fringilla mauris. Ut molestie eros mi, ut hendrerit nulla tempor et.\cite{Rudin-UnitBallCN}

\subsection{Subsection Heading}
Fusce eget tempus lectus, non porttitor tellus. Aliquam molestie sed urna quis convallis. Aenean nibh eros, aliquam non eros in, tempus lacinia justo. In magna sapien, blandit a faucibus ac, scelerisque nec purus. Praesent fermentum felis nec massa interdum, vel dapibus mi luctus. Cras id fringilla mauris. Ut molestie eros mi, ut hendrerit nulla tempor et. \cite{Rudin-UnitBallCN}

\section{Section Heading}
Nec accumsan turpis quam at neque. Ut pellentesque velit sed placerat cursus. Integer congue urna non massa dictum, a pellentesque arcu accumsan. Nulla posuere, elit accumsan eleifend elementum, ipsum massa tristique metus, in ornare neque nisl sed odio. Nullam eget elementum nisi. Duis a consectetur erat, sit amet malesuada sapien. Aliquam nec sapien et leo sagittis porttitor at ut lacus. Vivamus vulputate elit vitae libero condimentum dictum. Nulla facilisi. Quisque non nibh et massa ullamcorper iaculis. A random citation to demonstrate the bibliography; \cite{ConwayCSV1}.

\section{Algorithm Example}
Below is an example of an algorithm using algpseudocode. You can see how using a normal algorithm forces it into an object. 
\begin{algorithm}
\caption{This is the same algorithm with caption, taken from Overleaf, to show the use of the counterwithin command at the beginning of this chapter.}\label{alg:cap2}
\begin{algorithmic}
\Require $n \geq 0$
\Ensure $y = x^n$
\State $y \gets 1$
\State $X \gets x$
\State $N \gets n$
\While{$N \neq 0$}
\If{$N$ is even}
    \State $X \gets X \times X$
    \State $N \gets \frac{N}{2}$  
\ElsIf{$N$ is odd}
    \State $y \gets y \times X$
    \State $N \gets N - 1$
\EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}
%Please note that using more than one object in one chapter will require the addition of the following in all chapters that have them:

\section{Examples of Adding Graphics}
\label{Sec:addingGraphics}
All of the below code with subfigures A-Z was generated with:
\begin{verbatim}
\begin{multiFigure}
\addFigure{0.3}{./theworld.png}
\addFigure{0.2}{./theworld.png}
\addFigure{0.4}{./theworld.png}
\addFigure[Z]{0.6}{./theworld.png}
\captionof{figure}[This is a test caption.]{This is a test caption. 
This text has the bit for the whole figure. 
Meanwhile, subfigure A is weird looking map. 
Subfigure B is a smaller map. 
And Subfigure C is a bigger but still weird looking map. 
Moreover, I can override the map, which is why Z is 
another weird map that came after map C.}
\end{multiFigure}
\end{verbatim}

Note that \LaTeX{} can be pretty fickle when it comes to placing figures relative to text near the figure. Specifically, the ``Figure" environment is a `float' type, which is placed somewhere ``nearby" where it appears in the text, which can be pretty frustrating. For this reason I have circumvented the `float' part of the figure in order to allow more control over the figure placement. So if one uses the \verb|\begin{figure}\end{figure}| construction, the figure may appear in a slightly weird place, whereas you can use the \verb|\begin{multiFigure}\end{multiFigure}| even with only 1 figure, to force placement to work. The only caveat here is that captions need to be placed using the command \verb|\captionof{<NAME>}[<LIST-ENTRY>]{<CAPTION>}| where NAME is the type of caption, LIST-ENTRY is what appears in the `List of' at the beginning of the thesis, and CAPTION is the actual caption.

\begin{flushleft}
\begin{multiFigure}
\begin{center}
\addFigure{0.3}{Images/theworld.png}
\addFigure{0.2}{Images/theworld.png}
\addFigure{0.4}{Images/theworld.png}
\addFigure[Z]{0.6}{Images/theworld.png}
\end{center}
\captionof{figure}[This is a test caption.]{This is a test caption. This text has the bit for the whole figure. Meanwhile, A) is weird looking map. B) is a smaller map. And C) is a bigger but still weird looking map. Moreover, I can override the map, which is why Z is another weird map that came after map C.}

\end{multiFigure}
\end{flushleft}

\section{A Note on Graphics}
The command \verb|\addFigure| in the multiFigure environment, and/or the command \verb|\includegraphics| will take almost every type of graphic file currently in use as of the writing of this template. The only notable exception is the bitmap, ie .bmp file. Most software won't save to bitmap without specifically requesting it at this point, but if you have generated a .bmp file you can load it in most any graphic editor (eg MSpaint or photoshop) and save it as a different file type, such as .PNG which is significantly smaller file size as well.

Note that the commands typically require the file extension to be included, and it is case sensitive. Thus in the above \verb|\addFigure{0.2}{./theworld.png}| works but \verb|\addFigure{0.2}{./theworld.PNG}| would error and \verb|\addFigure{0.2}{./theworld}| may or may not work depending on which specific TeX editor you are using.

\begin{figure}
    \centering
    \includegraphics{Images/sharkimg.png}    
    \caption[This is another figure.] {This is another figure as an example. Please note that the Editorial Office does not allow borders around figures. Used with permission from Martin, R. Aidan.  2003. }
    \label{fig:my_label}
\end{figure}{}
\section{Placement Specifiers}

Floats are used to allow LaTeX to handle figures while maintaining the best possible presentation. However, there may be times when you disagree, and a typical example is with its positioning of figures. 
The placement specifier parameter exists as a compromise, and its purpose is to give the author a greater degree of control over where certain floats are placed. As you can see in Figure \ref{fig:my_label}A this is a shark. As you can see in \ref{tbl1} this is a something.

\begin{table}[H]
\caption{Specifier Table}
\begin{tabular}{l p{14cm} }
\hline
Specifier & Permission \\ \hline
h & Place the float here, i.e., approximately at the same point it occurs in the source text (however, not exactly at the spot) \\
\\
t & Position at the top of the page.  \\
\\
b & Position at the bottom of the page.  \\
\\
p & Put on a special page for floats only.  \\
\\
! & Override internal parameters LaTeX uses for determining "good" float positions. \\
\\
H & Places the float at precisely the location in the LaTeX code. \\
\hline
\end{tabular}
\end{table}

\newpage

%An example of a specifier parameter is shown below to force a figure into place where it is mentioned in text: 
%\begin{verbatim}
%\begin{figure}[h!]
%    \begin{center}
 %       \includegraphics[width=0.9\textwidth]{./theworld.png}
%    \end{center}
%     \caption[My short caption.]{My full caption in curly brackets.}
%\end{figure}
%\end{verbatim}

### 2025-02-08 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
up to 84

---
title: "thesis_draft"
author: "Eric"
date: "2025-02-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dung Beetle Corridor Draft 
### Pieces 
- abstract

- intro 
  - touch up writing from prop

- methods

### Site

  - description of srs
  - experimental design 
  - conditions during sample period
  - historical significance of site and experimental design 
  - justification for selected patches

### Dung beetle sampling

  - structure and arrangment of traps 
  - description of traps
  - bait 
  - sample period
  - ID 
  - biomass if we do biomass

### Analysis

  - lets hammer this out 
  - modeling? 
    - glmm with poisson dist reccomended by julian
  - beta, abundance, biomass? per site
  - species list by sampling blocks (anything with this?)
  - habitat preference 
  - rarefaction

- results
- discussion

### 2025-02-11 corridor_docs/eric_ms_thesis/02_data_analysis.R
added some summary code and prepping for rarefaction

@@ -57,12 +57,154 @@ total_N<-btl_data %>% summarise(n=sum(sum, na.rm=TRUE))
 # total N by species ------------------------------------------------------
 
 
+# total number of beetles sampled each round
+total_N_sample<-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(date) %>% 
+  summarise(n=sum(n, na.rm=TRUE)) %>% 
+  arrange(date)
+
+
+
+btl_order<-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n)))
+
+
+btl_data_sums<-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(species,patch) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(patch,desc(n))
+
+ggplot(btl_data_sums, aes(x=patch, y=n)) + 
+  geom_bar(stat = "identity")+
+  facet_wrap(vars(species),scales="free",ncol = 5) 
+
+
+
+ggplot(btl_data_sums, aes(x=species, y=n)) + 
+  geom_bar(stat = "identity")+
+  facet_wrap(vars(patch),scales="free",ncol = 4) 
+
+
+
+
+mega_spp<-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>% 
+  summarise(n=sum(n, na.rm=TRUE)) %>% 
+  filter(n>500) %>% 
+  ungroup() %>% 
+  select(species,n)%>% 
+  arrange(desc(n))
+
+
+rare_spp<-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>% 
+  summarise(n=sum(n, na.rm=TRUE)) %>% 
+  filter(n<50) %>% 
+  ungroup() %>% 
+  select(species,n) %>% 
+  arrange(desc(n))
+
+
+
+plot_n<-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
[…]

### 2025-02-19 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
updates

@@
 ---
-title: "thesis_draft"
-author: "Eric"
-date: "2025-02-05"
-output: pdf_document
+title: "Influence of Connectivity on Dung Beetle Communities"
+author: "ERIC IN ALL CAPS"
+
+# date: "\\vspace{-1em}"
+header-includes:
+ # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
+ - \usepackage{lineno}
+ - \linenumbers
+ - \usepackage[default]{sourcesanspro}
+ - \usepackage{fancyhdr} # modify header into left, center, right
+ - \pagestyle{fancy} # cues the use of fancy headers
+ - \fancyfoot{} # cues the use of fancy footer
+ - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
+ - \fancyhead[R]{p. \thepage} # name and page on header right
+# # - \fancyfoot[R]{ }
+# - \usepackage{setspace}{singlespace} # manipulate line spacing 
+# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
+# - \usepackage{xcolor} # changes colors of text
+# # http://latexcolor.com/ to choose colors
+# # define the colors used above with rgb codes
+# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
+# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
+# # to center to top level (#) headings (my name) and change color of font
+# # more info # https://www.ctan.org/pkg/sectsty
+# - \usepackage{sectsty}
+# - \sectionfont{\centering\color{darkmidnightblue}}
+# # to center to sub-sub-headings (###) and change color of font
+# - \usepackage{sectsty} 
+# - \subsectionfont{\centering\color{darkmidnightblue}}
+# - \usepackage{sectsty} 
+# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
+# - \usepackage{wrapfig} 
+# - \newcommand{\blandscape}{\begin{landscape}}
+# - \newcommand{\elandscape}{\end{landscape}}
+# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
+# - \usepackage{indentfirst}
+# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
+bibliography: ["eric-thesis.bib"]
+output:
+  pdf_document:
+    includes:
+      # in_header: style/preamble.tex
+      template: null
+      keep_tex: no
+      # number_sections: true
+# documentclass: style/nsf2
+biblio-style: apalike
+csl: ecology.csl
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+# Abstract
 
-## Dung Beetle Corridor Draft 
-### Pieces 
-- abstract
+Thesis abstracts should be 250 words or less.
 
-- intro 
-  - touch up writing from prop
+\newpage
 
-- methods
+# INTRODUCTION
 
-### Site
+Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes [@haddad_corridor_2003].
+These movements are hypothesized to prevent species diversity from declining in fragments, as well as help maintain the ecosystem services provided by these species (at both the patch- and landscape-level)[@burt_ants_2022].
+Although there is some evidence that animals disperse between patches via corridors, and that connected patches have higher species diversity than unconnected ones, little work to date has investigated the consequences of these corridor-driven patterns for ecosystem services.
+
+Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity and their ecosystem services [@ros_how_2012; @roslin_dung_2000].
+The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles such as enhanced nutrient cycling and soil quality and reduction of parasites on methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
+Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) [@WOS:000891747700001].
+Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].
+However, it remains unknown if corridors actually act to reduce the loss of dung beetle species from fragments, if such declines are influenced by inter-specific differences in dispersal capability, and what the consequences of these patterns are for the ecosystems services they provide.
+One major factor behind this lack of information is the challenge in finding locations where one can assess the role of corridors while also while controlling for confounding factors such as patch size, edge, and corridor length [@haddad_habitat_2015].
+
+**We sampled the commmunity of of dung beetles at the SRS Corridor Experiment to test the following prediction:** Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches.
+
+# Methods
+
+## Study site
 
@@
 
-### Dung beetle sampling
+## Dung beetle sampling
 
@@
 
-### Analysis
+### Analyses
+
+1. **Species Richness:** absolute number? non-parametric estimators?
+1. **Species Diversity:** what index should we compare?
+1. **Functional Diversity:** Need to assign each species to a functional group: roller, tunneler. dweller, others?
+
+>_Look through dung beetle pubs and see how/what people compare_
 
@@
 
-- results
-- discussion
+# RESULTS
+
+some summary statistics:
+
+1. total number of beetles _from_ total number of species 
+1. were all species found in all habitats? Were any species found in only 1 habitat?
+1. Were all species found in all blocks? Were any restricted to only 1-2 blocks?
+1. Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
+1. Number of species in each functional group
+1. most common 3-4 species
+1. any rare species?
+
+# DISCUSSION
+
+1. dont forget o discuss the basic biology...why might a species be so common? why might one be rare?
+
+# ACKNOWLEDGMENTS
+
+Acknowledgments must be written in complete sentences.
+Do not use direct address.
+For example, instead of Thanks, Mom and Dad!, you should say I thank my parents.
+The heading “ACKNOWLEDGMENTS” uses the 002 CHAPTER TITLE style..
+The paragraphs in this section should use the style called 006 Body Text
+
+# OTHER REQUIRED TEXT
+
+### Dedication 
+Your dedication is typed here.
+It should begin with the word “To.” (To my Mom is a typical dedication) If your dedication is longer than a single line it should be single-spaced, and centered vertically and horizontally
+
+### List of Abbreviations
+1. A word to be defined:	Write the definition here. 
+1. Another word:	And the list continues with another definition.
+
+# Biographical Sketch
+A biographical sketch is required of all candidates.
+The biographical sketch should be in narrative form.
+Third person, past tense, it typically includes the educational background of the candidate.
+The author should have replaced this paragraph with their own.

### 2025-02-25 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
questions for analyses and cleaning up data
tidbits
introduction outline, beetles by species filtered by block and patch

@@
 
+INTRODUCTION
+
 Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes [@haddad_corridor_2003].
@@
  **We sampled the commmunity of of dung beetles at the SRS Corridor Experiment to test the following prediction:** Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches.
+
+- First Paragraph: What is the topic of your introduction and why is it important/interesting/relevant?
+
+What?
+
+Connectivity and fragmentation
+
+understanding how movement corridors effect populations dung beetles
+
+Dispersal ability
+
+Why?
+
+Understanding responses to landscape change allows us to anticipate how environments might change with further disturbance.
+
+Patterns in population distributions might suggest how we can study focal organisms in the future.
+
+Could help understand how interactions may be changing.
+
+- Second Paragraph: What is known about this topic already?
+
+Effects of landscape change are already very well studied, accross different taxa
+
+effects of edge, patch size, and direct habitat loss
+
+using corridors to connect fragmented landscapes is also well studied and for many taxa
+
+dung beetles as a model system for studying movement, functional diversity, and ecosystem health.
+
+dung beetle populations in connected and fragmented habitats
+
+insect declines in fragmented landscapes
+
+studying coprophagous insects
+
+- Third Paragraph: What isn't known about this topic and why might it change how we think/act about the topic?
+
+How corridors directly impact dung beetle communitites
+
+are changes driven by dispersal ability
+
+at our study site we can directly compare patch connectivity and patch shape with fragmented landscapes to obtain a strong idea of what landscape features effect dung beetle collection.
+
+- Fourth Paragraph: Why hasn't this thing been studied/assessed/done before?
+
+other work at the corridor project but we are doing dung beetles instead
+
+- Fifth Paragraph: Literally the words "Here we..."
+
+Here sampled the commmunity of of dung beetles at the SRS Corridor Experiment to test the following prediction:
+
+Species richness, diversity, and functional diversity will be higher in patches connected by a movement corridor than in patches that are unconnected.
 
@@
 1. **Species Richness:** absolute number? non-parametric estimators?
+  - Michaelis-Menten
+  - choa?
 1. **Species Diversity:** what index should we compare?
+  - alpha diversity per patch type
+  - beta between patch types
+  - hill numbers
+
 1. **Functional Diversity:** Need to assign each species to a functional group: roller, tunneler. dweller, others? 
+  - habitat preference (forest, pasture, generalist)
 

### 2025-02-25 corridor_docs/eric_ms_thesis/02_data_analysis.R

@@ -78,6 +78,71 @@ btl_order<-btl_data %>%
 btl_order
 
 
+btl_order_block_53n <-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  filter(block == "53n") %>%
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n)))
+
+btl_order_block_8 <-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  filter(block == "8") %>%
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n)))
+
+btl_order_block_52 <-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  filter(block == "52") %>%
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n)))
+
+btl_order_block_54 <-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  filter(block == "54") %>%
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n)))
+
+btl_order_m <-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  filter(patch == "m") %>%
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n)))
+
+btl_order_c <-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  filter(patch == "c") %>%
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n)))
+
+btl_order_w <-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
[…]

### 2025-02-25 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
added methoods

@@
 
+Study site
+
+- description of srs
+
+Our study took place at the Savannah River Site(SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals(Tewksbury et al.
+2002).
+Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna.
+In each replicant the central patch (100 x 100 m) is always connected to one peripheral patch with identical dimensions by a 150 x 25 m corridor, this will hereafter be referred  to as the connected patch.
+The remaining patches are either "winged" or "rectangular".
+The winged patch is also 100 x 100 m, however they exhibit their characteristic wings in the form of two 75 x 25 m offshoots meant to account for the extra area and edge space the corridor provides.
+The rectangular patch is 100 x 137.5 m also the same area as the space of the connected patch plus the corridor.
+Each block has a duplicate of either the winged or rectangle patch, all peripheral patches being 150 m from the center patch.
+For this study sampling was done in one of each patch type and in one matrix plot per block, all matrix blocks were set up 150 m away from the center as well.
+
+Dung beetle sampling
+
+In the months of July and August 2024 dung beetles were sampled in 4 blocks spread across SRS, baited pitfall traps were placed in one of each patch type and in one matrix plot per block.
+Traps were placed in groups of 3 in the centers of each patch approximately 250 meters from the midpoint of the central patch 40 m from patch edge.
+Pitfalls were oriented in a triangular pattern with the bottom two traps positioned towards the center patch, each trap 20 m apart.
+Plots in the matrix were set up in a similar fashion with the center point 250 m from the center placed equidistant between adjacent patches.
+For each sample period, traps were baited with pig dung between 8-9 pm and picked up 12 hours later, all beetles captured were stored in ethanol for further processing.
+In total 16 sampling rounds were carried out with 4 rounds per block, 196 samples were collected.
+All dung beetles were counted and identified to species with the exception of beetles of the genus Aphodius, overall 15 species were identified and approximately 5300 individual beetles were collected.
+
 ## Study site

### 2025-02-26 corridor_docs/eric_ms_thesis/02_data_analysis.R
run data clean up without 4 hr data

@@ -315,9 +315,15 @@ rarecurve(rarefaction_data,
 # sandbox -----------------------------------------------------------------
 
 
+hill_num = renyi(data, hill= TRUE)
 
-
-
+btl_order<-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n)))
+btl_order
 
 
 

### 2025-02-26 corridor_docs/eric_ms_thesis/01_data_cleanup.R
hills and bray curtis

@@ -80,7 +80,7 @@ btl_counts<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw",
   rowwise() %>%
   mutate(sum = sum(c_across(pvin:ocon), na.rm = TRUE)) %>% 
   mutate(sum=ifelse(sum==0,NA,sum)) %>% 
-  select(-c("...18","...19")) %>% 
+#  select(-c("...18","...19")) %>% 
   replace(is.na(.), 0) %>% 
   left_join(smpl_pts) %>% 
   relocate(c(block,patch, point, month, day,year,date),.after = sample_id) 

### 2025-02-26 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
hills and bray curtis
not too sure its 5am

@@
 
-Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes [@haddad_corridor_2003].
  These movements are hypothesized to prevent species diversity from declining in fragments, as well as help maintain the ecosystem services provided by these species (at both the patch- and landscape-level)[@burt_ants_2022].
@@
 
+As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
+This degradation lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
+Effects from isolation can vary, however as habitats are broken down community structures are significantly altered [@WOS:000419965700012].
+Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003].
+
 What?
@@
 Connectivity and fragmentation 
-
 understanding how movement corridors effect populations dung beetles 
-
 Dispersal ability
@@
 Understanding responses to landscape change allows us to anticipate how environments might change with further disturbance. 
-
 Patterns in population distributions might suggest how we can study focal organisms in the future.
-
 Could help understand how interactions may be changing.
@@
 
-Effects of landscape change are already very well studied, accross different taxa
-
-effects of edge, patch size, and direct habitat loss
-
-using corridors to connect fragmented landscapes is also well studied and for many taxa
-
+Effects of landscape change are already very well studied, accross different taxa .
+effects of edge, patch size, and direct habitat loss.
+using corridors to connect fragmented landscapes is also well studied and for many taxa.
 dung beetles as a model system for studying movement, functional diversity, and ecosystem health.
-
-dung beetle populations in connected and fragmented habitats
-
-insect declines in fragmented landscapes
+dung beetle populations in connected and fragmented habitats .
+insect declines in fragmented landscapes.
 
@@
 
-Our study took place at the Savannah River Site(SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals(Tewksbury et al.
-2002).
+Our study took place at the Savannah River Site(SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals[@tewksbury_corridors_2002].
 Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna.
@@
 # RESULTS
-
+git config --global user.email "you@example.com"
 some summary statistics:

### 2025-02-26 corridor_docs/eric_ms_thesis/02_data_analysis.R

@@ -314,30 +314,58 @@ rarecurve(rarefaction_data,
 
 # sandbox -----------------------------------------------------------------
 
+#example hill function
+#hill_num = renyi(data, hill= TRUE)
 
-hill_num = renyi(data, hill= TRUE)
+#btl_order<-btl_data %>%
+  #pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  #group_by(species) %>%
+  #summarize(n=sum(n, na.rm=TRUE)) %>%
+  #arrange(desc(n)) %>%
+  #mutate(species = reorder(species, desc(n)))
+#btl_order
 
-btl_order<-btl_data %>%
-  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
-  group_by(species) %>%
-  summarize(n=sum(n, na.rm=TRUE)) %>%
-  arrange(desc(n)) %>%
-  mutate(species = reorder(species, desc(n)))
-btl_order
-
-
-
-
-
-
-
-
+#btl_data_sums<-btl_data %>%
+ # pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  #group_by(species,patch) %>%
+  #summarize(n=sum(n, na.rm=TRUE)) %>%
+  #arrange(patch,desc(n))
 
 
+#btl_data_sums<-btl_data %>%
+ # pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  #group_by(species,patch) %>%
+  #summarize(n=sum(n, na.rm=TRUE)) %>%
+  #arrange(patch,desc(n))
 
+# species list by patch type
+btl_spec_patch<- btl_data |>
+  pivot_longer(pvin:ostr, names_to = "species", values_to = "n") |>
+  group_by(species,patch) |>
+  summarise(n=sum(n, na.rm=TRUE)) |>
+  arrange(patch,desc(n)) |>
+  pivot_wider(values_from=n,names_from=species) |> 
+  column_to_rownames("patch")
+# hill number comparisson 
+hill_btl <- renyi(hill_btl_data, hill = TRUE)
 
+plot(hill_btl)
 
+# bray curtis
+#example
+#taxa_bray = vegan::vegdist(data, method = "bray", binary = FALSE)
+#taxa_bray
 
+taxa_bray <- vegan::vegdist(btl_spec_patch, method = "bray", binary = FALSE)
+taxa_bray
+
[…]

### 2025-02-26 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
thesisi planning and some rarefaction

@@
 Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003].
-
-What?
-
-Connectivity and fragmentation 
-understanding how movement corridors effect populations dung beetles 
-Dispersal ability
-
-Why?
-
-Understanding responses to landscape change allows us to anticipate how environments might change with further disturbance. 
-Patterns in population distributions might suggest how we can study focal organisms in the future.
-Could help understand how interactions may be changing.
+As disturbance continues to intensify, it is becoming increasingly more important to understand how different taxonomic groups.
+Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes.
 
@@
 How corridors directly impact dung beetle communitites
+are changes driven by dispersal ability
 
-are changes driven by dispersal ability
+Dung beetles are an incredibly well studied group of insects which play an important role in providng ecosystem services for dung removal, secondary seed distribution, and even suppressing populations of parasitic pests[@shepherd_dung_1998; @WOS:000369463100010].
+Studies have also focused on how landscape structures alter community compositions[@WOS:000398662200001], yet direect
 
@@
 All dung beetles were counted and identified to species with the exception of beetles of the genus Aphodius, overall 15 species were identified and approximately 5300 individual beetles were collected.
+
+Insert description above of individual trap
 
@@
 # RESULTS
-git config --global user.email "you@example.com"
 some summary statistics:
@@
 ### Dedication 
-Your dedication is typed here.
-It should begin with the word “To.” (To my Mom is a typical dedication) If your dedication is longer than a single line it should be single-spaced, and centered vertically and horizontally
+To my family who never stopped supporting me along this journey, my friends who kept me company along the way, and my mentors at VCU who believed in me before I did myself.
 

### 2025-02-26 corridor_docs/eric_ms_thesis/02_data_analysis.R

@@ -368,7 +368,32 @@ btl_spec_patch<- btl_data |>
   column_to_rownames("patch")
 
 
+rarefaction_data<-btl_data %>% 
+  select(-c(month,day,year,date)) %>% 
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(sample_id,species) %>% 
+  summarize(n=sum(n)) %>% 
+  pivot_wider(values_from=n,names_from=species) %>% 
+  column_to_rownames("sample_id") %>% 
+  filter((rowSums(rarefaction_data) > 10)) 
 
+s <- specnumber(btl_spec_patch)
+s
+(raremax <- min(rowSums(btl_spec_patch)))
+
+btl.rare <- rarefy(btl_spec_patch, raremax)
+btl.rare
+plot(s,btl.rare,xlab = "Observed No. of Species", ylab = "Rarefied No. of Species")
+
+
+rarecurve(btl_spec_patch, step = 20, sample = raremax, col = "blue", cex = 0.6)
+
+rarecurve(rarefaction_data, 
+          step = 20, 
+          sample = raremax, 
+          col = "blue", 
+          cex = 0.6,
+          label=FALSE)
 
 
 

### 2025-02-26 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
figures and tables

@@
 ---
+
+```{r GlobalOptions, include = FALSE}
+options(knitr.duplicate.label = 'allow')
+```
+
+```{r setup, include = FALSE}
+
+knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning=FALSE)
+library(here)
+library(tidyverse)
+library(kableExtra)
+# library(gridExtra)
+# library(egg)
+# library(magick)
+# library(citr)
+# library(devtools)
+# devtools::install_github("crsh/papaja")
+```
 
@@
 The author should have replaced this paragraph with their own.
+
+```{r Table1, echo=FALSE,message = FALSE,warning=FALSE}
+# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
+
+btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
+spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+
+btl_order<-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n))) %>% 
+  rename(sp_code=species)
+
+btl_sums_patch<-btl_data %>%
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(species,patch) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(patch,desc(n)) %>% 
+  rename(sp_code=species,
+         n_patch=n) %>% 
+  pivot_wider(
+    names_from = patch,
+    values_from = c(n_patch))
+
+btl_order<-full_join(btl_order,btl_sums_patch,by="sp_code")
+
+spp_table<-btl_order %>% 
+  left_join(spp_codes,by="sp_code") %>% 
+  unite("species",genus:species, sep=" ") %>% 
+  mutate(species=if_else(sp_code=="ap*", "Aphodius spp.",species)) %>% 
+  mutate(c=case_when(
+  c == 0 ~ " ",
+  c > 0 ~ "x",
+  .default = as.character(c)
+)) %>% 
+  mutate(m=case_when(
+  m == 0 ~ " ",
+  m > 0 ~ "x",
+  .default = as.character(m)
+)) %>% 
+  mutate(r=case_when(
+  r == 0 ~ " ",
+  r > 0 ~ "x",
+  .default = as.character(r)
+)) %>% 
+mutate(w=case_when(
+  w == 0 ~ " ",
+  w > 0 ~ "x",
+  .default = as.character(w)
+)) %>% 
+  select(species,n,m,c,w,r) %>% 
+  rename(Species=species,
+         N=n,
+         Matrix=m,
+         Corridor=c,
+         Winged=w,
+         Rectangular=r)
+
+spp_table<-knitr::kable(spp_table, 
+      digits = 2,
+      align="lrcccc",
+      format="latex",
+      row.names = FALSE,
+      booktabs=T,
+      linesep = "", #removes the blank line after every 5 lines
+      caption = "Dung beetle species sampled in the SRS site and their total abundance over the course of the study.") %>%
+  kable_styling(bootstrap_options = c("striped"),
+                full_width = F,
+                latex_options="scale_down",
+                font_size = 12,
+                position = "left") %>% 
+  column_spec(1, italic = T) %>% 
+  row_spec(0, bold = T) %>% 
+  row_spec(1:15, hline_after = T) 
+spp_table
+```
+
+```{r Table2, echo=FALSE,message = FALSE,warning=FALSE}
+# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
+
+btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
+spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+
+```

### 2025-02-26 corridor_docs/eric_ms_thesis/02_data_analysis.R

@@ -7,7 +7,8 @@ library(tidyverse)
 library(here)
 library(lubridate) # to deal with dates https://lubridate.tidyverse.org/
 library(vegan)
-
+library(permute)
+library(lattice)
 # load data ---------------------------------------------------------------
 
 btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
@@ -158,7 +159,7 @@ btl_sums_patch<-btl_data %>%
 # summaries by patch type -------------------------------------------------
 
 
-btl_data_sums<-btl_data %>%
+btl_sums_patch<-btl_data %>%
   pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
   group_by(species,patch) %>%
   summarize(n=sum(n, na.rm=TRUE)) %>%
@@ -344,13 +345,21 @@ btl_spec_patch<- btl_data |>
   group_by(species,patch) |>
   summarise(n=sum(n, na.rm=TRUE)) |>
   arrange(patch,desc(n)) |>
-  pivot_wider(values_from=n,names_from=species) |> 
+  pivot_wider(values_from=n,names_from=species) |>
+  mutate(patch=case_when(
+    patch == "c" ~ "Corridor",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) |>
   column_to_rownames("patch")
 # hill number comparisson 
-hill_btl <- renyi(hill_btl_data, hill = TRUE)
+hill_btl <- renyi(btl_spec_patch, hill = TRUE)
 
-plot(hill_btl)
-
+hill_plot<-plot(hill_btl)
+hill_plot
+# ggsave(hill_plot, 
+#        "hill_plot.png")
 # bray curtis
 #example
 #taxa_bray = vegan::vegdist(data, method = "bray", binary = FALSE)

### 2025-02-26 corridor_docs/eric_ms_thesis/02_data_analysis.R
fig2

@@ -153,7 +153,18 @@ btl_sums_patch<-btl_data %>%
   pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
   group_by(patch) %>%
   summarize(n=sum(n, na.rm=TRUE)) %>%
-  arrange(patch,desc(n))
+  arrange(patch,desc(n)) %>% 
+  mutate(patch=case_when(
+    patch == "c" ~ "Corridor",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) 
+
+ggplot(btl_sums_patch, aes(x=as.factor(patch),
+                           y=n)) +
+  geom_bar(stat = "identity") +
+  theme_classic()
 
 
 # summaries by patch type -------------------------------------------------

### 2025-02-26 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
added table and fig

@@
 
+btl_sums_patch<-btl_data %>%
+  select(-sum) %>% 
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(patch) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(patch,desc(n)) %>% 
+  mutate(patch=case_when(
+    patch == "c" ~ "Corridor",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch)))
+
+abund_patch<-knitr::kable(btl_sums_patch, 
+      digits = 2,
+      align="lr",
+      format="latex",
+      row.names = FALSE,
+      booktabs=T,
+      linesep = "", #removes the blank line after every 5 lines
+      caption = "Total dung beetles captured in all replicates of a patch type.") %>%
+  kable_styling(bootstrap_options = c("striped"),
+                full_width = F,
+                latex_options="scale_down",
+                font_size = 12,
+                position = "left") %>% 
+  # column_spec(1, italic = T) %>% 
+  row_spec(0, bold = T) 
+  # row_spec(1:15, hline_after = T) 
+abund_patch
+
 ```

### 2025-02-26 corridor_docs/eric_ms_thesis/02_data_analysis.R

@@ -150,6 +150,7 @@ btl_order_r <-btl_data %>%
 
 
 btl_sums_patch<-btl_data %>%
+  select(-sum) %>% 
   pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
   group_by(patch) %>%
   summarize(n=sum(n, na.rm=TRUE)) %>%
@@ -164,6 +165,57 @@ btl_sums_patch<-btl_data %>%
 ggplot(btl_sums_patch, aes(x=as.factor(patch),
                            y=n)) +
   geom_bar(stat = "identity") +
+  theme_classic()
+
+
+
+
+# avg abundance per patch type --------------------------------------------
+
+
+
+btl_avg_patch<-btl_data %>%
+  select(-sum) %>% 
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(block,patch) %>%
+  summarize(sum=sum(n, na.rm=TRUE)) %>%
+  group_by(patch) %>%
+  summarize(avg_n=mean(sum, na.rm=TRUE),
+            sd_n=sd(sum, na.rm=TRUE)) %>%
+  arrange(desc(avg_n)) %>% 
+  mutate(patch=case_when(
+    patch == "c" ~ "Corridor",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) 
+
+ggplot(btl_avg_patch, aes(x=as.factor(patch),
+                           y=avg_n)) +
+  geom_bar(stat = "identity") +
+  theme_classic()
+
+
+
+
+btl_avg_patch2<-btl_data %>%
+  select(-sum) %>% 
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(block,patch) %>%
+  summarize(sum=sum(n, na.rm=TRUE)) %>%
+  mutate(patch=case_when(
+    patch == "c" ~ "Corridor",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch)))
+
+
+
+ggplot(btl_avg_patch2, aes(x=as.factor(patch),
+                           y=sum)) +
+  geom_boxplot(stat = "boxplot") +
+  geom_jitter(color="blue", size=0.4, alpha=0.9) +
   theme_classic()
 
 

### 2025-02-28 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
inserted a bunch

@@
 title: "Influence of Connectivity on Dung Beetle Communities"
-author: "ERIC IN ALL CAPS"
+author: "ERIC ESCOBAR-CHENA"
 
@@
 
-Thesis abstracts should be 250 words or less.
+Habitat fragmentation threatens biodiversity across the globe as habitat loss, isolation, and edge effects become increasingly prevalent.
+Corridors have become an important tool in order to combat the negative effects of fragmentation, however they are difficult to study in natural systems without incurring confounding effects.
+To observe changes in insect community composition as an effect of landscape features we sampled dung beetles in a landscape scale experiment.
+We did not see a difference in species richness or diversity, but dung beetle abundances were higher in continuous forest habitat and open habitat patches connected by a corridors than in isolated patches.
 
@@
 
+As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
+This degradation lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
+Effects from isolation can vary, however as habitats are broken down community structures are significantly altered [@WOS:000419965700012].
+Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003].
+As disturbance continues to intensify, it is becoming increasingly more important to understand how different taxonomic groups.
+Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes.
+
+Dung beetles have emerged as a model system with which to test spatial ecology hypotheses and  [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which play an important role in providng ecosystem services for dung removal, secondary seed distribution, and even suppressing populations of parasitic pests[@shepherd_dung_1998; @WOS:000369463100010].
+Past studies have also focused on how landscape structure alters the community compositions of dung beetles[@WOS:000398662200001], yet large landscape scale experimental studies with carefully controlled and replicated treatments are non-existent for this model species.
+
+Here, we  aim to determine how connectivity and fragmentation affect Species Richness and Diversity, Abundance, and functional diversity.
+We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge.
+To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
+Additionally we asked (2) Are corridors benefiting any one functional trait over another?
+Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling.
+
+#### messy notes
 INTRODUCTION
@@
 
- Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity and their ecosystem services [@ros_how_2012; @roslin_dung_2000].
   The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles such as enhanced nutrient cycling and soil quality and reduction of parasites on methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
@@
 
-Dung beetles are an incredibly well studied group of insects which play an important role in providng ecosystem services for dung removal, secondary seed distribution, and even suppressing populations of parasitic pests[@shepherd_dung_1998; @WOS:000369463100010].
-Studies have also focused on how landscape structures alter community compositions[@WOS:000398662200001], yet direect
+Dung beetles have emerged as a model system with which to test spatial ecology hypotheses and  [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which play an important role in providng ecosystem services for dung removal, secondary seed distribution, and even suppressing populations of parasitic pests[@shepherd_dung_1998; @WOS:000369463100010].
+Past studies have also focused on how landscape structure alters the community compositions of dung beetles[@WOS:000398662200001], yet large landscape scale experimental studies with carefully controlled and replicated treatments are non-existent for this model species.
 
@@
 
+Here, we  aim to determine how connectivity and fragmentation affect Species Richness and Diversity, Abundance, and functional diversity.
+We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge.
+To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
+Additionally we asked (2) Are corridors benefiting any one functional trait over another?
+Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling.
+
 Here sampled the commmunity of of dung beetles at the SRS Corridor Experiment to test the following prediction:
@@
 
-- description of srs
-
-Our study took place at the Savannah River Site(SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals[@tewksbury_corridors_2002].
+Our study took place at the Savannah River Site(SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals(Tewksbury et al.
+2002).
 Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna.
@@
 
+Overall, 4995 total dung beetles were collected belonging to 15 species. The most common species were C. vigilans (n=1300), Ateuchus lecontei (n=1112), and Phanaeus igneus (n=919).
+
 1. total number of beetles _from_ total number of species 
@@
 
-Acknowledgments must be written in complete sentences.
-Do not use direct address.
-For example, instead of Thanks, Mom and Dad!, you should say I thank my parents.
-The heading “ACKNOWLEDGMENTS” uses the 002 CHAPTER TITLE style..
-The paragraphs in this section should use the style called 006 Body Text
+We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site, particularly Sabrie Breland, Ben Overly, and Eva Schwarz.
+We also thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry, and to Nico Acajabon, Jarrett Emory, and Cayla Garmen for their advice in putting together this manuscript.
 
@@
 ### List of Abbreviations
-1. A word to be defined:	Write the definition here. 
+
+1. SRS:	Savannah River Site. 
 1. Another word:	And the list continues with another definition.
@@
 # Biographical Sketch
-A biographical sketch is required of all candidates.
-The biographical sketch should be in narrative form.
-Third person, past tense, it typically includes the educational background of the candidate.
-The author should have replaced this paragraph with their own.
+
+Eric Escobar-Chena completed his Bachelors education at Virginia Commonwealth University in 2023.
+During his time there he developed a fondness for insects which  grew into a curiosity of the natural world.
+He later began to explore this curiosity deeper in beginning his graduate education at the University of Florida as a Master's Student under the supervision of Emilio Bruna.
 

### 2025-02-28 corridor_docs/eric_ms_thesis/thesis_planning.tex
inserted a bunch

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{lineno}
\linenumbers
\usepackage[default]{sourcesanspro}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyfoot{}
\fancyhead[L]{EEC MS Thesis Draft}
\fancyhead[R]{p. \thepage}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Influence of Connectivity on Dung Beetle Communities},
  pdfauthor={ERIC IN ALL CAPS},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Influence of Connectivity on Dung Beetle Communities}
\author{ERIC IN ALL CAPS}
\date{}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Thesis abstracts should be 250 words or less.

\newpage

\hypertarget{introduction}{%
\section{INTRODUCTION}\label{introduction}}

INTRODUCTION

These movements are hypothesized to prevent species diversity from
declining in fragments, as well as help maintain the ecosystem services
provided by these species (at both the patch- and landscape-level)(Burt
et al. 2022). Although there is some evidence that animals disperse
between patches via corridors, and that connected patches have higher
species diversity than unconnected ones, little work to date has
investigated the consequences of these corridor-driven patterns for
ecosystem services.

Dung beetles have emerged as a model system with which to test
hypotheses on how changes in landscape structure driven by human
activities influence biodiversity and their ecosystem services (Roslin
2000, Rös et al. 2012). The removal, breakdown, and burial of animal
feces is an important ecosystem service provided by dung beetles such as
enhanced nutrient cycling and soil quality and reduction of parasites on
methane emissions from dung (Iwasa et al. 2015, Slade et al. 2016).
Local assemblages of dung beetles can be species-rich with species
comprising a broad range of functional traits (e.g., size, foraging
style, resource-use) (deCastro-Arrazola et al. 2023). Previous studies
have shown that isolated patches of habitat frequently have lower dung
beetle diversity and abundance than areas of continuous habitat, as well
as documented their presence in linear strips of habitat that resemble
corridors (Gray et al. 2022). However, it remains unknown if corridors
actually act to reduce the loss of dung beetle species from fragments,
if such declines are influenced by inter-specific differences in
dispersal capability, and what the consequences of these patterns are
for the ecosystems services they provide. One major factor behind this
lack of information is the challenge in finding locations where one can
assess the role of corridors while also while controlling for
confounding factors such as patch size, edge, and corridor length
(Haddad 2015).

\textbf{We sampled the commmunity of of dung beetles at the SRS Corridor
Experiment to test the following prediction:} Species Richness, Species
Diversity, and Functional Diversity will be higher in patches connected
by corridors than in unconnected patches.

\begin{itemize}
\tightlist
\item
  First Paragraph: What is the topic of your introduction and why is it
  important/interesting/relevant?
\end{itemize}

As human disturbances continue to expand into natural landscapes, intact
habitats are becoming increasingly fragmented. This degradation lends to
loss in biodiversity on a global scale and interruptions in ecosystem
processes and functions (Haddad 2015). Effects from isolation can vary,
however as habitats are broken down community structures are
significantly altered (Laurance et al. 2018). Corridors have been shown
to be an important mechanism for facilitating the movement of organisms
through fragmented landscapes with the goal of minimizing negative
consequences of fragmentation(Haddad et al. 2003). As disturbance
continues to intensify, it is becoming increasingly more important to
understand how different taxonomic groups. Here, we aim to gain an
understanding of how dung beetles, a group of insects well known for
strong dispersal ability in order to compete for ephemeral
resources(Hanski and Cambefort 1991), interact with corridors in their
landscapes.

\begin{itemize}
\tightlist
\item
  Second Paragraph: What is known about this topic already?
\end{itemize}

Effects of landscape change are already very well studied, accross
different taxa . effects of edge, patch size, and direct habitat loss.
using corridors to connect fragmented landscapes is also well studied
and for many taxa. dung beetles as a model system for studying movement,
functional diversity, and ecosystem health. dung beetle populations in
connected and fragmented habitats . insect declines in fragmented
landscapes.

studying coprophagous insects

\begin{itemize}
\tightlist
\item
  Third Paragraph: What isn't known about this topic and why might it
  change how we think/act about the topic?
\end{itemize}

How corridors directly impact dung beetle communitites are changes
driven by dispersal ability

Dung beetles are an incredibly well studied group of insects which play
an important role in providng ecosystem services for dung removal,
secondary seed distribution, and even suppressing populations of
parasitic pests(Shepherd and Chapman 1998, Manning et al. 2016). Studies
have also focused on how landscape structures alter community
compositions(Costa et al. 2017), yet direect

at our study site we can directly compare patch connectivity and patch
shape with fragmented landscapes to obtain a strong idea of what
landscape features effect dung beetle collection.

\begin{itemize}
\tightlist
\item
  Fourth Paragraph: Why hasn't this thing been studied/assessed/done
  before?
\end{itemize}

other work at the corridor project but we are doing dung beetles instead

\begin{itemize}
\tightlist
\item
  Fifth Paragraph: Literally the words ``Here we\ldots{}''
\end{itemize}

Here sampled the commmunity of of dung beetles at the SRS Corridor
Experiment to test the following prediction:

Species richness, diversity, and functional diversity will be higher in
patches connected by a movement corridor than in patches that are
unconnected.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Study site

\begin{itemize}
\tightlist
\item
  description of srs
\end{itemize}

Our study took place at the Savannah River Site(SRS), a National
Environmental Research Park in southern South Carolina, US(33.208 N,
81.408 W) in four of seven experimental landscapes designed for the
purposes of directly observing the impacts of corridors and patch shape
on the movements of plants and animals(Tewksbury et al. 2002). Each
experimental landscape, termed blocks, consists of four patches of open
habitat around a central patch all together within a matrix of pine
savanna. In each replicant the central patch (100 x 100 m) is always
connected to one peripheral patch with identical dimensions by a 150 x
25 m corridor, this will hereafter be referred to as the connected
patch. The remaining patches are either ``winged'' or ``rectangular''.
The winged patch is also 100 x 100 m, however they exhibit their
characteristic wings in the form of two 75 x 25 m offshoots meant to
account for the extra area and edge space the corridor provides. The
rectangular patch is 100 x 137.5 m also the same area as the space of
the connected patch plus the corridor. Each block has a duplicate of
either the winged or rectangle patch, all peripheral patches being 150 m
from the center patch. For this study sampling was done in one of each
patch type and in one matrix plot per block, all matrix blocks were set
up 150 m away from the center as well.

Dung beetle sampling

In the months of July and August 2024 dung beetles were sampled in 4
blocks spread across SRS, baited pitfall traps were placed in one of
each patch type and in one matrix plot per block. Traps were placed in
groups of 3 in the centers of each patch approximately 250 meters from
the midpoint of the central patch 40 m from patch edge. Pitfalls were
oriented in a triangular pattern with the bottom two traps positioned
towards the center patch, each trap 20 m apart. Plots in the matrix were
set up in a similar fashion with the center point 250 m from the center
placed equidistant between adjacent patches. For each sample period,
traps were baited with pig dung between 8-9 pm and picked up 12 hours
later, all beetles captured were stored in ethanol for further
processing. In total 16 sampling rounds were carried out with 4 rounds
per block, 196 samples were collected. All dung beetles were counted and
identified to species with the exception of beetles of the genus
Aphodius, overall 15 species were identified and approximately 5300
individual beetles were collected.

Insert description above of individual trap

\hypertarget{study-site}{%
\subsection{Study site}\label{study-site}}

\begin{itemize}
\tightlist
\item
  description of srs
\item
  experimental design
\item
  conditions during sample period
\item
  historical significance of site and experimental design
\item
  justification for selected patches
\end{itemize}

\hypertarget{dung-beetle-sampling}{%
\subsection{Dung beetle sampling}\label{dung-beetle-sampling}}

\begin{itemize}
\tightlist
\item
  structure and arrangment of traps
\item
  description of traps
\item
  bait
\item
  sample period
\item
  ID
\item
  biomass if we do biomass
\end{itemize}

\hypertarget{analyses}{%
\subsubsection{Analyses}\label{analyses}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Species Richness:} absolute number? non-parametric estimators?
\end{enumerate}

\begin{itemize}
\tightlist
\item
  Michaelis-Menten
\item
  choa?
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Species Diversity:} what index should we compare?
\end{enumerate}

\begin{itemize}
\tightlist
\item
  alpha diversity per patch type
\item
  beta between patch types
\item
  hill numbers
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Functional Diversity:} Need to assign each species to a
  functional group: roller, tunneler. dweller, others?
\end{enumerate}

\begin{itemize}
\tightlist
\item
  habitat preference (forest, pasture, generalist)
\end{itemize}

\begin{quote}
\emph{Look through dung beetle pubs and see how/what people compare}
\end{quote}

\begin{itemize}
\tightlist
\item
  lets hammer this out
\item
  modeling?

  \begin{itemize}
  \tightlist
  \item
    glmm with poisson dist reccomended by julian
  \end{itemize}
\item
  beta, abundance, biomass? per site
\item
  species list by sampling blocks (anything with this?)
\item
  habitat preference
\item
  rarefaction
\end{itemize}

\hypertarget{results}{%
\section{RESULTS}\label{results}}

some summary statistics:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  total number of beetles \emph{from} total number of species
\item
  were all species found in all habitats? Were any species found in only
  1 habitat?
\item
  Were all species found in all blocks? Were any restricted to only 1-2
  blocks?
\item
  Were all species found in the matrix? (expect that so, since it is the
  `baseline' or `source' habitat)
\item
  Number of species in each functional group
\item
  most common 3-4 species
\item
  any rare species?
\end{enumerate}

\hypertarget{discussion}{%
\section{DISCUSSION}\label{discussion}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  dont forget o discuss the basic biology\ldots why might a species be
  so common? why might one be rare?
\end{enumerate}

\hypertarget{acknowledgments}{%
\section{ACKNOWLEDGMENTS}\label{acknowledgments}}

Acknowledgments must be written in complete sentences. Do not use direct
address. For example, instead of Thanks, Mom and Dad!, you should say I
thank my parents. The heading ``ACKNOWLEDGMENTS'' uses the 002 CHAPTER
TITLE style.. The paragraphs in this section should use the style called
006 Body Text

\hypertarget{other-required-text}{%
\section{OTHER REQUIRED TEXT}\label{other-required-text}}

\hypertarget{dedication}{%
\subsubsection{Dedication}\label{dedication}}

To my family who never stopped supporting me along this journey, my
friends who kept me company along the way, and my mentors at VCU who
believed in me before I did myself.

\hypertarget{list-of-abbreviations}{%
\subsubsection{List of Abbreviations}\label{list-of-abbreviations}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A word to be defined: Write the definition here.
\item
  Another word: And the list continues with another definition.
\end{enumerate}

\hypertarget{biographical-sketch}{%
\section{Biographical Sketch}\label{biographical-sketch}}

A biographical sketch is required of all candidates. The biographical
sketch should be in narrative form. Third person, past tense, it
typically includes the educational background of the candidate. The
author should have replaced this paragraph with their own.

\begin{table}

\caption{\label{tab:Table1}Dung beetle species sampled in the SRS site and their total abundance over the course of the study.}
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{>{}lrcccc}
\toprule
\textbf{Species} & \textbf{N} & \textbf{Matrix} & \textbf{Corridor} & \textbf{Winged} & \textbf{Rectangular}\\
\midrule
\em{Canthon vigilans} & 1300 & x & x & x & x\\
\midrule
\em{Ateuchus lecontei} & 1112 & x & x & x & x\\
\midrule
\em{Phanaeus igneus} & 919 & x & x & x & x\\
\midrule
\em{Aphodius spp.} & 614 & x & x & x & x\\
\midrule
\em{Dichotomius carolinus} & 547 & x & x & x & x\\
\midrule
\em{Onthophagus pennsylvanicus} & 202 & x & x & x & x\\
\midrule
\em{Phanaeus vindex} & 131 & x & x & x & x\\
\midrule
\em{Melanocanthon bispinatus} & 75 & x & x & x & x\\
\midrule
\em{Boreocanthon probus} & 47 & x & x & x & x\\
\midrule
\em{Copris minutus} & 24 & x & x & x & x\\
\midrule
\em{Deltochilum gibbosum} & 14 & x & x & x & x\\
\midrule
\em{Melanocanthon vulturnatus} & 7 & x & x &  & x\\
\midrule
\em{Onthophagus striatulus} & 3 & x &  &  & x\\
\midrule
\em{Onthophagus concinnus} & 2 & x &  & x & \\
\midrule
\em{Geotrupes blackburnii} & 1 &  &  & x & \\
\midrule
\em{Onthophagus tuberculifrons} & 1 & x &  &  & \\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-burt_ants_2022}{}}%
Burt, M. A., J. Resasco, N. M. Haddad, and S. R. Whitehead. 2022.
\href{https://doi.org/10.1002/ecs2.4324}{Ants disperse seeds farther in
habitat patches with corridors}. Ecosphere 13:e4324.

\leavevmode\vadjust pre{\hypertarget{ref-WOS:000398662200001}{}}%
Costa, C., V. H. F. Oliveira, R. Maciel, W. Beiroz, V. Korasaki, and J.
Louzada. 2017. \href{https://doi.org/10.7717/peerj.3125}{Variegated
tropical landscapes conserve diverse dung beetle communities}. {PEERJ}
5.

\leavevmode\vadjust pre{\hypertarget{ref-WOS:000891747700001}{}}%
deCastro-Arrazola, I., N. R. Andrew, M. P. Berg, A. Curtsdotter, J.-P.
Lumaret, R. Menendez, M. Moretti, B. Nervo, E. S. Nichols, F.
Sanchez-Pinero, A. M. C. Santos, K. S. Sheldon, E. M. Slade, and J.
Hortal. 2023. \href{https://doi.org/10.1111/1365-2656.13829}{A
trait-based framework for dung beetle functional ecology}. {JOURNAL}
{OF} {ANIMAL} {ECOLOGY} 92:44--65.

\leavevmode\vadjust pre{\hypertarget{ref-WOS:000707546700001}{}}%
Gray, R. E. J., L. F. Rodriguez, O. T. Lewis, A. Y. C. Chung, O.
Ovaskainen, and E. M. Slade. 2022.
\href{https://doi.org/10.1111/1365-2664.14049}{Movement of
forest-dependent dung beetles through riparian buffers in bornean oil
palm plantations}. {JOURNAL} {OF} {APPLIED} {ECOLOGY} 59:238--250.

\leavevmode\vadjust pre{\hypertarget{ref-haddad_habitat_2015}{}}%
Haddad, N. M. 2015, March 20. Habitat fragmentation and its lasting
impact on earth's ecosystems {\textbar} science advances.
\url{https://www-science-org.lp.hscl.ufl.edu/doi/10.1126/sciadv.1500052}.

\leavevmode\vadjust pre{\hypertarget{ref-haddad_corridor_2003}{}}%
Haddad, N. M., D. R. Bowne, A. Cunningham, B. J. Danielson, D. J. Levey,
S. Sargent, and T. Spira. 2003.
\href{https://doi.org/10.1890/0012-9658(2003)084\%5B0609:CUBDT\%5D2.0.CO;2}{{CORRIDOR}
{USE} {BY} {DIVERSE} {TAXA}}. Ecology 84:609--615.

\leavevmode\vadjust pre{\hypertarget{ref-WOS:000350444300013}{}}%
Iwasa, M., Y. Moki, and J. Takahashi. 2015.
\href{https://doi.org/10.1093/ee/nvu023}{Effects of the activity of
coprophagous insects on greenhouse gas emissions from cattle dung pats
and changes in amounts of nitrogen, carbon, and energy}. {ENVIRONMENTAL}
{ENTOMOLOGY} 44:106--113.

\leavevmode\vadjust pre{\hypertarget{ref-WOS:000419965700012}{}}%
Laurance, W. F., J. L. C. Camargo, P. M. Fearnside, T. E. Lovejoy, G. B.
Williamson, R. C. G. Mesquita, C. F. J. Meyer, P. E. D. Bobrowiec, and
S. G. W. Laurance. 2018. \href{https://doi.org/10.1111/brv.12343}{An
amazonian rainforest and its fragments as a laboratory of global
change}. {BIOLOGICAL} {REVIEWS} 93:223--247.

\leavevmode\vadjust pre{\hypertarget{ref-WOS:000369463100010}{}}%
Manning, P., E. M. Slade, S. A. Beynon, and O. T. Lewis. 2016.
\href{https://doi.org/10.1016/j.agee.2015.11.007}{Functionally rich dung
beetle assemblages are required to provide multiple ecosystem services}.
{AGRICULTURE} {ECOSYSTEMS} \& {ENVIRONMENT} 218:87--94.

\leavevmode\vadjust pre{\hypertarget{ref-ros_how_2012}{}}%
Rös, M., F. Escobar, and G. Halffter. 2012.
\href{https://doi.org/10.1111/j.1472-4642.2011.00834.x}{How dung beetles
respond to a human-modified variegated landscape in mexican cloud
forest: A study of biodiversity integrating ecological and
biogeographical perspectives}. Diversity and Distributions 18:377--389.

\leavevmode\vadjust pre{\hypertarget{ref-roslin_dung_2000}{}}%
Roslin, T. 2000.
\href{https://doi.org/10.1034/j.1600-0706.2000.910213.x}{Dung beetle
movements at two spatial scales}. Oikos 91:323--335.

\leavevmode\vadjust pre{\hypertarget{ref-shepherd_dung_1998}{}}%
Shepherd, V. E., and C. A. Chapman. 1998.
\href{https://doi.org/10.1017/S0266467498000169}{Dung beetles as
secondary seed dispersers: Impact on seed predation and germination}.
Journal of Tropical Ecology 14:199--215.

\leavevmode\vadjust pre{\hypertarget{ref-slade_disentangling_2016}{}}%
Slade, E. M., T. Roslin, M. Santalahti, and T. Bell. 2016.
\href{https://doi.org/10.1111/oik.02640}{Disentangling the ``brown
world' faecal-detritus interaction web: Dung beetle effects on soil
microbial properties}. Oikos (Copenhagen, Denmark) 125:629--635.

\leavevmode\vadjust pre{\hypertarget{ref-tewksbury_corridors_2002}{}}%
Tewksbury, J. J., D. J. Levey, N. M. Haddad, S. Sargent, J. L. Orrock,
A. Weldon, B. J. Danielson, J. Brinkerhoff, E. I. Damschen, and P.
Townsend. 2002. \href{https://doi.org/10.1073/pnas.202242699}{Corridors
affect plants, animals, and their interactions in fragmented
landscapes}. Proceedings of the National Academy of Sciences
99:12923--12926.

\end{CSLReferences}

\end{document}

### 2025-02-28 corridor_docs/eric_ms_thesis/02_data_analysis.R
pushing stuff

@@ -9,6 +9,7 @@ library(lubridate) # to deal with dates https://lubridate.tidyverse.org/
 library(vegan)
 library(permute)
 library(lattice)
+library(ggmosaic)
 # load data ---------------------------------------------------------------
 
 btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
@@ -477,6 +478,20 @@ N_by_spp<-btl_data %>%
   
 
 
+#df = structure(list(effect = structure(c(2L, 2L, 1L, 1L), .Label = c("yes", 
+#                                                                     "no"), class = "factor"), sex = structure(c(1L, 2L, 1L, 2L), 
+#                                                                                                               .Label = c("f",  "m"), class = "factor"), n = c(8, 3, 8, 12)), 
+#               row.names = c(NA, -4L), class = "data.frame")
+
+#ggplot(df, aes(effect, y=n, fill = sex)) +
+#  geom_col(position="fill")
+
+head(btl_data_sums)
+
+mosaic_data <- filter(btl_data_sums, .by )
+
+ggplot(btl_data_sums, aes(species, y = n, fill = patch)) +
+  geom_col(position = "fill")
 
 # total N by date type --------------------------------------------------
 

### 2025-02-28 corridor_docs/eric_ms_thesis/02_data_analysis.R
doin' Eric a solid

@@ -15,13 +15,114 @@ library(ggmosaic)
 btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
 
 
-# quick summaries to look over the sampling point information ---------------
-# how many points in each block, how many points per treatment, etc.
 
-# names(btl)
-# 
-# smpl_pts %>% 
-#   group_by(block) %>% 
+# TO DO -------------------------------------------------------------------
+
+# total beetles captured --------------------------------------------------
+total_N<-btl_data %>% 
+  summarise(n=sum(sum, na.rm=TRUE))
+total_N
+# How many species total?
+total_spp<-btl_data %>% 
+  select(-sum) %>% 
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n))) %>% 
+  summarise(n_spp=n_distinct(species))
+total_spp
+
+# Number captured per species (all data pooled) ---------------------------
+btl_order<-btl_data %>%
+  select(-sum) %>% 
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n)))
+btl_order
+
+
+# Number captured per block -----------------------------------------------
+N_block<-btl_data %>%
+  select(-sum) %>% 
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  group_by(block) %>%
+  summarize(n=sum(n, na.rm=TRUE))
+N_block
+
+# number species per block ------------------------------------------------
+
+N_spp_block<-btl_data %>%
+  select(-sum) %>% 
+  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  filter(n>0) %>% 
+  group_by(block) %>%
+  summarize(n_spp=n_distinct(species, na.rm=TRUE))
+N_spp_block
+
+
+# count per species per block ---------------------------------------------
+
+btl_order_block<-btl_data %>%
[…]

### 2025-03-02 corridor_docs/eric_ms_thesis/01_data_cleanup.R
last changes for raw data

@@ -78,7 +78,7 @@ write_csv(smpl_pts,"./corridor_docs/eric_ms_thesis/ms_data/data_clean/clean_samp
 btl_counts<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","beetle_counts.csv")) %>% 
   mutate(sample_id=as.factor(sample_id)) %>% 
   rowwise() %>%
-  mutate(sum = sum(c_across(pvin:ocon), na.rm = TRUE)) %>% 
+  mutate(sum = sum(c_across(pvin:osyl), na.rm = TRUE)) %>% 
   mutate(sum=ifelse(sum==0,NA,sum)) %>% 
 #  select(-c("...18","...19")) %>% 
   replace(is.na(.), 0) %>% 

### 2025-03-02 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
huhh

@@
 btl_order<-btl_data %>%
-  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
   group_by(species) %>%
@@
 btl_sums_patch<-btl_data %>%
-  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
   group_by(species,patch) %>%
@@
   select(-sum) %>% 
-  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
   group_by(patch) %>%
@@
 
-abund_patch<-knitr::kable(btl_sums_patch, 
+abund_patch <- knitr::kable(btl_sums_patch, 
       digits = 2,

### 2025-03-02 corridor_docs/eric_ms_thesis/02_data_analysis.R

@@ -25,7 +25,7 @@ total_N
 # How many species total?
 total_spp<-btl_data %>% 
   select(-sum) %>% 
-  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
   group_by(species) %>%
   summarize(n=sum(n, na.rm=TRUE)) %>%
   arrange(desc(n)) %>%
@@ -36,7 +36,7 @@ total_spp
 # Number captured per species (all data pooled) ---------------------------
 btl_order<-btl_data %>%
   select(-sum) %>% 
-  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
   group_by(species) %>%
   summarize(n=sum(n, na.rm=TRUE)) %>%
   arrange(desc(n)) %>%
@@ -47,7 +47,7 @@ btl_order
 # Number captured per block -----------------------------------------------
 N_block<-btl_data %>%
   select(-sum) %>% 
-  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
   group_by(block) %>%
   summarize(n=sum(n, na.rm=TRUE))
 N_block
@@ -56,7 +56,7 @@ N_block
 
 N_spp_block<-btl_data %>%
   select(-sum) %>% 
-  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
   filter(n>0) %>% 
   group_by(block) %>%
   summarize(n_spp=n_distinct(species, na.rm=TRUE))
@@ -67,7 +67,7 @@ N_spp_block
 
 btl_order_block<-btl_data %>%
   select(-sum) %>% 
-  pivot_longer(pvin:ostr,names_to = "species",values_to = "n") %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
   group_by(block,species) %>%
   summarize(n=sum(n, na.rm=TRUE)) %>%
   arrange(block,desc(n)) %>%
@@ -81,7 +81,7 @@ btl_order_block
 
[…]

### 2025-03-02 corridor_docs/eric_ms_thesis/02_data_analysis.R
diversity and richness metrics, methods

@@ -10,6 +10,7 @@ library(vegan)
 library(permute)
 library(lattice)
 library(ggmosaic)
+library(iNEXT)
 # load data ---------------------------------------------------------------
 
 btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
@@ -427,14 +428,14 @@ rarefaction_data<-btl_data %>%
 #rarecurve(BCI, step = 20, sample = Raremax, col = "blue", cex = 0.6)
 
 # subsetting data so that there are no zeros 
-rarefaction_data<-btl_data %>% 
-  select(-c(month,day,year,date)) %>% 
-  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
-  group_by(sample_id,species) %>% 
-  summarize(n=sum(n)) %>% 
-  pivot_wider(values_from=n,names_from=species) %>% 
-  column_to_rownames("sample_id") %>% 
-  filter((rowSums(rarefaction_data) > 10)) 
+#rarefaction_data<-btl_data %>% 
+  #select(-c(month,day,year,date)) %>% 
+  #pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  #group_by(sample_id,species) %>% 
+  #summarize(n=sum(n)) %>% 
+  #pivot_wider(values_from=n,names_from=species) %>% 
+  #column_to_rownames("sample_id") %>% 
+  #filter((rowSums(rarefaction_data) > 10)) 
 
 s <- specnumber(rarefaction_data)
 s
@@ -575,7 +576,187 @@ mosaic_data <- filter(btl_data_sums, .by )
 ggplot(btl_data_sums, aes(species, y = n, fill = patch)) +
   geom_col(position = "fill")
 
-# total N by date type --------------------------------------------------
+# functions for ES TES and plot.TES
+
+ES <-  function (x,m=1,method=c("a","b"))
+{ 
+  method <- match.arg(method, c("a", "b"))
+  
+  if (m<1){warning("m must be a positive value");break}
+  if (m%%1!=0)warning("results may be meaningless because m is not an integer")
+  if (any(x < 0, na.rm = TRUE)) 
+  {warning("data have negative entries");break}
+  if (any(is.na(x))) 
[…]

### 2025-03-03 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
WRITINNG

@@
 In total 16 sampling rounds were carried out with 4 rounds per block, 196 samples were collected.
-All dung beetles were counted and identified to species with the exception of beetles of the genus Aphodius, overall 15 species were identified and approximately 5300 individual beetles were collected.
+All beetles of subfamily Scarabaeinae  were counted and identified to species using keys from @nemes_illustrated_2015 ### update bib.
+Overall 16 species were identified and approximately 5200 in total individual beetles were collected.
+
+Analyses
+
+Biodiversity between patch types was compared using Hill numbers (Jost 2006).
+We looked at community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
+Diversity numbers were calculated in R studio using package iNEXT (Chao et al.
+2016).
+
+Dung beetles were assigned traits by
 
@@
 ```
+
+```{r Table3, echo=FALSE,message = FALSE,warning=FALSE}
+library(iNEXT)
+
+btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
+spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+
+btl_sums_patch<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(species,patch) %>%
+  mutate(patch=case_when(
+    patch == "c" ~ "Corridor",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(patch,desc(n)) %>% 
+  rename(sp_code=species,
+         n_patch=n) %>% 
+  pivot_wider(
+    names_from = patch,
+    values_from = c(n_patch))
+
+div.data <- btl_sums_patch |>
+  column_to_rownames(var = "sp_code")
+
+div.btl <- iNEXT(div.data, q = c(0,1,2), datatype = "abundance")
+
+z <- iNEXT(div.data, q=c(0,1,2), datatype="abundance")
+ggiNEXT(z, facet.var="Order.q", color.var="Assemblage") + 
+  theme_bw(base_size=18)
+
+```

### 2025-03-03 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
writing

@@
 
-Dung beetles have emerged as a model system with which to test spatial ecology hypotheses and  [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which play an important role in providng ecosystem services for dung removal, secondary seed distribution, and even suppressing populations of parasitic pests[@shepherd_dung_1998; @WOS:000369463100010].
+Dung beetles have emerged as a model system with which to test spatial ecology hypotheses  [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al.
+2024).
+The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles such as enhanced nutrient cycling and soil quality and reduction of parasites on methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
+Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) [@WOS:000891747700001].
+Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].
 Past studies have also focused on how landscape structure alters the community compositions of dung beetles[@WOS:000398662200001], yet large landscape scale experimental studies with carefully controlled and replicated treatments are non-existent for this model species.
@@
 Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling.
-
-#### messy notes
-INTRODUCTION
-
- These movements are hypothesized to prevent species diversity from declining in fragments, as well as help maintain the ecosystem services provided by these species (at both the patch- and landscape-level)[@burt_ants_2022].
-Although there is some evidence that animals disperse between patches via corridors, and that connected patches have higher species diversity than unconnected ones, little work to date has investigated the consequences of these corridor-driven patterns for ecosystem services.
-
-  The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles such as enhanced nutrient cycling and soil quality and reduction of parasites on methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
-Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) [@WOS:000891747700001].
-Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].
-However, it remains unknown if corridors actually act to reduce the loss of dung beetle species from fragments, if such declines are influenced by inter-specific differences in dispersal capability, and what the consequences of these patterns are for the ecosystems services they provide.
-One major factor behind this lack of information is the challenge in finding locations where one can assess the role of corridors while also while controlling for confounding factors such as patch size, edge, and corridor length [@haddad_habitat_2015].
-
- **We sampled the commmunity of of dung beetles at the SRS Corridor Experiment to test the following prediction:** Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches.
-
-- First Paragraph: What is the topic of your introduction and why is it important/interesting/relevant?
-
-As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
-This degradation lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
-Effects from isolation can vary, however as habitats are broken down community structures are significantly altered [@WOS:000419965700012].
-Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003].
-As disturbance continues to intensify, it is becoming increasingly more important to understand how different taxonomic groups.
-Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes.
-
-- Second Paragraph: What is known about this topic already?
-
-Effects of landscape change are already very well studied, accross different taxa .
-effects of edge, patch size, and direct habitat loss.
-using corridors to connect fragmented landscapes is also well studied and for many taxa.
-dung beetles as a model system for studying movement, functional diversity, and ecosystem health.
-dung beetle populations in connected and fragmented habitats .
-insect declines in fragmented landscapes.
-
-studying coprophagous insects
-
-- Third Paragraph: What isn't known about this topic and why might it change how we think/act about the topic?
-
-How corridors directly impact dung beetle communitites
-are changes driven by dispersal ability
-
-Dung beetles have emerged as a model system with which to test spatial ecology hypotheses and  [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which play an important role in providng ecosystem services for dung removal, secondary seed distribution, and even suppressing populations of parasitic pests[@shepherd_dung_1998; @WOS:000369463100010].
-Past studies have also focused on how landscape structure alters the community compositions of dung beetles[@WOS:000398662200001], yet large landscape scale experimental studies with carefully controlled and replicated treatments are non-existent for this model species.
-
-at our study site we can directly compare patch connectivity and patch shape with fragmented landscapes to obtain a strong idea of what landscape features effect dung beetle collection.
-
-- Fourth Paragraph: Why hasn't this thing been studied/assessed/done before?
-
-other work at the corridor project but we are doing dung beetles instead
-
-- Fifth Paragraph: Literally the words "Here we..."
-
-Here, we  aim to determine how connectivity and fragmentation affect Species Richness and Diversity, Abundance, and functional diversity.
-We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge.
-To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
-Additionally we asked (2) Are corridors benefiting any one functional trait over another?
-Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling.
-
-Here sampled the commmunity of of dung beetles at the SRS Corridor Experiment to test the following prediction:
-
-Species richness, diversity, and functional diversity will be higher in patches connected by a movement corridor than in patches that are unconnected.
 
@@
 2016).
-
-Dung beetles were assigned traits by
+Dung beetles were assigned traits by waste removal guild and habitat preference.
 
@@
 # RESULTS
-some summary statistics:
-
-Overall, 4995 total dung beetles were collected belonging to 15 species. The most common species were C. vigilans (n=1300), Ateuchus lecontei (n=1112), and Phanaeus igneus (n=919).
-
-1. total number of beetles _from_ total number of species 
-1. were all species found in all habitats? Were any species found in only 1 habitat?
-1. Were all species found in all blocks? Were any restricted to only 1-2 blocks?
-1. Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
-1. Number of species in each functional group
-1. most common 3-4 species
-1. any rare species?
-
-# DISCUSSION
-
-1. dont forget o discuss the basic biology...why might a species be so common? why might one be rare?
-
-# ACKNOWLEDGMENTS
-
-We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site, particularly Sabrie Breland, Ben Overly, and Eva Schwarz.
-We also thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry, and to Nico Acajabon, Jarrett Emory, and Cayla Garmen for their advice in putting together this manuscript.
-
-# OTHER REQUIRED TEXT
-
-### Dedication 
-To my family who never stopped supporting me along this journey, my friends who kept me company along the way, and my mentors at VCU who believed in me before I did myself.
-
-### List of Abbreviations
-
-1. SRS:	Savannah River Site. 
-1. Another word:	And the list continues with another definition.
-
-# Biographical Sketch
-
-Eric Escobar-Chena completed his Bachelors education at Virginia Commonwealth University in 2023.
-During his time there he developed a fondness for insects which  grew into a curiosity of the natural world.
-He later began to explore this curiosity deeper in beginning his graduate education at the University of Florida as a Master's Student under the supervision of Emilio Bruna.
 
@@
 
+Overall, 5,213 total dung beetles were collected belonging to 16 species.
+The most common species were C.
+vigilans (n=1473), Ateuchus lecontei (n=1115), and Phanaeus igneus (n=958).
+All but one species was found in matrix plots at fifteen species, with the exception of Geotrupes blackburnii, one of our only singletons along with Onthophagous turberculifrons.
+The winged patch had the next most species at fourteen, rectangular followed with thirteen, and connected patches had the fewest species at twelve (Table 1).
+Beetle abundance was significantly higher in matrix plots with 33% (n = 1713) of captures followed by 26% (n = 1359) caught in connected patches, 23% (n = 1199) in winged, and lastly 18% (n = 942) from rectanglar patches (Table 2).
+
 ```{r Table2, echo=FALSE,message = FALSE,warning=FALSE}
@@
 
+Assemblage composition was mostly uniform across patch types, however
+
+1. total number of beetles _from_ total number of species 
+1. were all species found in all habitats? Were any species found in only 1 habitat?
+1. Were all species found in all blocks? Were any restricted to only 1-2 blocks?
+1. Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
+1. Number of species in each functional group
+1. most common 3-4 species
+1. any rare species?
+
+# DISCUSSION
+
+1. dont forget o discuss the basic biology...why might a species be so common? why might one be rare?
+
+# ACKNOWLEDGMENTS
+
+We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site, particularly Sabrie Breland, Ben Overly, and Eva Schwarz.
+We also thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry, and to Nico Acajabon, Jarrett Emory, and Cayla Garmen for their advice in putting together this manuscript.
+
+# OTHER REQUIRED TEXT
+
+### Dedication 
+To my family who never stopped supporting me along this journey, my friends who kept me company along the way, and my mentors at VCU who believed in me before I did myself.
+
+### List of Abbreviations
+
+1. SRS:	Savannah River Site. 
+1. Another word:	And the list continues with another definition.
+
+# Biographical Sketch
+
+Eric Escobar-Chena completed his Bachelors education at Virginia Commonwealth University in 2023.
+During his time there he developed a fondness for insects which  grew into a curiosity of the natural world.
+He later began to explore this curiosity deeper in beginning his graduate education at the University of Florida as a Master's Student under the supervision of Emilio Bruna.
+
 ```{r Table3, echo=FALSE,message = FALSE,warning=FALSE}

### 2025-03-03 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
pushibg

@@
 
-```{r Table3, echo=FALSE,message = FALSE,warning=FALSE}
+```{r Figure 1, echo=FALSE,message = FALSE,warning=FALSE}
 library(iNEXT)
@@
 ```
+
+```{r Table 3, echo=FALSE,message = FALSE,warning=FALSE}
+btl_spec_patch<- btl_data |>
+  pivot_longer(pvin:osyl, names_to = "species", values_to = "n") |>
+  group_by(species,patch) |>
+  summarise(n=sum(n, na.rm=TRUE)) |>
+  arrange(patch,desc(n)) |>
+  pivot_wider(values_from=n,names_from=species) |>
+  mutate(patch=case_when(
+    patch == "c" ~ "Corridor",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) |>
+  column_to_rownames("patch")
+
+taxa_bray <- vegan::vegdist(btl_spec_patch, method = "bray", binary = FALSE)
+taxa_bray
+
+```

### 2025-03-03 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
results

@@
 2016).
+Bray-Curtis dissimiolarity values were calculated using package Vegan in R studio.
 Dung beetles were assigned traits by waste removal guild and habitat preference.
@@
 Insert description above of individual trap
-
-## Study site
-
-  - description of srs
-  - experimental design 
-  - conditions during sample period
-  - historical significance of site and experimental design 
-  - justification for selected patches
-
-## Dung beetle sampling
-
-  - structure and arrangment of traps 
-  - description of traps
-  - bait 
-  - sample period
-  - ID 
-  - biomass if we do biomass
-
-### Analyses
-
-1. **Species Richness:** absolute number? non-parametric estimators?
-  - Michaelis-Menten
-  - choa?
-1. **Species Diversity:** what index should we compare?
-  - alpha diversity per patch type
-  - beta between patch types
-  - hill numbers
 
@@
 Overall, 5,213 total dung beetles were collected belonging to 16 species.
-The most common species were C.
-vigilans (n=1473), Ateuchus lecontei (n=1115), and Phanaeus igneus (n=958).
-All but one species was found in matrix plots at fifteen species, with the exception of Geotrupes blackburnii, one of our only singletons along with Onthophagous turberculifrons.
-The winged patch had the next most species at fourteen, rectangular followed with thirteen, and connected patches had the fewest species at twelve (Table 1).
-Beetle abundance was significantly higher in matrix plots with 33% (n = 1713) of captures followed by 26% (n = 1359) caught in connected patches, 23% (n = 1199) in winged, and lastly 18% (n = 942) from rectanglar patches (Table 2).
+The most dominant species were Canthon vigilans (n=1473), Ateuchus lecontei (n=1115), and Phanaeus igneus (n=958) (Table 1).
+All but one species was found in matrix plots.
+Four species (Onthophagus striatulus, Onthophagus concinnus, Geotrupes Blackburnii, Onthophagus truberculifrons) were the only species not representative in every patch type.
+We had two singleton species, Geotrupes Blackburnii appearing only once in winged, and Onthophagus tuberculiofrons  in matrix.
 
@@
 
-Assemblage composition was mostly uniform across patch types, however
-
-1. total number of beetles _from_ total number of species 
-1. were all species found in all habitats? Were any species found in only 1 habitat?
-1. Were all species found in all blocks? Were any restricted to only 1-2 blocks?
-1. Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
-1. Number of species in each functional group
-1. most common 3-4 species
-1. any rare species?
-
-# DISCUSSION
-
-1. dont forget o discuss the basic biology...why might a species be so common? why might one be rare?
-
-# ACKNOWLEDGMENTS
-
-We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site, particularly Sabrie Breland, Ben Overly, and Eva Schwarz.
-We also thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry, and to Nico Acajabon, Jarrett Emory, and Cayla Garmen for their advice in putting together this manuscript.
-
-# OTHER REQUIRED TEXT
-
-### Dedication 
-To my family who never stopped supporting me along this journey, my friends who kept me company along the way, and my mentors at VCU who believed in me before I did myself.
-
-### List of Abbreviations
-
-1. SRS:	Savannah River Site. 
-1. Another word:	And the list continues with another definition.
-
-# Biographical Sketch
-
-Eric Escobar-Chena completed his Bachelors education at Virginia Commonwealth University in 2023.
-During his time there he developed a fondness for insects which  grew into a curiosity of the natural world.
-He later began to explore this curiosity deeper in beginning his graduate education at the University of Florida as a Master's Student under the supervision of Emilio Bruna.
+Beetle abundance was significantly higher in matrix plots with 33% (n = 1713) of captures followed by 26% (n = 1359) caught in connected patches, 23% (n = 1199) in winged, and lastly 18% (n = 942) from rectangular patches (Table 2).
 
@@
 
+Species richness was distinctly different between patch types with matrix plots exhibited the highest richness while connected patches were lowest (Fig.
+1).
+Estimations of Shannon diversity index were fairly uniform between patch types.
+Simpson diversity indices were similarly even however winged patches showed a lower value.
+
 ```{r Table 3, echo=FALSE,message = FALSE,warning=FALSE}
@@
 ```
+
+Notable Bray-Curtis dissimilarity values were between connected and matrix (BC = 0.127), connected and winged (BC = 0.077),  matrix and rectangle (BC = 0.318), and matrix and winged (BC = 0.189) (Table 3).
+
+# DISCUSSION
+
+1. dont forget o discuss the basic biology...why might a species be so common? why might one be rare?
+
+# ACKNOWLEDGMENTS
+
+We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site, particularly Sabrie Breland, Ben Overly, and Eva Schwarz.
+We also thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry, and to Nico Acajabon, Jarrett Emory, and Cayla Garmen for their advice in putting together this manuscript.
+
+# OTHER REQUIRED TEXT
+
+### Dedication 
+To my family who never stopped supporting me along this journey, my friends who kept me company along the way, and my mentors at VCU who believed in me before I did myself.
+
+### List of Abbreviations
+
+1. SRS:	Savannah River Site. 
+1. Another word:	And the list continues with another definition.
+
+# Biographical Sketch
+
+Eric Escobar-Chena completed his Bachelors education at Virginia Commonwealth University in 2023.
+During his time there he developed a fondness for insects which  grew into a curiosity of the natural world.
+He later began to explore this curiosity deeper in beginning his graduate education at the University of Florida as a Master's Student under the supervision of Emilio Bruna.

### 2025-03-03 corridor_docs/eric_ms_thesis/02_data_analysis.R

@@ -515,6 +515,8 @@ hill_plot
 taxa_bray <- vegan::vegdist(btl_spec_patch, method = "bray", binary = FALSE)
 taxa_bray
 
+
+
 btl_spec_patch<- btl_data |>
   pivot_longer(pvin:osyl, names_to = "species", values_to = "n") |>
   group_by(species,patch) |>
@@ -735,12 +737,27 @@ plot.TES(TES_r)
 
 # iNEXT example
 
-data(spider)
-out1 <- iNEXT(spider, q=c(0,1,2), datatype="abundance")
+library(iNEXT)
 
-data(bird)
-out2 <- iNEXT(bird, q=0, datatype="abundance")
-out2
+btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
+spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+
+btl_sums_patch<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(species,patch) %>%
+  mutate(patch=case_when(
+    patch == "c" ~ "Corridor",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(patch,desc(n)) %>% 
+  rename(sp_code=species,
+         n_patch=n) %>% 
+  pivot_wider(
+    names_from = patch,
+    values_from = c(n_patch))
 
 div.data <- btl_sums_patch |>
   column_to_rownames(var = "sp_code") 
@@ -748,14 +765,9 @@ div.data <- btl_sums_patch |>
 
 div.btl <- iNEXT(div.data, q = c(0,1,2), datatype = "abundance")
 
-
 z <- iNEXT(div.data, q=c(0,1,2), datatype="abundance")
-ggiNEXT(z, facet.var="Order.q", color.var="Assemblage")
-ggiNEXT(z, type = 2, se = TRUE)
-ggiNEXT(z, facet.var="Both", color.var="Both")
-
-div.btl$DataInfo
-div.btl$AsyEst
+ggiNEXT(z, facet.var="Order.q", color.var="Assemblage") + 
+  theme_bw(base_size=18)
 # total NTES_c# total N by date type --------------------------------------------------
 
 

### 2025-03-03 corridor_docs/eric_ms_thesis/02_data_analysis.R
s

@@ -437,6 +437,7 @@ rarefaction_data<-btl_data %>%
   #column_to_rownames("sample_id") %>% 
   #filter((rowSums(rarefaction_data) > 10)) 
 
+
 s <- specnumber(rarefaction_data)
 s
 (raremax <- min(rowSums(rarefaction_data)))
@@ -514,8 +515,6 @@ hill_plot
 
 taxa_bray <- vegan::vegdist(btl_spec_patch, method = "bray", binary = FALSE)
 taxa_bray
-
-
 
 btl_spec_patch<- btl_data |>
   pivot_longer(pvin:osyl, names_to = "species", values_to = "n") |>
@@ -737,20 +736,16 @@ plot.TES(TES_r)
 
 # iNEXT example
 
-library(iNEXT)
+data(spider)
+out1 <- iNEXT(spider, q=c(0,1,2), datatype="abundance")
 
-btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
-spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+data(bird)
+out2 <- iNEXT(bird, q=0, datatype="abundance")
+out2
 
 btl_sums_patch<-btl_data %>%
   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
   group_by(species,patch) %>%
-  mutate(patch=case_when(
-    patch == "c" ~ "Corridor",
-    patch == "m" ~ "Matrix",
-    patch == "w" ~ "Winged",
-    patch == "r" ~ "Rectangle",
-    .default = as.character(patch))) %>%
   summarize(n=sum(n, na.rm=TRUE)) %>%
   arrange(patch,desc(n)) %>% 
   rename(sp_code=species,
@@ -765,9 +760,14 @@ div.data <- btl_sums_patch |>
 
 div.btl <- iNEXT(div.data, q = c(0,1,2), datatype = "abundance")
 
+
 z <- iNEXT(div.data, q=c(0,1,2), datatype="abundance")
-ggiNEXT(z, facet.var="Order.q", color.var="Assemblage") + 
-  theme_bw(base_size=18)
+ggiNEXT(z, facet.var="Order.q", color.var="Assemblage")
+ggiNEXT(z, type = 2, se = TRUE)
+ggiNEXT(z, facet.var="Both", color.var="Both")
+
+div.btl$DataInfo
[…]

### 2025-03-03 corridor_docs/eric_ms_thesis/thesis_planning.Rmd

@@
 In total 16 sampling rounds were carried out with 4 rounds per block, 196 samples were collected.
-All beetles of subfamily Scarabaeinae  were counted and identified to species using keys from @nemes_illustrated_2015 ### update bib.
-Overall 16 species were identified and approximately 5200 in total individual beetles were collected.
+All dung beetles were counted and identified to species with the exception of beetles of the genus Aphodius, overall 15 species were identified and approximately 5300 individual beetles were collected.
 
@@
 
+1. total number of beetles _from_ total number of species 
+1. were all species found in all habitats? Were any species found in only 1 habitat?
+1. Were all species found in all blocks? Were any restricted to only 1-2 blocks?
+1. Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
+1. Number of species in each functional group
+1. most common 3-4 species
+1. any rare species?
+
 # DISCUSSION

### 2025-03-03 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
outlined discussion

@@
 
-1. dont forget o discuss the basic biology...why might a species be so common? why might one be rare?
+Recap main goals into findings
+ abundances in matrix vs connected patch and why this could be happening
+ - source pop to habiitat edge
+
+ species richness again supporting that matrix is more of an ideal habitat for dung beetle community
+
+ diversity indices community structures weren't highly different between patch types
+
+ bray curtis hinting at similar land uses between corridor and winged patch
+ - like julians paper corridors benefit certain populations and more fit populations are able to make better use
+
+Interpretation of results
+ Abundance 
+ - reasin for highest abundances in matrix and connected patch
+ - lower abundance in rectangle hinting at fragmentation effects
+ Species Richness & Diversity
+ - the role of hbitat connectivity in shaping community structure or lack thereof
+ - why do we think connnected had lowest species richness??
+ Functional traits and ecological impacts
+ - did corridoors favor a functional trait
+ - why might there be a trait response 
+ - implications for ecosystem processes
+ - like duung removal papers seed dispersal and yep
+
+Comparing to previous studies
+- how are things aligning
+- think about the biology
+
+Limitations and future work
+- potential confounding factors (seasonality, distance from edge, sampling methodology and temmporal variation)
+- other directions to go (dispersal -> radar, secifically measuring changes in ecosystem services)
+
+Takeaways for conservation and management
+- dung beetles are robust
+- what do think about corridor design and considerations for fragmented lanndscapes
+- practical applications think about the beetles
 

### 2025-03-03 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
new pages and pitfall tidbit
added inline code

@@
 # devtools::install_github("crsh/papaja")
+```
+
+```{r load-data, cached=TRUE, echo = FALSE, warning=FALSE, message = FALSE}
+
+btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
+spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+
+# total beetles captured 
+total_N<-btl_data %>% 
+  summarise(n=sum(sum, na.rm=TRUE))
+
+# total no of spp
+total_spp<-btl_data %>% 
+  select(-sum) %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n))) %>% 
+  summarise(n_spp=n_distinct(species))
 ```
@@
 
-spp_table<-btl_order %>% 
+spp_table_data<-btl_order %>% 
   left_join(spp_codes,by="sp_code") %>% 
@@
 )) %>% 
-  select(species,n,m,c,w,r) %>% 
+  select(species,Guild,n,m,c,w,r) %>% 
   rename(Species=species,
@@
 
-spp_table<-knitr::kable(spp_table, 
+spp_table<-knitr::kable(spp_table_data, 
       digits = 2,
-      align="lrcccc",
+      align="lcrcccc",
       format="latex",
@@
 spp_table
+
+most_common<-spp_table_data %>% select(Species) %>% slice(1)
+most_common_n<-spp_table_data %>% select(N) %>% slice(1)
 ```
 
-Overall, 5,213 total dung beetles were collected belonging to 16 species.
-The most dominant species were Canthon vigilans (n=1473), Ateuchus lecontei (n=1115), and Phanaeus igneus (n=958) (Table 1).
+Overall, `r total_N` total dung beetles were collected belonging to N = `r total_spp` species.
+The most dominant species were _`r most_common`_ (N = `r most_common_n`), Ateuchus lecontei (n=1115), and Phanaeus igneus (n=958) (Table 1).
 All but one species was found in matrix plots.

### 2025-03-05 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
updates

--- 2025-03-03 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
@@
   summarise(n_spp=n_distinct(species))
+
+# 
+# most_common<-spp_table_data %>% select(Species) %>% slice(1)
+# most_common_n<-spp_table_data %>% select(N) %>% slice(1)
+
 ```
@@
 
+\newpage
+
 # Methods
 
-Study site
+## Study site
 
@@
 
-Dung beetle sampling
+## Dung beetle sampling
 
@@
 Plots in the matrix were set up in a similar fashion with the center point 250 m from the center placed equidistant between adjacent patches.
+Individual pitfall traps consisted of two components, a 10cm tall by 8 cm wide cylinder base topped with a funnel with a 10cm wide rim.
+Pig feces was suspended above the funnel by a 6.5 cm by 6.5 cm mesh square.
 For each sample period, traps were baited with pig dung between 8-9 pm and picked up 12 hours later, all beetles captured were stored in ethanol for further processing.
 In total 16 sampling rounds were carried out with 4 rounds per block, 196 samples were collected.
+
 All dung beetles were counted and identified to species with the exception of beetles of the genus Aphodius, overall 15 species were identified and approximately 5300 individual beetles were collected.
 
-Analyses
+## Analyses
+
+- species richness
+
+  > glmm with poisson dist reccomended by julian
+
+- species diversity
+
+  > glmm with poisson dist reccomended by julian
+
+- functional diversity
+
+  > 1st pass: roller, tunneler, dweller  
+  > after all done: morphological measurements like Alonso et al 2022 (see table from decastro 2023, Giménez Gómez 2025)
+
+- abundance (or biomass)
+
+  > abundance of the most common species?  
+  > abundance of species caught in all sites?  
+  > changes in rank-abundance.
+
+_Reason for this breakdown: _
+
+- richness and diversity are "community level" response. 
+- abundance is a species-level response. 
+- changes in abundance or rank-abundance  could have implications for ecosystem services if it is connected to finctional group.
 
@@
 
-  - lets hammer this out 
-  - modeling? 
-    - glmm with poisson dist reccomended by julian
-  - beta, abundance, biomass? per site
-  - species list by sampling blocks (anything with this?)
-  - habitat preference 
-  - rarefaction
+<!--   - lets hammer this out  -->
+<!--   - modeling?  -->
+<!--     - glmm with poisson dist reccomended by julian -->
+<!--   - beta, abundance, biomass? per site -->
+<!--   - ~species list by sampling blocks (anything with this?)~ -->
+<!--   - habitat preference  -->
+<!--   - rarefaction -->
+
+\newpage
 
@@
 
+## Overview
+
+- how many beetles total
+- most common species overall
+- was. the same species the most common one in all habitats?
+- were there any that were caught in all habitats? Any restricted to only one?
+- any invasive / exotic species?
+
+Overall, `r total_N` total dung beetles were collected belonging to N = `r total_spp` species.
+The most dominant species were ______  (N = _____), Ateuchus lecontei (n=1115), and Phanaeus igneus (n=958) (Table 1).
+All but one species was found in matrix plots.
+Four species (Onthophagus striatulus, Onthophagus concinnus, Geotrupes Blackburnii, Onthophagus truberculifrons) were the only species not representative in every patch type.
+We had two singleton species, Geotrupes Blackburnii appearing only once in winged, and Onthophagus tuberculiofrons  in matrix.
+
+Beetle abundance was significantly higher in matrix plots with 33% (n = 1713) of captures followed by 26% (n = 1359) caught in connected patches, 23% (n = 1199) in winged, and lastly 18% (n = 942) from rectangular patches (Table 2).
+
+Species richness was distinctly different between patch types with matrix plots exhibited the highest richness while connected patches were lowest (Fig.
+1).
+Estimations of Shannon diversity index were fairly uniform between patch types.
+Simpson diversity indices were similarly even however winged patches showed a lower value.
+
+Notable Bray-Curtis dissimilarity values were between connected and matrix (BC = 0.127), connected and winged (BC = 0.077),  matrix and rectangle (BC = 0.318), and matrix and winged (BC = 0.189) (Table 3).
+
+1. total number of beetles _from_ total number of species 
+1. were all species found in all habitats? Were any species found in only 1 habitat?
+1. Were all species found in all blocks? Were any restricted to only 1-2 blocks?
+1. Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
+1. Number of species in each functional group
+1. most common 3-4 species
+1. any rare species?
+
+\newpage
+
+# DISCUSSION
+
+Recap main goals into findings
+ abundances in matrix vs connected patch and why this could be happening
+ - source pop to habiitat edge
+
+ species richness again supporting that matrix is more of an ideal habitat for dung beetle community
+
+ diversity indices community structures weren't highly different between patch types
+
+ bray curtis hinting at similar land uses between corridor and winged patch
+ - like julians paper corridors benefit certain populations and more fit populations are able to make better use
+
+Interpretation of results
+ Abundance 
+ - reasin for highest abundances in matrix and connected patch
+ - lower abundance in rectangle hinting at fragmentation effects
+ Species Richness & Diversity
+ - the role of hbitat connectivity in shaping community structure or lack thereof
+ - why do we think connnected had lowest species richness??
+ Functional traits and ecological impacts
+ - did corridoors favor a functional trait
+ - why might there be a trait response 
+ - implications for ecosystem processes
+ - like duung removal papers seed dispersal and yep
+
+Comparing to previous studies
+- how are things aligning
+- think about the biology
+
+Limitations and future work
+- potential confounding factors (seasonality, distance from edge, sampling methodology and temmporal variation)
+- other directions to go (dispersal -> radar, secifically measuring changes in ecosystem services)
+
+Takeaways for conservation and management
+- dung beetles are robust
+- what do think about corridor design and considerations for fragmented lanndscapes
+- practical applications think about the beetles
+
+# ACKNOWLEDGMENTS
+
+We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site, particularly Sabrie Breland, Ben Overly, and Eva Schwarz.
+We also thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry, and to Nico Acajabon, Jarrett Emory, and Cayla Garmen for their advice in putting together this manuscript.
+
+\newpage
+
+# REFERENCES
+
+\smallskip  
+::: {#refs}
+:::
+
+\newpage
+# FIGURES
+
+<!-- ```{r  label = pitfalls, echo = TRUE, fig.cap = "Experimental landscapes indicating location of the pitfall traps.", out.width = '50%'} -->
+<!-- knitr::include_graphics('./corridor_docs/eric_ms_thesis/pitfalls.png') -->
+<!-- ``` -->
+
+![Experimental landscapes indicating location of the pitfall traps.](pitfalls.png)
+
+```{r Figure_hill, echo=FALSE,message = FALSE,warning=FALSE}
+library(iNEXT)
+#
+# btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
+# spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+
+btl_sums_patch<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
+  group_by(species,patch,block) %>%
+  mutate(patch=case_when(
+    patch == "c" ~ "Corridor",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(patch,desc(n)) %>%
+  rename(sp_code=species,
+         n_patch=n) %>%
+  mutate(pb=paste(patch,block,sep="-")) %>% 
+  ungroup() %>% 
+  select(-patch,-block) %>% 
+  pivot_wider(
+    names_from = pb,
+    values_from = c(n_patch))
+
+div.data <- btl_sums_patch |>
+  column_to_rownames(var = "sp_code")
+
+div.btl <- iNEXT(div.data, q = c(0,1,2), datatype = "abundance")
+
+z <- iNEXT(div.data, q=c(0,1,2), datatype="abundance")
+ggiNEXT(z, facet.var="Order.q", color.var="Assemblage") +
+  theme_bw(base_size=18)
+
+```
+
+\newpage
+
 ```{r Table1, echo=FALSE,message = FALSE,warning=FALSE}
 # https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
-
-btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
-spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
 
@@
 
-most_common<-spp_table_data %>% select(Species) %>% slice(1)
-most_common_n<-spp_table_data %>% select(N) %>% slice(1)
 ```
 
-Overall, `r total_N` total dung beetles were collected belonging to N = `r total_spp` species.
-The most dominant species were _`r most_common`_ (N = `r most_common_n`), Ateuchus lecontei (n=1115), and Phanaeus igneus (n=958) (Table 1).
-All but one species was found in matrix plots.
-Four species (Onthophagus striatulus, Onthophagus concinnus, Geotrupes Blackburnii, Onthophagus truberculifrons) were the only species not representative in every patch type.
-We had two singleton species, Geotrupes Blackburnii appearing only once in winged, and Onthophagus tuberculiofrons  in matrix.
+\newpage
 
@@
 
-Beetle abundance was significantly higher in matrix plots with 33% (n = 1713) of captures followed by 26% (n = 1359) caught in connected patches, 23% (n = 1199) in winged, and lastly 18% (n = 942) from rectangular patches (Table 2).
-
-```{r Figure 1, echo=FALSE,message = FALSE,warning=FALSE}
-library(iNEXT)
-
-btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
-spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
-
-btl_sums_patch<-btl_data %>%
-  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
-  group_by(species,patch) %>%
-  mutate(patch=case_when(
-    patch == "c" ~ "Corridor",
-    patch == "m" ~ "Matrix",
-    patch == "w" ~ "Winged",
-    patch == "r" ~ "Rectangle",
-    .default = as.character(patch))) %>%
-  summarize(n=sum(n, na.rm=TRUE)) %>%
-  arrange(patch,desc(n)) %>% 
-  rename(sp_code=species,
-         n_patch=n) %>% 
-  pivot_wider(
-    names_from = patch,
-    values_from = c(n_patch))
-
-div.data <- btl_sums_patch |>
-  column_to_rownames(var = "sp_code")
-
-div.btl <- iNEXT(div.data, q = c(0,1,2), datatype = "abundance")
-
-z <- iNEXT(div.data, q=c(0,1,2), datatype="abundance")
-ggiNEXT(z, facet.var="Order.q", color.var="Assemblage") + 
-  theme_bw(base_size=18)
-
-```
-
-Species richness was distinctly different between patch types with matrix plots exhibited the highest richness while connected patches were lowest (Fig.
-1).
-Estimations of Shannon diversity index were fairly uniform between patch types.
-Simpson diversity indices were similarly even however winged patches showed a lower value.
+\newpage
 
@@
 ```
-
-Notable Bray-Curtis dissimilarity values were between connected and matrix (BC = 0.127), connected and winged (BC = 0.077),  matrix and rectangle (BC = 0.318), and matrix and winged (BC = 0.189) (Table 3).
-
-1. total number of beetles _from_ total number of species 
-1. were all species found in all habitats? Were any species found in only 1 habitat?
-1. Were all species found in all blocks? Were any restricted to only 1-2 blocks?
-1. Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
-1. Number of species in each functional group
-1. most common 3-4 species
-1. any rare species?
-
-# DISCUSSION
-
-Recap main goals into findings
- abundances in matrix vs connected patch and why this could be happening
- - source pop to habiitat edge
-
- species richness again supporting that matrix is more of an ideal habitat for dung beetle community
-
- diversity indices community structures weren't highly different between patch types
-
- bray curtis hinting at similar land uses between corridor and winged patch
- - like julians paper corridors benefit certain populations and more fit populations are able to make better use
-
-Interpretation of results
- Abundance 
- - reasin for highest abundances in matrix and connected patch
- - lower abundance in rectangle hinting at fragmentation effects
- Species Richness & Diversity
- - the role of hbitat connectivity in shaping community structure or lack thereof
- - why do we think connnected had lowest species richness??
- Functional traits and ecological impacts
- - did corridoors favor a functional trait
- - why might there be a trait response 
- - implications for ecosystem processes
- - like duung removal papers seed dispersal and yep
-
-Comparing to previous studies
-- how are things aligning
-- think about the biology
-
-Limitations and future work
-- potential confounding factors (seasonality, distance from edge, sampling methodology and temmporal variation)
-- other directions to go (dispersal -> radar, secifically measuring changes in ecosystem services)
-
-Takeaways for conservation and management
-- dung beetles are robust
-- what do think about corridor design and considerations for fragmented lanndscapes
-- practical applications think about the beetles
-
-# ACKNOWLEDGMENTS
-
-We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site, particularly Sabrie Breland, Ben Overly, and Eva Schwarz.
-We also thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry, and to Nico Acajabon, Jarrett Emory, and Cayla Garmen for their advice in putting together this manuscript.
+\newpage
 

### 2025-03-05 corridor_docs/eric_ms_thesis/02_data_analysis.R
inext edits

@@ -739,13 +739,17 @@ plot.TES(TES_r)
 # iNEXT example
 
 library(iNEXT)
-
+library(tidyverse)
+library(here)
 btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
 spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
 
+# I edited the code so that it calculates the values for EACH LOCATION
+# to do this I needed to make a hybrid name of the plot and patch
+
 btl_sums_patch<-btl_data %>%
-  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
-  group_by(species,patch) %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
+  group_by(species,patch,block) %>%
   mutate(patch=case_when(
     patch == "c" ~ "Corridor",
     patch == "m" ~ "Matrix",
@@ -753,23 +757,17 @@ btl_sums_patch<-btl_data %>%
     patch == "r" ~ "Rectangle",
     .default = as.character(patch))) %>%
   summarize(n=sum(n, na.rm=TRUE)) %>%
-  arrange(patch,desc(n)) %>% 
+  arrange(patch,desc(n)) %>%
   rename(sp_code=species,
-         n_patch=n) %>% 
+         n_patch=n) %>%
+  mutate(pb=paste(patch,block,sep="-")) %>% 
+  ungroup() %>% 
+  select(-patch,-block) %>% 
   pivot_wider(
-    names_from = patch,
-    values_from = c(n_patch))
+    names_from = pb,
+    values_from = c(n_patch)) 
 
-btl_sums_patch<-btl_data %>%
-  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
-  group_by(species,patch) %>%
-  summarize(n=sum(n, na.rm=TRUE)) %>%
-  arrange(patch,desc(n)) %>% 
-  rename(sp_code=species,
-         n_patch=n) %>% 
-  pivot_wider(
-    names_from = patch,
-    values_from = c(n_patch))
+
 
 div.data <- btl_sums_patch |>
   column_to_rownames(var = "sp_code") 
@@ -777,9 +775,65 @@ div.data <- btl_sums_patch |>
 
[…]

### 2025-03-12 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
thesis step-by-step

@@
 library(kableExtra)
+library(grateful)
 # library(gridExtra)
@@
 # devtools::install_github("crsh/papaja")
+cite_packages(citation.style = "peerj", out.dir = ".")
 ```
@@
 
-As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
+**_Note from EB:_** _I don't want to see the words 'dung beetles' until paragraph 5_.
+
+Paragraph 1: Overall Topic & Why it is interesting and Important.
+
+>Topic Sentence:
+
+Paragraph 2: What we know.
+
+>Topic Sentence:
+
+Paragraph 3: What we don't know.
+
+>Topic Sentence:
+
+Paragraph 4: Why we don't know it. _(hint: ecosystem services)_
+
+>Topic Sentence:
+
+Paragraph 5: Here is what we did to address this unknown.
+
+>Topic Sentence:
+
+---
+
+> As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
 This degradation lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
@@
 
-Dung beetles have emerged as a model system with which to test spatial ecology hypotheses  [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al.
+> Dung beetles have emerged as a model system with which to test spatial ecology hypotheses  [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al.
 2024).
@@
 
-Here, we  aim to determine how connectivity and fragmentation affect Species Richness and Diversity, Abundance, and functional diversity.
+> Here, we  aim to determine how connectivity and fragmentation affect Species Richness and Diversity, Abundance, and functional diversity.
 We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge.
@@
 All dung beetles were counted and identified to species with the exception of beetles of the genus Aphodius, overall 15 species were identified and approximately 5300 individual beetles were collected.
+ - what guides/keys did you do to ID? 
+ - did you deposit vouchers and if so where are they available?
 
@@
 
-- species richness
+### Overview
 
-  > glmm with poisson dist reccomended by julian
+1. how many beetles total
+1. most common species overall. 
+1. was. the same species the most common one in all habitats?
+1. were there any that were caught in all habitats? Any restricted to only one?
+1. any invasive / exotic species?
+1. Table: count by species by habitat
 
-- species diversity
+### Species richness
 
-  > glmm with poisson dist reccomended by julian
+1. glmm with poisson dist reccomended by julian
+1. table: results of analyses. 
+1. visualization: hill's plots? rank-abundance plots? rarefaction curve?
 
-- functional diversity
+### Species diversity
 
-  > 1st pass: roller, tunneler, dweller  
-  > after all done: morphological measurements like Alonso et al 2022 (see table from decastro 2023, Giménez Gómez 2025)
+1. glmm with poisson dist reccomended by julian
+1. table: results of analyses
+1. visualization: hill's plots? rarefaction curve?
 
-- abundance (or biomass)
+### Functional diversity
 
-  > abundance of the most common species?  
-  > abundance of species caught in all sites?  
-  > changes in rank-abundance.
+1. 1nd pass: roller, tunneler, dweller
+1. 2nd: morphological measurements like Alonso et al 2022 (see table from decastro 2023, Giménez Gómez 2025)
+1. table: results of analyses 
+1. visualization: biomass of different (R/T/D) functional groups?
 
-_Reason for this breakdown: _
+_Reason for this breakdown: _ - richness and diversity are "community level" response.
+abundance is a species-level response.
+changes in abundance or rank-abundance  could have implications for ecosystem services if it is connected to functional group.
 
-- richness and diversity are "community level" response. 
-- abundance is a species-level response. 
-- changes in abundance or rank-abundance  could have implications for ecosystem services if it is connected to finctional group.
+---
 
-Biodiversity between patch types was compared using Hill numbers (Jost 2006).
+> Biodiversity between patch types was compared using Hill numbers (Jost 2006).
 We looked at community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
@@
 
-Insert description above of individual trap
-
-1. **Functional Diversity:** Need to assign each species to a functional group: roller, tunneler. dweller, others? 
-  - habitat preference (forest, pasture, generalist)
+> Functional Diversity: Need to assign each species to a functional group: roller, tunneler. dweller, others?  - habitat preference (forest, pasture, generalist)
+ - what stats / computer packages did you use to analyze and visualize the data?
 
@@
 
-## Overview
-
-- how many beetles total
-- most common species overall
-- was. the same species the most common one in all habitats?
-- were there any that were caught in all habitats? Any restricted to only one?
-- any invasive / exotic species?
-
-Overall, `r total_N` total dung beetles were collected belonging to N = `r total_spp` species.
-The most dominant species were ______  (N = _____), Ateuchus lecontei (n=1115), and Phanaeus igneus (n=958) (Table 1).
-All but one species was found in matrix plots.
-Four species (Onthophagus striatulus, Onthophagus concinnus, Geotrupes Blackburnii, Onthophagus truberculifrons) were the only species not representative in every patch type.
-We had two singleton species, Geotrupes Blackburnii appearing only once in winged, and Onthophagus tuberculiofrons  in matrix.
-
-Beetle abundance was significantly higher in matrix plots with 33% (n = 1713) of captures followed by 26% (n = 1359) caught in connected patches, 23% (n = 1199) in winged, and lastly 18% (n = 942) from rectangular patches (Table 2).
-
-Species richness was distinctly different between patch types with matrix plots exhibited the highest richness while connected patches were lowest (Fig.
-1).
-Estimations of Shannon diversity index were fairly uniform between patch types.
-Simpson diversity indices were similarly even however winged patches showed a lower value.
-
-Notable Bray-Curtis dissimilarity values were between connected and matrix (BC = 0.127), connected and winged (BC = 0.077),  matrix and rectangle (BC = 0.318), and matrix and winged (BC = 0.189) (Table 3).
+### Overview
 
@@
 1. any rare species? 
+1. any invasive / exotic species?
+
+### Species richness
+
+### Species diversity
+
+### Functional diversity
+
+---
+
+> Overall, `r total_N` total dung beetles were collected belonging to N = `r total_spp` species.
+The most dominant species were ______  (N = _____), _Ateuchus lecontei_ (n=_____), and _Phanaeus igneus_ (n=_958___) (Table 1).
+All but one species was found in matrix plots.
+Four species (_Onthophagus striatulus_, _Onthophagus concinnus_, _Geotrupes blackburnii_, _Onthophagus truberculifrons_) were the only species not representative in every patch type.
+We had two singleton species, _Geotrupes blackburnii_ appearing only once in winged, and _Onthophagus tuberculiofrons_ in matrix.
+
+> Beetle abundance was significantly higher in matrix plots with 33% (n = 1713) of captures followed by 26% (n = 1359) caught in connected patches, 23% (n = 1199) in winged, and lastly 18% (n = 942) from rectangular patches (Table 2).
+
+> Species richness was distinctly different between patch types with matrix plots exhibited the highest richness while connected patches were lowest (Fig.
+1).
+Estimations of Shannon diversity index were fairly uniform between patch types.
+Simpson diversity indices were similarly even however winged patches showed a lower value.
+
+> Notable Bray-Curtis dissimilarity values were between connected and matrix (BC = 0.127), connected and winged (BC = 0.077),  matrix and rectangle (BC = 0.318), and matrix and winged (BC = 0.189) (Table 3).
 
@@
 \newpage
-# FIGURES
+# FIGURES & TABLES
+
+1. ~~Table: count by species by habitat.~~
+1. table: results of species richness analyses.
+1. Figure of species richness results.
+1. table: results of species diversity analyses.
+1. Figure of species diversity results.
+1. table: results of functional diversity analyses.
+1. Figure of biomass of different (R/T/D) functional groups.
+1. Map of where field site is located.
+1. ~~Figure of the patch layout with pitfalls~~.
+1. Pictures of different species, sampling, field sites?
+
+\newpage
 
@@
 
-![Experimental landscapes indicating location of the pitfall traps.](pitfalls.png)
+![Experimental landscapes indicating location of the pitfall traps.](pitfalls.png){height=50%}
 
-```{r Figure_hill, echo=FALSE,message = FALSE,warning=FALSE}
-library(iNEXT)
-#
-# btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
-# spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+<!-- ```{r Figure_hill, echo=FALSE,message = FALSE,warning=FALSE} -->
+<!-- library(iNEXT) -->
+<!-- # -->
+<!-- # btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv")) -->
+<!-- # spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv")) -->
 
-btl_sums_patch<-btl_data %>%
-  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
-  group_by(species,patch,block) %>%
-  mutate(patch=case_when(
-    patch == "c" ~ "Corridor",
-    patch == "m" ~ "Matrix",
-    patch == "w" ~ "Winged",
-    patch == "r" ~ "Rectangle",
-    .default = as.character(patch))) %>%
-  summarize(n=sum(n, na.rm=TRUE)) %>%
-  arrange(patch,desc(n)) %>%
-  rename(sp_code=species,
-         n_patch=n) %>%
-  mutate(pb=paste(patch,block,sep="-")) %>% 
-  ungroup() %>% 
-  select(-patch,-block) %>% 
-  pivot_wider(
-    names_from = pb,
-    values_from = c(n_patch))
+<!-- btl_sums_patch<-btl_data %>% -->
+<!--   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% -->
+<!--   group_by(species,patch,block) %>% -->
+<!--   mutate(patch=case_when( -->
+<!--     patch == "c" ~ "Corridor", -->
+<!--     patch == "m" ~ "Matrix", -->
+<!--     patch == "w" ~ "Winged", -->
+<!--     patch == "r" ~ "Rectangle", -->
+<!--     .default = as.character(patch))) %>% -->
+<!--   summarize(n=sum(n, na.rm=TRUE)) %>% -->
+<!--   arrange(patch,desc(n)) %>% -->
+<!--   rename(sp_code=species, -->
+<!--          n_patch=n) %>% -->
+<!--   mutate(pb=paste(patch,block,sep="-")) %>%  -->
+<!--   ungroup() %>%  -->
+<!--   select(-patch,-block) %>%  -->
+<!--   pivot_wider( -->
+<!--     names_from = pb, -->
+<!--     values_from = c(n_patch))  -->
 
-div.data <- btl_sums_patch |>
-  column_to_rownames(var = "sp_code")
+<!-- div.data <- btl_sums_patch |> -->
+<!--   column_to_rownames(var = "sp_code") -->
 
-div.btl <- iNEXT(div.data, q = c(0,1,2), datatype = "abundance")
+<!-- div.btl <- iNEXT(div.data, q = c(0,1,2), datatype = "abundance") -->
 
-z <- iNEXT(div.data, q=c(0,1,2), datatype="abundance")
-ggiNEXT(z, facet.var="Order.q", color.var="Assemblage") +
-  theme_bw(base_size=18)
+<!-- z <- iNEXT(div.data, q=c(0,1,2), datatype="abundance") -->
+<!-- ggiNEXT(z, facet.var="Order.q", color.var="Assemblage") + -->
+<!--   theme_bw(base_size=18) -->
 
-```
+<!-- ``` -->
 
@@
                 font_size = 12,
-                position = "left") %>% 
+                position = "center") %>% 
   column_spec(1, italic = T) %>% 
@@
 ```
+---
 
@@
                 font_size = 12,
-                position = "left") %>% 
+                position = "center") %>%
   # column_spec(1, italic = T) %>%
@@
 ```
-
-\newpage
 
@@
 taxa_bray <- vegan::vegdist(btl_spec_patch, method = "bray", binary = FALSE)
-taxa_bray
+# taxa_bray
 
 ```
+---
+
 \newpage
@@
 
-### Dedication 
+## Dedication 
 To my family who never stopped supporting me along this journey, my friends who kept me company along the way, and my mentors at VCU who believed in me before I did myself.
 
-### List of Abbreviations
+## List of Abbreviations
 
@@
 
-# Biographical Sketch
+## Biographical Sketch
 

### 2025-03-25 corridor_docs/eric_ms_thesis/btl_drying.R
updated beetle wts

@@ -1,0 +1,29 @@
+# Script to determine if weights of btls in drying overn have stabilized  
+# date: 19 March 2025
+
+# load libraries ----------------------------------------------------------
+
+library(tidyverse)
+library(here)
+
+
+# load data ---------------------------------------------------------------
+
+btl_wt<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","btl_weight_equilibrium.csv"))
+
+btl_wt<-btl_wt %>% 
+  pivot_longer(wet_weight:weight_4,names_to = "interval") %>% 
+  mutate(interval=if_else(interval=="wet_weight","weight_0",interval)) %>% 
+  arrange(sample_id,sp_code,interval) %>% 
+  group_by(sample_id,sp_code) %>% 
+  mutate(perc_loss=((lag(value)-(value))/lag(value)*100)) %>% 
+  mutate(sample=paste(sp_code,sample_id,sep="-"))
+  
+btl_wt
+
+
+btl_wt %>% 
+  # filter(interval!="weight_1") %>%
+  # filter(interval!="weight_0") %>%
+  ggplot( aes(x=interval, y=value, group=sample, color=sample)) +
+  geom_line()

### 2025-03-26 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
inserting code to rmd

@@
 
+```{r data-load, include = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
+
+btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
+spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+# total beetles captured --------------------------------------------------
+total_N<-btl_data %>% 
+  summarise(n=sum(sum, na.rm=TRUE))
+
+# How many species total?
+total_spp<-btl_data %>% 
+  select(-sum) %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n))) %>% 
+  summarise(n_spp=n_distinct(species))
+
+# most common species
+
+how_many_common<-3
+mega_spp<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>% 
+  summarise(n=sum(n, na.rm=TRUE)) %>% 
+  filter(n>500) %>% 
+  ungroup() %>% 
+  select(species,n)%>% 
+  arrange(desc(n)) %>% 
+  rename("sp_code"="species") %>% 
+  left_join(spp_codes) %>% 
+  mutate(species=tolower(species)) %>% 
+  mutate(species=paste(genus,species,sep=" ")) %>% 
+  select(-c(subspecies,notes,ordway,srs,Guild,family,sp_code,genus)) %>% 
+  slice(1:how_many_common)
+
+# rare species
+rare_spp<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>% 
+  summarise(n=sum(n, na.rm=TRUE)) %>% 
+  filter(n<50) %>% 
+  ungroup() %>% 
+  select(species,n) %>% 
+  arrange(n)
+
+```
+
 ### Overview
 
-1. total number of beetles _from_ total number of species 
+1. total number of beetles from total number of species: `r total_N` beetles from `r total_spp` species.
 1. were all species found in all habitats? Were any species found in only 1 habitat?
@@
 1. Number of species in each functional group
-1. most common 3-4 species
+1. most common 3-4 species: The `r how_many_common` most common species were \textit{`r mega_spp[1,2]`}, \textit{`r mega_spp[2,2]`}, and \textit{`r mega_spp[3,2]`}.
 1. any rare species? 

### 2025-04-01 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
edits to text (r code)

--- 2025-03-12 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
@@
 
-Our study took place at the Savannah River Site(SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals(Tewksbury et al.
+Our study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals (Tewksbury et al.
 2002).
@@
 
+```{r data-load, include = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
+
+btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
+spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+# total beetles captured --------------------------------------------------
+total_N<-btl_data %>% 
+  summarise(n=sum(sum, na.rm=TRUE))
+
+## How many species total?
+total_spp<-btl_data %>% 
+  select(-sum) %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n))) %>% 
+  summarise(n_spp=n_distinct(species))
+
+## Abundance by species
+spp_abund<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>% 
+  summarise(n=sum(n, na.rm=TRUE)) %>% 
+  # filter(n>500) %>% 
+  ungroup() %>% 
+  select(species,n)%>% 
+  arrange(desc(n)) %>% 
+  rename("sp_code"="species") %>% 
+  left_join(spp_codes) %>% 
+  mutate(species=tolower(species)) %>% 
+  mutate(species=paste(genus,species,sep=" ")) %>% 
+  select(-c(subspecies,notes,ordway,srs,Guild,family,sp_code,genus))
+
+## Common species
+
+# by number (i.e., N most common)
+how_many_common<-3
+most_common_spp<-spp_abund %>% 
+  slice_head(n = how_many_common)
+
+# by abundance (i.e., over NNNN sampled)
+high_abund_cutoff<-500
+most_common_count<-spp_abund %>% 
+  filter(n>high_abund_cutoff)
+
+## Rare species
+
+# by number (i.e., N most common)
+how_many_rare<-4
+rare_spp<-spp_abund %>% 
+  slice_tail(n=how_many_rare)
+
+# by abundance (i.e., over NNNN sampled)
+low_abund_cutoff<-10
+rare_spp_count<-spp_abund %>% 
+  filter(n<low_abund_cutoff)
+
+## Beetles by species and plot type (for Table 1)
+btl_sums_patch<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(species,patch) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(patch,desc(n)) %>% 
+  rename(sp_code=species,
+         n_patch=n) %>% 
+  pivot_wider(
+    names_from = patch,
+    values_from = c(n_patch)) %>% 
+  left_join(spp_codes) %>% 
+  mutate(species=tolower(species)) %>% 
+  mutate(species=paste(genus,species,sep=" ")) %>% 
+  ungroup() %>% 
+  select(species,Guild,m,c,w,r) %>% 
+  rename(Species=species)
+
+# which ones NOT found in all habitats?
+
+not_in_all_hab<-btl_sums_patch %>% 
+  rename(species=Species) %>% 
+  left_join(spp_abund) %>% 
+  select(-n,-Guild) %>% 
+  filter(m<1|c<1|w<1|r<1)
+
+# This finalizes the table for Table 1
+spp_table_data<-btl_sums_patch %>% 
+  rename(species=Species) %>% 
+  left_join(spp_abund) %>% 
+  relocate(n,.after="Guild") %>% 
+  mutate(c=case_when(
+  c == 0 ~ " ",
+  c > 0 ~ "x",
+  .default = as.character(c)
+)) %>% 
+  mutate(m=case_when(
+  m == 0 ~ " ",
+  m > 0 ~ "x",
+  .default = as.character(m)
+)) %>% 
+  mutate(r=case_when(
+  r == 0 ~ " ",
+  r > 0 ~ "x",
+  .default = as.character(r)
+)) %>% 
+mutate(w=case_when(
+  w == 0 ~ " ",
+  w > 0 ~ "x",
+  .default = as.character(w)
+)) %>% 
+  mutate(Guild=case_when(
+  Guild == "r" ~ "roll",
+  Guild == "t" ~ "tunnell",
+  Guild == "d" ~ "dwell",
+  .default = as.character(Guild)
+)) %>% 
+  rename(Species=species,
+         N=n,
+         Matrix=m,
+         Corridor=c,
+         Winged=w,
+         Rectangular=r)
+
+```
+
 ### Overview
 
-1. total number of beetles _from_ total number of species 
+1. total number of beetles from total number of species: `r total_N` beetles from `r total_spp` species.
 1. were all species found in all habitats? Were any species found in only 1 habitat?
@@
 1. Number of species in each functional group
-1. most common 3-4 species
+1.
+most common 3-4 species: The `r how_many_common` most common species were \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), and \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`).
 1. any rare species? 
@@
 
-btl_order<-btl_data %>%
-  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
-  group_by(species) %>%
-  summarize(n=sum(n, na.rm=TRUE)) %>%
-  arrange(desc(n)) %>%
-  mutate(species = reorder(species, desc(n))) %>% 
-  rename(sp_code=species)
-
-btl_sums_patch<-btl_data %>%
-  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
-  group_by(species,patch) %>%
-  summarize(n=sum(n, na.rm=TRUE)) %>%
-  arrange(patch,desc(n)) %>% 
-  rename(sp_code=species,
-         n_patch=n) %>% 
-  pivot_wider(
-    names_from = patch,
-    values_from = c(n_patch))
-
-btl_order<-full_join(btl_order,btl_sums_patch,by="sp_code")
-
-spp_table_data<-btl_order %>% 
-  left_join(spp_codes,by="sp_code") %>% 
-  unite("species",genus:species, sep=" ") %>% 
-  mutate(species=if_else(sp_code=="ap*", "Aphodius spp.",species)) %>% 
-  mutate(c=case_when(
-  c == 0 ~ " ",
-  c > 0 ~ "x",
-  .default = as.character(c)
-)) %>% 
-  mutate(m=case_when(
-  m == 0 ~ " ",
-  m > 0 ~ "x",
-  .default = as.character(m)
-)) %>% 
-  mutate(r=case_when(
-  r == 0 ~ " ",
-  r > 0 ~ "x",
-  .default = as.character(r)
-)) %>% 
-mutate(w=case_when(
-  w == 0 ~ " ",
-  w > 0 ~ "x",
-  .default = as.character(w)
-)) %>% 
-  select(species,Guild,n,m,c,w,r) %>% 
-  rename(Species=species,
-         N=n,
-         Matrix=m,
-         Corridor=c,
-         Winged=w,
-         Rectangular=r)
-
 spp_table<-knitr::kable(spp_table_data, 

### 2025-04-02 corridor_docs/eric_ms_thesis/btl_drying.R
last beetle weights and partitioning some of paragraph 1 and 2

@@ -12,7 +12,7 @@ library(here)
 btl_wt<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","btl_weight_equilibrium.csv"))
 
 btl_wt<-btl_wt %>% 
-  pivot_longer(wet_weight:weight_4,names_to = "interval") %>% 
+  pivot_longer(wet_weight:weight_5,names_to = "interval") %>% 
   mutate(interval=if_else(interval=="wet_weight","weight_0",interval)) %>% 
   arrange(sample_id,sp_code,interval) %>% 
   group_by(sample_id,sp_code) %>% 

### 2025-04-02 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
last beetle weights and partitioning some of paragraph 1 and 2
GLMMs

@@
   select(-n,-Guild) %>% 
-  filter(m<1|c<1|w<1|r<1)
+  mutate(row_sums=m+c+r+w) %>% 
+  filter(row_sums<4)
+
+# row_sums==1 for any restricted to only one habitat 
+# row_sums==2 for any restricted to only two habitats...
+
+# filter(m<1|c<1|w<1|r<1)
 
 # This finalizes the table for Table 1
+
 spp_table_data<-btl_sums_patch %>% 

### 2025-04-02 corridor_docs/eric_ms_thesis/02_data_analysis.R

@@ -16,8 +16,7 @@ library(iNEXT)
 btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
 
 
-
-# TO DO -------------------------------------------------------------------
+# SUMMARY STATS -----------------------------------------------------------
 
 # total beetles captured --------------------------------------------------
 total_N<-btl_data %>% 
@@ -370,6 +369,246 @@ plot_btl_n <- function(data) {
 }
 
 
+
+# GLMM OF DIV AND RICHNESS ------------------------------------------------
+
+
+# iNEXT example
+
+library(iNEXT)
+library(tidyverse)
+library(here)
+btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
+spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+
+
+# using hillR
+# https://github.com/daijiang/hillR/blob/master/README.md
+# https://daijiang.r-universe.dev/hillR
+library(hillR)
+
+
+hill_data<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
+  group_by(species,patch,block) %>%
+  mutate(patch=case_when(
+    patch == "c" ~ "Connected",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(patch,desc(n)) %>%
+  rename(sp_code=species,
+         n_patch=n) %>%
+  mutate(patch_block=paste(patch,block,sep="_")) %>% 
+  ungroup() %>% 
+  select(-patch,-block) %>% 
+  pivot_wider(
+    names_from = sp_code,
+    values_from = n_patch) 
+
+
+
+spp_abund<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
+  mutate(patch=case_when(
+    patch == "c" ~ "Connected",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) %>%
[…]

### 2025-04-09 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
some writing and restructuring

@@
 Paragraph 1: Overall Topic & Why it is interesting and Important.     
+  As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.Effects from isolation can vary, however as habitats are broken down community structures are significantly altered [@WOS:000419965700012].
+This degradation of structure lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
 
->Topic Sentence:
+>Topic Sentence:  Fragmentation is an extensive problem across the globe and it is vital to understand shifts in biodiversity and community structures as a result of trophic cascades.
 
 Paragraph 2: What we know. 
+  Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003].
 
->Topic Sentence:
+>Topic Sentence: Habitat connectivity is generally viewed as beneficial for ecosystems as a whole.
+We know that patch shape and size can be an influental factor for how populations are configured.
+Wildlife corridors
 
@@
 
->Topic Sentence:
+>Topic Sentence:  Changes in ecosystem fuction as a result of shifts in community structure
 
@@
 
->Topic Sentence:
+>Topic Sentence: Is this where we talk about challenges in large scale experimental design and the corridor project or is this where we talk about ecosystem services and biodiversity?
 
@@
 
->Topic Sentence:
+Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al.
+2024).
+The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]
+Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes.
+We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge.
+To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
+Additionally we asked (2) Are corridors benefiting any one functional trait over another?
+Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling.
+
+>Topic Sentence:  Here we talk about dung beetles how they fit into the broader scope of our question and directly state what questions we are asking.
 
@@
 2024).
-The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles such as enhanced nutrient cycling and soil quality and reduction of parasites on methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
+The removal, breakdown, and burial of animal feces drive important ecostystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
 Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) [@WOS:000891747700001].

### 2025-04-09 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
quick question
update

@@
 Paragraph 2: What we know.
-  Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003].
+
+Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003], including decreases in the abundance and diversity of species.
 
@@
 
->Topic Sentence:  Changes in ecosystem fuction as a result of shifts in community structure
+<!-- stuff that these inviduals or species do: -->
+<!-- 1) their interactions with other species -->
+<!-- - predation -->
+<!-- - pollination -->
+<!-- - seed dispersal -->
+<!-- 2) stuff on their own that influences ecosystems -->
+<!-- - ecosystem engineering (beavers, leaf-cutter ants) -->
+<!-- - nutrient cycling -->
+<!-- - decomposition -->
 
-Paragraph 4: Why we don't know it. _(hint: ecosystem services)_
+>Topic Sentence:  Changes in ecosystem fuction as a result of shifts in community structure (EB but we may not want to emphasize this given we didn't measure directly)
 
->Topic Sentence: Is this where we talk about challenges in large scale experimental design and the corridor project or is this where we talk about ecosystem services and biodiversity?
+Potential alternative routes:
+
+1) We know far less on how corridors influence functional diversity.
+Functional diversity is important for the maintenance of ecosystems services, etc, so corridors cvould help ensure that.
+(in context of dung beetles that would be feeding guilds (simplest), morphology etc.
+in deCastro Arrazola et al)
+
+2) Studies investigating corridors rarely look at whats i the matrix.
+The matrix could be a source of diversity that moves into patches and then through corridors or a completely different subset of species.
+(any other options?) Understanding whichh of these it is is important because it leads to differnt conclusions abiout role of corridors and conservation strategies
+
+3) regardless, not a bad idea to discuss both matrix and functional group in diuscussion.
+
+4) and the alternative approach: revamp into around spillover from matrix into patches and if depednent on corridors or edge
+
+<!-- Paragraph 4: Why we don't know it. _(hint: ecosystem services)_   -->
+
+<!-- >Topic Sentence: Is this where we talk about challenges in large scale experimental design and the corridor project or is this where we talk about ecosystem services and biodiversity?  -->
 
@@
 Additionally we asked (2) Are corridors benefiting any one functional trait over another?
-Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling.
+<!-- Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling. -->
 
@@
 
-All dung beetles were counted and identified to species with the exception of beetles of the genus Aphodius, overall 15 species were identified and approximately 5300 individual beetles were collected.
+All dung beetles were counted and identified to species. 
+ - We also dried and weighed individuals to estimate biomass 
  - what guides/keys did you do to ID? 
@@
 ### Overview
-
+_Don't have to describe these, just do it and report in results. this is just to remind us of what to do_
 1. how many beetles total
@@
 
+To Test for differences in species richness, we....
+
 1. glmm with poisson dist reccomended by julian
@@
 ### Species diversity
+
+We used a similar approach to compare species diversity in winged, isolated, and connected patches....
 

### 2025-04-09 corridor_docs/eric_ms_thesis/02_data_analysis.R

@@ -118,6 +118,32 @@ btl_order_patch<-btl_data %>%
 btl_order_patch
 
 
+# FIGURE
+
+btl_n_patch_block<-btl_data %>%
+  select(-sum) %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(patch,species,block) %>%
+  # group_by(patch,species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  pivot_wider(names_from = patch, 
+              values_from = n) %>% 
+  arrange(species,block) %>% 
+  pivot_longer(c:w,names_to = "patch",values_to = "n") 
+
+
+
+ggplot(data=btl_n_patch_block, aes(x=patch, y=n, group=block)) +
+  geom_line(size=0.5) + 
+  geom_point(size=2, aes(colour=block, shape=block))+
+  facet_wrap(vars(species), nrow = 2)+
+  ylab("No. of Beetles Captured")+
+  xlab("Patch Type")+
+  theme_classic()
+
+
+
+
 
 # total number of sample points -------------------------------------------
 

### 2025-04-11 corridor_docs/eric_ms_thesis/02_data_analysis.R
model tries

@@ -531,6 +531,23 @@ M0 <- glmer(n ~ patch_type * sp_code + (1 + patch_type * sp_code | block),
             family = poisson)
 summary(M0)
 
+# Model for richness 
+M_rich <- glmer(h_rich ~ patch_type + (1 + patch_type | block),
+                data = h_rich,
+                family = poisson)
+summary(M_rich)
+
+# Model for shannons 
+h_shannon
+
+?glmer
+
+M_shannon <- glmer(h_shannon ~ patch_type + (1 + patch_type | block),
+                data = h_shannon,
+                family = gaussian)
+summary(M_shannon)
+
+
 # residuals
 simulationOutput <- 
   simulateResiduals(fittedModel = M0, plot = TRUE)
@@ -543,7 +560,7 @@ simulationOutput <-
 # species richness (q = 0)
 # Shannon diversity (q = 1), the exponential of Shannon entropy) and 
 # Simpson diversity (q = 2, the inverse of Simpson
-#                                                                                                                                                                          concentration). For each diversity measure, iNEXT uses the observed sample of abundance or incidence
+# concentration). For each diversity measure, iNEXT uses the observed sample of abundance or incidence
 # data (called the “reference sample”) to compute diversity estimates and 
 # the associated 95% confidence intervals for the following two types of 
 # rarefaction and extrapolation (R/E):

### 2025-04-16 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
merge again
update
fixed merge error

@@
 author: "ERIC ESCOBAR-CHENA"
-
 # date: "\\vspace{-1em}"
@@
 <!-- - decomposition -->
+
+>Topic Sentence: Is this where we talk about challenges in large scale experimental design and the corridor project or is this where we talk about ecosystem services and biodiversity?
+Or sensitivity to ecosystem change and how that might impact functionality?
 
@@
 \newpage
-# FIGURES & TABLES
+# TABLES
 
@@
 1. table: results of species richness analyses.
-1. Figure of species richness results.
 1. table: results of species diversity analyses.
-1. Figure of species diversity results.
 1. table: results of functional diversity analyses.
-1. Figure of biomass of different (R/T/D) functional groups.
+
+# FIGURES
 1. Map of where field site is located.
@@
 1. Pictures of different species, sampling, field sites?
+
+1. Figure of species richness results.
+1. Figure of species diversity results.
+1. Figure of biomass of different (R/T/D) functional groups.
+1. Figure of biomass by species.
+1. Figure of abundance by species
 

### 2025-04-17 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
paragrapgh 1 and updating thesis bib.

--- 2025-04-09 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
@@
 Paragraph 1: Overall Topic & Why it is interesting and Important.     
-  As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.Effects from isolation can vary, however as habitats are broken down community structures are significantly altered [@WOS:000419965700012].
+  As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
+Effects from isolation can vary, however as habitats are broken down community structures are significantly altered [@WOS:000419965700012].
 This degradation of structure lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
+
+Like many ecological processes, fragmentation is a complex and multifaceted phenomenon
 
@@
 Paragraph 3: What we don't know.
+
+With experimental designs we are able to isolate factors might be contributing to ecological patterns and processes.
 

### 2025-04-22 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
p1

--- 2025-04-17 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
@@
 Paragraph 1: Overall Topic & Why it is interesting and Important.
+
   As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
-Effects from isolation can vary, however as habitats are broken down community structures are significantly altered [@WOS:000419965700012].
-This degradation of structure lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
-
-Like many ecological processes, fragmentation is a complex and multifaceted phenomenon
+Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003] [@fletcher_is_2018].
+However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012].
+This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
+Researchers today continue to search for strategies to mitigate the negative effects resulting from changes in landscape structure [@noh_forest_2022].
 

### 2025-04-23 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
some writing

@@
 
-Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003], including decreases in the abundance and diversity of species.
+Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003].
+
+Corridors are good and generally understood as beneficial for
 

### 2025-04-23 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
some changes

@@
 This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
-Researchers today continue to search for strategies to mitigate the negative effects resulting from changes in landscape structure [@noh_forest_2022].
+
+# Researchers today continue to search for strategies to mitigate the negative effects resulting from changes in landscape structure [@noh_forest_2022].
 

### 2025-04-24 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
formatting
removed sentence

@@
 
-Corridors are good and generally understood as beneficial for
-
 >Topic Sentence: Habitat connectivity is generally viewed as beneficial for ecosystems as a whole.
@@
 All dung beetles were counted and identified to species.
+
  - We also dried and weighed individuals to estimate biomass  
@@
 ### Overview
+
 _Don't have to describe these, just do it and report in results. this is just to remind us of what to do_
+
 1. how many beetles total  
@@
 1. any invasive / exotic species?  
-1. Table: count by species by habitat
+1. Table: count by species by habitat.
 

### 2025-04-24 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
updates

@@
 
+# TO DO LIST
+
+### Methods: 
+1. add a few sentences on drying, keys used to ID, and depositing vouchers
+
+### Analyses
+1. Write a short paragraph on how you calculated and compared species richness and diversity (hill numbers, glmm). 
+1. Write a short paragraph on how you calculated and compared functional diversity (how you measured it, how you compared it)
+
+### Results
+1. answer the overview questions, them stich them together into a paragraph
+ - ~~any species restricted to 1-2 habitat types~~
+ - Were all species found in all blocks? Were any restricted to only 1-2 blocks?
+ - Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
+ - Number of species in each functional group
+ - ~~most common 3-4 species~~
+ - any rare species? 
+ - any invasive / exotic species?
+
+### Figures
+
+1. Figure of species richness results.
+1. Figure of species diversity results.
+1. Figure of biomass of different (R/T/D) functional groups.
+1. Figure of abundance of individual species.
+1. ~~Figure of the patch layout with pitfalls~~.
+
+### Tables
+
+1. ~~Table: count by species by habitat.~~
+1. table: results of species richness analyses.
+1. table: results of species diversity analyses.
+1. table: results of functional diversity analyses.
+
+### Misc.
+
+1. Map of where field site is located.
+
+1. Pictures of different species, sampling, field sites
+
 \newpage
@@
 
-# Researchers today continue to search for strategies to mitigate the negative effects resulting from changes in landscape structure [@noh_forest_2022].
+>_Researchers today continue to search for strategies to mitigate the negative effects resulting from changes in landscape structure [@noh_forest_2022]. _
 
@@
 
-All dung beetles were counted and identified to species.
+All dung beetles were counted and identified to species
 
@@
  - what guides/keys did you do to ID?  
- - did you deposit vouchers and if so where are they available?
+ - "Voucher specimens of species I collected will be deposited at the Florida State Collection of Arthropods" (https://thefsca.org/services/specimen-donations/)
 
@@
 
-_Don't have to describe these, just do it and report in results. this is just to remind us of what to do_
+<!-- _Don't have to describe these, just do it and report in results. this is just to remind us of what to do_ -->
 
-1. how many beetles total  
-1. most common species overall.  
-1. was. the same species the most common one in all habitats?  
-1. were there any that were caught in all habitats? Any restricted to only one?  
-1. any invasive / exotic species?  
-1. Table: count by species by habitat.
+<!-- 1. how many beetles total   -->
+<!-- 1. most common species overall.   -->
+<!-- 1. was. the same species the most common one in all habitats?   -->
+<!-- 1. were there any that were caught in all habitats? Any restricted to only one?   -->
+<!-- 1. any invasive / exotic species?   -->
+<!-- 1. Table: count by species by habitat.  -->
 
@@
 1. total number of beetles from total number of species: `r total_N` beetles from `r total_spp` species.
-1. were all species found in all habitats? Were any species found in only 1 habitat?
+
+1.
+All but four species species were found in every patch type.
+_Onthophagus concinnus_ was found in the matrix and winged patches, while _Onthophagus striatulus_ was found in matrix habitat and rectanular patches.
+_Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively).
+
 1. Were all species found in all blocks? Were any restricted to only 1-2 blocks?

### 2025-04-25 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
some writing

@@
 
-Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003].
+Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003].
+By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy (insert richard forman 1995).
+It is because of this dynamic
 
@@
 
-All dung beetles were counted and identified to species
+All dung beetles were counted and identified to species as described in
 

### 2025-04-29 corridor_docs/eric_ms_thesis/04_GLMM.R
added some packages and worked on visualizatiopn

@@ -1,5 +1,5 @@
 library(tidyverse)
-
+library(here)
 
 
 # load and prep data ------------------------------------------------------
@@ -104,7 +104,7 @@ library(lme4)
 library(car)
 
 # LINEAR MEXED MODEL
-model <- lmer(h_rich ~ patch_type + (1|block), data = hill_results) # You should only use the 'aov()' for one-way ANOVAs with equal sample sizes
+model <- lmer(h_rich ~ patch_type + (1|block), data = hill_results)# You should only use the 'aov()' for one-way ANOVAs with equal sample sizes
 summary(model)
 Anova(model)
 shapiro.test(resid(model))
@@ -165,21 +165,36 @@ M0 <- glmer(n ~ patch_type * sp_code + (1 + patch_type * sp_code | block),
             family = poisson)
 summary(M0)
 
+Anova(M0)
+
 # Model for richness 
 M_rich <- glmer(h_rich ~ patch_type + (1 + patch_type | block),
                 data = h_rich,
                 family = poisson)
 summary(M_rich)
 
+Anova(M_rich)
+
+library(broom.mixed)
+tidy_model <- tidy(M_rich, effects = "fixed")
+
+library(gt)
+table <- tidy_model |>
+  gt() 
+
+library(flextable)
+ft <- flextable(tidy_model)
+save_as_image(ft, path = "glmer_table.png")
+
 # Model for shannons 
 h_shannon
 
 ?glmer
 
-M_shannon <- glmer(h_shannon ~ patch_type + (1 + patch_type | block),
-                   data = h_shannon,
-                   family = gaussian)
-summary(M_shannon)
+M_simpson <- glmer(h_simpson ~ patch_type + (1 + patch_type | block),
+                   data = h_simpson,
+                   family = poisson)
+summary(M_simpson)
 
 
 # residuals
@@ -211,11 +226,36 @@ ggplot(aes(x = block,
 hill_results %>% 
   ggplot(aes(x = patch_type, 
              # y = n,
-             # y = h_rich,
-             y = h_shannon,
+             y = h_rich,
+             # y = h_shannon,
              color = block)) +
   scale_color_viridis_d(option = "turbo") + 
   geom_point(position="jitter") + 
   theme_bw()+
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.7, hjust = 0.9))
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.7, hjust = 0.9))+
[…]

### 2025-05-02 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
p2 and 3 and brushed up 4

@@
  - \usepackage{lineno}
- - \linenumbers
+# - \linenumbers
  - \usepackage[default]{sourcesanspro}
@@
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy (insert richard forman 1995).
-It is because of this dynamic
+Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity.
 
@@
 
-With experimental designs we are able to isolate factors might be contributing to ecological patterns and processes.
+By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes.
+Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected.
+Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services.
 
@@
 To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
-Additionally we asked (2) Are corridors benefiting any one functional trait over another?
+Additionally we asked are corridors benefiting any one functional trait over another?
 <!-- Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling. -->
@@
 
-All dung beetles were counted and identified to species as described in
+All dung beetles were counted and identified to species as described in @nemes_illustrated_2015-1 and @edmonds_taxonomic_2023.
+Fifteen individuals of each species with adequate captures was dried to equalibrium and weighed for biomass measurements.
 

### 2025-05-02 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
added extra methods bit.

@@
 All dung beetles were counted and identified to species as described in @nemes_illustrated_2015-1 and @edmonds_taxonomic_2023.
-Fifteen individuals of each species with adequate captures was dried to equalibrium and weighed for biomass measurements.
+Fifteen individuals of each species with adequate captures were dried to equalibrium and weighed for biomass measurements.
+Voucher specimcens for each species will be deposited at the Florida State Collection of Arthropods.
 

### 2025-05-02 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
added discription of analyses and crossed off the list

@@
 ### Methods: 
-1. add a few sentences on drying, keys used to ID, and depositing vouchers
+1. ~~add a few sentences on drying, keys used to ID, and depositing vouchers~~
 
 ### Analyses
-1. Write a short paragraph on how you calculated and compared species richness and diversity (hill numbers, glmm). 
+1. ~~Write a short paragraph on how you calculated and compared species richness and diversity (hill numbers, glmm). ~~
 1. Write a short paragraph on how you calculated and compared functional diversity (how you measured it, how you compared it)
@@
  - ~~any species restricted to 1-2 habitat types~~
- - Were all species found in all blocks? Were any restricted to only 1-2 blocks?
+ - ~~Were all species found in all blocks? Were any restricted to only 1-2 blocks?~~
  - Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
@@
 Voucher specimcens for each species will be deposited at the Florida State Collection of Arthropods.
+
+## Analysis
+
+Biodiversity between patch types was compared using Hill numbers (Jost 2006).
+We looked at community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
+Diversity numbers were calculated in R studio using package hillR and modeled using package lme4.
+For the purposes of this project richness was modeled using a generalized linear mixed model with a poisson distribution.
+A qqplot of our data shows that our residuals are normally distributed (Figure 1).
 

### 2025-05-02 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
updates

@@
 
+Next assignment:
+
+1) Are any of the Dung Beetles we collected considered "open habitat" specialists? (might want to look at the work on pastures in the southeast)
+
+2) Is there anything about these species that might be interesting in the context of corridors?
+  - are any species known to be good fliers/movers?
+  - specialists on deer dung?
+  - navigate by starlight?
+  - feeding preferences?
+
+3) Did you get any species not captured in other long-leaf/SE USA forest sites?
+
+4) What are hill numbers?
+
+5) What is the advantage of useing hill numbers to estimate diversity / richness?
+
+6) Are the diversity and richness estimates calculated with hill numbers sensitive to anything, e.g., rare species/abundance, etc:
+
+7) Find me the best example paragraph you think you read describing the use of hill numbers to estimate diversity/richness (field survey paper): _________________
+
+8) Find me the best TWO description of using GLMM to compare abundance/diversity/richness at the Corridor Project: _________________
+
 ### Methods: 
@@
 Paragraph 1: Overall Topic & Why it is interesting and Important.
+_Fragmentation is an extensive problem across the globe and it is vital to understand shifts in biodiversity and community structures as a result of trophic cascades._
 
-  As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
+>As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
 Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003] [@fletcher_is_2018].
@@
 
->_Researchers today continue to search for strategies to mitigate the negative effects resulting from changes in landscape structure [@noh_forest_2022]. _
+Paragraph 2: What we know.
+_Habitat connectivity is generally viewed as beneficial for ecosystems as a whole.
+We know that patch shape and size can be an influental factor for how populations are configured.
+Wildlife corridors_
 
->Topic Sentence:  Fragmentation is an extensive problem across the globe and it is vital to understand shifts in biodiversity and community structures as a result of trophic cascades.
-
-Paragraph 2: What we know.
-
-Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003].
+>Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003].
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy (insert richard forman 1995).
@@
 
->Topic Sentence: Habitat connectivity is generally viewed as beneficial for ecosystems as a whole.
-We know that patch shape and size can be an influental factor for how populations are configured.
-Wildlife corridors
+Paragraph 3: What we don't know. _Changes in ecosystem fuction as a result of shifts in community structure (EB but we may not want to emphasize this given we didn't measure directly_
 
-Paragraph 3: What we don't know.
-
-By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes.
+>By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes.
 Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected.
@@
 
->Topic Sentence:  Changes in ecosystem fuction as a result of shifts in community structure (EB but we may not want to emphasize this given we didn't measure directly)
+<!-- Potential alternative routes:  -->
 
-Potential alternative routes:
-
-1) We know far less on how corridors influence functional diversity.
+<!-- 1) We know far less on how corridors influence functional diversity.
 Functional diversity is important for the maintenance of ecosystems services, etc, so corridors cvould help ensure that.
 (in context of dung beetles that would be feeding guilds (simplest), morphology etc.
-in deCastro Arrazola et al)
+in deCastro Arrazola et al) -->
 
-2) Studies investigating corridors rarely look at whats i the matrix.
+<!-- 2) Studies investigating corridors rarely look at whats i the matrix.
 The matrix could be a source of diversity that moves into patches and then through corridors or a completely different subset of species.
-(any other options?) Understanding whichh of these it is is important because it leads to differnt conclusions abiout role of corridors and conservation strategies
+(any other options?) Understanding whichh of these it is is important because it leads to differnt conclusions abiout role of corridors and conservation strategies -->
 
-3) regardless, not a bad idea to discuss both matrix and functional group in diuscussion.
+<!-- 3) regardless, not a bad idea to discuss both matrix and functional group in diuscussion. -->
 
-4) and the alternative approach: revamp into around spillover from matrix into patches and if depednent on corridors or edge
+<!-- 4) and the alternative approach: revamp into around spillover from matrix into patches and if depednent on corridors or edge  -->
 
@@
 
-Paragraph 5: Here is what we did to address this unknown.
+Paragraph 4: Study System and why it is so excellent for this issue.
+_Here we talk about dung beetles how they fit into the broader scope of our question and directly state what questions we are asking._
 
-Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al.
+>Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al.
 2024).
@@
 Additionally we asked are corridors benefiting any one functional trait over another?
+
 <!-- Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling. -->
 
->Topic Sentence:  Here we talk about dung beetles how they fit into the broader scope of our question and directly state what questions we are asking.
-
----
-
-> As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
-This degradation lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
-Effects from isolation can vary, however as habitats are broken down community structures are significantly altered [@WOS:000419965700012].
-Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003].
-As disturbance continues to intensify, it is becoming increasingly more important to understand how different taxonomic groups.
-Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes.
-
-> Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al.
-2024).
-The removal, breakdown, and burial of animal feces drive important ecostystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
-Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) [@WOS:000891747700001].
-Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].
-Past studies have also focused on how landscape structure alters the community compositions of dung beetles[@WOS:000398662200001], yet large landscape scale experimental studies with carefully controlled and replicated treatments are non-existent for this model species.
-
-> Here, we  aim to determine how connectivity and fragmentation affect Species Richness and Diversity, Abundance, and functional diversity.
-We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge.
-To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
-Additionally we asked (2) Are corridors benefiting any one functional trait over another?
-Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling.
+Paragraph 5: Here is what we did to address this unknown. _The hypotheses we tested_
 
@@
 Biodiversity between patch types was compared using Hill numbers (Jost 2006).
-We looked at community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
+Hill numbers are....
+They are an alternative to...We looked at community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
 Diversity numbers were calculated in R studio using package hillR and modeled using package lme4.
@@
 1. Were all species found in all blocks? Were any restricted to only 1-2 blocks?
+
 1. Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
+
 1. Number of species in each functional group
+
 1.
@@
 ### Functional diversity
-
----
-
-> Overall, `r total_N` total dung beetles were collected belonging to N = `r total_spp` species.
-The most dominant species were ______  (N = _____), _Ateuchus lecontei_ (n=_____), and _Phanaeus igneus_ (n=_958___) (Table 1).
-All but one species was found in matrix plots.
-Four species (_Onthophagus striatulus_, _Onthophagus concinnus_, _Geotrupes blackburnii_, _Onthophagus truberculifrons_) were the only species not representative in every patch type.
-We had two singleton species, _Geotrupes blackburnii_ appearing only once in winged, and _Onthophagus tuberculiofrons_ in matrix.
-
-> Beetle abundance was significantly higher in matrix plots with 33% (n = 1713) of captures followed by 26% (n = 1359) caught in connected patches, 23% (n = 1199) in winged, and lastly 18% (n = 942) from rectangular patches (Table 2).
-
-> Species richness was distinctly different between patch types with matrix plots exhibited the highest richness while connected patches were lowest (Fig.
-1).
-Estimations of Shannon diversity index were fairly uniform between patch types.
-Simpson diversity indices were similarly even however winged patches showed a lower value.
-
-> Notable Bray-Curtis dissimilarity values were between connected and matrix (BC = 0.127), connected and winged (BC = 0.077),  matrix and rectangle (BC = 0.318), and matrix and winged (BC = 0.189) (Table 3).
 
@@
 He later began to explore this curiosity deeper in beginning his graduate education at the University of Florida as a Master's Student under the supervision of Emilio Bruna.
+
+\newpage
+
+## Text Sandbox
+
+As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
+This degradation lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
+Effects from isolation can vary, however as habitats are broken down community structures are significantly altered [@WOS:000419965700012].
+Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003].
+As disturbance continues to intensify, it is becoming increasingly more important to understand how different taxonomic groups.
+Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes.
+
+Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al.
+2024).
+The removal, breakdown, and burial of animal feces drive important ecostystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
+Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) [@WOS:000891747700001].
+Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].
+Past studies have also focused on how landscape structure alters the community compositions of dung beetles[@WOS:000398662200001], yet large landscape scale experimental studies with carefully controlled and replicated treatments are non-existent for this model species.
+
+Here, we  aim to determine how connectivity and fragmentation affect Species Richness and Diversity, Abundance, and functional diversity.
+We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge.
+To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
+Additionally we asked (2) Are corridors benefiting any one functional trait over another?
+Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling.

### 2025-05-02 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -55,12 +55,30 @@ spp_abund<-btl_data %>%
   mutate(patch_block=paste(patch,block,sep="_")) %>% 
   ungroup() %>% 
   select(-patch,-block) %>% 
+  separate(patch_block,c("patch_type","block"),sep="_",remove=TRUE) 
+
+spp_abund_top6<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
+  mutate(patch=case_when(
+    patch == "c" ~ "Connected",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) %>%
+  group_by(species,patch,block) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(patch,desc(n)) %>%
+  rename(sp_code=species) %>%
+  mutate(patch_block=paste(patch,block,sep="_")) %>% 
+  ungroup() %>% 
+  select(-patch,-block) %>% 
   separate(patch_block,c("patch_type","block"),sep="_",remove=TRUE) %>% 
   filter(sp_code=="cvig"|
            sp_code=="alec"|
            sp_code=="pign"|
+           sp_code=="dcar"|
+           sp_code=="open"|
            sp_code=="aaeg")
-
 
 
 hill_data<-hill_data %>% column_to_rownames("patch_block")
@@ -145,8 +163,8 @@ hist(hill_results$h_rich,
      main = "")
 
 
-
-ggplot(spp_abund, 
+# foo<-spp_abund %>% group_by(patch_type,block) %>% summarize(n=sum(n))
+ggplot(spp_abund_top6, 
        aes(x = patch_type, 
            y = n,
            # y = h_rich,
@@ -159,13 +177,40 @@ ggplot(spp_abund,
   theme_bw()
 
 
-# Global model: Take 1
-M0 <- glmer(n ~ patch_type * sp_code + (1 + patch_type * sp_code | block), 
+
+# Global model: Take 1 abundance, indep of species ID
+M0 <- glmer(n ~ patch_type + (1 + patch_type | block), 
             data = spp_abund, 
             family = poisson)
 summary(M0)
 
 Anova(M0)
+
+# Model 2
+M2 <- glmer(n ~ sp_code + (1 | block), 
+            data = spp_abund, 
+            family = poisson)
+summary(M2)
+
+Anova(M2)
+
+
+
[…]

### 2025-05-21 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
answered some questions

@@
 
+cvig - open sand chaparral, open grassland preference nealis 1977
+mbis - open sand chapparal preference nealis 1977
+open - slight preference towards open chap and shaded chap, little open grassland nealis 1977
+
+pign - sharing forest habitat but need more info conover 2019
+pvin - open habitat but share forest with pign conover 2019
+alec - forest preference conover 2019
+
 2) Is there anything about these species that might be interesting in the context of corridors?
   - are any species known to be good fliers/movers?
+    - Ospina 2018 -> wing shape py habitat preference and differences in wing shape determined by species groupings
+    ^ beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies)
+    ^ large body beetles tending towards lower energy flight strategies. maybe justification for corridor preference?
+
+    = stanbrook and king 2022 -> tunnelers prefering open habitat and tunnelers also contributing more towards dung removal.
+
   - specialists on deer dung?
+    - Conover 2019 -> dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types
+
   - navigate by starlight?

### 2025-05-22 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
hill numbers

@@
 
+hill numbers are a set of biodiversity metrics designed to quantify biodiversity and richness differently depending on how we want to emphasize abundance while also accounting for differences in sampling sizes and efforts (read more about this Eric)
+
 5) What is the advantage of useing hill numbers to estimate diversity / richness?
+
+- Chao et al.
+2014 -> incorporating abundance into calculating diversity metrics is important for studying functionality in particular since some species that may not be as abundant as others may not contribute to an ecosystem function as much as others.
+
+- ^ At the same time rare species may be filling a specific niche or we might be concerned about exotics or at risk species.
 

### 2025-05-22 corridor_docs/eric_ms_thesis/04_GLMM.R
stats with miguel!

@@ -158,7 +158,7 @@ shelf(tidyverse,
       quiet = TRUE)
 
 
-hist(hill_results$h_rich,
+ hist(hill_results$h_rich,
      xlab = "h_rich",
      main = "")
 
@@ -186,6 +186,8 @@ summary(M0)
 
 Anova(M0)
 
+plot_model(M0, type = "pred")
+
 # Model 2
 M2 <- glmer(n ~ sp_code + (1 | block), 
             data = spp_abund, 
@@ -195,14 +197,29 @@ summary(M2)
 Anova(M2)
 
 
+plot_model(M2, type = "pred")
 
 # Model 3
+# sjplot 
+library(sjPlot)
+
 M3 <- glmer(n ~ sp_code * patch_type + (1 | block), 
             data = spp_abund_top6, 
             family = poisson)
 summary(M3)
 
 Anova(M3)
+
+plot_model(M3, type = "pred", terms = c("sp_code", "patch_type"))
+
+plot_model(M3, type = "pred", terms = c("patch_type", "sp_code"))
+
+
+?interaction.plot
+
+interaction.plot(M3)
+
+?plot_model
 
 # Global model: Take 1
 M1 <- glmer(n ~ patch_type * sp_code + (1 + patch_type * sp_code | block), 
@@ -220,6 +237,8 @@ summary(M_rich)
 
 Anova(M_rich)
 
+plot_model(M_rich, type = "pred")
+
 library(broom.mixed)
 tidy_model <- tidy(M_rich, effects = "fixed")
 
@@ -236,9 +255,10 @@ h_shannon
 
 ?glmer
 
-M_simpson <- glmer(h_simpson ~ patch_type + (1 + patch_type | block),
-                   data = h_simpson,
-                   family = poisson)
+h_simpson
+
+M_simpson <- lmer(h_simpson ~ patch_type + (1 | block),
+                   data = h_simpson)
 summary(M_simpson)
 
 
@@ -281,6 +301,10 @@ hill_results %>%
   labs(x = "Patch Type", y = "Species Richness")
 
 
+# figure attempts 
+
+
+# tables for model results 
 spp_table<-knitr::kable(spp_table_data, 
                         digits = 2,
                         align="lcrcccc",

### 2025-05-28 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
slight tweaks

@@
 
-    = stanbrook and king 2022 -> tunnelers prefering open habitat and tunnelers also contributing more towards dung removal.
+    = stanbrook and king 2022 -> tunnelers preferring open habitat and tunnelers also contributing more towards dung removal.
 
@@
     - Conover 2019 -> dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types 
+    - Gimenez Gomez et al 2021 -> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used.
+However this study was done in a more tropical ecosystem so its hard to say if the same would apply to our more temperate system so more work is needed in this specific avenue.
 
@@
 
-hill numbers are a set of biodiversity metrics designed to quantify biodiversity and richness differently depending on how we want to emphasize abundance while also accounting for differences in sampling sizes and efforts (read more about this Eric)
+hill numbers are a set of biodiversity metrics designed to standardize metrics and quantify biodiversity and richness differently depending on how we want to emphasize abundance while also accounting for differences in sampling sizes and efforts (read more about this Eric)
 
@@
 1. Figure of biomass of different (R/T/D) functional groups.
-1. Figure of abundance of individual species.
-1. ~~Figure of the patch layout with pitfalls~~.
+1. ~~Figure of abundance of individual species.~~
+1. ~~Figure of the patch layout with pitfalls.~~
 

### 2025-05-29 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
two paragrapghs baybee

@@
 Biodiversity between patch types was compared using Hill numbers (Jost 2006).
-Hill numbers are....
-They are an alternative to...We looked at community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
-Diversity numbers were calculated in R studio using package hillR and modeled using package lme4.
-For the purposes of this project richness was modeled using a generalized linear mixed model with a poisson distribution.
-A qqplot of our data shows that our residuals are normally distributed (Figure 1).
+Hill numbers are metrics developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity.
+In our work, we use Hill numbers to explore the impacts of abundance in our diversity measurements.
+They are an alternative to more specialized metrics which are less intuitive for interpretation.
+We looked at community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
+Diversity numbers and species richness were calculated in R studio using package hill.
+
+To test for the effects of connectivity we modeled species abundance, richness, and species diversity using Shannon and Simpson indexes.
+For abundance and richness we used generalized linear mixed models fitted to a poisson distribution.
+We looked at the abundance of the top 6 most common species collected and overall species richness per patch type using our sampling blocks as random effects.
+To model our diversity metrics we took a similar approach but this time using glmms with a Gaussian distribution.
+All models were fit using lme4 package and we used qqplots to evaluate the the sutability of our data using DARMa package.
+
+modeled using package lme4.
+
+For the purposes of this project richness was modeled using a generalized linear mixed model with a poisson distribution. A qqplot of our data shows that our residuals are normally distributed.
 

### 2025-05-29 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
super minor tweaks.

--- 2025-05-28 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
@@
 Biodiversity between patch types was compared using Hill numbers (Jost 2006).
-Hill numbers are....
-They are an alternative to...We looked at community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
-Diversity numbers were calculated in R studio using package hillR and modeled using package lme4.
-For the purposes of this project richness was modeled using a generalized linear mixed model with a poisson distribution.
-A qqplot of our data shows that our residuals are normally distributed (Figure 1).
+Hill numbers are metrics developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity.
+[delete: In our work, we use Hill numbers to explore the impacts of abundance in our diversity measurements].
+They are an alternative to more specialized metrics such as _________, which are less intuitive for interpretation.
+[maybe a sentence on how / why hill numbers are easier to interpret?]  We looked at [compared] community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
+[can you add a sentence about how these are calculated?
+by whihc I mean - are the calculations different than for richness, shannon, etc.?] Diversity numbers and species richness were calculated in R studio using package `hill`.
+
+To test for the effects of connectivity on species abundance, richness, and diversity [we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix].
+For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution.
+We looked at [compared...looked at sounds squishy.
+try to use more active verbs like compared, quantified, assessed, etc.] (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type.
+We included the identity of the sampling block as a random effects.
+To model our diversity metrics we took a similar approach, but this time using GLMMs with a Gaussian distribution.
+All models were fit using `lme4` package.
+Prior to conducting our modeling we used _qqplots_ (`DARMa` package) to evaluate the the suitability of our data.
 

### 2025-05-29 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
editing figures

@@
 
+results text blah blah (@sp_richness_patches)
+
+```{r  label = sp_richness_patches, echo = FALSE, fig.cap = "Dung beetle secies richness in three different.....and the forest matrix surrounding patches.", out.width = '75%'}
+knitr::include_graphics("corridor_docs/eric_ms_thesis/images/sp_richness_patches.png")
+```
 ### Species diversity

### 2025-05-29 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -1,6 +1,6 @@
 library(tidyverse)
 library(here)
-
+library(kableExtra)
 
 # load and prep data ------------------------------------------------------
 
@@ -273,6 +273,7 @@ simulationOutput <-
 
 # RICHNESS
 
+
 hill_results %>% 
 ggplot(aes(x = block, 
            # y = n,
@@ -288,17 +289,60 @@ ggplot(aes(x = block,
   
 
 
+# Emilio's suggestions for making snazzy figure. --------------------------
+
+# also includes code block to insert in Rmd
+
+sp_richness_patches<-  
 hill_results %>% 
+  mutate(patch_type = factor(patch_type,
+                             levels = c("Matrix", "Rectangle", "Winged", "Connected"))) %>%   # this changes the order of things on the x axis. suggest this order (Matrixm, then least to most)
   ggplot(aes(x = patch_type, 
              # y = n,
              y = h_rich,
              # y = h_shannon,
-             color = block)) +
+             color = block,
+             shape = block)) +
+  scale_shape_manual(values=c(15:18))+ # change shapes of points
   scale_color_viridis_d(option = "turbo") + 
-  geom_point(position="jitter") + 
-  theme_bw()+
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.7, hjust = 0.9))+
-  labs(x = "Patch Type", y = "Species Richness")
+  geom_point(position=position_jitter(width=0.16, height=0.15))+ # manipulate spread of jitter
+  theme_classic()+
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust = 0.95,size = 10))+
+  labs(x = "Patch Type", 
+       y = "Species Richness",
+       color="Block",   # edit legend title. be sure to do both shape & color
+       shape="Block")+ 
+  # theme(
+  #   legend.box.background = element_rect(color="black", linewidth = 0.1),
+  #   # legend.box.margin = margin(116, 6, 6, 6)
+  # )+
+  theme(
+    legend.box.background = element_rect(color="gray", size=1),
+    legend.box.margin = margin(0.1, 0.1, 0.1, 0.1),
+    legend.position = "top",
+    legend.text = element_text(size = 10, colour = "black"),
+    legend.title = element_text(size = 10, colour = "black")
+  )+
+  theme(axis.title.y = element_text(size = 12,face="bold"))+
+  theme(axis.title.x =element_text(size = 12,face="bold"))+
[…]

### 2025-05-29 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
image in rmd

--- 2025-05-29 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
@@
 
+![Dung beetle secies richness in three different --- and the forest matrix surrounding patches.](images/sp_richness_patches.png){height=100%}
+
+<!-- ```{r sp_richness_patches, echo = FALSE, fig.cap = "Dung beetle secies richness in three different --- and the forest matrix surrounding patches.", out.width = '100%'} -->
+<!-- knitr::include_graphics("./corridor_beetles/corridor_docs/eric_ms_thesis/images/sp_richness_patches.png") -->
+<!-- ``` -->
+
+\newpage
 ```{r Table1, echo=FALSE,message = FALSE,warning=FALSE}

### 2025-05-29 corridor_docs/eric_ms_thesis/04_GLMM.R
making table preety table

@@ -372,3 +372,34 @@ tidy_model <- tidy(M0, effects = "fixed")
 library(knitr)
 M0_table <- kable(tidy_model, digits = 3)
 M0_table
+
+
+
+
+# emilio's suggestion for stats table -------------------------------------
+
+# broom gets you almost all the way to your final table
+
+M0_table %>% 
+kable(digits = 2,
+      format = "latex",
+      caption = "this is where your caption foes",
+      align = "llcccc",
+      escape = FALSE,
+      row.names = FALSE,
+      booktabs = T,
+      linesep = ""
+) %>%
+  kable_styling(
+    bootstrap_options = c("hover"),
+    # full_width = F,
+    latex_options = c("scale_down","hold_position"),
+    font_size = 12,
+    position = "center"
+  ) 
+
+
+
+# end of emilio's stats table ---------------------------------------------
+
+

### 2025-05-31 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
tweaks and starts to results

@@
 
- - We also dried and weighed individuals to estimate biomass  
- - what guides/keys did you do to ID?  
- - "Voucher specimens of species I collected will be deposited at the Florida State Collection of Arthropods" (https://thefsca.org/services/specimen-donations/)
-
-## Analyses
-
 ### Overview
@@
 ```
+
+## Community Composition and Abundance
+
+Overall, 5213 total dung beetles were collected belonging to 16 species.
+The most dominant species were _Canthon vigilans_ (n = 1300) , _Ateuchus lecontei_ (n = 1112), and _Phanaeus igneus_ ( n =  958) (Table 1).
+All but four species were found in every patch type.
+_Onthophagus concinnus_ was found in the matrix and winged patches, while _Onthophagus striatulus_ was found in matrix habitat and rectangular patches.
+_Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively).
 

### 2025-05-31 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
decent chunk and some questions

@@
 Overall, 5213 total dung beetles were collected belonging to 16 species.
-The most dominant species were _Canthon vigilans_ (n = 1300) , _Ateuchus lecontei_ (n = 1112), and _Phanaeus igneus_ ( n =  958) (Table 1).
+Of the most dominant species, the top 6 comprised of 94% of total captures, those being _Canthon vigilans_ (n = 1300) , _Ateuchus lecontei_ (n = 1112), _Phanaeus igneus_ ( n =  958), _Dichotomius carolinus_ (n = 556), _Aphodeus alloblackburneus_ (n = 585), and _Onthophagus pennsylvanicus_ (n = 207) (Table 1).
 All but four species were found in every patch type.
@@
 _Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively).
+
+When taking into account abundance independent of species ID, patch type was not significantly related to the total beetle counts aside from the rectangle patch where fewer individuals were typically captured.
+Any further statistical analysis relating to abundance focused on the six most abundant species.
+
+### _QUESTION FOR EMILIO_ <-------
+And then here I want to talk about model M3 in 04_GLMM but this one is tricky there is a lot going on when looking at the model summary what should be the important info to report?
+
+What i think is significant 
+when looking at counts for our most abundant species most follow a similar pattern where the most individuals of a given species are most frequent in the matrix then corridor and matrix with a dip at the rectangle patch.
+One difference to note is that _Aphodeus alloblackburneus_ had much higher counts in matrix patches in contrast to other patch types.
+
+## Species Richness
+
+### Another _QUESTION FOR EMILIO_ <-------
+How best can I expand upon this more and same question for hill numbers, maybe means?
+
+Differences in species richness showed no significance across patch types.
+
+## Biodiversity
 

### 2025-06-02 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
some responses from EB

--- 2025-05-29 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
@@
 # by number (i.e., N most common)
-how_many_common<-3
+how_many_common<-6
 most_common_spp<-spp_abund %>% 
   slice_head(n = how_many_common)
+count_most_common<-sum(most_common_spp$n)
 
@@
 ```
+
+## Community Composition and Abundance
+
+Overall, I collected N = `r total_N` dung beetles belonging to N = `r total_spp` species.
+The N = `r how_many_common` species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: _Canthon vigilans_ (N = `r most_common_spp[1,1]`), _Ateuchus lecontei_ (N = `r most_common_spp[2,1]`), _Phanaeus igneus_ (N = `r most_common_spp[3,1]`), _Dichotomius carolinus_ (N = `r most_common_spp[4,1]`), _Aphodeus alloblackburneus_ (N = `r most_common_spp[5,1]`), and _Onthophagus pennsylvanicus_ (N = `r most_common_spp[6,1]`; Table 1).
+All but four species were captured in every patch type.
+_Onthophagus concinnus_ was only found in the matrix and winged patches, while _Onthophagus striatulus_ was only captured in matrix habitat and rectangular patches.
+_Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively).
+
+Wegen comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangulat patches.
+Statistical analysis of abundance relating to abundance focused on the six most abundant species.
+
+### _QUESTION FOR EMILIO_ <-------
+And then here I want to talk about model M3 in 04_GLMM but this one is tricky there is a lot going on when looking at the model summary what should be the important info to report?
+
+FROM EB: THE BIG BICTURE - sometimes the big picture is a clear effect (abundance of all species was higher in the winged patches).
+SOmetimes it is a split result (eg abundance of species a higher in winged, abund of species b higher in connected).
+And sometimes it is messy.
+that looks like the case here, so say that in thesis language: "the results were highly variable, with some species higher in A (eg, spec 1, 2) whiole others were higher in b (sp 3,4).
+For some species there was no effect of patch type on abundance (sp 5).
+Complete results are in table X.
+in other words in some cases "a lot going on" IS the message.
+
+What i think is significant:  
+when looking at counts for our most abundant species most follow a similar pattern where the most individuals of a given species are most frequent in the matrix then corridor and matrix with a dip at the rectangle patch.
+One difference to note is that _Aphodeus alloblackburneus_ had much higher counts in matrix patches in contrast to other patch types.
+
+EB: there you go! That's how its done.
+
+## Species Richness
+
+### Another _QUESTION FOR EMILIO_ <-------
+How best can I expand upon this more and same question for hill numbers, maybe means?
+
+Differences in species richness showed no significance across patch types.
+
+FROM EB: THIS IS WHAT YOU NEED TO SAY IN THE RESULTS.
+you can add a little.
+("The hill numbers ranged from X in habitat Y to X in habitat z.
+However, there was no significant...") You expand on what it *means* in the discussion.
+Here just report the facts.
+
+## Biodiversity
 

### 2025-06-02 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
clean up doc

@@
  - \usepackage{lineno}
-# - \linenumbers
+ - \linenumbers
  - \usepackage[default]{sourcesanspro}
@@
 
-1) Are any of the Dung Beetles we collected considered "open habitat" specialists? (might want to look at the work on pastures in the southeast)
+~~1) Are any of the Dung Beetles we collected considered "open habitat" specialists? (might want to look at the work on pastures in the southeast)~~
 
-cvig - open sand chaparral, open grassland preference nealis 1977
-mbis - open sand chapparal preference nealis 1977
-open - slight preference towards open chap and shaded chap, little open grassland nealis 1977
+- cvig - open sand chaparral, open grassland preference nealis 1977
+- mbis - open sand chapparal preference nealis 1977
+- open - slight preference towards open chap and shaded chap, little open grassland nealis 1977
 
-pign - sharing forest habitat but need more info conover 2019
-pvin - open habitat but share forest with pign conover 2019
-alec - forest preference conover 2019
+- pign - sharing forest habitat but need more info conover 2019
+- pvin - open habitat but share forest with pign conover 2019
+- alec - forest preference conover 2019
 
-2) Is there anything about these species that might be interesting in the context of corridors?
-  - are any species known to be good fliers/movers?
-    - Ospina 2018 -> wing shape py habitat preference and differences in wing shape determined by species groupings
-    ^ beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies)
-    ^ large body beetles tending towards lower energy flight strategies. maybe justification for corridor preference?
+~~2) Is there anything about these species that might be interesting in the context of corridors? Are any species known to be good fliers/movers? Specialists on deer dung? navigate by starlight?~~
 
-    = stanbrook and king 2022 -> tunnelers preferring open habitat and tunnelers also contributing more towards dung removal.
-
-  - specialists on deer dung?
-    - Conover 2019 -> dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types 
+  - Ospina 2018: wing shape py habitat preference and differences in wing shape determined by species groupings.
+Beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies).
+Large body beetles tending towards lower energy flight strategies.
+maybe justification for corridor preference?
+    - stanbrook and king 2022: tunnelers preferring open habitat and tunnelers also contributing more towards dung removal. 
+    - Conover 2019: dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types 
     - Gimenez Gomez et al 2021 -> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used.
@@
 
-  - navigate by starlight?
-  - feeding preferences?
-
 3) Did you get any species not captured in other long-leaf/SE USA forest sites?
 
-4) What are hill numbers?
+~~4) What are hill numbers?~~
 
-hill numbers are a set of biodiversity metrics designed to standardize metrics and quantify biodiversity and richness differently depending on how we want to emphasize abundance while also accounting for differences in sampling sizes and efforts (read more about this Eric)
+- hill numbers are a set of biodiversity metrics designed to standardize metrics and quantify biodiversity and richness differently depending on how we want to emphasize abundance while also accounting for differences in sampling sizes and efforts (read more about this Eric)
 
-5) What is the advantage of useing hill numbers to estimate diversity / richness?
+~~5) What is the advantage of useing hill numbers to estimate diversity / richness?~~
 
 - Chao et al.
-2014 -> incorporating abundance into calculating diversity metrics is important for studying functionality in particular since some species that may not be as abundant as others may not contribute to an ecosystem function as much as others.
-
-- ^ At the same time rare species may be filling a specific niche or we might be concerned about exotics or at risk species.
+2014: incorporating abundance into calculating diversity metrics is important for studying functionality in particular since some species that may not be as abundant as others may not contribute to an ecosystem function as much as others.
+At the same time rare species may be filling a specific niche or we might be concerned about exotics or at risk species.
 
@@
 ### Methods:
+
 1. ~~add a few sentences on drying, keys used to ID, and depositing vouchers~~
@@
 ### Analyses
-1. ~~Write a short paragraph on how you calculated and compared species richness and diversity (hill numbers, glmm). ~~
+
+1. ~~Write a short paragraph on how you calculated and compared species richness and diversity (hill numbers, glmm).~~
+
 1. Write a short paragraph on how you calculated and compared functional diversity (how you measured it, how you compared it)
@@
 ### Results
+
 1. answer the overview questions, them stich them together into a paragraph
+
  - ~~any species restricted to 1-2 habitat types~~
  - ~~Were all species found in all blocks? Were any restricted to only 1-2 blocks?~~
- - Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
+ - ~~Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)~~
  - Number of species in each functional group
@@
 
-**_Note from EB:_** _I don't want to see the words 'dung beetles' until paragraph 5_.
-
-Paragraph 1: Overall Topic & Why it is interesting and Important.
-_Fragmentation is an extensive problem across the globe and it is vital to understand shifts in biodiversity and community structures as a result of trophic cascades._
-
->As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
+As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
 Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003] [@fletcher_is_2018].
@@
 
-Paragraph 2: What we know.
-_Habitat connectivity is generally viewed as beneficial for ecosystems as a whole.
-We know that patch shape and size can be an influental factor for how populations are configured.
-Wildlife corridors_
-
->Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003].
+Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003].
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy (insert richard forman 1995).
@@
 
-Paragraph 3: What we don't know. _Changes in ecosystem fuction as a result of shifts in community structure (EB but we may not want to emphasize this given we didn't measure directly_
-
->By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes.
+By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes.
 Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected.
@@
 
-Paragraph 4: Study System and why it is so excellent for this issue.
-_Here we talk about dung beetles how they fit into the broader scope of our question and directly state what questions we are asking._
-
->Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al.
+Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al.
 2024).
@@
 Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna.
-In each replicant the central patch (100 x 100 m) is always connected to one peripheral patch with identical dimensions by a 150 x 25 m corridor, this will hereafter be referred  to as the connected patch.
+In each block the central patch (100 x 100 m) is always connected to one peripheral patch with identical dimensions by a 150 x 25 m corridor, this will hereafter be referred  to as the connected patch.
 The remaining patches are either "winged" or "rectangular".
@@
 
-## Analysis
+## Analyses
 
@@
 [maybe a sentence on how / why hill numbers are easier to interpret?]  We looked at [compared] community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
-[can you add a sentence about how these are calculated?
-by whihc I mean - are the calculations different than for richness, shannon, etc.?] Diversity numbers and species richness were calculated in R studio using package `hill`.
+_[can you add a sentence about how these are calculated?
+by which I mean - are the calculations different than for richness, shannon, etc.?]_ Diversity numbers and species richness were calculated in R studio using package `hill`.Diversity numbers were calculated in R studio using package iNEXT (Chao et al.
+2016).
+Bray-Curtis dissimiolarity values were calculated using package Vegan in R studio.
+Dung beetles were assigned traits by waste removal guild and habitat preference.
 
-To test for the effects of connectivity on species abundance, richness, and diversity [we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix].
+To test for the effects of connectivity on  abundance, speciesrichness, and species diversity [we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix].
 For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution.
@@
 
-### Overview
+<!-- ### Functional diversity -->
 
-<!-- _Don't have to describe these, just do it and report in results. this is just to remind us of what to do_ -->
+<!-- 1. 1nd pass: roller, tunneler, dweller -->
+<!-- 1. 2nd: morphological measurements like Alonso et al 2022 (see table from decastro 2023, Giménez Gómez 2025) -->
+<!-- 1. table: results of analyses  -->
+<!-- 1. visualization: biomass of different (R/T/D) functional groups?    -->
 
-<!-- 1. how many beetles total   -->
-<!-- 1. most common species overall.   -->
-<!-- 1. was. the same species the most common one in all habitats?   -->
-<!-- 1. were there any that were caught in all habitats? Any restricted to only one?   -->
-<!-- 1. any invasive / exotic species?   -->
-<!-- 1. Table: count by species by habitat.  -->
-
-### Species richness
-
-To Test for differences in species richness, we....
-
-1. glmm with poisson dist reccomended by julian
-1. table: results of analyses. 
-1. visualization: hill's plots? rank-abundance plots? rarefaction curve?
-
-### Species diversity
-
-We used a similar approach to compare species diversity in winged, isolated, and connected patches....
-
-1. glmm with poisson dist reccomended by julian
-1. table: results of analyses
-1. visualization: hill's plots? rarefaction curve?
-
-### Functional diversity
-
-1. 1nd pass: roller, tunneler, dweller
-1. 2nd: morphological measurements like Alonso et al 2022 (see table from decastro 2023, Giménez Gómez 2025)
-1. table: results of analyses 
-1. visualization: biomass of different (R/T/D) functional groups?
-
-_Reason for this breakdown: _ - richness and diversity are "community level" response.
+<!-- _Reason for this breakdown: _ - richness and diversity are "community level" response.
 abundance is a species-level response.
 changes in abundance or rank-abundance  could have implications for ecosystem services if it is connected to functional group.
+-->
 
----
-
-> Biodiversity between patch types was compared using Hill numbers (Jost 2006).
-We looked at community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
-Diversity numbers were calculated in R studio using package iNEXT (Chao et al.
-2016).
-Bray-Curtis dissimiolarity values were calculated using package Vegan in R studio.
-Dung beetles were assigned traits by waste removal guild and habitat preference.
-
-> Functional Diversity: Need to assign each species to a functional group: roller, tunneler. dweller, others?  - habitat preference (forest, pasture, generalist)
- - what stats / computer packages did you use to analyze and visualize the data?
-
->_Look through dung beetle pubs and see how/what people compare_
-
-<!--   - lets hammer this out  -->
-<!--   - modeling?  -->
-<!--     - glmm with poisson dist reccomended by julian -->
-<!--   - beta, abundance, biomass? per site -->
-<!--   - ~species list by sampling blocks (anything with this?)~ -->
-<!--   - habitat preference  -->
-<!--   - rarefaction -->
+<!-- Functional Diversity: Need to assign each species to a functional group: roller, tunneler.
+dweller, others?
+- habitat preference (forest, pasture, generalist).
+Look through dung beetle pubs and see how/what people compare_ what stats / computer packages did you use to analyze and visualize the data?
+-->
 
@@
 
-## Community Composition and Abundance
+## Overview: Community Composition and Abundance
 
 Overall, I collected N = `r total_N` dung beetles belonging to N = `r total_spp` species.
-The N = `r how_many_common` species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: _Canthon vigilans_ (N = `r most_common_spp[1,1]`), _Ateuchus lecontei_ (N = `r most_common_spp[2,1]`), _Phanaeus igneus_ (N = `r most_common_spp[3,1]`), _Dichotomius carolinus_ (N = `r most_common_spp[4,1]`), _Aphodeus alloblackburneus_ (N = `r most_common_spp[5,1]`), and _Onthophagus pennsylvanicus_ (N = `r most_common_spp[6,1]`; Table 1).
+The N = `r how_many_common` species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1).
 All but four species were captured in every patch type.
@@
 
-Wegen comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangulat patches.
+- any rare species? 
+- any invasive / exotic species?
+
+When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangulat patches.
 Statistical analysis of abundance relating to abundance focused on the six most abundant species.
@@
 
-## Species Richness
+### Species Richness
 
@@
 
-## Biodiversity
+### Species Diversity
 
-### Overview
-
-1. total number of beetles from total number of species: `r total_N` beetles from `r total_spp` species.
-
-1.
-All but four species species were found in every patch type.
-_Onthophagus concinnus_ was found in the matrix and winged patches, while _Onthophagus striatulus_ was found in matrix habitat and rectanular patches.
-_Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively).
-
-1. Were all species found in all blocks? Were any restricted to only 1-2 blocks?
-
-1. Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
-
-1. Number of species in each functional group
-
-1.
-most common 3-4 species: The `r how_many_common` most common species were \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), and \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`).
-1. any rare species? 
-1. any invasive / exotic species?
-
-### Species richness
-
-### Species diversity
-
-### Functional diversity
+### Functional Diversity
 
@@
 
-Recap main goals into findings
- abundances in matrix vs connected patch and why this could be happening
+## Recap main goals into findings
+
+ 1. Abundances in matrix vs connected patch and why this could be happening
  - source pop to habiitat edge
 
- species richness again supporting that matrix is more of an ideal habitat for dung beetle community
+2. species richness again supporting that matrix is more of an ideal habitat for dung beetle community
 
- diversity indices community structures weren't highly different between patch types
+3. diversity indices community structures weren't highly different between patch types
 
- bray curtis hinting at similar land uses between corridor and winged patch
+4. bray curtis hinting at similar land uses between corridor and winged patch
  - like julians paper corridors benefit certain populations and more fit populations are able to make better use
 
-Interpretation of results
- Abundance 
- - reasin for highest abundances in matrix and connected patch
+## Interpretation of results
+
+1. Abundance 
+ - reason for highest abundances in matrix and connected patch
  - lower abundance in rectangle hinting at fragmentation effects
- Species Richness & Diversity
+2. Species Richness & Diversity
  - the role of hbitat connectivity in shaping community structure or lack thereof
  - why do we think connnected had lowest species richness??
- Functional traits and ecological impacts
+3. Functional traits and ecological impacts
  - did corridoors favor a functional trait
@@
  - like duung removal papers seed dispersal and yep
-
-Comparing to previous studies
+4. Comparing to previous studies
 - how are things aligning
 - think about the biology 
-
-Limitations and future work
+5. Limitations and future work
 - potential confounding factors (seasonality, distance from edge, sampling methodology and temmporal variation)
 - other directions to go (dispersal -> radar, secifically measuring changes in ecosystem services)
-
-Takeaways for conservation and management
+6. Takeaways for conservation and management
 - dung beetles are robust

### 2025-06-02 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
reorg and some minor additions

@@
 _Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively).
+All species were within their native ranges.
 
@@
 
-When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangulat patches.
+When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches.
 Statistical analysis of abundance relating to abundance focused on the six most abundant species.
+When looking at counts for our most abundant species, results were variable but overall follow a similar pattern where the most individuals of a given species are more frequent in the matrix then corridor and winged patches with a dip at the rectangle patch.
+One difference to note is that _Aphodius alloblackburneus_ had much higher counts in matrix patches in contrast to other patch types.
 
@@
 What i think is significant:
-when looking at counts for our most abundant species most follow a similar pattern where the most individuals of a given species are most frequent in the matrix then corridor and matrix with a dip at the rectangle patch.
-One difference to note is that _Aphodeus alloblackburneus_ had much higher counts in matrix patches in contrast to other patch types.
 

### 2025-06-02 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
richhness

@@
 
-To test for the effects of connectivity on  abundance, speciesrichness, and species diversity [we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix].
+To test for the effects of connectivity on  abundance, species richness, and species diversity [we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix].
 For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution.
@@
 
-What i think is significant:
-
 EB: there you go! That's how its done.
@@
 Differences in species richness showed no significance across patch types.
+The total number of species captured in each treatment varied by sampling block, however species totals generally ranged from 8 to 13 species.
+The patch with the most species captured was in the matrix at block 53n while the fewest were captured at the rectangle patch in block 8 (sp_richness.png).
 

### 2025-06-03 corridor_docs/eric_ms_thesis/04_GLMM.R
added model for shannons div and visual

@@ -265,6 +265,16 @@ save_as_image(ft, path = "glmer_table.png")
 # Model for shannons 
 h_shannon
 
+M_shannon
+
+M_shannon <- lmer(h_shannon ~ patch_type + (1 | block),
+                  data = h_shannon)
+
+summary(M_shannon)
+
+plot_model(M_shannon, type = "pred")
+
+
 ?glmer
 
 h_simpson
@@ -273,6 +283,7 @@ M_simpson <- lmer(h_simpson ~ patch_type + (1 | block),
                    data = h_simpson)
 summary(M_simpson)
 
+plot_model(M_simpson, type = "pred")
 
 # residuals
 simulationOutput <- 

### 2025-06-05 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
bblurb for diversity
some organizing

@@
 
-As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
+As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented(here, fragmentation).
 Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003] [@fletcher_is_2018].
@@
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy (insert richard forman 1995).
-Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity.
+Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity(here, trophic cascades).
 
-By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes.
+By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes(experimental designs, maybe that one haddad).
 Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected.
@@
 
+Values for Hill numbers also showed no significance between patch types.
+Modeled values for both Shannon and Simpson's diversity had relatively high variance but averaged out at approximately qD = 5.29 and qD = 4.07 respectively.
+
 ### Functional Diversity

### 2025-06-05 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
inserted a citation

@@
 Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003].
-By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy (insert richard forman 1995).
+By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy [@forman_general_1995].
 Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity(here, trophic cascades).

### 2025-06-06 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
added citations and refreshed bib

@@
 
-As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented(here, fragmentation).
+As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented(Diaz 2019).
 Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003] [@fletcher_is_2018].
@@
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy [@forman_general_1995].
-Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity(here, trophic cascades).
+Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity(debinski and holt).
 
-By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes(experimental designs, maybe that one haddad).
+By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes[@fletcher_jr_landscape_2023; @resasco_contribution_2017].
 Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected.
-Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services.
+Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services(hevia 2017).
 
@@
 Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes.
-We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge.
+We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge[@tewksbury_corridors_2002].
 To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
@@
 All dung beetles were counted and identified to species as described in @nemes_illustrated_2015-1 and @edmonds_taxonomic_2023.
-Fifteen individuals of each species with adequate captures were dried to equalibrium and weighed for biomass measurements.
-Voucher specimcens for each species will be deposited at the Florida State Collection of Arthropods.
+Fifteen individuals of each species with adequate captures were dried to equilibrium and weighed for biomass measurements.
+Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods.
 
@@
 
-Biodiversity between patch types was compared using Hill numbers (Jost 2006).
-Hill numbers are metrics developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity.
-[delete: In our work, we use Hill numbers to explore the impacts of abundance in our diversity measurements].
+Biodiversity between patch types was compared using Hill numbers, a set of indeces developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity [@jost_entropy_2006].
 They are an alternative to more specialized metrics such as _________, which are less intuitive for interpretation.

### 2025-06-06 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
citations and tweaks to hills methods

@@
 
-Biodiversity between patch types was compared using Hill numbers, a set of indeces developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity [@jost_entropy_2006].
-They are an alternative to more specialized metrics such as _________, which are less intuitive for interpretation.
-[maybe a sentence on how / why hill numbers are easier to interpret?]  We looked at [compared] community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
+Biodiversity between patch types was compared using Hill numbers, a set of indexes developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity [@jost_entropy_2006].
+They are an alternative to more specialized metrics such as alpha, beta, and gamma diversities while being more standardized than other indexes such as Renyi or HCDT entropies, of which both groups of metrics are less intuitive for interpretation.
+_Hill numbers become the fitting choice for understanding community dynamics because they are extrapolated from the same equation, manipulating a single parameter (q) to reach our estimates, and also because when we manipulate q we can gain an understanding of compositional shifts otherwise obscured while using species richness[@chao_rarefaction_2014]_.
+We compared community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
 _[can you add a sentence about how these are calculated?
-by which I mean - are the calculations different than for richness, shannon, etc.?]_ Diversity numbers and species richness were calculated in R studio using package `hill`.Diversity numbers were calculated in R studio using package iNEXT (Chao et al.
-2016).
+by which I mean - are the calculations different than for richness, shannon, etc.?]_ Diversity numbers and species richness were calculated in R studio using package `hill`.Diversity numbers were calculated in R studio using package iNEXT [@hsieh_inext_2016].
 Bray-Curtis dissimiolarity values were calculated using package Vegan in R studio.

### 2025-06-06 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
pepo

@@
 For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution.
-We looked at [compared...looked at sounds squishy.
-try to use more active verbs like compared, quantified, assessed, etc.] (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type.
+Compared (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type.
 We included the identity of the sampling block as a random effects.

### 2025-06-06 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
last methods and made figures

@@
 
-- any rare species? 
-- any invasive / exotic species?
-
 When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches.
-Statistical analysis of abundance relating to abundance focused on the six most abundant species.
-When looking at counts for our most abundant species, results were variable but overall follow a similar pattern where the most individuals of a given species are more frequent in the matrix then corridor and winged patches with a dip at the rectangle patch.
+Statistical analysis of abundance focused on the six most abundant species.
+When looking at counts for our most abundant species, results were variable but overall follow a consistent pattern where the most individuals of a given species are more frequent in the matrix then corridor and winged patches with a dip at the rectangle patch.
 One difference to note is that _Aphodius alloblackburneus_ had much higher counts in matrix patches in contrast to other patch types.
-
-### _QUESTION FOR EMILIO_ <-------
-And then here I want to talk about model M3 in 04_GLMM but this one is tricky there is a lot going on when looking at the model summary what should be the important info to report?
-
-FROM EB: THE BIG BICTURE - sometimes the big picture is a clear effect (abundance of all species was higher in the winged patches).
-SOmetimes it is a split result (eg abundance of species a higher in winged, abund of species b higher in connected).
-And sometimes it is messy.
-that looks like the case here, so say that in thesis language: "the results were highly variable, with some species higher in A (eg, spec 1, 2) whiole others were higher in b (sp 3,4).
-For some species there was no effect of patch type on abundance (sp 5).
-Complete results are in table X.
-in other words in some cases "a lot going on" IS the message.
-
-EB: there you go! That's how its done.
 
@@
 
-### Another _QUESTION FOR EMILIO_ <-------
-How best can I expand upon this more and same question for hill numbers, maybe means?
-
-Differences in species richness showed no significance across patch types.
-The total number of species captured in each treatment varied by sampling block, however species totals generally ranged from 8 to 13 species.
-The patch with the most species captured was in the matrix at block 53n while the fewest were captured at the rectangle patch in block 8 (sp_richness.png).
-
-FROM EB: THIS IS WHAT YOU NEED TO SAY IN THE RESULTS.
-you can add a little.
-("The hill numbers ranged from X in habitat Y to X in habitat z.
-However, there was no significant...") You expand on what it *means* in the discussion.
-Here just report the facts.
+The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N).
+Species richness was relatively even across patch types with matrix and connected patches averaging out to approximately N = 11 species and rectangle and winged patches averaging to approximately N = 10 species.
+Overall there were no significant effects, however species richness patterns were variable between sampling blocks.
 
@@
 
-Values for Hill numbers also showed no significance between patch types.
-Modeled values for both Shannon and Simpson's diversity had relatively high variance but averaged out at approximately qD = 5.29 and qD = 4.07 respectively.
+Values for Hill numbers showed no patterns for shifts in species composition by patch type.
+Data for Shannon diversity showed no differences in the distribution of rare species between patch types.
+On average connected patches had a value of D = 5.29, while mean estimates for all other patches remained close with +0.03% for rectangle, -0.02% for matrix, and -0.06% for winged patches.
+There was no significance between patch types but data was highly variable between patch types and sampling blocks.
+
+Values for Simpson Diversity also showed no differences but instead for more abundant species composing the dung beetle assemblages.
+Simpson diversity had a mean value of D = 4.07 in the connected patch, also with other values in close proximity with matrix +0.02%, rectangle +0.04%, and winged at -0.1%.
+There was no significant difference between patch types, however this data was also highly variable between patch type and sampling blocks.
 

### 2025-06-06 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -370,6 +370,84 @@ ggsave("corridor_docs/eric_ms_thesis/images/sp_richness_patches.png", width = 4,
 
 # figure attempts 
 
+sp_shannon_patches<-  
+  hill_results %>% 
+  mutate(patch_type = factor(patch_type,
+                             levels = c("Matrix", "Rectangle", "Winged", "Connected"))) %>%   # this changes the order of things on the x axis. suggest this order (Matrixm, then least to most)
+  ggplot(aes(x = patch_type, 
+             # y = n,
+             # y = h_rich,
+             y = h_shannon,
+             color = block,
+             shape = block)) +
+  scale_shape_manual(values=c(15:18))+ # change shapes of points
+  scale_color_viridis_d(option = "turbo") + 
+  geom_point(position=position_jitter(width=0.16, height=0.15))+ # manipulate spread of jitter
+  theme_classic()+
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust = 0.95,size = 10))+
+  labs(x = "Patch Type", 
+       y = "Shannon Diversity Index",
+       color="Block",   # edit legend title. be sure to do both shape & color
+       shape="Block")+ 
+  # theme(
+  #   legend.box.background = element_rect(color="black", linewidth = 0.1),
+  #   # legend.box.margin = margin(116, 6, 6, 6)
+  # )+
+  theme(
+    legend.box.background = element_rect(color="gray", size=1),
+    legend.box.margin = margin(0.1, 0.1, 0.1, 0.1),
+    legend.position = "top",
+    legend.text = element_text(size = 10, colour = "black"),
+    legend.title = element_text(size = 10, colour = "black")
+  )+
+  theme(axis.title.y = element_text(size = 12,face="bold"))+
+  theme(axis.title.x =element_text(size = 12,face="bold"))+
+  theme(axis.text.y = element_text(size = 10))+
+  scale_y_continuous(limits = c(1, 10), breaks = seq(1, 10, by = 2)) # y axis limit and breaks
+
+
+# This saves a png version of file to the images folder
+ggsave("corridor_docs/eric_ms_thesis/images/sp_shannon_patches.png", width = 4, height = 4, units = "in")
+
+sp_simpson_patches<-  
+  hill_results %>% 
+  mutate(patch_type = factor(patch_type,
[…]

### 2025-06-09 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
update with images and tables

--- 2025-06-06 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
@@
  - \fancyhead[R]{p. \thepage} # name and page on header right
+ - \usepackage{subcaption}
+ - \usepackage{float}
 # # - \fancyfoot[R]{ }
@@
       keep_tex: no
+      extra_dependencies: ["float"]
       # number_sections: true
@@
 options(knitr.duplicate.label = 'allow')
+options(knitr.graphics.error = FALSE)  # This allows for the include_graphics to not use the path from root 
+knitr::opts_chunk$set(echo = TRUE, fig.align="center")
+knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
+knitr::opts_chunk$set(tab.pos = "H", out.extra = "")
 ```
@@
 library(grateful)
+library(knitr)
+# library(grid)
 # library(gridExtra)
+# library(png)
 # library(egg)
@@
 
-1. Figure of species richness results.
+1. ~~Figure of species richness results.~~
 1. Figure of species diversity results.
@@
 
-As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented(Diaz 2019).
+As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented(here, fragmentation).
 Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003] [@fletcher_is_2018].
@@
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy [@forman_general_1995].
-Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity(debinski and holt).
+Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity(here, trophic cascades).
 
-By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes[@fletcher_jr_landscape_2023; @resasco_contribution_2017].
+By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes(experimental designs, maybe that one haddad).
 Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected.
-Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services(hevia 2017).
+Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services.
 
@@
 Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes.
-We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge[@tewksbury_corridors_2002].
+We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge.
 To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
@@
 All dung beetles were counted and identified to species as described in @nemes_illustrated_2015-1 and @edmonds_taxonomic_2023.
-Fifteen individuals of each species with adequate captures were dried to equilibrium and weighed for biomass measurements.
-Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods.
+Fifteen individuals of each species with adequate captures were dried to equalibrium and weighed for biomass measurements.
+Voucher specimcens for each species will be deposited at the Florida State Collection of Arthropods.
 
@@
 
-Biodiversity between patch types was compared using Hill numbers, a set of indexes developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity [@jost_entropy_2006].
-They are an alternative to more specialized metrics such as alpha, beta, and gamma diversities while being more standardized than other indexes such as Renyi or HCDT entropies, of which both groups of metrics are less intuitive for interpretation.
-_Hill numbers become the fitting choice for understanding community dynamics because they are extrapolated from the same equation, manipulating a single parameter (q) to reach our estimates, and also because when we manipulate q we can gain an understanding of compositional shifts otherwise obscured while using species richness[@chao_rarefaction_2014]_.
-We compared community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
+Biodiversity between patch types was compared using Hill numbers (Jost 2006).
+Hill numbers are metrics developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity.
+[delete: In our work, we use Hill numbers to explore the impacts of abundance in our diversity measurements].
+They are an alternative to more specialized metrics such as _________, which are less intuitive for interpretation.
+[maybe a sentence on how / why hill numbers are easier to interpret?]  We looked at [compared] community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
 _[can you add a sentence about how these are calculated?
-by which I mean - are the calculations different than for richness, shannon, etc.?]_ Diversity numbers and species richness were calculated in R studio using package `hill`.Diversity numbers were calculated in R studio using package iNEXT [@hsieh_inext_2016].
+by which I mean - are the calculations different than for richness, shannon, etc.?]_ Diversity numbers and species richness were calculated in R studio using package `hill`.Diversity numbers were calculated in R studio using package iNEXT (Chao et al.
+2016).
 Bray-Curtis dissimiolarity values were calculated using package Vegan in R studio.
@@
 1. table: results of species richness analyses.
-1. Figure of species richness results.
+1. ~~Figure of species richness results.~~
 1. table: results of species diversity analyses.
@@
 1. Figure of biomass of different (R/T/D) functional groups.
-1. Map of where field site is located.
+1. ~~Map of where field site is located~~
 1. ~~Figure of the patch layout with pitfalls~~.
@@
 
+<!-- (text - Nathan Fig of plots) -->
 \newpage
 
-<!-- ```{r  label = pitfalls, echo = TRUE, fig.cap = "Experimental landscapes indicating location of the pitfall traps.", out.width = '50%'} -->
-<!-- knitr::include_graphics('./corridor_docs/eric_ms_thesis/pitfalls.png') -->
+<!-- ```{r, label = pitfalls, echo = FALSE, message = FALSE, warning = FALSE,fig.cap = 'Experimental landscapes indicating location of the pitfall traps.',  out.width="25%", fig.align='center'} -->
+
+<!-- a<-readPNG("images/pitfalls.png") -->
+<!-- b<-readPNG("images/srs-map.png") -->
+<!-- grid.arrange(rasterGrob(a),rasterGrob(b),ncol=1) -->
+
 <!-- ``` -->
 
-![Experimental landscapes indicating location of the pitfall traps.](pitfalls.png){height=50%}
+```{r, label = srs-map, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Experimental block with the different patches and distance between them.", out.width="40%", fig.align='center'}
+
+knitr::include_graphics("images/srs-map.png")
+
+```
+
+```{r, label = pitfalls, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Experimental block with the different patches and distance between them.", out.width="50%", fig.align='center'}
+
+knitr::include_graphics("images/pitfalls.png")
+
+```
+
+```{r, label = patches, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Experimental block with the different patches and distance between them.", out.width="50%", fig.align='center'}
+
+knitr::include_graphics("images/patches-pic.png")
+
+```
 
@@
 
-![Dung beetle secies richness in three different --- and the forest matrix surrounding patches.](images/sp_richness_patches.png){height=100%}
+```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle secies richness in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+
+knitr::include_graphics("images/sp_richness_patches.png")
+
+```
 
@@
                 full_width = F,
-                latex_options="scale_down",
+                latex_options=c("scale_down","hold_position"),
                 font_size = 12,
@@
   column_spec(1, italic = T) %>% 
-  row_spec(0, bold = T) %>% 
-  row_spec(1:15, hline_after = T) 
+  row_spec(0, bold = T) 
+# %>% 
+  # row_spec(1:15, hline_after = T) 
 spp_table
@@
 ```
----
 
@@
 ```
----
+
+\newpage
+
+```{r Table aovMO, echo=FALSE,message = FALSE,warning=FALSE}
+
+aovM0<-read_csv("./corridor_docs/eric_ms_thesis/tables/aovM0.csv") %>% 
+  kable(
+      digits = 3,
+      align="lcc",
+      format="latex",
+      row.names = FALSE,
+      booktabs=T,
+      linesep = "", #removes the blank line after every 5 lines
+      caption = "AOV TABLE CAPTION") %>%
+  kable_styling(bootstrap_options = c("striped"),
+                full_width = F,
+                latex_options=c("scale_down","hold_position"),
+                font_size = 12,
+                position = "center") %>%
+  # column_spec(1, italic = T) %>%
+  row_spec(0, bold = T)
+  # row_spec(1:15, hline_after = T)
+aovM0
+```
 

### 2025-06-09 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -139,6 +139,9 @@ isSingular(glmerLaplace, tol = 1e-4)
 summary(glmerLaplace)
 
 
+
+
+
 # GLMM 2 ------------------------------------------------------------------
 
 
@@ -182,17 +185,21 @@ ggplot(spp_abund_top6,
 M0 <- glmer(n ~ patch_type + (1 + patch_type | block), 
             data = spp_abund, 
             family = poisson)
-summary(M0)
 
 Anova(M0)
+aovM0<-Anova(M0)
 
 plot_model(M0, type = "pred")
 
+
+
+# EMILIO ADDITION - EXPORTING TABLE RESULTS TO UPLOAD TO THESIS DOC
+
 # REQUIRES CLEANUP, BUT HERE IS HOW TO GET THE STATS TABLES IN KABLEEXTRA FOR YOUR THESIS
-# M0 %>% broom::tidy()
-# tidy(M0) %>%
-#   kbl() %>%
-#   kable_styling(font_size = 8)
+tidyM0<-M0 %>% broom::tidy()
+aovM0<-aovM0 %>% broom::tidy()
+write_csv(tidyM0,"./corridor_docs/eric_ms_thesis/tables/m0.csv")
+write_csv(aovM0,"./corridor_docs/eric_ms_thesis/tables/aovM0.csv")
 # library(rempsyc)
 # (stats.table <- tidy(M0, conf.int = TRUE))
 # nice_table(stats.table, broom = "glmm")
@@ -208,8 +215,11 @@ summary(M2)
 
 Anova(M2)
 
-
 plot_model(M2, type = "pred")
+
+
+M2 %>% broom::tidy()
+
 
 # Model 3
 # sjplot 

### 2025-06-10 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
first chunk of discussion

@@
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy [@forman_general_1995].
-Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity(debinski and holt).
+Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity[@debinski_survey_2000].
 
@@
 Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected.
-Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services(hevia 2017).
+Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services[@hevia_trait-based_2017].
 
@@
 
-Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al.
-2024).
+Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions [@hasan_dung_2024].
 The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]
-Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes.
+Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources[@hanski_dung_1991], interact with corridors in their landscapes.
 We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge[@tewksbury_corridors_2002].
@@
 
-Our study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals (Tewksbury et al.
-2002).
+Our study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals [@tewksbury_corridors_2002].
 Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna.
@@
 ## Recap main goals into findings
+
+This study shows that while habitat type may have an impact on the species composition of dung beetles, the arrangement of individuals show little difference as a result of the presence of movement corridors.
+Past studies have shown similar trends where habitat type and forest regeneration stages are key in the partitioning of dung beetle assemblages[@conover_phenology_2019; @WOS:000386714700011; @arellano_response_2008].
+This is further supported by the fact that capture rates were lowest in the rectangular patch, the treatment with the least edge to area ratio, suggesting that dung beetles in the matrix act as a source population which feeds into open patches through habitat edge.
+
+ Richness was the same
+
+ The total number of dung beetle species captured N = `r total_spp` and the total individuals N = `r total_N` are similar to those in studies conducted in similar habitats and regions of the US [@conover_phenology_2019; @nealis_habitat_1977; @stanbrook_dung_2022].
 

### 2025-06-10 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -186,6 +186,7 @@ M0 <- glmer(n ~ patch_type + (1 + patch_type | block),
             data = spp_abund, 
             family = poisson)
 
+summary(M0)
 Anova(M0)
 aovM0<-Anova(M0)
 

### 2025-06-10 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
good chunk of writing

@@
 
-This study shows that while habitat type may have an impact on the species composition of dung beetles, the arrangement of individuals show little difference as a result of the presence of movement corridors.
+This study shows that while habitat type may have an impact on the individual abundances of dung beetles, the arrangement of individuals show little difference as a result of the presence of movement corridors.
 Past studies have shown similar trends where habitat type and forest regeneration stages are key in the partitioning of dung beetle assemblages[@conover_phenology_2019; @WOS:000386714700011; @arellano_response_2008].
@@
 
- Richness was the same
+While we did not detect any patterns of species richness between patch types in our modeling,
+
+_i want to look back at rarefied richness, was sampling complete in each patch?_
+
+A comparison of total species richness per patch type per block shows that the sampling plot with the most total species captured was in the matrix.
+
+Analysis of Hill numbers indicated that species compositions were even between patch types.
+Comparing values of Simpson's Diversity between patches determined that dominant species were not prevalent in any one treatment type.
+Likewise, the Shannon index values insisted that rare species throughout the study population were even.
+
+_here write about low sample sizes and reasons for why dung beetle communities would be so similar across treatment types.
+Dispersal, scale (roslin), maybe again a good reason to expect matrix as a source pop._
 
  The total number of dung beetle species captured N = `r total_spp` and the total individuals N = `r total_N` are similar to those in studies conducted in similar habitats and regions of the US [@conover_phenology_2019; @nealis_habitat_1977; @stanbrook_dung_2022].
+
+ revisit bray curtis
 

### 2025-06-10 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
update may 10

@@
 
-As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented(Diaz 2019).
-Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003] [@fletcher_is_2018].
-However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012].
+As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented **(Diaz 2019)**.
+Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003; @fletcher_is_2018].
+However, as habitats are broken down community structures are significantly altered **[@WOS:000419965700012]**.
 This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions  [@haddad_habitat_2015].
 
-Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003].
+Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation [@haddad_corridor_2003].
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy  [@forman_general_1995].
-Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity[@debinski_survey_2000].
+Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
 
-By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes[@fletcher_jr_landscape_2023; @resasco_contribution_2017].
+By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes [@fletcher_jr_landscape_2023; @resasco_contribution_2017].
 Past studies have measured changes in biodiversity for many different taxa  [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected.
-Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services[@hevia_trait-based_2017].
+Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services [@hevia_trait-based_2017].
 
@@
 The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]
-Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources[@hanski_dung_1991], interact with corridors in their landscapes.
-We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge[@tewksbury_corridors_2002].
+Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources [@hanski_dung_1991], interact with corridors in their landscapes.
+We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge [@tewksbury_corridors_2002].
 To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
@@
 
-Paragraph 5: Here is what we did to address this unknown. _The hypotheses we tested_
+**Paragraph 5:** Here is what we did to address this unknown. _The hypotheses we tested_
 
@@
 
-Our study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals [@tewksbury_corridors_2002].
+Our study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, US (33.208$^\circ$ N, 81.408$^\circ$ W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals [@tewksbury_corridors_2002].
 Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna.
-In each block the central patch (100 x 100 m) is always connected to one peripheral patch with identical dimensions by a 150 x 25 m corridor, this will hereafter be referred  to as the connected patch.
+In each block the central patch (100 $\times$ 100 m) is always connected to one peripheral patch with identical dimensions by a 150 $\times$ 25 m corridor, this will hereafter be referred  to as the connected patch.
 The remaining patches are either "winged" or "rectangular".
-The winged patch is also 100 x 100 m, however they exhibit their characteristic wings in the form of two 75 x 25 m offshoots meant to account for the extra area and edge space the corridor provides.
-The rectangular patch is 100 x 137.5 m also the same area as the space of the connected patch plus the corridor.
+The winged patch is also 100 $\times$ 100 m, however they exhibit their characteristic wings in the form of two 75 $\times$ 25 m offshoots meant to account for the extra area and edge space the corridor provides.
+The rectangular patch is 100 $\times$ 137.5 m also the same area as the space of the connected patch plus the corridor.
 Each block has a duplicate of either the winged or rectangle patch, all peripheral patches being 150 m from the center patch.
@@
 They are an alternative to more specialized metrics such as alpha, beta, and gamma diversities while being more standardized than other indexes such as Renyi or HCDT entropies, of which both groups of metrics are less intuitive for interpretation.
-_Hill numbers become the fitting choice for understanding community dynamics because they are extrapolated from the same equation, manipulating a single parameter (q) to reach our estimates, and also because when we manipulate q we can gain an understanding of compositional shifts otherwise obscured while using species richness[@chao_rarefaction_2014]_.
-We compared community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity).
-_[can you add a sentence about how these are calculated?
-by which I mean - are the calculations different than for richness, shannon, etc.?]_ Diversity numbers and species richness were calculated in R studio using package `hill`.Diversity numbers were calculated in R studio using package iNEXT [@hsieh_inext_2016].
-Bray-Curtis dissimiolarity values were calculated using package Vegan in R studio.
+Hill numbers are now the preferred metric for describing community dynamics for two reasons.
+First, they are extrapolated from the same equation, manipulating a single parameter (i.e., _q_) to arrive at estimates of richness and diversity.
+Second, by  manipulating _q_ we can gain an understanding of compositional shifts otherwise obscured while using species richness [@chao_rarefaction_2014].
+We compared community composition by increasing magnitudes of diversity components (i.e., _qD_) of _0D_ (i.e., species richness), _1D_ (i.e., Shannon entropy), and _2D_ (i.e, Simpson Diversity).
+<!--[can you add a sentence about how these are calculated?
+by which I mean - are the calculations different than for richness, shannon, etc.?]--> Diversity numbers and species richness were calculated using the package `hill` **(citation)** for the R statistical programming language  **(citation)**.
+Diversity numbers were calculated using package `iNEXT` [@hsieh_inext_2016].
+Bray-Curtis dissimilarity values were calculated using package `Vegan` **(citation)**.
 Dung beetles were assigned traits by waste removal guild and habitat preference.
 
-To test for the effects of connectivity on  abundance, species richness, and species diversity [we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix].
-For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution.
+To test for the effects of connectivity on  abundance, species richness, and species diversity we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix.
+For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution  **(bolker citation?)**.
 Compared (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type.
 We included the identity of the sampling block as a random effects.
-To model our diversity metrics we took a similar approach, but this time using GLMMs with a Gaussian distribution.
-All models were fit using `lme4` package.
-Prior to conducting our modeling we used _qqplots_ (`DARMa` package) to evaluate the the suitability of our data.
+To model our diversity metrics we took a similar approach, but this time using GLMMs with a Gaussian distribution  **(citation)**.
+All models were fit using `lme4` package  **(citation)**.
+Prior to conducting our modeling we evaluated the the suitability of our data with _qqplots_ generated with the `DARMa` package  **(citation)**.
 
@@
 -->
-
-\newpage
 

### 2025-06-11 corridor_docs/eric_ms_thesis/04_GLMM.R
rewwrote questions and poked around at tables

@@ -244,6 +244,15 @@ interaction.plot(M3)
 
 ?plot_model
 
+# Eric trying table
+
+tidyM3<-M3 %>% broom::tidy()
+aovM3<-aovM0 %>% broom::tidy()
+write_csv(tidyM0,"./corridor_docs/eric_ms_thesis/tables/m0.csv")
+write_csv(aovM0,"./corridor_docs/eric_ms_thesis/tables/aovM0.csv")
+
+
+
 # Global model: Take 1
 M1 <- glmer(n ~ patch_type * sp_code + (1 + patch_type * sp_code | block), 
             data = spp_abund, 
@@ -495,7 +504,7 @@ M0_table
 M0_table %>% 
 kable(digits = 2,
       format = "latex",
-      caption = "this is where your caption foes",
+      caption = "this is where your caption goes",
       align = "llcccc",
       escape = FALSE,
       row.names = FALSE,

### 2025-06-11 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
bib for packages

@@
 Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources [@hanski_dung_1991], interact with corridors in their landscapes.
-We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge [@tewksbury_corridors_2002].
-To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
+We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch area to edge ratio and distance to edge [@tewksbury_corridors_2002].
+To ask the questions: (1)
+How are species abundances of dung beetles distributed within isolated and connected patches and what are their relationships with assemblages in matrix habitat?
+(2) Does species richness vary when movement between patches is facilitated by movement corridors?
+(3) Are there shifts in species diversity/composition and what are the implications for ecosystem service functionality?
+
+<!--how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
 Additionally we asked are corridors benefiting any one functional trait over another?
+-->
 
@@
 Individual pitfall traps consisted of two components, a 10cm tall by 8 cm wide cylinder base topped with a funnel with a 10cm wide rim.
-Pig feces was suspended above the funnel by a 6.5 cm by 6.5 cm mesh square.
-For each sample period, traps were baited with pig dung between 8-9 pm and picked up 12 hours later, all beetles captured were stored in ethanol for further processing.
+We sourced pig feces from the University of Florida Swine Barn Unit.
+Bait was processed into 5cm wide balls and wrapped in a layer of coffee filter material.
+For each sample period, traps were buried flush with the ground and baited with pig dung between 8-9 pm and picked up 12 hours later, all beetles captured were stored in ethanol for further processing.
 In total 16 sampling rounds were carried out with 4 rounds per block, 196 samples were collected.
@@
 
-When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches.
+When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches(P < 2e-6).
 Statistical analysis of abundance focused on the six most abundant species.

### 2025-06-11 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
cite

@@
 <!--[can you add a sentence about how these are calculated?
-by which I mean - are the calculations different than for richness, shannon, etc.?]--> Diversity numbers and species richness were calculated using the package `hill` **(citation)** for the R statistical programming language  **(citation)**.
+by which I mean - are the calculations different than for richness, shannon, etc.?]--> Diversity numbers and species richness were calculated using the package `hill` [@li_hillr_2018] for the R statistical programming language  **(citation)**.
 Diversity numbers were calculated using package `iNEXT` [@hsieh_inext_2016].
-Bray-Curtis dissimilarity values were calculated using package `Vegan` **(citation)**.
+Bray-Curtis dissimilarity values were calculated using package `Vegan` [@oksanen_vegan_2025].
 Dung beetles were assigned traits by waste removal guild and habitat preference.

### 2025-06-12 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
small changes and bib

@@
 To model our diversity metrics we took a similar approach, but this time using GLMMs with a Gaussian distribution  **(citation)**.
-All models were fit using `lme4` package  **(citation)**.
+All models were fit using `lme4` package  [@bates_fitting_2015].
 Prior to conducting our modeling we evaluated the the suitability of our data with _qqplots_ generated with the `DARMa` package  **(citation)**.
@@
 
+```{r, label = shannon, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Shannon diversity in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+
+knitr::include_graphics("images/sp_shannon_patches.png")
+
+```
+
+```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+
+knitr::include_graphics("images/sp_simpson_patches.png")
+
+```
+
 <!-- ```{r sp_richness_patches, echo = FALSE, fig.cap = "Dung beetle secies richness in three different --- and the forest matrix surrounding patches.", out.width = '100%'} -->
@@
 
-aovM0<-read_csv("./corridor_docs/eric_ms_thesis/tables/aovM0.csv") %>% 
+aovM0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM0.csv") %>% 
   kable(

### 2025-06-13 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
inserted stats calues in methods and a bunch of tables. exported bray curtis im so tired

@@
 <!--[can you add a sentence about how these are calculated?
-by which I mean - are the calculations different than for richness, shannon, etc.?]--> Diversity numbers and species richness were calculated using the package `hill` [@li_hillr_2018] for the R statistical programming language  **(citation)**.
+by which I mean - are the calculations different than for richness, shannon, etc.?]--> Diversity numbers and species richness were calculated using the package `hill` [@li_hillr_2018] for the R statistical programming language  [@posit_team_rstudio_2025].
 Diversity numbers were calculated using package `iNEXT` [@hsieh_inext_2016].
@@
 All models were fit using `lme4` package  [@bates_fitting_2015].
-Prior to conducting our modeling we evaluated the the suitability of our data with _qqplots_ generated with the `DARMa` package  **(citation)**.
+Prior to conducting our modeling we evaluated the the suitability of our data with _qqplots_ generated with the `DARMa` package  [@hartig_dharma_2024].
 
@@
 
-When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches(P < 2e-6).
+When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches(GLMM, P < 2e-6, Table 4).
 Statistical analysis of abundance focused on the six most abundant species.
-When looking at counts for our most abundant species, results were variable but overall follow a consistent pattern where the most individuals of a given species are more frequent in the matrix then corridor and winged patches with a dip at the rectangle patch.
-One difference to note is that _Aphodius alloblackburneus_ had much higher counts in matrix patches in contrast to other patch types.
+When looking at counts for our most abundant species, results were variable but overall follow a consistent pattern where the most individuals of a given species are more frequent in the matrix then corridor and winged patches with a dip at the rectangle patch (Table 6).
+One difference to note is that _Aphodius alloblackburneus_ had much higher counts in matrix patches in contrast to other patch types (Figure 6).
 
@@
 The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N).
-Species richness was relatively even across patch types with matrix and connected patches averaging out to approximately N = 11 species and rectangle and winged patches averaging to approximately N = 10 species.
+Species richness was relatively even across patch types with matrix and connected patches averaging out to approximately N = 11 species and rectangle and winged patches averaging to approximately N = 10 species (Figure 4).
 Overall there were no significant effects, however species richness patterns were variable between sampling blocks.
@@
 On average connected patches had a value of D = 5.29, while mean estimates for all other patches remained close with +0.03% for rectangle, -0.02% for matrix, and -0.06% for winged patches.
-There was no significance between patch types but data was highly variable between patch types and sampling blocks.
+There was no significance between patch types but data was highly variable between patch types and sampling blocks (Figure 5).
 
@@
 Simpson diversity had a mean value of D = 4.07 in the connected patch, also with other values in close proximity with matrix +0.02%, rectangle +0.04%, and winged at -0.1%.
-There was no significant difference between patch types, however this data was also highly variable between patch type and sampling blocks.
+There was no significant difference between patch types, however this data was also highly variable between patch type and sampling blocks (Figure 7).
 
@@
 
+```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle abundance in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+
+knitr::include_graphics("images/sp_abund_patch.png")
+
+```
+
 ```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
@@
 
+```{r Table bray_patches, echo=FALSE,message = FALSE,warning=FALSE}
+
+bray<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/bray_patches.csv") %>% 
+  kable(
+      digits = 3,
+      align="lcc",
+      format="latex",
+      row.names = FALSE,
+      booktabs=T,
+      linesep = "", #removes the blank line after every 5 lines
+      caption = "Bray-Curtis dissimilarity comparing composition of dung beetles between patches.") %>%
+  kable_styling(bootstrap_options = c("striped"),
+                full_width = F,
+                latex_options=c("scale_down","hold_position"),
+                font_size = 12,
+                position = "center") %>%
+  # column_spec(1, italic = T) %>%
+  row_spec(0, bold = T)
+  # row_spec(1:15, hline_after = T)
+bray
+```
+
 \newpage
@@
 aovM0
+```
+
+```{r Table MO, echo=FALSE,message = FALSE,warning=FALSE}
+
+M0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M0.csv") %>% 
+  kable(
+      digits = 3,
+      align="lcc",
+      format="latex",
+      row.names = FALSE,
+      booktabs=T,
+      linesep = "", #removes the blank line after every 5 lines
+      caption = "AOV TABLE CAPTION") %>%
+  kable_styling(bootstrap_options = c("striped"),
+                full_width = F,
+                latex_options=c("scale_down","hold_position"),
+                font_size = 12,
+                position = "center") %>%
+  # column_spec(1, italic = T) %>%
+  row_spec(0, bold = T)
+  # row_spec(1:15, hline_after = T)
+M0
+```
+
+```{r Table aovMO3, echo=FALSE,message = FALSE,warning=FALSE}
+
+aovM3<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM3.csv") %>% 
+  kable(
+      digits = 3,
+      align="lcc",
+      format="latex",
+      row.names = FALSE,
+      booktabs=T,
+      linesep = "", #removes the blank line after every 5 lines
+      caption = "AOV TABLE CAPTION") %>%
+  kable_styling(bootstrap_options = c("striped"),
+                full_width = F,
+                latex_options=c("scale_down","hold_position"),
+                font_size = 12,
+                position = "center") %>%
+  # column_spec(1, italic = T) %>%
+  row_spec(0, bold = T)
+  # row_spec(1:15, hline_after = T)
+aovM3
+```
+
+```{r Table M3, echo=FALSE,message = FALSE,warning=FALSE}
+
+M3<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M3.csv") %>% 
+  kable(
+      digits = 3,
+      align="lcc",
+      format="latex",
+      row.names = FALSE,
+      booktabs=T,
+      linesep = "", #removes the blank line after every 5 lines
+      caption = "AOV TABLE CAPTION") %>%
+  kable_styling(bootstrap_options = c("striped"),
+                full_width = F,
+                latex_options=c("scale_down","hold_position"),
+                font_size = 12,
+                position = "center") %>%
+  # column_spec(1, italic = T) %>%
+  row_spec(0, bold = T)
+  # row_spec(1:15, hline_after = T)
+M3
 ```

### 2025-06-13 corridor_docs/eric_ms_thesis/02_data_analysis.R

@@ -735,8 +735,18 @@ hill_plot
 #taxa_bray = vegan::vegdist(data, method = "bray", binary = FALSE)
 #taxa_bray
 
+library(readr)
+
 taxa_bray <- vegan::vegdist(btl_spec_patch, method = "bray", binary = FALSE)
 taxa_bray
+
+
+matrix_bray<-as.matrix(taxa_bray)
+
+df_bray <- as.data.frame(matrix_bray)
+
+write_csv(df_bray,"./corridor_docs/eric_ms_thesis/tables/bray_patches.csv")
+
 
 
 

### 2025-06-13 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -129,6 +129,8 @@ shapiro.test(resid(model))
 qqnorm(resid(model)) # plot of the normal theoretical quantiles against the exponential data quantiles
 qqline(resid(model)) # the residuals should fall along this line if they are normally distributed
 
+?qqnorm
+
 # GLMM
 glmerLaplace <- glmer(formula = h_rich ~ patch_type + (1 | block),
                       data = hill_results,
@@ -160,6 +162,7 @@ shelf(tidyverse,
       lib = tempdir(),
       quiet = TRUE)
 
+library(DHARMa)
 
  hist(hill_results$h_rich,
      xlab = "h_rich",
@@ -231,11 +234,10 @@ M3 <- glmer(n ~ sp_code * patch_type + (1 | block),
             family = poisson)
 summary(M3)
 
-Anova(M3)
+aovM3 <- Anova(M3)
 
 plot_model(M3, type = "pred", terms = c("sp_code", "patch_type"))
 
-plot_model(M3, type = "pred", terms = c("patch_type", "sp_code"))
 
 
 ?interaction.plot
@@ -247,9 +249,9 @@ interaction.plot(M3)
 # Eric trying table
 
 tidyM3<-M3 %>% broom::tidy()
-aovM3<-aovM0 %>% broom::tidy()
-write_csv(tidyM0,"./corridor_docs/eric_ms_thesis/tables/m0.csv")
-write_csv(aovM0,"./corridor_docs/eric_ms_thesis/tables/aovM0.csv")
+aovM3<-aovM3 %>% broom::tidy()
+write_csv(tidyM3,"./corridor_docs/eric_ms_thesis/tables/m3.csv")
+write_csv(aovM3,"./corridor_docs/eric_ms_thesis/tables/aovM3.csv")
 
 
 
@@ -468,6 +470,40 @@ sp_simpson_patches<-
 
 # This saves a png version of file to the images folder
 ggsave("corridor_docs/eric_ms_thesis/images/sp_simpson_patches.png", width = 4, height = 4, units = "in")
+
+# Figure for abundance model 
+
+?plot_model
+
+btl_abund_patch <- plot_model(M3, type = "pred", terms = c("patch_type", "sp_code"), title = NULL)
+btl_abund_patch
+
+btl_abund_patch + 
+  theme_minimal() +
+  labs( title = NULL ,
+       x = "Patch Type",
+       y = "Number of Beetles") +
+  theme(axis.text = element_text(size = 12)) +
[…]

### 2025-06-14 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
big ol writing session 1

@@
 
-This study shows that while habitat type may have an impact on the individual abundances of dung beetles, the arrangement of individuals show little difference as a result of the presence of movement corridors.
+This study advances our understanding of the factors shaping dung beetle community composition in temperate regions of the southeastern United States.
+In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches.
+Our main findings emphasized: (1) Dominant species were most abundant in forested matrix, but not equally for all species, some were represented proportionally higher in matrix than in open patch habitat.
+(2) Species richness and diversity were relatively consistent across all treatments, however variation across sampling blocks is evident.
+
+While habitat type may have an impact on the individual abundances of dung beetles, the arrangement of individuals show little difference as a result of the presence of movement corridors.
 Past studies have shown similar trends where habitat type and forest regeneration stages are key in the partitioning of dung beetle assemblages[@conover_phenology_2019; @WOS:000386714700011; @arellano_response_2008].
-This is further supported by the fact that capture rates were lowest in the rectangular patch, the treatment with the least edge to area ratio, suggesting that dung beetles in the matrix act as a source population which feeds into open patches through habitat edge.
+This is further supported by the fact that capture rates were lowest in the rectangular patch, the treatment with the least edge to area ratio, suggesting that dung beetles in the matrix may act as a source population which feeds into open patches through habitat edge.
 
-While we did not detect any patterns of species richness between patch types in our modeling,
-
-_i want to look back at rarefied richness, was sampling complete in each patch?_
-
-A comparison of total species richness per patch type per block shows that the sampling plot with the most total species captured was in the matrix.
+We did not detect any patterns of species richness between patch types in our modeling.
+Yet in comparing raw species counts the sampling plot with the most total species captured was in the matrix while the fewest species sampled were in a rectangular patch.
+_here look at other srs papers and see whats going on with edges_.
 
@@
  - lower abundance in rectangle hinting at fragmentation effects
+
 2. Species Richness & Diversity
@@
  - why do we think connnected had lowest species richness??
+
 3. Functional traits and ecological impacts
@@
  - like duung removal papers seed dispersal and yep
+
 4. Comparing to previous studies
@@
 - think about the biology
+
 5. Limitations and future work
@@
 - other directions to go (dispersal -> radar, secifically measuring changes in ecosystem services)
+
 6. Takeaways for conservation and management
@@
 
-We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site, particularly Sabrie Breland, Ben Overly, and Eva Schwarz.
-We also thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry, and to Nico Acajabon, Jarrett Emory, and Cayla Garmen for their advice in putting together this manuscript.
+We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site.
+We also thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry.
 
@@
 
+```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+
+knitr::include_graphics("images/sp_simpson_patches.png")
+
+```
+
 ```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle secies richness in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
@@
 knitr::include_graphics("images/sp_abund_patch.png")
-
-```
-
-```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
-
-knitr::include_graphics("images/sp_simpson_patches.png")
 

### 2025-06-14 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -273,6 +273,8 @@ Anova(M_rich)
 
 plot_model(M_rich, type = "pred")
 
+
+
 library(broom.mixed)
 tidy_model <- tidy(M_rich, effects = "fixed")
 

### 2025-06-14 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
gargantuan writing effort p2

@@
 Overall, I collected N = `r total_N` dung beetles belonging to N = `r total_spp` species.
-The N = `r how_many_common` species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1).
-All but four species were captured in every patch type.
+The N = `r how_many_common` most dominant species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1).
+All but six species were captured in every patch type.
 _Onthophagus concinnus_ was only found in the matrix and winged patches, while _Onthophagus striatulus_ was only captured in matrix habitat and rectangular patches.
@@
 
-When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches(GLMM, P < 2e-6, Table 4).
+When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches(GLMM, P < 0.001, Table 4).
+
 Statistical analysis of abundance focused on the six most abundant species.
-When looking at counts for our most abundant species, results were variable but overall follow a consistent pattern where the most individuals of a given species are more frequent in the matrix then corridor and winged patches with a dip at the rectangle patch (Table 6).
-One difference to note is that _Aphodius alloblackburneus_ had much higher counts in matrix patches in contrast to other patch types (Figure 6).
+A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 7).
+The baseline abundance (intercept) corresponds to the abundance of _Aphodius alloblackburneus_ in connected patches.
+Compared to this baseline, _Canthon vigilans_, _Ateuchus lecontei_, and _Phanaeus igneus_ showed significantly higher abundance  (β = 1.05, 0.73, 0.66 respectively; all p < 0.001), while _Onthophagus pennsylvanicus_ had lower abundance (β = –0.80, p < 0.001).
+
+Species-patch interactions indicated indicated species specific responses to treatments.
+In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix β = –0.59, p < 0.001).
+_Ateuchus lecontei_ and _Phanaeus igneus_ showed increased abundance in rectangle patches relative to expectations (β = 0.40 and 0.47, respectively; p < 0.05).
+Random effects indicated moderate variability among blocks (standard deviation of intercept = 0.51).
 
@@
 Our main findings emphasized: (1) Dominant species were most abundant in forested matrix, but not equally for all species, some were represented proportionally higher in matrix than in open patch habitat.
-(2) Species richness and diversity were relatively consistent across all treatments, however variation across sampling blocks is evident.
+(2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level.
+(3) Species diversity metrics were also relatively even across patch types however varied widely by block.
+These results suggest that dung beetle species are fully capable of permeating fragmented landscapes and that habitat type and connectivity shape community compositions, but landscape effects of a larger scale are driving changes in biodiversity.
+
+## Abundance
 
@@
 
+## Richness
+
 We did not detect any patterns of species richness between patch types in our modeling.
@@
 _here look at other srs papers and see whats going on with edges_.
+
+Diversity
 

### 2025-06-14 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
writing extravanamza p3

@@
 The baseline abundance (intercept) corresponds to the abundance of _Aphodius alloblackburneus_ in connected patches.
-Compared to this baseline, _Canthon vigilans_, _Ateuchus lecontei_, and _Phanaeus igneus_ showed significantly higher abundance  (β = 1.05, 0.73, 0.66 respectively; all p < 0.001), while _Onthophagus pennsylvanicus_ had lower abundance (β = –0.80, p < 0.001).
+Compared to this baseline, _Canthon vigilans_, _Ateuchus lecontei_, and _Phanaeus igneus_ showed significantly higher abundance  ($\beta$ = 1.05, 0.73, 0.66 respectively; all p < 0.001), while _Onthophagus pennsylvanicus_ had lower abundance ($\beta$ = –0.80, p < 0.001).
 
 Species-patch interactions indicated indicated species specific responses to treatments.
-In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix β = –0.59, p < 0.001).
-_Ateuchus lecontei_ and _Phanaeus igneus_ showed increased abundance in rectangle patches relative to expectations (β = 0.40 and 0.47, respectively; p < 0.05).
+In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix $\beta$ = –0.59, p < 0.001).
+_Ateuchus lecontei_ and _Phanaeus igneus_ showed increased abundance in rectangle patches relative to expectations ($\beta$ = 0.40 and 0.47, respectively; p < 0.05).
 Random effects indicated moderate variability among blocks (standard deviation of intercept = 0.51).
@@
 
-The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N).
-Species richness was relatively even across patch types with matrix and connected patches averaging out to approximately N = 11 species and rectangle and winged patches averaging to approximately N = 10 species (Figure 4).
-Overall there were no significant effects, however species richness patterns were variable between sampling blocks.
+Plotting species richness by patch type reveals consistent richness across patch types with some variation between sampling blocks (Figure 5).The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N).
+Modeling the effect of patch type on species richness with block as a random effect determined there was no significant differences among patch types.
+Comparing treatments using connected patches as a baseline resulted in no significant differences in matrix ($\beta$ = 0.05, p = 0.83), rectangle ($\beta$ = -0.07, p = 0.74), and winged patches ($\beta$ = -0.05, p = 0.83).
 
@@
 
-Values for Hill numbers showed no patterns for shifts in species composition by patch type.
-Data for Shannon diversity showed no differences in the distribution of rare species between patch types.
-On average connected patches had a value of D = 5.29, while mean estimates for all other patches remained close with +0.03% for rectangle, -0.02% for matrix, and -0.06% for winged patches.
-There was no significance between patch types but data was highly variable between patch types and sampling blocks (Figure 5).
+We used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation.
+In both models, the reference level for patch type was Connected.
 
-Values for Simpson Diversity also showed no differences but instead for more abundant species composing the dung beetle assemblages.
-Simpson diversity had a mean value of D = 4.07 in the connected patch, also with other values in close proximity with matrix +0.02%, rectangle +0.04%, and winged at -0.1%.
-There was no significant difference between patch types, however this data was also highly variable between patch type and sampling blocks (Figure 7).
+For Shannon Diversity, the estimated mean in Connected patches was 5.29 (SE = 0.66, t = 7.95).
+None of the alternative patch types showed statistically significant differences compared to Connected: Matrix ($\beta$ = –0.085, SE = 0.50, t = –0.17), Rectangle ($\beta$ = 0.15, SE = 0.50, t = 0.30), or Winged ($\beta$ = –0.34, SE = 0.50, t = –0.68).
 
-### Functional Diversity
+Similarly, for Simpson's Diversity, the average value in Connected patches was 4.07 (SE = 0.64, t = 6.37).
+Again, none of the other patch types had significant effects: Matrix ($\beta$ = 0.11, SE = 0.41, t = 0.26), Rectangle ($\beta$ = 0.17, SE = 0.41, t = 0.40), and Winged ($\beta$ = –0.37, SE = 0.41, t = –0.91).
+
+Across both models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.
 
@@
 
+```{r Table M_rich, echo=FALSE,message = FALSE,warning=FALSE}
+
+M_rich<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_rich.csv") %>% 
+  kable(
+      digits = 3,
+      align="lcc",
+      format="latex",
+      row.names = FALSE,
+      booktabs=T,
+      linesep = "", #removes the blank line after every 5 lines
+      caption = "Richness Model") %>%
+  kable_styling(bootstrap_options = c("striped"),
+                full_width = F,
+                latex_options=c("scale_down","hold_position"),
+                font_size = 12,
+                position = "center") %>%
+  # column_spec(1, italic = T) %>%
+  row_spec(0, bold = T)
+  # row_spec(1:15, hline_after = T)
+M_rich
+```
+
+```{r Table M_shan, echo=FALSE,message = FALSE,warning=FALSE}
+
+M_shan<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_shan.csv") %>% 
+  kable(
+      digits = 3,
+      align="lcc",
+      format="latex",
+      row.names = FALSE,
+      booktabs=T,
+      linesep = "", #removes the blank line after every 5 lines
+      caption = "Shannon Diversity Model") %>%
+  kable_styling(bootstrap_options = c("striped"),
+                full_width = F,
+                latex_options=c("scale_down","hold_position"),
+                font_size = 12,
+                position = "center") %>%
+  # column_spec(1, italic = T) %>%
+  row_spec(0, bold = T)
+  # row_spec(1:15, hline_after = T)
+M_shan
+```
+
+```{r Table M_simp, echo=FALSE,message = FALSE,warning=FALSE}
+
+M_simp<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_simp.csv") %>% 
+  kable(
+      digits = 3,
+      align="lcc",
+      format="latex",
+      row.names = FALSE,
+      booktabs=T,
+      linesep = "", #removes the blank line after every 5 lines
+      caption = "Simpson's Index Model") %>%
+  kable_styling(bootstrap_options = c("striped"),
+                full_width = F,
+                latex_options=c("scale_down","hold_position"),
+                font_size = 12,
+                position = "center") %>%
+  # column_spec(1, italic = T) %>%
+  row_spec(0, bold = T)
+  # row_spec(1:15, hline_after = T)
+M_simp
+```
+
 \newpage

### 2025-06-14 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -286,6 +286,9 @@ library(flextable)
 ft <- flextable(tidy_model)
 save_as_image(ft, path = "glmer_table.png")
 
+write_csv(tidy_model,"./corridor_docs/eric_ms_thesis/tables/M_rich.csv")
+
+
 # Model for shannons 
 h_shannon
 
@@ -298,6 +301,10 @@ summary(M_shannon)
 
 plot_model(M_shannon, type = "pred")
 
+tidyM_shan<-M_shannon %>% broom::tidy()
+write_csv(tidyM_shan,"./corridor_docs/eric_ms_thesis/tables/m_shan.csv")
+
+
 
 ?glmer
 
@@ -308,6 +315,9 @@ M_simpson <- lmer(h_simpson ~ patch_type + (1 | block),
 summary(M_simpson)
 
 plot_model(M_simpson, type = "pred")
+
+tidyMsimp<-M_simpson %>% broom::tidy()
+write_csv(tidyMsimp,"./corridor_docs/eric_ms_thesis/tables/m_simp.csv")
 
 # residuals
 simulationOutput <- 
@@ -560,5 +570,8 @@ kable(digits = 2,
 
 
 # end of emilio's stats table ---------------------------------------------
+
+
+
 
 

### 2025-06-16 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
start of erics big ol writing session 2 electic boogaloo

@@
 The N = `r how_many_common` most dominant species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1).
-All but six species were captured in every patch type.
+All but four species were captured in every patch type.
 _Onthophagus concinnus_ was only found in the matrix and winged patches, while _Onthophagus striatulus_ was only captured in matrix habitat and rectangular patches.
@@
 In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches.
-Our main findings emphasized: (1) Dominant species were most abundant in forested matrix, but not equally for all species, some were represented proportionally higher in matrix than in open patch habitat.
+Our main findings emphasized: (1) Habitat type and pathc shape were the main drivng factors for dung beetle species abundances were composed, however effects were species specific.
 (2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level.
-(3) Species diversity metrics were also relatively even across patch types however varied widely by block.
+(3) Species diversity metrics were also relatively even across patch types however varied widely by sampling blocks.
 These results suggest that dung beetle species are fully capable of permeating fragmented landscapes and that habitat type and connectivity shape community compositions, but landscape effects of a larger scale are driving changes in biodiversity.
@@
 
-While habitat type may have an impact on the individual abundances of dung beetles, the arrangement of individuals show little difference as a result of the presence of movement corridors.
+The total number of dung beetle species captured N = `r total_spp` and the total individuals N = `r total_N` are similar to those in studies conducted in similar habitats and regions of the US [@conover_phenology_2019; @nealis_habitat_1977; @stanbrook_dung_2022].
+
+Compared to the other treatment types the greatest number of individual beetles were captured from matrix habitat.
+Likewise, for each of the most dominant species the highest counts came from the matrix.
+
 Past studies have shown similar trends where habitat type and forest regeneration stages are key in the partitioning of dung beetle assemblages[@conover_phenology_2019; @WOS:000386714700011; @arellano_response_2008].
@@
 
-Diversity
+##Diversity
 
@@
 Dispersal, scale (roslin), maybe again a good reason to expect matrix as a source pop._
-
- The total number of dung beetle species captured N = `r total_spp` and the total individuals N = `r total_N` are similar to those in studies conducted in similar habitats and regions of the US [@conover_phenology_2019; @nealis_habitat_1977; @stanbrook_dung_2022].
 
@@
 1. ~~Table: count by species by habitat.~~
-1. table: results of species richness analyses.
+1. ~~table: results of species richness analyses.~~
 1. ~~Figure of species richness results.~~
-1. table: results of species diversity analyses.
-1. Figure of species diversity results.
+1. ~~table: results of species diversity analyses.~~
+1. ~~Figure of species diversity results.~~
 1. table: results of functional diversity analyses.
@@
 1. ~~Figure of the patch layout with pitfalls~~.
-1. Pictures of different species, sampling, field sites?
+1. ~~Pictures of different species, sampling, field sites?~~
 

### 2025-06-16 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
im so tired push 2

@@
 In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches.
-Our main findings emphasized: (1) Habitat type and pathc shape were the main drivng factors for dung beetle species abundances were composed, however effects were species specific.
+Our main findings emphasized: (1) Habitat type and patch shape were the main driving factors for dung beetle species abundances were composed, however effects were species specific.
 (2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level.
@@
 
-Compared to the other treatment types the greatest number of individual beetles were captured from matrix habitat.
-Likewise, for each of the most dominant species the highest counts came from the matrix.
-
-Past studies have shown similar trends where habitat type and forest regeneration stages are key in the partitioning of dung beetle assemblages[@conover_phenology_2019; @WOS:000386714700011; @arellano_response_2008].
-This is further supported by the fact that capture rates were lowest in the rectangular patch, the treatment with the least edge to area ratio, suggesting that dung beetles in the matrix may act as a source population which feeds into open patches through habitat edge.
+Dung beetle abundances were significantly lowest in rectangular patches.
+Matrix patches had the most total beetles collected and consistently had the highest counts for the most dominant species.
+However abundances were not significantly different between matrix, connected, and winged patches.
+This may suggest that isolation is leading to a decrease in abundances, but it is more likely that populations in the matrix are acting as a source, and since connected and winged patches have higher edge to area ratios dung beetles were more likely to move into those patches.
+Past studies have shown trends where habitat type and forest regeneration stages are key in the partitioning of dung beetle assemblages[@conover_phenology_2019; @WOS:000386714700011; @arellano_response_2008] which supports why we might see these differences in local populations.
+In addition distance from habitat edge has been attributed to differences in spillover from source populations[@WOS:000707546700001], emphasizing the importance of habitat edge in the context of our study.
 
@@
 Yet in comparing raw species counts the sampling plot with the most total species captured was in the matrix while the fewest species sampled were in a rectangular patch.
+
 _here look at other srs papers and see whats going on with edges_.
@@
 We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site.
-We also thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry.
+I also wanted to specificallly thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry.
 

### 2025-06-17 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
writing numba 3 p1

@@
 
-We did not detect any patterns of species richness between patch types in our modeling.
-Yet in comparing raw species counts the sampling plot with the most total species captured was in the matrix while the fewest species sampled were in a rectangular patch.
+While abundances were different between patch types, we did not detect any patterns of species richness in our modeling.
+Total species counts were very even across patch types and sampling blocks.
+This difference indicates that even though habitat type or landscape features are informing occupational preferences, land use is more or less the same across the study site.
+This differs from past work where species richness is lower in forest fragments [@WOS:000178929600002]
 

### 2025-06-17 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
pushy wushy

@@
 
+We also observed that patch effects were not equally proportional for all species.
+The abundances of all dominant species were positively influenced by matrix habitat, but _Aphoidius alloblackburneus_ responded more positively than any other species.
+_Phanaeus igneus_ and _Ateuchus lecontei_
+
 ## Richness
@@
 Total species counts were very even across patch types and sampling blocks.
-This difference indicates that even though habitat type or landscape features are informing occupational preferences, land use is more or less the same across the study site.
+This difference indicates that even though habitat type or landscape features are informing occupational preferences, land use is more or less the same between patches and matrix.
 This differs from past work where species richness is lower in forest fragments [@WOS:000178929600002]
@@
 Likewise, the Shannon index values insisted that rare species throughout the study population were even.
+However, for both metrics, values varied greatly by sampling blocks.
+Similar to our richness results this suggests that variation between patch and matrix is not distinct enough to limit the land use of dung beetles within our study site.
+
+In contrast to what we observed with species richness it appears that assemblages larger landscape patterns are effecting the composition of assemblages across the study site.
 

### 2025-06-17 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
discisuition

@@
 In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix $\beta$ = –0.59, p < 0.001).
-_Ateuchus lecontei_ and _Phanaeus igneus_ showed increased abundance in rectangle patches relative to expectations ($\beta$ = 0.40 and 0.47, respectively; p < 0.05).
+_Ateuchus lecontei_ and _Phanaeus igneus_ showed increased effect on abundance in rectangle patches relative to expectations ($\beta$ = 0.40 and 0.47, respectively; p < 0.05).
 Random effects indicated moderate variability among blocks (standard deviation of intercept = 0.51).
@@
 The abundances of all dominant species were positively influenced by matrix habitat, but _Aphoidius alloblackburneus_ responded more positively than any other species.
-_Phanaeus igneus_ and _Ateuchus lecontei_
+_Phanaeus igneus_ and _Ateuchus lecontei_ also had higher positive effects in rectangle patches.
+.
 
@@
 
+ notes that are important for discussion
+
+ - habitat preference and seasonality (conover 2019, nealis 1977)
+ - habitat again but  only part of variation form habitat specifically aphodius also phenology(roslin 2001)
+ - spatial scales (roslin 2000)
+ - morphology (ospina 2018)
+ - functionality especially gas emissions (slade 2016)
+ - species id and funcitonality (slade 2017)
+ - faovored species dominating in fragments (resasco 2014)
+ - wind direction in corridors (damschen 2014)
+
  revisit bray curtis
@@
 We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site.
-I also wanted to specificallly thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry.
+I also wanted to specifically thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry.
 

### 2025-06-17 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
small change

@@
 
-When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches(GLMM, P < 0.001, Table 4).
+When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches($\beta$ = -0.402, P < 0.001, Table 4).
 

### 2025-06-17 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
update for final push

@@
 # # - \fancyfoot[R]{ }
-# - \usepackage{setspace}{singlespace} # manipulate line spacing 
+ - \usepackage{setspace}
+ # - doublespacing # manipulate line spacing 
 # - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
@@
 library(kableExtra)
-library(grateful)
+# library(grateful)
 library(knitr)
@@
 # devtools::install_github("crsh/papaja")
-cite_packages(citation.style = "peerj", out.dir = ".")
+# cite_packages(citation.style = "peerj", out.dir = ".")
 ```
@@
 
-# Abstract
+# TO DO LIST
+
+1) **Results:** Did you get any species not captured in other long-leaf/SE USA forest sites?
+
+2) **Methods:** Find me the best example paragraph you think you read describing the use of hill numbers to estimate diversity/richness (field survey paper)
+
+3) **Methods:** Find me the best TWO descriptions of using GLMM to compare abundance/diversity/richness at the Corridor Project
+
+4) **Methods:** Write a short paragraph on how you calculated and compared functional diversity (how you measured it, how you compared it)
+
+5) **Figure: ** biomass of different (R/T/D) functional groups.
+
+6) **Table:** Results of functional diversity analyses.
+
+\newpage
+
+YOUR TITLE GOES HERE 
+IN UPPER CASE SINGLE SPACE
+
+By
+
+YOUR NAME GOES HERE IN ALL CAPITAL LETTERS
+
+A THESIS PRESENTED TO THE GRADUATE SCHOOL OF THE UNIVERSITY OF FLORIDA IN PARTIAL FULFILLMENT OF THE REQUIREMENTS FOR THE DEGREE OF MASTER OF SCIENCE
+
+UNIVERSITY OF FLORIDA
+
+2025
+\newpage
+
+© 2025 Your Name Goes Here in Title Case
+
+\newpage
+
+_Your dedication is typed here. It should begin with the word “To.”
+(To my Mom is a typical dedication) If your dedication is longer than a single line it should be single-spaced, and centered vertically and horizontally_
+
+To my family who never stopped supporting me along this journey, my friends who kept me company along the way, and my mentors at VCU who believed in me before I did myself.
+
+\newpage
+
+ACKNOWLEDGMENTS
+_Acknowledgments must be written in complete sentences. Do not use direct address. For example, instead of Thanks, Mom and Dad!, you should say I thank my parents._
+
+We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site.
+I also wanted to specifically thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry.
+
+\newpage
+
+LIST OF ABBREVIATIONS
+
+A word to be defined: Write the definition here.
+
+Another word:	Another def.
+
+Another word:	Another def.
+
+SRS:	Savannah River Site.
+
+\newpage
+
+Abstract of Thesis Presented to the Graduate School  
+of the University of Florida in Partial Fulfillment of the  
+Requirements for the Degree of Master of Science.
+
+TITLE OF THE WORK, CENTERED, SINGLE-SPACED,
+IN ALL CAPITAL LETTERS, EXACTLY AS ON TITLE PAGE
+
+By
+
+Your Name as on the Title Page, but in Title Case.
+
+Month and year of graduation
+
+Chair: Name (Do not put Dr. before the name, and do not put degrees after it)  
+Cochair: Name (if any, otherwise delete this line)  
+Major: See your Editorial Document Management record for your exact major
 
@@
 
-# TO DO LIST
-
-Next assignment:
-
-~~1) Are any of the Dung Beetles we collected considered "open habitat" specialists? (might want to look at the work on pastures in the southeast)~~
-
-- cvig - open sand chaparral, open grassland preference nealis 1977
-- mbis - open sand chapparal preference nealis 1977
-- open - slight preference towards open chap and shaded chap, little open grassland nealis 1977
-
-- pign - sharing forest habitat but need more info conover 2019
-- pvin - open habitat but share forest with pign conover 2019
-- alec - forest preference conover 2019
-
-~~2) Is there anything about these species that might be interesting in the context of corridors? Are any species known to be good fliers/movers? Specialists on deer dung? navigate by starlight?~~
-
-  - Ospina 2018: wing shape py habitat preference and differences in wing shape determined by species groupings.
-Beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies).
-Large body beetles tending towards lower energy flight strategies.
-maybe justification for corridor preference?
-    - stanbrook and king 2022: tunnelers preferring open habitat and tunnelers also contributing more towards dung removal. 
-    - Conover 2019: dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types 
-    - Gimenez Gomez et al 2021 -> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used.
-However this study was done in a more tropical ecosystem so its hard to say if the same would apply to our more temperate system so more work is needed in this specific avenue.
-
-3) Did you get any species not captured in other long-leaf/SE USA forest sites?
-
-~~4) What are hill numbers?~~
-
-- hill numbers are a set of biodiversity metrics designed to standardize metrics and quantify biodiversity and richness differently depending on how we want to emphasize abundance while also accounting for differences in sampling sizes and efforts (read more about this Eric)
-
-~~5) What is the advantage of useing hill numbers to estimate diversity / richness?~~
-
-- Chao et al.
-2014: incorporating abundance into calculating diversity metrics is important for studying functionality in particular since some species that may not be as abundant as others may not contribute to an ecosystem function as much as others.
-At the same time rare species may be filling a specific niche or we might be concerned about exotics or at risk species.
-
-6) Are the diversity and richness estimates calculated with hill numbers sensitive to anything, e.g., rare species/abundance, etc:
-
-7) Find me the best example paragraph you think you read describing the use of hill numbers to estimate diversity/richness (field survey paper): _________________
-
-8) Find me the best TWO description of using GLMM to compare abundance/diversity/richness at the Corridor Project: _________________
-
-### Methods:
-
-1. ~~add a few sentences on drying, keys used to ID, and depositing vouchers~~
-
-### Analyses
-
-1. ~~Write a short paragraph on how you calculated and compared species richness and diversity (hill numbers, glmm).~~
-
-1. Write a short paragraph on how you calculated and compared functional diversity (how you measured it, how you compared it)
-
-### Results
-
-1. answer the overview questions, them stich them together into a paragraph
-
- - ~~any species restricted to 1-2 habitat types~~
- - ~~Were all species found in all blocks? Were any restricted to only 1-2 blocks?~~
- - ~~Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)~~
- - Number of species in each functional group
- - ~~most common 3-4 species~~
- - any rare species? 
- - any invasive / exotic species?
-
-### Figures
-
-1. ~~Figure of species richness results.~~
-1. Figure of species diversity results.
-1. Figure of biomass of different (R/T/D) functional groups.
-1. ~~Figure of abundance of individual species.~~
-1. ~~Figure of the patch layout with pitfalls.~~
-
-### Tables
-
-1. ~~Table: count by species by habitat.~~
-1. table: results of species richness analyses.
-1. table: results of species diversity analyses.
-1. table: results of functional diversity analyses.
-
-### Misc.
-
-1. Map of where field site is located.
-
-1. Pictures of different species, sampling, field sites
-
-\newpage
-
-# INTRODUCTION
+\newpage
+
+\doublespacing
+
+## CHAPTER 1
+
+## TITLE OF CHAPTER ONE
+
+## INTRODUCTION
 
@@
 
-**Paragraph 5:** Here is what we did to address this unknown. _The hypotheses we tested_
-
-\newpage
-
-# Methods
-
-## Study site
+## Methods
+
+### Study site
 
@@
 
-## Dung beetle sampling
+### Dung beetle sampling
 
@@
 
-## Analyses
+### Analyses
 
@@
 
-# RESULTS
+## RESULTS
 
@@
 
-## Overview: Community Composition and Abundance
+### Overview: Community Composition and Abundance
 
@@
 
-\newpage
-
-# DISCUSSION
-
-## Recap main goals into findings
+## DISCUSSION
+
+_Don't repeat results...tell us:
+
+1) what your OVERALL conclusions are, 
+2) HOW your results compare to those of other studies, 
+3) WHY they are similar and / or different
+4) BIOLOGICAL INSIGHTS you might have gained about partiicular species
+5) any CAVEATS to your results or shortcomings of your study
+6) the NEXT STEPS you recommend be taken by researchers interested in this toopic or system
 
@@
 
-## Abundance
+### Abundance
 
@@
 _Phanaeus igneus_ and _Ateuchus lecontei_ also had higher positive effects in rectangle patches.
-.
-
-## Richness
+
+### Richness
 
@@
 
-##Diversity
+### Diversity
 
@@
  - habitat preference and seasonality (conover 2019, nealis 1977)
- - habitat again but  only part of variation form habitat specifically aphodius also phenology(roslin 2001)
+
+> - cvig - open sand chaparral, open grassland preference nealis 1977
+> - mbis - open sand chapparal preference nealis 1977
+> - open - slight preference towards open chap and shaded chap, little open grassland nealis 1977
+
+> - pign - sharing forest habitat but need more info conover 2019
+> - pvin - open habitat but share forest with pign conover 2019
+> - alec - forest preference conover 2019
+
+ - habitat again but  only part of variation form habitat specifically aphodius also phenology (roslin 2001)
  - spatial scales (roslin 2000)
- - morphology (ospina 2018)
+ - morphology
+
+> - Ospina 2018: wing shape py habitat preference and differences in wing shape determined by species groupings.
+Beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies).
+Large body beetles tending towards lower energy flight strategies.
+maybe justification for corridor preference?
+> - stanbrook and king 2022: tunnelers preferring open habitat and tunnelers also contributing more towards dung removal. 
+> - Conover 2019: dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types 
+> - Gimenez Gomez et al 2021 -> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used.
+However this study was done in a more tropical ecosystem so its hard to say if the same would apply to our more temperate system so more work is needed in this specific avenue.
+
  - functionality especially gas emissions (slade 2016)
@@
 
-## Interpretation of results
+---
+
+Discussion Outline & Notes
 
@@
 
-# ACKNOWLEDGMENTS
-
-We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site.
-I also wanted to specifically thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry.
-
-\newpage
-
-# REFERENCES
+\newpage
+
+## REFERENCES
 
@@
 \newpage
-# FIGURES & TABLES
-
-1. ~~Table: count by species by habitat.~~
-1. ~~table: results of species richness analyses.~~
-1. ~~Figure of species richness results.~~
-1. ~~table: results of species diversity analyses.~~
-1. ~~Figure of species diversity results.~~
-1. table: results of functional diversity analyses.
-1. Figure of biomass of different (R/T/D) functional groups.
-1. ~~Map of where field site is located~~
-1. ~~Figure of the patch layout with pitfalls~~.
-1. ~~Pictures of different species, sampling, field sites?~~
+## FIGURES & TABLES
 
@@
 
-bray<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/bray_patches.csv") %>% 
-  kable(
+# bray<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/bray_patches.csv") 
+bray<-read_csv(here("corridor_docs","eric_ms_thesis","tables","bray_patches.csv")) %>% kable(
       digits = 3,
@@
 ```{r Table aovMO, echo=FALSE,message = FALSE,warning=FALSE}
-
-aovM0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM0.csv") %>% 
+aovM0<-read_csv(here("corridor_docs","eric_ms_thesis","tables","aovM0.csv")) %>%
+# aovM0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM0.csv") %>% 
   kable(
@@
 ```{r Table MO, echo=FALSE,message = FALSE,warning=FALSE}
-
-M0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M0.csv") %>% 
+M0<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M0.csv")) %>%
+# M0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M0.csv") %>% 
   kable(
@@
 ```{r Table aovMO3, echo=FALSE,message = FALSE,warning=FALSE}
-
-aovM3<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM3.csv") %>% 
+aovM3<-read_csv(here("corridor_docs","eric_ms_thesis","tables","aovM3.csv")) %>%
+# aovM3<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM3.csv") %>% 
   kable(
@@
 ```{r Table M3, echo=FALSE,message = FALSE,warning=FALSE}
-
-M3<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M3.csv") %>% 
+M3<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M3.csv")) %>%
+# M3<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M3.csv") %>% 
   kable(
@@
 ```{r Table M_rich, echo=FALSE,message = FALSE,warning=FALSE}
-
-M_rich<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_rich.csv") %>% 
+M_rich<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M_rich.csv")) %>%
+# M_rich<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_rich.csv") %>% 
   kable(
@@
 ```{r Table M_shan, echo=FALSE,message = FALSE,warning=FALSE}
-
-M_shan<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_shan.csv") %>% 
+M_shan<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M_shan.csv")) %>%
+# M_shan<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_shan.csv") %>% 
   kable(
@@
 ```{r Table M_simp, echo=FALSE,message = FALSE,warning=FALSE}
-
-M_simp<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_simp.csv") %>% 
+M_simp<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M_simp.csv")) %>%
+# M_simp<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_simp.csv") %>% 
   kable(
@@
 
-# OTHER REQUIRED TEXT
-
-## Dedication 
-To my family who never stopped supporting me along this journey, my friends who kept me company along the way, and my mentors at VCU who believed in me before I did myself.
-
-## List of Abbreviations
-
-1. SRS:	Savannah River Site. 
-1. Another word:	And the list continues with another definition.
-
-## Biographical Sketch
+# BIOGRAPHICAL SKETCH
+
+_A biographical sketch is required of all candidates.
+The biographical sketch should be in narrative form.
+Third person, past tense, it typically includes the educational background of the candidate.
+The author should have replaced this paragraph with their own._
 
@@
 The removal, breakdown, and burial of animal feces drive important ecostystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
-Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) [@WOS:000891747700001].
+Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) **(citation)**.
 Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].

### 2025-06-17 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -2,6 +2,10 @@ library(tidyverse)
 library(here)
 library(kableExtra)
 library(sjPlot)
+
+# READ THIS ABOUT GLMM REFERENCE LEVELS
+# https://stats.stackexchange.com/questions/628348/when-i-change-my-reference-level-on-my-glmer-in-r-why-do-the-p-values-change-an
+
 # load and prep data ------------------------------------------------------
 
 
@@ -239,12 +243,12 @@ aovM3 <- Anova(M3)
 plot_model(M3, type = "pred", terms = c("sp_code", "patch_type"))
 
 
-
-?interaction.plot
-
-interaction.plot(M3)
-
-?plot_model
+# 
+# ?interaction.plot
+# 
+# interaction.plot(M3)
+# 
+# ?plot_model
 
 # Eric trying table
 
@@ -254,14 +258,14 @@ write_csv(tidyM3,"./corridor_docs/eric_ms_thesis/tables/m3.csv")
 write_csv(aovM3,"./corridor_docs/eric_ms_thesis/tables/aovM3.csv")
 
 
-
-# Global model: Take 1
-M1 <- glmer(n ~ patch_type * sp_code + (1 + patch_type * sp_code | block), 
-            data = spp_abund, 
-            family = poisson)
-summary(M1)
-
-Anova(M1)
+# 
+# # Global model: Take 1
+# M1 <- glmer(n ~ patch_type * sp_code + (1 + patch_type * sp_code | block), 
+#             data = spp_abund, 
+#             family = poisson)
+# summary(M1)
+# 
+# Anova(M1)
 
 # Model for richness 
 M_rich <- glmer(h_rich ~ patch_type + (1 + patch_type | block),
@@ -292,7 +296,7 @@ write_csv(tidy_model,"./corridor_docs/eric_ms_thesis/tables/M_rich.csv")
 # Model for shannons 
 h_shannon
 
-M_shannon
+
 
 M_shannon <- lmer(h_shannon ~ patch_type + (1 | block),
                   data = h_shannon)

### 2025-06-17 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
beuroucracyhowdoyouspellthis

@@
 
-YOUR TITLE GOES HERE 
-IN UPPER CASE SINGLE SPACE
+THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES
 
@@
 
-YOUR NAME GOES HERE IN ALL CAPITAL LETTERS
+ERIC ESCOBAR-CHENA
 
@@
 
-© 2025 Your Name Goes Here in Title Case
+© 2025 Eric Escobar-Chena
 
@@
 
-TITLE OF THE WORK, CENTERED, SINGLE-SPACED,
-IN ALL CAPITAL LETTERS, EXACTLY AS ON TITLE PAGE
+THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES
 
@@
 
-Your Name as on the Title Page, but in Title Case.
-
-Month and year of graduation
-
-Chair: Name (Do not put Dr. before the name, and do not put degrees after it)  
-Cochair: Name (if any, otherwise delete this line)  
-Major: See your Editorial Document Management record for your exact major
+Eric Escobar-Chena
+
+August 2025
+
+Chair: Emilio Bruna  
+Major: Wildlife Ecology and Conservation
 

### 2025-06-18 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
added notes for eric

--- 2025-06-17 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
@@
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy  [@forman_general_1995].
-Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
-
-By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes [@fletcher_jr_landscape_2023; @resasco_contribution_2017].
-Past studies have measured changes in biodiversity for many different taxa  [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected.
-Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services [@hevia_trait-based_2017].
+[**EB edit:** break this next sentence into 3 to carry reader though the logic: changes in species diverstiy are important to understand because species are involved in interactions.
+If we gain or lose species we could gain or lose interactions and resulting ecosystems processes.
+there is some evidence of this, but...
+look at the green text notes below] Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
 
@@
 
-<!-- Potential alternative routes:  -->
+By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes [@fletcher_jr_landscape_2023; @resasco_contribution_2017].
+Past studies have measured changes in biodiversity for many different taxa  [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected.
+Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services [@hevia_trait-based_2017].
+[**EB edit:** these last sentences don't really make the case for experiments, whcih were the initial emphasis of the paragraph.
+Might want to opt for another option, like the ones in the green notes below]
 
@@
 The matrix could be a source of diversity that moves into patches and then through corridors or a completely different subset of species.
-(any other options?) Understanding whichh of these it is is important because it leads to differnt conclusions abiout role of corridors and conservation strategies -->
-
-<!-- 3) regardless, not a bad idea to discuss both matrix and functional group in diuscussion. -->
-
-<!-- 4) and the alternative approach: revamp into around spillover from matrix into patches and if depednent on corridors or edge  -->
-
-<!-- Paragraph 4: Why we don't know it. _(hint: ecosystem services)_   -->
-
-<!-- >Topic Sentence: Is this where we talk about challenges in large scale experimental design and the corridor project or is this where we talk about ecosystem services and biodiversity?  -->
+(any other options?) Understanding whichh of these it is is important because it leads to differnt conclusions abiout role of corridors and conservation strategies.
+can also be spillover from matrix into patches and if depednent on corridors or edge -->
 
@@
 Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources [@hanski_dung_1991], interact with corridors in their landscapes.
+
 We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch area to edge ratio and distance to edge [@tewksbury_corridors_2002].
-To ask the questions: (1)
-How are species abundances of dung beetles distributed within isolated and connected patches and what are their relationships with assemblages in matrix habitat?
+To ask the questions: (1) How are species abundances of dung beetles distributed within isolated and connected patches and what are their relationships with assemblages in matrix habitat?
 (2) Does species richness vary when movement between patches is facilitated by movement corridors?
@@
 
-Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al.
-2024).
-The removal, breakdown, and burial of animal feces drive important ecostystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
-Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) **(citation)**.
-Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].
-Past studies have also focused on how landscape structure alters the community compositions of dung beetles[@WOS:000398662200001], yet large landscape scale experimental studies with carefully controlled and replicated treatments are non-existent for this model species.
-
 Here, we  aim to determine how connectivity and fragmentation affect Species Richness and Diversity, Abundance, and functional diversity.

### 2025-06-18 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
update for eric with some edits and assignment for tomorrow

@@
 bibliography: ["eric-thesis.bib"]
-output:
-  pdf_document:
-    includes:
-      # in_header: style/preamble.tex
-      template: null
-      keep_tex: no
-      extra_dependencies: ["float"]
+output: bookdown::pdf_document2
+  # pdf_document:
+  #   includes:
+  #     # in_header: style/preamble.tex
+  #     template: null
+  #     keep_tex: no
+  #     extra_dependencies: ["float"]
       # number_sections: true
@@
 
+FOR TOMORROW:
+
+1) Finish adding citations
+
+2) Stat results: biology first, then stats (e.g., abundance ranged form xxx in rectangular to xxx in winged)
+
+3) Review other papers to find:
+
+- those with similar results
+- those with contrary results
+- do this with dung bettles AND other insect taxa
+
+DOWN THE ROAD
+
 1) **Results:** Did you get any species not captured in other long-leaf/SE USA forest sites?
@@
 2025
-\newpage
+
+-----
 
@@
 
-\newpage
-
-_Your dedication is typed here. It should begin with the word “To.”
-(To my Mom is a typical dedication) If your dedication is longer than a single line it should be single-spaced, and centered vertically and horizontally_
+-----
+
+<!-- _Your dedication is typed here.
+It should begin with the word “To.” (To my Mom is a typical dedication) If your dedication is longer than a single line it should be single-spaced, and centered vertically and horizontally_ -->
 
@@
 
-\newpage
-
-ACKNOWLEDGMENTS
-_Acknowledgments must be written in complete sentences. Do not use direct address. For example, instead of Thanks, Mom and Dad!, you should say I thank my parents._
+----
+
+**ACKNOWLEDGMENTS**
+<!-- _Acknowledgments must be written in complete sentences. Do not use direct address. For example, instead of Thanks, Mom and Dad!, you should say I thank my parents._  -->
 
@@
 
-\newpage
-
-LIST OF ABBREVIATIONS
-
-A word to be defined: Write the definition here.
-
-Another word:	Another def.
-
-Another word:	Another def.
+----
+
+**LIST OF ABBREVIATIONS**
 
@@
 
-\newpage
+----
 
@@
 
-## CHAPTER 1
-
-## TITLE OF CHAPTER ONE
+**CHAPTER 1**
+
+**Influence of Connectivity on Dung Beetle Communities**
 
@@
 Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions [@hasan_dung_2024].
-The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]
+The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
+[**EB edit:** I added the following to justify dung veetles more, and then made the questions a second paragraph] Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) **(citation)**.
+Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].
+Past studies have also focused on how landscape structure alters the community compositions of dung beetles[@WOS:000398662200001], yet large landscape scale experimental studies with carefully controlled and replicated treatments are non-existent for this model species.
+
 Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources [@hanski_dung_1991], interact with corridors in their landscapes.
-
 We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch area to edge ratio and distance to edge [@tewksbury_corridors_2002].
@@
 
-Our study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, US (33.208$^\circ$ N, 81.408$^\circ$ W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals [@tewksbury_corridors_2002].
-Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna.
+Our study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, USA (33.208$^\circ$ N, 81.408$^\circ$ W, Figure \@ref(fig:srs-map)).
+in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals [@tewksbury_corridors_2002].
+Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna (Figure \@ref(fig:patches)).
 In each block the central patch (100 $\times$ 100 m) is always connected to one peripheral patch with identical dimensions by a 150 $\times$ 25 m corridor, this will hereafter be referred  to as the connected patch.
@@
 
-In the months of July and August 2024 dung beetles were sampled in 4 blocks spread across SRS, baited pitfall traps were placed in one of each patch type and in one matrix plot per block.
+In the months of July and August 2024 dung beetles were sampled in 4 blocks spread across SRS, baited pitfall traps were placed in one of each patch type and in one matrix plot per block (Figure \@ref(fig:pitfalls)).
 Traps were placed in groups of 3 in the centers of each patch approximately 250 meters from the midpoint of the central patch 40 m from patch edge.
@@
 
-### Overview: Community Composition and Abundance
-
 Overall, I collected N = `r total_N` dung beetles belonging to N = `r total_spp` species.
@@
 
+### Beetle Abundance
+
 When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches($\beta$ = -0.402, P < 0.001, Table 4).
@@
 
-### Species Richness
+### Beetle Richness and Diversity
 
@@
 
-### Species Diversity
-
-We used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation.
+[**EB:** edit: shouldn't this go in the methods?] We used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation.
 In both models, the reference level for patch type was Connected.
@@
 
-_Don't repeat results...tell us:
+_Don't repeat results...tell us:_
 
@@
 This differs from past work where species richness is lower in forest fragments [@WOS:000178929600002]
-
-_here look at other srs papers and see whats going on with edges_.
 
@@
 
+```{r, label = patches, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Experimental block with the different patches and distance between them.", out.width="50%", fig.align='center'}
+
+knitr::include_graphics("images/patches-pic.png")
+
+```
+
 ```{r, label = pitfalls, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Experimental block with the different patches and distance between them.", out.width="50%", fig.align='center'}
@@
 knitr::include_graphics("images/pitfalls.png")
-
-```
-
-```{r, label = patches, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Experimental block with the different patches and distance between them.", out.width="50%", fig.align='center'}
-
-knitr::include_graphics("images/patches-pic.png")
 
@@
 \newpage
-```{r Table1, echo=FALSE,message = FALSE,warning=FALSE}
+```{r summary-table, echo=FALSE,message = FALSE,warning=FALSE}
 # https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
@@
 
-```{r Table2, echo=FALSE,message = FALSE,warning=FALSE}
+```{r abundance-patch, echo=FALSE,message = FALSE,warning=FALSE}
 # https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
@@
 
-```{r Table 3, echo=FALSE,message = FALSE,warning=FALSE}
+\newpage
+
+```{r bray-curtis, echo=FALSE,message = FALSE,warning=FALSE}
 btl_spec_patch<- btl_data |>
@@
 # taxa_bray
-
-```
-
-```{r Table bray_patches, echo=FALSE,message = FALSE,warning=FALSE}
 
@@
 
-```{r Table aovMO, echo=FALSE,message = FALSE,warning=FALSE}
+```{r Table-aovMO, echo=FALSE,message = FALSE,warning=FALSE}
 aovM0<-read_csv(here("corridor_docs","eric_ms_thesis","tables","aovM0.csv")) %>%
+  mutate(term=case_when(
+    term == "patch_type" ~ "Patch type",
+    .default = as.character(term))) |>
 # aovM0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM0.csv") %>% 
@@
 
-```{r Table MO, echo=FALSE,message = FALSE,warning=FALSE}
+\newpage
+
+```{r Table-MO, echo=FALSE,message = FALSE,warning=FALSE}
 M0<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M0.csv")) %>%
+  replace_na(list(group=" ")) %>% 
+  # need to round numbers above to 4, convert to character, then na_replace
+  # replace_na(list(group=" ", std.error="",statistic=" ",p.value = "")) %>% 
 # M0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M0.csv") %>% 
@@
 
-```{r Table aovMO3, echo=FALSE,message = FALSE,warning=FALSE}
+\newpage
+
+```{r Table-aovMO3, echo=FALSE,message = FALSE,warning=FALSE}
 aovM3<-read_csv(here("corridor_docs","eric_ms_thesis","tables","aovM3.csv")) %>%
+    mutate(term=case_when(
+    term == "sp_code" ~ "Species",
+    term == "patch_type" ~ "Patch type",
+    term == "sp_code:patch_type" ~ "Species x Patch type",
+    .default = as.character(term))) |>
 # aovM3<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM3.csv") %>% 
@@
 
-```{r Table M3, echo=FALSE,message = FALSE,warning=FALSE}
+\newpage
+
+```{r Table-M3, echo=FALSE,message = FALSE,warning=FALSE}
 M3<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M3.csv")) %>%
+  replace_na(list(group=" ")) %>% 
+  # need to round numbers above to 4, convert to character, then na_replace
+  # replace_na(list(group=" ", std.error="",statistic=" ",p.value = "")) %>% 
 # M3<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M3.csv") %>% 
@@
 
-```{r Table M_rich, echo=FALSE,message = FALSE,warning=FALSE}
+\newpage
+
+```{r Table-M-rich, echo=FALSE,message = FALSE,warning=FALSE}
 M_rich<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M_rich.csv")) %>%
+   mutate(term=case_when(
+    term == "patch_typeMatrix" ~ "Patch Type: Matrix",
+    term == "patch_typeRectangle" ~ "Patch Type: Rectangle",
+    term == "patch_typeWinged" ~ "Patch Type: Winged",
+    .default = as.character(term))) |>
 # M_rich<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_rich.csv") %>% 
@@
 
-```{r Table M_shan, echo=FALSE,message = FALSE,warning=FALSE}
+\newpage
+
+```{r Table-M-shan, echo=FALSE,message = FALSE,warning=FALSE}
 M_shan<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M_shan.csv")) %>%
+  replace_na(list(group=" ")) %>% 
+  # need to round numbers above to 4, convert to character, then na_replace
+  # replace_na(list(group=" ", std.error="",statistic=" ",p.value = "")) %>% 
 # M_shan<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_shan.csv") %>% 
@@
 
-```{r Table M_simp, echo=FALSE,message = FALSE,warning=FALSE}
+\newpage
+
+```{r Table-M-simp, echo=FALSE,message = FALSE,warning=FALSE}
 M_simp<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M_simp.csv")) %>%
+  replace_na(list(group=" ")) %>% 
+  # need to round numbers above to 4, convert to character, then na_replace
+  # replace_na(list(group=" ", std.error="",statistic=" ",p.value = "")) %>% 
 # M_simp<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_simp.csv") %>% 
@@
 
-_A biographical sketch is required of all candidates.
+<!-- _A biographical sketch is required of all candidates.
 The biographical sketch should be in narrative form.
 Third person, past tense, it typically includes the educational background of the candidate.
-The author should have replaced this paragraph with their own._
+The author should have replaced this paragraph with their own._ -->
 

### 2025-06-20 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
citations and bib

@@
 
-As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented **(Diaz 2019)**.
+As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented [@diaz_pervasive_2019; @ma_global_2023; @taubert_global_2018].
 Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003; @fletcher_is_2018].
-However, as habitats are broken down community structures are significantly altered **[@WOS:000419965700012]**.
+However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012; @haddad_corridor_2003; @jennings_composition_2006].
 This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
@@
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy  [@forman_general_1995].
+
 [**EB edit:** break this next sentence into 3 to carry reader though the logic: changes in species diverstiy are important to understand because species are involved in interactions.
@@
 there is some evidence of this, but...
-look at the green text notes below] Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
+look at the green text notes below]
+
+Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
 
@@
 To test for the effects of connectivity on  abundance, species richness, and species diversity we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix.
-For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution  **(bolker citation?)**.
+For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution [@bolker_generalized_2009].
 Compared (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type.
 We included the identity of the sampling block as a random effects.
-To model our diversity metrics we took a similar approach, but this time using GLMMs with a Gaussian distribution  **(citation)**.
+To model our diversity metrics we took a similar approach, but this time using GLMMs with a Gaussian distribution [@chao_rarefaction_2014].
 All models were fit using `lme4` package  [@bates_fitting_2015].

### 2025-06-20 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
formatting

@@
 # date: "\\vspace{-1em}"
+# header-includes:
+ # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
+#  - \usepackage{lineno}
+#  - \linenumbers
+#  - \usepackage[default]{sourcesanspro}
+#  - \usepackage{fancyhdr} # modify header into left, center, right
+#  - \pagestyle{fancy} # cues the use of fancy headers
+#  - \fancyfoot{} # cues the use of fancy footer
+#  - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
+#  - \fancyhead[R]{p. \thepage} # name and page on header right
+#  - \usepackage{subcaption}
+#  - \usepackage{float}
+# # # - \fancyfoot[R]{ }
+#  - \usepackage{setspace}
+
 header-includes:
- # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
- - \usepackage{lineno}
- - \linenumbers
- - \usepackage[default]{sourcesanspro}
- - \usepackage{fancyhdr} # modify header into left, center, right
- - \pagestyle{fancy} # cues the use of fancy headers
- - \fancyfoot{} # cues the use of fancy footer
- - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
- - \fancyhead[R]{p. \thepage} # name and page on header right
- - \usepackage{subcaption}
- - \usepackage{float}
-# # - \fancyfoot[R]{ }
+  - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
+  - \usepackage{float} #to keep figure in the location placed instead of end
+  - \floatplacement{figure}{H}
+  - \usepackage{booktabs}
+  - \raggedbottom
+  - \usepackage{fancyhdr}
+  - \pagestyle{plain}
+  - \fancyfoot[C]{\thepage}
+  - \fancyhead[L]{}
+  - \fancyhead[R]{\small{EEC MS Thesis Draft}}
+  - \thispagestyle{fancy}
+  - \usepackage{tabu}
+  - \usepackage{makecell}
+  - \usepackage{pdflscape}
+  - \usepackage{longtable}
+  - \newcommand{\blandscape}{\begin{landscape}}
+  - \newcommand{\elandscape}{\end{landscape}}
+  - \setlength\parindent{22pt}
+  - \usepackage{indentfirst}
+  - \renewenvironment{abstract}{}{} #suppress the abstract
   - \usepackage{setspace}
+
  # - doublespacing # manipulate line spacing 
@@
 
-5) **Figure: ** biomass of different (R/T/D) functional groups.
+5)  **Figure:** biomass of different (R/T/D) functional groups.
 
@@
 
+\vspace{8cm}
+
+\begin{center}
 THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES
-
+\end{center}
+
+\vspace{6cm}
+
+\begin{center}
 By
-
+\end{center}
+
+\vspace{3cm}
+
+\begin{center}
 ERIC ESCOBAR-CHENA
-
+\end{center}
+
+\vspace{5cm}
+
+\begin{center}
 A THESIS PRESENTED TO THE GRADUATE SCHOOL OF THE UNIVERSITY OF FLORIDA IN PARTIAL FULFILLMENT OF THE REQUIREMENTS FOR THE DEGREE OF MASTER OF SCIENCE
-
-UNIVERSITY OF FLORIDA
-
+\end{center}
+
+\vspace{2cm}
+
+\begin{center}
 2025
-
------
-
+\end{center}
+
+\newpage
+
+\vspace{20cm}
+
+\begin{center}
 © 2025 Eric Escobar-Chena
-
------
-
+\end{center}
+
+\newpage
+
+\vspace{15cm} 
 <!-- _Your dedication is typed here.
@@
 
+\begin{center}
 To my family who never stopped supporting me along this journey, my friends who kept me company along the way, and my mentors at VCU who believed in me before I did myself.
-
-----
-
-**ACKNOWLEDGMENTS**
+\end{center}
+
+\newpage
+
+\begin{center}
+
+ACKNOWLEDGMENTS
+
+\end{center}
 <!-- _Acknowledgments must be written in complete sentences. Do not use direct address. For example, instead of Thanks, Mom and Dad!, you should say I thank my parents._  -->
+\begin{center}
 
@@
 
-----
-
-**LIST OF ABBREVIATIONS**
+\end{center}
+
+\newpage
+
+\begin{center}
+
+LIST OF ABBREVIATIONS
+
+\end{center}
 
@@
 
-----
+\newpage
+
+\begin{center}
 
 Abstract of Thesis Presented to the Graduate School
+
 of the University of Florida in Partial Fulfillment of the
+
 Requirements for the Degree of Master of Science.
+
+\end{center}
+
+\newpage
 
@@
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy [@forman_general_1995].
-
-[**EB edit:** break this next sentence into 3 to carry reader though the logic: changes in species diverstiy are important to understand because species are involved in interactions.
+*[**EB edit:** break this next sentence into 3 to carry reader though the logic: changes in species diverstiy are important to understand because species are involved in interactions.
 If we gain or lose species we could gain or lose interactions and resulting ecosystems processes.
 there is some evidence of this, but...
-look at the green text notes below]
-
-Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
+look at the green text notes below]* Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
 
 <!-- stuff that these inviduals or species do: -->
+
 <!-- 1) their interactions with other species -->
+
 <!-- - predation -->
+
 <!-- - pollination -->
+
 <!-- - seed dispersal -->
+
 <!-- 2) stuff on their own that influences ecosystems -->
+
 <!-- - ecosystem engineering (beavers, leaf-cutter ants) -->
+
 <!-- - nutrient cycling -->
+
 <!-- - decomposition -->
@@
 Hill numbers are now the preferred metric for describing community dynamics for two reasons.
-First, they are extrapolated from the same equation, manipulating a single parameter (i.e., _q_) to arrive at estimates of richness and diversity.
-Second, by  manipulating _q_ we can gain an understanding of compositional shifts otherwise obscured while using species richness [@chao_rarefaction_2014].
-We compared community composition by increasing magnitudes of diversity components (i.e., _qD_) of _0D_ (i.e., species richness), _1D_ (i.e., Shannon entropy), and _2D_ (i.e, Simpson Diversity).
+First, they are extrapolated from the same equation, manipulating a single parameter (i.e., *q*) to arrive at estimates of richness and diversity.
+Second, by manipulating *q* we can gain an understanding of compositional shifts otherwise obscured while using species richness [@chao_rarefaction_2014].
+We compared community composition by increasing magnitudes of diversity components (i.e., *qD*) of *0D* (i.e., species richness), *1D* (i.e., Shannon entropy), and *2D* (i.e, Simpson Diversity).
 <!--[can you add a sentence about how these are calculated?
@@
 All models were fit using `lme4` package [@bates_fitting_2015].
-Prior to conducting our modeling we evaluated the the suitability of our data with _qqplots_ generated with the `DARMa` package  [@hartig_dharma_2024].
+Prior to conducting our modeling we evaluated the the suitability of our data with *qqplots* generated with the `DARMa` package [@hartig_dharma_2024].
 
@@
 <!-- 1. 1nd pass: roller, tunneler, dweller -->
+
 <!-- 1. 2nd: morphological measurements like Alonso et al 2022 (see table from decastro 2023, Giménez Gómez 2025) -->
+
 <!-- 1. table: results of analyses  -->
+
 <!-- 1. visualization: biomass of different (R/T/D) functional groups?    -->
@@
 All but four species were captured in every patch type.
-_Onthophagus concinnus_ was only found in the matrix and winged patches, while _Onthophagus striatulus_ was only captured in matrix habitat and rectangular patches.
-_Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively).
+*Onthophagus concinnus* was only found in the matrix and winged patches, while *Onthophagus striatulus* was only captured in matrix habitat and rectangular patches.
+*Geotrupes blackburnii* and *Onthophagus tuberculifrons* were the only species restricted to one patch type (winged and matrix, respectively).
 All species were within their native ranges.
@@
 
-When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches($\beta$ = -0.402, P < 0.001, Table 4).
+When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches($\beta$ = -0.402, P \< 0.001, Table 4).
 
@@
 A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 7).
-The baseline abundance (intercept) corresponds to the abundance of _Aphodius alloblackburneus_ in connected patches.
-Compared to this baseline, _Canthon vigilans_, _Ateuchus lecontei_, and _Phanaeus igneus_ showed significantly higher abundance  ($\beta$ = 1.05, 0.73, 0.66 respectively; all p < 0.001), while _Onthophagus pennsylvanicus_ had lower abundance ($\beta$ = –0.80, p < 0.001).
+The baseline abundance (intercept) corresponds to the abundance of *Aphodius alloblackburneus* in connected patches.
+Compared to this baseline, *Canthon vigilans*, *Ateuchus lecontei*, and *Phanaeus igneus* showed significantly higher abundance ($\beta$ = 1.05, 0.73, 0.66 respectively; all p \< 0.001), while *Onthophagus pennsylvanicus* had lower abundance ($\beta$ = –0.80, p \< 0.001).
 
 Species-patch interactions indicated indicated species specific responses to treatments.
-In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix $\beta$ = –0.59, p < 0.001).
-_Ateuchus lecontei_ and _Phanaeus igneus_ showed increased effect on abundance in rectangle patches relative to expectations ($\beta$ = 0.40 and 0.47, respectively; p < 0.05).
+In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix $\beta$ = –0.59, p \< 0.001).
+*Ateuchus lecontei* and *Phanaeus igneus* showed increased effect on abundance in rectangle patches relative to expectations ($\beta$ = 0.40 and 0.47, respectively; p \< 0.05).
 Random effects indicated moderate variability among blocks (standard deviation of intercept = 0.51).
@@
 None of the alternative patch types showed statistically significant differences compared to Connected: Matrix ($\beta$ = –0.085, SE = 0.50, t = –0.17), Rectangle ($\beta$ = 0.15, SE = 0.50, t = 0.30), or Winged ($\beta$ = –0.34, SE = 0.50, t = –0.68).
-
 Similarly, for Simpson's Diversity, the average value in Connected patches was 4.07 (SE = 0.64, t = 6.37).
@@
 
-_Don't repeat results...tell us:_
+*Don't repeat results...tell us:*
 
@@
 We also observed that patch effects were not equally proportional for all species.
-The abundances of all dominant species were positively influenced by matrix habitat, but _Aphoidius alloblackburneus_ responded more positively than any other species.
-_Phanaeus igneus_ and _Ateuchus lecontei_ also had higher positive effects in rectangle patches.
+The abundances of all dominant species were positively influenced by matrix habitat, but *Aphoidius alloblackburneus* responded more positively than any other species.
+*Phanaeus igneus* and *Ateuchus lecontei* also had higher positive effects in rectangle patches.
 
@@
 
-_here write about low sample sizes and reasons for why dung beetle communities would be so similar across treatment types.
-Dispersal, scale (roslin), maybe again a good reason to expect matrix as a source pop._
+*here write about low sample sizes and reasons for why dung beetle communities would be so similar across treatment types.
+Dispersal, scale (roslin), maybe again a good reason to expect matrix as a source pop.*
 
@@
 > -   Conover 2019: dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types
-> - Gimenez Gomez et al 2021 -> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used.
+> -   Gimenez Gomez et al 2021 -\> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used.
 However this study was done in a more tropical ecosystem so its hard to say if the same would apply to our more temperate system so more work is needed in this specific avenue.
@@
 1.  Abundances in matrix vs connected patch and why this could be happening
+
 -   source pop to habiitat edge
@@
 4.  bray curtis hinting at similar land uses between corridor and winged patch
+
 -   like julians paper corridors benefit certain populations and more fit populations are able to make better use
 
----
+------------------------------------------------------------------------
 
@@
 1.  Abundance
+
 -   reason for highest abundances in matrix and connected patch
@@
 2.  Species Richness & Diversity
+
 -   the role of hbitat connectivity in shaping community structure or lack thereof
@@
 3.  Functional traits and ecological impacts
+
 -   did corridoors favor a functional trait
@@
 4.  Comparing to previous studies
+
 -   how are things aligning
@@
 5.  Limitations and future work
+
 -   potential confounding factors (seasonality, distance from edge, sampling methodology and temmporal variation)
-- other directions to go (dispersal -> radar, secifically measuring changes in ecosystem services)
+-   other directions to go (dispersal -\> radar, secifically measuring changes in ecosystem services)
 
 6.  Takeaways for conservation and management
+
 -   dung beetles are robust
@@
 \smallskip
+
 ::: {#refs}
@@
 \newpage
+
 ## FIGURES & TABLES
@@
 <!-- (text - Nathan Fig of plots) -->
+
 \newpage
@@
 <!-- a<-readPNG("images/pitfalls.png") -->
+
 <!-- b<-readPNG("images/srs-map.png") -->
+
 <!-- grid.arrange(rasterGrob(a),rasterGrob(b),ncol=1) -->
@@
 <!-- ```{r Figure_hill, echo=FALSE,message = FALSE,warning=FALSE} -->
+
 <!-- library(iNEXT) -->
+
 <!-- # -->
+
 <!-- # btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv")) -->
+
 <!-- # spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv")) -->
@@
 <!-- btl_sums_patch<-btl_data %>% -->
+
 <!--   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% -->
+
 <!--   group_by(species,patch,block) %>% -->
+
 <!--   mutate(patch=case_when( -->
+
 <!--     patch == "c" ~ "Corridor", -->
+
 <!--     patch == "m" ~ "Matrix", -->
+
 <!--     patch == "w" ~ "Winged", -->
+
 <!--     patch == "r" ~ "Rectangle", -->
+
 <!--     .default = as.character(patch))) %>% -->
+
 <!--   summarize(n=sum(n, na.rm=TRUE)) %>% -->
+
 <!--   arrange(patch,desc(n)) %>% -->
+
 <!--   rename(sp_code=species, -->
+
 <!--          n_patch=n) %>% -->
+
 <!--   mutate(pb=paste(patch,block,sep="-")) %>%  -->
+
 <!--   ungroup() %>%  -->
+
 <!--   select(-patch,-block) %>%  -->
+
 <!--   pivot_wider( -->
+
 <!--     names_from = pb, -->
+
 <!--     values_from = c(n_patch))  -->
@@
 <!-- div.data <- btl_sums_patch |> -->
+
 <!--   column_to_rownames(var = "sp_code") -->
@@
 <!-- z <- iNEXT(div.data, q=c(0,1,2), datatype="abundance") -->
+
 <!-- ggiNEXT(z, facet.var="Order.q", color.var="Assemblage") + -->
+
 <!--   theme_bw(base_size=18) -->
@@
 <!-- ```{r sp_richness_patches, echo = FALSE, fig.cap = "Dung beetle secies richness in three different --- and the forest matrix surrounding patches.", out.width = '100%'} -->
+
 <!-- knitr::include_graphics("./corridor_beetles/corridor_docs/eric_ms_thesis/images/sp_richness_patches.png") -->
+
 <!-- ``` -->
@@
 \newpage
+
 ```{r summary-table, echo=FALSE,message = FALSE,warning=FALSE}

### 2025-06-20 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
some writing

@@
 there is some evidence of this, but...
-look at the green text notes below]* Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
+look at the green text notes below]*
+
+The resulting changes in species composition are important to identify because
+
+Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
 

### 2025-06-20 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
formatting

--- 2025-06-20 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
@@
   - \usepackage{setspace}
+  - \setcounter{secnumdepth}{0} # removes numbering from sections and subsections
+  - \usepackage{sectsty}\allsectionsfont{\raggedright}
+  - \usepackage{sectsty}\sectionfont{\centering}{\normalfont} # center # headings
+  - \usepackage{sectsty}\subsectionfont{\centering}{\normalfont} # center ## headings
+  - \usepackage{sectsty}\subsubsectionfont{\normalfont} # center ## headings
+  # - \usepackage{sectsty}\subsubsectionfont{\itshape} # no centering but italic subsu
+  # - \usepackage{sectsty}\subsubsectionfont{\centering}{ital} # center ### headings
+  # - \usepackage{sectsty}\sectionfont{\centering\color{darkmidnightblue}}
 
@@
 
-\vspace{8cm}
+\vspace*{5cm}
 
@@
 
-\vspace{6cm}
+\vspace{5cm}
 
@@
 
-\vspace{3cm}
+\vspace{1cm}
 
@@
 \begin{center}
-A THESIS PRESENTED TO THE GRADUATE SCHOOL OF THE UNIVERSITY OF FLORIDA IN PARTIAL FULFILLMENT OF THE REQUIREMENTS FOR THE DEGREE OF MASTER OF SCIENCE
-\end{center}
-
-\vspace{2cm}
+A THESIS PRESENTED TO THE GRADUATE SCHOOL\\  
+OF THE UNIVERSITY OF FLORIDA IN PARTIAL FULFILLMENT\\ 
+OF THE REQUIREMENTS FOR THE DEGREE OF\\  
+MASTER OF SCIENCE\\
+\end{center}
+
+\begin{center}
+UNIVERSITY OF FLORIDA
+\end{center}
 
@@
 
-\vspace{20cm}
+\vspace*{11cm}
 
@@
 
-\vspace{15cm} 
+\vspace*{10cm} 
 <!-- _Your dedication is typed here.
@@
 
-ACKNOWLEDGMENTS
+\textbf{ACKNOWLEDGMENTS}
 
@@
 <!-- _Acknowledgments must be written in complete sentences. Do not use direct address. For example, instead of Thanks, Mom and Dad!, you should say I thank my parents._  -->
-\begin{center}
+\doublespacing
 
@@
 
-\end{center}
-
-\newpage
-
-\begin{center}
-
-LIST OF ABBREVIATIONS
-
-\end{center}
+\newpage
+
+\begin{center}  
+\textbf{LIST OF ABBREVIATIONS}
+\end{center}
+
+\vspace{1cm}
 
@@
 
-\begin{center}
-
-Abstract of Thesis Presented to the Graduate School
-
-of the University of Florida in Partial Fulfillment of the
-
-Requirements for the Degree of Master of Science.
-
-\end{center}
-
-\newpage
-
+\vspace*{5cm}
+
+\singlespacing
+
+\begin{center}  
+Abstract of Thesis Presented to the Graduate School\\  
+of the University of Florida in Partial Fulfillment of the\\  
+Requirements for the Degree of Master of Science.\\   
+\end{center}
+
+\vspace*{1cm}
+
+\begin{center}
 THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES
-
+\end{center}
+
+\begin{center}
 By
-
+\end{center}
+
+\begin{center}
 Eric Escobar-Chena   
-
+\end{center}
+
+\begin{center}
 August 2025   
-
-Chair: Emilio Bruna 
-Major: Wildlife Ecology and Conservation
+\end{center}
+
+\begin{flushleft}
+Chair: Emilio Bruna\\  
+Major: Wildlife Ecology and Conservation\\   
+\end{flushleft}
+
+\doublespacing
 
@@
 
+\begin{center}
+CHAPTER 1\\  
+INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES\\  
+\end{center}
+
 \doublespacing
-
-**CHAPTER 1**
-
-**Influence of Connectivity on Dung Beetle Communities**
 
@@
 
-# BIOGRAPHICAL SKETCH
+\vspace*{1cm}
+
+\begin{center}
+\textbf{BIOGRAPHICAL SKETCH}
+\end{center}
 
@@
 The author should have replaced this paragraph with their own._ -->
+
+\doublespacing
 

### 2025-06-20 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
formatting

@@
   - \usepackage{sectsty}\subsectionfont{\centering}{\normalfont} # center ## headings
-  - \usepackage{sectsty}\subsubsectionfont{\normalfont} # center ## headings
+  - \usepackage{sectsty}\subsubsectionfont{\bfseries} # center ## headings
   # - \usepackage{sectsty}\subsubsectionfont{\itshape} # no centering but italic subsu

### 2025-06-20 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
writing

@@
   - \usepackage{setspace}
-  - \setcounter{secnumdepth}{0} # removes numbering from sections and subsections
-  - \usepackage{sectsty}\allsectionsfont{\raggedright}
-  - \usepackage{sectsty}\sectionfont{\centering}{\normalfont} # center # headings
-  - \usepackage{sectsty}\subsectionfont{\centering}{\normalfont} # center ## headings
-  - \usepackage{sectsty}\subsubsectionfont{\bfseries} # center ## headings
-  # - \usepackage{sectsty}\subsubsectionfont{\itshape} # no centering but italic subsu
-  # - \usepackage{sectsty}\subsubsectionfont{\centering}{ital} # center ### headings
-  # - \usepackage{sectsty}\sectionfont{\centering\color{darkmidnightblue}}
 
@@
 
-\vspace*{5cm}
+\vspace{8cm}
 
@@
 
-\vspace{5cm}
+\vspace{6cm}
 
@@
 
-\vspace{1cm}
+\vspace{3cm}
 
@@
 \begin{center}
-A THESIS PRESENTED TO THE GRADUATE SCHOOL\\  
-OF THE UNIVERSITY OF FLORIDA IN PARTIAL FULFILLMENT\\ 
-OF THE REQUIREMENTS FOR THE DEGREE OF\\  
-MASTER OF SCIENCE\\
+A THESIS PRESENTED TO THE GRADUATE SCHOOL OF THE UNIVERSITY OF FLORIDA IN PARTIAL FULFILLMENT OF THE REQUIREMENTS FOR THE DEGREE OF MASTER OF SCIENCE
 \end{center}
 
-\begin{center}
-UNIVERSITY OF FLORIDA
-\end{center}
+\vspace{2cm}
 
@@
 
-\vspace*{11cm}
+\vspace{20cm}
 
@@
 
-\vspace*{10cm} 
+\vspace{15cm} 
 <!-- _Your dedication is typed here.
@@
 
-\textbf{ACKNOWLEDGMENTS}
+ACKNOWLEDGMENTS
 
@@
 <!-- _Acknowledgments must be written in complete sentences. Do not use direct address. For example, instead of Thanks, Mom and Dad!, you should say I thank my parents._  -->
-\doublespacing
+\begin{center}
 
@@
 
+\end{center}
+
 \newpage
@@
 \begin{center}
-\textbf{LIST OF ABBREVIATIONS}
+
+LIST OF ABBREVIATIONS
+
 \end{center}
 
-\vspace{1cm}
-
 SRS: Savannah River Site.  
-
-\newpage
-
-\vspace*{5cm}
-
-\singlespacing
+GLMM: Generalized Linear Mixed Model
+
+\newpage
 
 \begin{center}
-Abstract of Thesis Presented to the Graduate School\\  
-of the University of Florida in Partial Fulfillment of the\\  
-Requirements for the Degree of Master of Science.\\   
+
+Abstract of Thesis Presented to the Graduate School
+
+of the University of Florida in Partial Fulfillment of the
+
+Requirements for the Degree of Master of Science.
+
 \end{center}
 
-\vspace*{1cm}
-
-\begin{center}
+\newpage
+
 THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES
-\end{center}
-
-\begin{center}
+
 By
-\end{center}
-
-\begin{center}
+
 Eric Escobar-Chena
-\end{center}
-
-\begin{center}
+
 August 2025
-\end{center}
-
-\begin{flushleft}
-Chair: Emilio Bruna\\  
-Major: Wildlife Ecology and Conservation\\   
-\end{flushleft}
-
-\doublespacing
+
+Chair: Emilio Bruna 
+Major: Wildlife Ecology and Conservation
 
@@
 
-\begin{center}
-CHAPTER 1\\  
-INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES\\  
-\end{center}
-
 \doublespacing
+
+**CHAPTER 1**
+
+**Influence of Connectivity on Dung Beetle Communities**
 
@@
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy [@forman_general_1995].
+The resulting changes in species composition are important to identify because any species impacted would have corresponding effects depending on how they interact with the ecosysem **(Zhou 2023)**.
+Any gain or loss in key members of an community could disrupt processes which on their own could shape ecosystems**(Cuke 2016)**, or effect other organisms which rely on said interaction **(Wu 2011)**.
+Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
+
 *[**EB edit:** break this next sentence into 3 to carry reader though the logic: changes in species diverstiy are important to understand because species are involved in interactions.
@@
 
-The resulting changes in species composition are important to identify because
-
-Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
-
 <!-- stuff that these inviduals or species do: -->
@@
 By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes [@fletcher_jr_landscape_2023; @resasco_contribution_2017].
+
 Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected.
@@
 Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions [@hasan_dung_2024].
-The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013].
-[**EB edit:** I added the following to justify dung veetles more, and then made the questions a second paragraph] Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) **(citation)**.
+The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013 **nichols 2008**].
+[**EB edit:** I added the following to justify dung veetles more, and then made the questions a second paragraph] Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) **(ospina 2018, decastro 2023)**.
 Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].
@@
 We compared community composition by increasing magnitudes of diversity components (i.e., *qD*) of *0D* (i.e., species richness), *1D* (i.e., Shannon entropy), and *2D* (i.e, Simpson Diversity).
-<!--[can you add a sentence about how these are calculated?
-by which I mean - are the calculations different than for richness, shannon, etc.?]--> Diversity numbers and species richness were calculated using the package `hill` [@li_hillr_2018] for the R statistical programming language [@posit_team_rstudio_2025].
+Diversity numbers and species richness were calculated using the package `hill` [@li_hillr_2018] for the R statistical programming language [@posit_team_rstudio_2025].
 Diversity numbers were calculated using package `iNEXT` [@hsieh_inext_2016].
@@
 We included the identity of the sampling block as a random effects.
-To model our diversity metrics we took a similar approach, but this time using GLMMs with a Gaussian distribution [@chao_rarefaction_2014].
+To model our diversity metrics we took a similar approach, but this time using linear mixed effects models with a Gaussian distribution [@chao_rarefaction_2014].
 All models were fit using `lme4` package [@bates_fitting_2015].
 Prior to conducting our modeling we evaluated the the suitability of our data with *qqplots* generated with the `DARMa` package [@hartig_dharma_2024].
+
+We used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation.
+In both models, the reference level for patch type was Connected.
+Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.
 
@@
 
-[**EB:** edit: shouldn't this go in the methods?] We used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation.
-In both models, the reference level for patch type was Connected.
-
 For Shannon Diversity, the estimated mean in Connected patches was 5.29 (SE = 0.66, t = 7.95).
@@
 Again, none of the other patch types had significant effects: Matrix ($\beta$ = 0.11, SE = 0.41, t = 0.26), Rectangle ($\beta$ = 0.17, SE = 0.41, t = 0.40), and Winged ($\beta$ = –0.37, SE = 0.41, t = –0.91).
-
-Across both models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.
 
@@
 
-\vspace*{1cm}
-
-\begin{center}
-\textbf{BIOGRAPHICAL SKETCH}
-\end{center}
+# BIOGRAPHICAL SKETCH
 
@@
 The author should have replaced this paragraph with their own._ -->
-
-\doublespacing
 

### 2025-06-23 corridor_docs/eric_ms_thesis/MS_thesis.Rmd

--- 2025-06-20 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
@@
 SRS: Savannah River Site.  
+GLMM: Generalized Linear Mixed Model
 

### 2025-06-23 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
updating outline

@@
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy [@forman_general_1995].
-The resulting changes in species composition are important to identify because any species impacted would have corresponding effects depending on how they interact with the ecosysem **(Zhou 2023)**.
-Any gain or loss in key members of an community could disrupt processes which on their own could shape ecosystems**(Cuke 2016)**, or effect other organisms which rely on said interaction **(Wu 2011)**.
+The resulting changes in species composition are important to identify because any species impacted would have corresponding effects depending on how they interact with the ecosystem [@zhouLinkingEcosystemServices2023].
+Any gain or loss in key members of an community could disrupt processes which on their own could shape ecosystems [@WOS:000374962000009], or effect other organisms which rely on said interaction [@wuBrownworldCascadeDung2011].
 Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
@@
 Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions [@hasan_dung_2024].
-The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013 **nichols 2008**].
-[**EB edit:** I added the following to justify dung veetles more, and then made the questions a second paragraph] Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) **(ospina 2018, decastro 2023)**.
+The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013; @nicholsEcologicalFunctionsEcosystem2008].
+[**EB edit:** I added the following to justify dung veetles more, and then made the questions a second paragraph] Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits [e.g., size, foraging style, resource-use, @ospina-garcesDungBeetlesShow2018; @decastro-arrazolaTraitbasedFrameworkDung2023].
 Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].
@@
 
-*Don't repeat results...tell us:*
-
-1)  what your OVERALL conclusions are,
-2)  HOW your results compare to those of other studies,
-3)  WHY they are similar and / or different
-4)  BIOLOGICAL INSIGHTS you might have gained about partiicular species
-5)  any CAVEATS to your results or shortcomings of your study
-6)  the NEXT STEPS you recommend be taken by researchers interested in this toopic or system
+\singlespacing
+
+\begin{flushleft}  
+\textbf{NEED}
+\end{flushleft}
+
+1. rarefaction curves  
+1. add biomass to table 1  
+1. add totals of habitat in table 2 to table 1.
+
+\begin{flushleft}  
+\textbf{OUTLINE}
+\end{flushleft}
+
+\begin{flushleft} 
+\textbf{Paragraph 1: }Overview: While there may indeed be effects of patch area and connectivity on dung beetle abundance on community composition, other (landscape-scale) drivers appear to be more important.
+\end{flushleft}
+
+\begin{flushleft} 
+\textbf{Paragraph 2: }Effects of Edge vs. Connectivity
+\end{flushleft}
+
+- corridors vs. drift fence; predicted effects of body size and habitat preference
+
+- important difference from previous drift fence studies: ours has both passive drift effects AND directed movement towards dung
+
+- by laying out the same amount of dung in traps in all patch types, we eliminated the potential confounding effects of resource levels and abundance (sensu nick butterfly paper).
+Our results therefore reflect the effects of patch area / isolateion/ block rather than background abundnance.
+
+\begin{flushleft} 
+\textbf{Paragraph 3: }What then are the drivers of dung beetle diversity and abundance?
+\end{flushleft}
+
+- if not area, edge, or connectivity...what are we (and maybe others) missing? Block effects suggest something at much larger scales.  Potential hypotheses:
+
+  > a) _landscape-scale variation in abundance of large mammals (deer, hogs) at SRS._ Primary dung sources likely to be deer (?), wild hogs (?), maybe gopher tortoise, coyote, and fox (?).
+Is anything known about the abundance of these at the SRS site, wither in patches or at the block level?
+
+  > b) _matrix permeability/structure/canopy cover_ Is the matrix denser around patches where captures were lower?
+
+  > c) _gradients in soil moisture/compaction/quality_ (related to breeding)
+
+  > d) _land-use history / legacy of ag / road proximity_ (looping around to mammal abund)
+
+\begin{flushleft} 
+\textbf{Paragraph 4: }That said, there were some interspecific differences in abundance. 
+\end{flushleft}
+
+- What are they and what could be behind them?
+
+\begin{flushleft} 
+\textbf{Paragraph 5:} Our results largely echo/constrast with previous ones from the SRS and elsewhere 
+\end{flushleft}
+
+- example 1:
+
+- example 2:
+
+- example 3:
+
+\begin{flushleft} 
+\textbf{Paragraph 6: }Caveats
+\end{flushleft}
+
+- potential effects of seasonality
+
+- Sampling wasn't designed to look for potential effects of edge proximity, and so we couldn't test for previously observed pattern of generalist/opportunistic species more common near edges.
+
+- used great bait, but possible some species (eg specialists) not attracted
+
+\begin{flushleft} 
+\textbf{Paragraph 7: }Moving forward
+\end{flushleft}
+
+- tough to compare when most of the work has been done in the tropics; surprisingly few studies of this species-rich and ecolgically vital guild in the SE US.
+"Despite ample work documenting their ecological importance, there is a suprising lack of research on dung beetle diversity and corridors outside of the tropics.
+That said, tropical studies find stark contrast between fragments/pasture.
+They have been found to use corridors, but it is...."
+
+- experimental approaches to track movement and distinguish beetles already in pacthes vs. those coming from outside. Our preliminary data with MRR indicates individuals do move through corridors.
+
+\begin{flushleft} 
+\textbf{Paragraph 8: } Close with takeaways for conservation and management
+\end{flushleft}
+
+-   dung beetles are robust
+
+-   implications for ecosystem processes.
+
+-   what do think about corridor design and considerations for fragmented landscapes
+
+-   practical applications think about the beetles
+
+\begin{flushleft} 
+\textbf{A few things to consider incorporating}
+\end{flushleft}
+
+- ellen wind paper/ connectivity
 
@@
 -   habitat preference and seasonality (conover 2019, nealis 1977)
-
 > -   cvig - open sand chaparral, open grassland preference nealis 1977
@@
 > -   open - slight preference towards open chap and shaded chap, little open grassland nealis 1977
-
 > -   pign - sharing forest habitat but need more info conover 2019
@@
 -   like julians paper corridors benefit certain populations and more fit populations are able to make better use
-
-------------------------------------------------------------------------
-
-Discussion Outline & Notes
-
-1.  Abundance
-
--   reason for highest abundances in matrix and connected patch
--   lower abundance in rectangle hinting at fragmentation effects
-
-2.  Species Richness & Diversity
-
--   the role of hbitat connectivity in shaping community structure or lack thereof
--   why do we think connnected had lowest species richness??
-
-3.  Functional traits and ecological impacts
-
--   did corridoors favor a functional trait
--   why might there be a trait response
--   implications for ecosystem processes
--   like duung removal papers seed dispersal and yep
-
-4.  Comparing to previous studies
-
--   how are things aligning
--   think about the biology
-
-5.  Limitations and future work
-
--   potential confounding factors (seasonality, distance from edge, sampling methodology and temmporal variation)
--   other directions to go (dispersal -\> radar, secifically measuring changes in ecosystem services)
-
-6.  Takeaways for conservation and management
-
--   dung beetles are robust
--   what do think about corridor design and considerations for fragmented lanndscapes
--   practical applications think about the beetles
 
@@
 
-```{r Table-aovMO, echo=FALSE,message = FALSE,warning=FALSE}
-aovM0<-read_csv(here("corridor_docs","eric_ms_thesis","tables","aovM0.csv")) %>%
-  mutate(term=case_when(
-    term == "patch_type" ~ "Patch type",
-    .default = as.character(term))) |>
-# aovM0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM0.csv") %>% 
-  kable(
-      digits = 3,
-      align="lcc",
-      format="latex",
-      row.names = FALSE,
-      booktabs=T,
-      linesep = "", #removes the blank line after every 5 lines
-      caption = "AOV TABLE CAPTION") %>%
-  kable_styling(bootstrap_options = c("striped"),
-                full_width = F,
-                latex_options=c("scale_down","hold_position"),
-                font_size = 12,
-                position = "center") %>%
-  # column_spec(1, italic = T) %>%
-  row_spec(0, bold = T)
-  # row_spec(1:15, hline_after = T)
-aovM0
-```
+<!-- ```{r Table-aovMO, echo=FALSE,message = FALSE,warning=FALSE} -->
+<!-- aovM0<-read_csv(here("corridor_docs","eric_ms_thesis","tables","aovM0.csv")) %>% -->
+<!--   mutate(term=case_when( -->
+<!--     term == "patch_type" ~ "Patch type", -->
+<!--     .default = as.character(term))) |> -->
+<!-- # aovM0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM0.csv") %>%  -->
+<!--   kable( -->
+<!--       digits = 3, -->
+<!--       align="lcc", -->
+<!--       format="latex", -->
+<!--       row.names = FALSE, -->
+<!--       booktabs=T, -->
+<!--       linesep = "", #removes the blank line after every 5 lines -->
+<!--       caption = "AOV TABLE CAPTION") %>% -->
+<!--   kable_styling(bootstrap_options = c("striped"), -->
+<!--                 full_width = F, -->
+<!--                 latex_options=c("scale_down","hold_position"), -->
+<!--                 font_size = 12, -->
+<!--                 position = "center") %>% -->
+<!--   # column_spec(1, italic = T) %>% -->
+<!--   row_spec(0, bold = T) -->
+<!--   # row_spec(1:15, hline_after = T) -->
+<!-- aovM0 -->
+<!-- ``` -->
 
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "AOV TABLE CAPTION") %>%
+      caption = "aovM3 patch type effects on...") %>%
   kable_styling(bootstrap_options = c("striped"),
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "AOV TABLE CAPTION") %>%
+      caption = "M3 patch type effects on...") %>%
   kable_styling(bootstrap_options = c("striped"),

### 2025-06-23 corridor_docs/eric_ms_thesis/btl_drying.R

@@ -27,3 +27,30 @@ btl_wt %>%
   # filter(interval!="weight_0") %>%
   ggplot( aes(x=interval, y=value, group=sample, color=sample)) +
   geom_line()
+
+
+
+#  btl weights ------------------------------------------------------------
+
+
+plate_weight<-btl_wt %>% 
+  ungroup() %>% 
+  filter(interval=="weight_5") %>% 
+  filter(sp_code=="plate") %>% 
+  select(value)  
+
+
+
+corrected_btl_wt<-
+btl_wt %>% 
+  filter(interval=="weight_5") %>% 
+  mutate(plate_weight=plate_weight$value) %>% 
+  filter(sp_code!="plate") %>% 
+  group_by(sp_code) %>% 
+  mutate(corrected_wt=value-plate_weight) %>% 
+  summarize(mean_wt=mean(corrected_wt))
+
+
+btl_wt_clean<-write_csv(corrected_btl_wt,here("corridor_docs","eric_ms_thesis","ms_data","data_clean","corrected_btl_wt.csv"))
+
+

### 2025-06-24 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
fixed citations

@@
 By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy [@forman_general_1995].
-The resulting changes in species composition are important to identify because any species impacted would have corresponding effects depending on how they interact with the ecosystem [@zhouLinkingEcosystemServices2023].
-Any gain or loss in key members of an community could disrupt processes which on their own could shape ecosystems [@WOS:000374962000009], or effect other organisms which rely on said interaction [@wuBrownworldCascadeDung2011].
+The resulting changes in species composition are important to identify because any species impacted would have corresponding effects depending on how they interact with the ecosystem [@zhou_linking_2023].
+Any gain or loss in key members of an community could disrupt processes which on their own could shape ecosystems [@WOS:000374962000009], or effect other organisms which rely on said interaction [@wu_brown-world_2011].
 Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].
@@
 Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions [@hasan_dung_2024].
-The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013; @nicholsEcologicalFunctionsEcosystem2008].
-[**EB edit:** I added the following to justify dung veetles more, and then made the questions a second paragraph] Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits [e.g., size, foraging style, resource-use, @ospina-garcesDungBeetlesShow2018; @decastro-arrazolaTraitbasedFrameworkDung2023].
+The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013; @nichols_ecological_2008].
+[**EB edit:** I added the following to justify dung veetles more, and then made the questions a second paragraph] Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits [e.g., size, foraging style, resource-use, [@decastro-arrazola_trait-based_2023; @ospina-garces_dung_2018]].
 Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].

### 2025-06-28 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
changes to modeling baseline and rewording first paragrapgh of discussion

@@
 A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 7).
-The baseline abundance (intercept) corresponds to the abundance of *Aphodius alloblackburneus* in connected patches.
+The baseline abundance corresponds to the abundance of *Aphodius alloblackburneus* in connected patches.
 Compared to this baseline, *Canthon vigilans*, *Ateuchus lecontei*, and *Phanaeus igneus* showed significantly higher abundance ($\beta$ = 1.05, 0.73, 0.66 respectively; all p \< 0.001), while *Onthophagus pennsylvanicus* had lower abundance ($\beta$ = –0.80, p \< 0.001).
 
-Species-patch interactions indicated indicated species specific responses to treatments.
+Species-patch interactions indicated species specific responses to treatments.
 In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix $\beta$ = –0.59, p \< 0.001).
@@
 \end{flushleft}
+
+This study advances our understanding of the factors shaping dung beetle community composition in temperate regions of the southeastern United States.
+In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches.
+Our main findings emphasized: (1) Habitat type and patch shape were the main driving factors for dung beetle species abundances were composed, however effects were species specific.
+(2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level.
+(3) Species diversity metrics were also relatively even across patch types however varied widely by sampling blocks.
+These results suggest that while there may indeed be effects of patch structure and connectivity on dung beetle abundances and community composition, other landscape scale drivers appear to be more prominent for species richness and diversity.
 

### 2025-06-28 corridor_docs/eric_ms_thesis/02_data_analysis.R

@@ -468,6 +468,11 @@ h_rich<-as.data.frame(h_rich) |>
   rownames_to_column() %>% 
   separate(rowname,c("patch_type","block"),sep="_",remove=TRUE)
 
+h_rich<-h_rich %>% 
+  mutate(patch_type=as.factor(patch_type)) 
+h_rich$patch_type<-relevel(h_rich$patch_type,"Matrix")
+
+
 # taxonomic shannon diversity
 h_shannon<-hill_taxa(hill_data, q = 1)
 
@@ -475,12 +480,20 @@ h_shannon<-as.data.frame(h_shannon) |>
   rownames_to_column() %>% 
   separate(rowname,c("patch_type","block"),sep="_",remove=TRUE)
 
+h_shannon<-h_shannon %>% 
+  mutate(patch_type=as.factor(patch_type)) 
+h_shannon$patch_type<-relevel(h_shannon$patch_type,"Matrix")
+
 # taxonomic simpson diversity
 h_simpson<-hill_taxa(hill_data, q = 2)
 
 h_simpson<-as.data.frame(h_simpson) |>
   rownames_to_column() %>% 
   separate(rowname,c("patch_type","block"),sep="_",remove=TRUE)
+
+h_simpson<-h_simpson %>% 
+  mutate(patch_type=as.factor(patch_type)) 
+h_simpson$patch_type<-relevel(h_simpson$patch_type,"Matrix")
 
 
 hill_results<-h_rich %>% 

### 2025-06-28 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -2,7 +2,7 @@ library(tidyverse)
 library(here)
 library(kableExtra)
 library(sjPlot)
-
+library(lme4)
 # READ THIS ABOUT GLMM REFERENCE LEVELS
 # https://stats.stackexchange.com/questions/628348/when-i-change-my-reference-level-on-my-glmer-in-r-why-do-the-p-values-change-an
 
@@ -61,6 +61,11 @@ spp_abund<-btl_data %>%
   select(-patch,-block) %>% 
   separate(patch_block,c("patch_type","block"),sep="_",remove=TRUE) 
 
+#releveling
+spp_abund<-spp_abund %>% 
+  mutate(patch_type=as.factor(patch_type)) 
+spp_abund$patch_type<-relevel(spp_abund$patch_type,"Matrix")
+
 spp_abund_top6<-btl_data %>%
   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
   mutate(patch=case_when(
@@ -84,6 +89,9 @@ spp_abund_top6<-btl_data %>%
            sp_code=="open"|
            sp_code=="aaeg")
 
+spp_abund_top6<-spp_abund_top6 %>% 
+  mutate(patch_type=as.factor(patch_type)) 
+spp_abund_top6$patch_type<-relevel(spp_abund_top6$patch_type,"Matrix")
 
 hill_data<-hill_data %>% column_to_rownames("patch_block")
 
@@ -188,6 +196,8 @@ ggplot(spp_abund_top6,
 
 
 
+
+
 # Global model: Take 1 abundance, indep of species ID
 M0 <- glmer(n ~ patch_type + (1 + patch_type | block), 
             data = spp_abund, 
@@ -213,7 +223,26 @@ write_csv(aovM0,"./corridor_docs/eric_ms_thesis/tables/aovM0.csv")
 # nice_table(stats.table, broom = "glmm")
 # nice_table(stats.table, highlight = TRUE)
 
+#re orienting models to use matrix as a baseline factor
 
+set.seed(123)
+x <- rnorm(100)
+DF <- data.frame(x = x,
+                 y = 4 + (1.5*x) + rnorm(100, sd = 2),
+                 b = gl(5, 20))
+head(DF)
+str(DF)
+
+m1 <- lm(y ~ x + b, data = DF)
+summary(m1)
+
+DF <- within(DF, b <- relevel(b, ref = 3))
[…]

### 2025-07-04 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
fixing models and figures, rewriting most things before discussion, misc

@@
 ---
-title: "Influence of Connectivity on Dung Beetle Communities"
+title: "The Influence of Connectivity on Dung Beetle Communities"
 author: "ERIC ESCOBAR-CHENA"
@@
 \begin{center}
-To my family who never stopped supporting me along this journey, my friends who kept me company along the way, and my mentors at VCU who believed in me before I did myself.
+To my mom, dad, and two sisters. This thesis is a chapter in our immigrant story. 
 \end{center}
@@
 CHAPTER 1\\  
-INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES\\  
+THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES\\  
 \end{center}
@@
 By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes [@fletcher_jr_landscape_2023; @resasco_contribution_2017].
-
-Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected.
-Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services [@hevia_trait-based_2017].
-[**EB edit:** these last sentences don't really make the case for experiments, whcih were the initial emphasis of the paragraph.
-Might want to opt for another option, like the ones in the green notes below]
+Past studies have endeavored to experimentally measure changes in community compositions as a result of connecting habitats with corridors [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002] **here fix citations**, yet very few have directly compared matrix and patch populations.
+**examples of dispersal concepts (drift fence etc** Land use is different from one species to another so it is vital to understand where compositions are distributed and what processes might be driving population differences.
 
@@
 To model our diversity metrics we took a similar approach, but this time using linear mixed effects models with a Gaussian distribution [@chao_rarefaction_2014].
+In all models, the reference level for patch type was Matrix.
 All models were fit using `lme4` package [@bates_fitting_2015].
@@
 We used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation.
-In both models, the reference level for patch type was Connected.
 Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.
@@
 \newpage
+```{r, label = patch, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle abundance in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+
+knitr::include_graphics("images/sp_abund.png")
+
+```
+
+```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle abundance in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+
+knitr::include_graphics("images/sp_abund_patch.png")
+
+```
+
+```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle secies richness in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+
+knitr::include_graphics("images/sp_richness_patches.png")
+```
+
+```{r, label = shannon, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Shannon diversity in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+
+knitr::include_graphics("images/sp_shannon_patches.png")
+
+```
 
@@
 knitr::include_graphics("images/sp_simpson_patches.png")
-
-```
-
-```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle secies richness in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
-
-knitr::include_graphics("images/sp_richness_patches.png")
-
-```
-
-```{r, label = shannon, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Shannon diversity in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
-
-knitr::include_graphics("images/sp_shannon_patches.png")
-
-```
-
-```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle abundance in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
-
-knitr::include_graphics("images/sp_abund_patch.png")
 
@@
 Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling.
+
+Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services [@hevia_trait-based_2017].
+[**EB edit:** these last sentences don't really make the case for experiments, whcih were the initial emphasis of the paragraph.
+Might want to opt for another option, like the ones in the green notes below]

### 2025-07-04 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -517,9 +517,38 @@ sp_simpson_patches<-
 # This saves a png version of file to the images folder
 ggsave("corridor_docs/eric_ms_thesis/images/sp_simpson_patches.png", width = 4, height = 4, units = "in")
 
-# Figure for abundance model 
 
-?plot_model
+btl_abund <- plot_model(M0, type = "pred", title = NULL)
+btl_abund
+
+btl_abund + 
+  theme_minimal() +
+  labs( title = NULL ,
+        x = "Patch Type",
+        y = "Number of Beetles") +
+  theme(axis.text = element_text(size = 12)) +
+  scale_shape_manual(values=c(15:18))+ # change shapes of points
+  scale_color_viridis_d(option = "turbo") + 
+  #geom_point(position=position_jitter(width=0.16, height=0.15))+ # manipulate spread of jitter
+  theme_classic()+
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust = 0.95,size = 10)) +
+  theme(
+    legend.box.background = element_rect(color="gray", size=1),
+    legend.box.margin = margin(0.1, 0.1, 0.1, 0.1),
+    legend.position = "top",
+    legend.text = element_text(size = 10, colour = "black"),
+    legend.title = element_text(size = 10, colour = "black")
+  )+
+  theme(axis.title.y = element_text(size = 12,face="bold"))+
+  theme(axis.title.x =element_text(size = 12,face="bold"))+
+  theme(axis.text.y = element_text(size = 10))
+#(limits = c(1, 10), breaks = seq(1, 10, by = 2))
+
+ggsave("corridor_docs/eric_ms_thesis/images/sp_abund.png", width = 4, height = 4, units = "in")
+
+
+# Figure for abundance_sp_type model 
+
 
 btl_abund_patch <- plot_model(M3, type = "pred", terms = c("patch_type", "sp_code"), title = NULL)
 btl_abund_patch

### 2025-07-05 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
buuuuuuuuu
hummmmm

@@
 
+In comparing patc
+
 - corridors vs. drift fence; predicted effects of body size and habitat preference
@@
 
-```{r, label = srs-map, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Experimental block with the different patches and distance between them.", out.width="40%", fig.align='center'}
+```{r echo=FALSE, fig.align='center', fig.cap="Experimental block with the different patches and distance between them.", message=FALSE, warning=FALSE, label=srs-map, out.width="40%"}
 

### 2025-07-06 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
the end pt 1

@@
 Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003; @fletcher_is_2018].
-However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012; @haddad_corridor_2003; @jennings_composition_2006].
+However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012; @haddad_corridor_2003; @jennings_composition_2006; **harrison and bruna**].
 This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
@@
 The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013; @nichols_ecological_2008].
-[**EB edit:** I added the following to justify dung veetles more, and then made the questions a second paragraph] Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits [e.g., size, foraging style, resource-use, [@decastro-arrazola_trait-based_2023; @ospina-garces_dung_2018]].
+[**EB edit:** I added the following to justify dung beetles more, and then made the questions a second paragraph] Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits [e.g., size, foraging style, resource-use, [@decastro-arrazola_trait-based_2023; @ospina-garces_dung_2018]].
 Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].
@@
 
-When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches($\beta$ = -0.402, P \< 0.001, Table 4).
+When comparing the overall abundance of beetles (all species combined) across patch types, matrix plots had the highest captures, followed by connected patches, then winged, with the fewest in rectangular patches.
+Abundances from connected patches were not significantly different from those in the matrix while rectangle and winged patches had significantly fewer than the matrix ($\beta$ = -0.528 and -0.303 respectively, P \< 0.001, Table 4).
 
@@
 A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 7).
-The baseline abundance corresponds to the abundance of *Aphodius alloblackburneus* in connected patches.
-Compared to this baseline, *Canthon vigilans*, *Ateuchus lecontei*, and *Phanaeus igneus* showed significantly higher abundance ($\beta$ = 1.05, 0.73, 0.66 respectively; all p \< 0.001), while *Onthophagus pennsylvanicus* had lower abundance ($\beta$ = –0.80, p \< 0.001).
-
-Species-patch interactions indicated species specific responses to treatments.
-In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix $\beta$ = –0.59, p \< 0.001).
-*Ateuchus lecontei* and *Phanaeus igneus* showed increased effect on abundance in rectangle patches relative to expectations ($\beta$ = 0.40 and 0.47, respectively; p \< 0.05).
-Random effects indicated moderate variability among blocks (standard deviation of intercept = 0.51).
+The baseline abundance corresponds to the abundance of *Aphodius alloblackburneus* in matrix patches.
+Compared to this baseline, results were highly variable, emphasizing species specific responses to patch type.
+Similarly, species showed to have disproportionate responses to patch since interaction terms varied widely (Table 6).
 
@@
 Modeling the effect of patch type on species richness with block as a random effect determined there was no significant differences among patch types.
-Comparing treatments using connected patches as a baseline resulted in no significant differences in matrix ($\beta$ = 0.05, p = 0.83), rectangle ($\beta$ = -0.07, p = 0.74), and winged patches ($\beta$ = -0.05, p = 0.83).
-
-For Shannon Diversity, the estimated mean in Connected patches was 5.29 (SE = 0.66, t = 7.95).
-None of the alternative patch types showed statistically significant differences compared to Connected: Matrix ($\beta$ = –0.085, SE = 0.50, t = –0.17), Rectangle ($\beta$ = 0.15, SE = 0.50, t = 0.30), or Winged ($\beta$ = –0.34, SE = 0.50, t = –0.68).
-Similarly, for Simpson's Diversity, the average value in Connected patches was 4.07 (SE = 0.64, t = 6.37).
-Again, none of the other patch types had significant effects: Matrix ($\beta$ = 0.11, SE = 0.41, t = 0.26), Rectangle ($\beta$ = 0.17, SE = 0.41, t = 0.40), and Winged ($\beta$ = –0.37, SE = 0.41, t = –0.91).
+Comparing treatments using matrix patches as a baseline resulted in no significant differences in connected ($\beta$ = 0.216, p = 0.839; Table 7), rectangle ($\beta$ = -0.548, p = 0.583; Table 7), and winged patches ($\beta$ = -0.095, p = 0.663; Table 7).
+
+Biodiversity was also even among patch types, however metric values were varied by sampling blocks.
+For Shannon Diversity, the estimated mean in matrix patches was 5.202 (SE = 0.665, t = 7.825).
+None of the alternative patch types showed statistically significant differences compared to matrix: connected ($\beta$ = –0.085, SE = 0.495, t = 0.17), rectangle ($\beta$ = 0.234, SE = 0.495, t = 0.472), or winged ($\beta$ = –0.251, SE = 0.495, t = –0.506).
+Similarly, for Simpson's Diversity, the average value in matrix patches was 4.181 (SE = 0.640, t = 6.535).
+Again, none of the other patch types had significant effects: matrix ($\beta$ = -0.107, SE = 0.411, t = -0.259), rectangle ($\beta$ = 0.059, SE = 0.411, t = 0.143), and winged ($\beta$ = –0.480, SE = 0.411, t = –1.169).
 
@@
 In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches.
-Our main findings emphasized: (1) Habitat type and patch shape were the main driving factors for dung beetle species abundances were composed, however effects were species specific.
+Our main findings emphasized: (1) Habitat type and patch shape were the main driving factors for determining how dung beetle species abundances were composed, however effects were species specific.
 (2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level.
@@
 
-In comparing patc
+In our comparison of dung beetle compositions, connectivity and habitat edge accounted for differences in dung beetle abundances, yet species richness and diversity were even across patch types and in  the matrix.
+Total beetle counts were consistently highest in the matrix, in comparison abundances is connected patches were not significantly different while those of the winged and rectangular patches were lower.
+Additionally, we also observed that patch effects were not equally proportional for all species.
+This suggests that the species captured could be exhibiting habitat preferences between the open patches versus forested matrix.
+Another potential explanation is that species in the matrix are acting as a source population which feeds into patches with edge acting as a drift fence for directing movement.
+The latter example seems more likely because flies, which have similar resource dependencies were also found to have a similar interaction ** fried **, and also studies on dung beetles in the tropics found more dramatic differences in dung beetle populations in fragments and matrix [@bustamante-sanchez_dung_2004; @WOS:000180797500010; @WOS:000288810500026].
 
@@
 \end{flushleft}
+
+Although richness and diversity were the same among treatments, there was notable variability between sampling blocks.
 

### 2025-07-07 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
beeg push

@@
 
-- corridors vs. drift fence; predicted effects of body size and habitat preference
-
-- important difference from previous drift fence studies: ours has both passive drift effects AND directed movement towards dung
-
-- by laying out the same amount of dung in traps in all patch types, we eliminated the potential confounding effects of resource levels and abundance (sensu nick butterfly paper).
-Our results therefore reflect the effects of patch area / isolateion/ block rather than background abundnance.
-
 \begin{flushleft} 
@@
 Although richness and diversity were the same among treatments, there was notable variability between sampling blocks.
-
-- if not area, edge, or connectivity...what are we (and maybe others) missing? Block effects suggest something at much larger scales.  Potential hypotheses:
-
-  > a) _landscape-scale variation in abundance of large mammals (deer, hogs) at SRS._ Primary dung sources likely to be deer (?), wild hogs (?), maybe gopher tortoise, coyote, and fox (?).
-Is anything known about the abundance of these at the SRS site, wither in patches or at the block level?
-
-  > b) _matrix permeability/structure/canopy cover_ Is the matrix denser around patches where captures were lower?
-
-  > c) _gradients in soil moisture/compaction/quality_ (related to breeding)
-
-  > d) _land-use history / legacy of ag / road proximity_ (looping around to mammal abund)
+Block 8 generally had the lowest species richness and biodiversity while 53n had the highest.
+While the design of the experimental blocks control for as much as possible, landscape features across the sight are different enough to influence species composition.
+Several factors may explain this pattern, including soil quality which influences where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019] as well as matrix permeability, canopy cover, and different land-use types (e.g., roadways, water sources)** harvey **[@WOS:000651446900001].
+Not only do these impact dung beetle movement, but also these landscape features can affect mammal movement, which in turn limits dung availability for beetles.
 
@@
 
-- What are they and what could be behind them?
+Species specific differences are apparent but do not follow any particular trend.
+_Aphodius alloblackburneus_ had a disproportionate positive effect to being captured in matrix patches as compared to other species.
+Beetles from the genus _Aphodius_ do tend to show patterns of habitat specificity[@roslin_distribution_2001], so it is not unexpected that they might show a preference towards either forested matrix or open patches.
+We did not detect that any species was more positively associated with open patches despite expectations that some species ( eg.
+_Canthon vigilans_, _Melanocanthon bispinatus_) would prefer open habitat [@conover_phenology_2019; @nealis_habitat_1977].
+Altogether this could be another sign pointing towards matrix acting as a source population, and since open habitat was much less dominant in our experimental system beetles could be moving into patches from habitat edge.
 
@@
 
-- example 1:
-
-- example 2:
-
-- example 3:
+Despite dung beetles being a well studied model system, most of the current work in fragmented habitats come from tropical climates[@nichols_global_2007], with any work in temperate locations coming from Europe where _Aphodius_ beetles dominate[@roslin_distribution_2001].
+Many studies in the tropics have found patterns of habitat preference between open fields and continuous forest similar to what we did in this study[@arellano_response_2008; @WOS:000349242000004].
+However, differences in composition were much more distinctly different with some species being exclusive to habitat types unlike we see in this paper.
+Of the limited work out of the south eastern United States we did recover assemblages similar to other work [@conover_phenology_2019; @nealis_habitat_1977; @young_pines_nodate], and even returned results to another study where differences between treatment types were non-significant while variation on a larger scale was more apparent[@young_dung_2023].
+In comparison to other studies from SRS our results were similar to thise  of small mammals which were more likely to be using matrix land *mabry*.
+We did expect some species to be more dominant in patches with higher edge ratios as was found with ants at SRS [@resasco_landscape_2014], but species were evenly present in all patches.
 

### 2025-07-07 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
holy canoli

@@
 
-# TO DO LIST
-
-FOR TOMORROW:
-
-1)  Finish adding citations
-
-2)  Stat results: biology first, then stats (e.g., abundance ranged form xxx in rectangular to xxx in winged)
-
-3)  Review other papers to find:
-
--   those with similar results
--   those with contrary results
--   do this with dung bettles AND other insect taxa
-
-DOWN THE ROAD
-
-1)  **Results:** Did you get any species not captured in other long-leaf/SE USA forest sites?
-
-2)  **Methods:** Find me the best example paragraph you think you read describing the use of hill numbers to estimate diversity/richness (field survey paper)
-
-3)  **Methods:** Find me the best TWO descriptions of using GLMM to compare abundance/diversity/richness at the Corridor Project
-
-4)  **Methods:** Write a short paragraph on how you calculated and compared functional diversity (how you measured it, how you compared it)
-
-5)  **Figure:** biomass of different (R/T/D) functional groups.
-
-6)  **Table:** Results of functional diversity analyses.
-
-\newpage
-
 \vspace*{5cm}
@@
 Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003; @fletcher_is_2018].
-However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012; @haddad_corridor_2003; @jennings_composition_2006; **harrison and bruna**].
+However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012; @haddad_corridor_2003; @jennings_composition_2006; [@harrison_habitat_1999]].
 This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
@@
 
-*[**EB edit:** break this next sentence into 3 to carry reader though the logic: changes in species diverstiy are important to understand because species are involved in interactions.
-If we gain or lose species we could gain or lose interactions and resulting ecosystems processes.
-there is some evidence of this, but...
-look at the green text notes below]*
-
 <!-- stuff that these inviduals or species do: -->
@@
 By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes [@fletcher_jr_landscape_2023; @resasco_contribution_2017].
-Past studies have endeavored to experimentally measure changes in community compositions as a result of connecting habitats with corridors [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002] **here fix citations**, yet very few have directly compared matrix and patch populations.
-**examples of dispersal concepts (drift fence etc** Land use is different from one species to another so it is vital to understand where compositions are distributed and what processes might be driving population differences.
+Past studies have endeavored to experimentally measure changes in community compositions as a result of connecting habitats with corridors [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002]  Yet very few have directly compared matrix and patch populations.
+Land use is different from one species to another so it is vital to understand where compositions are distributed and what processes might be driving population differences[@haddad_corridor_1999].
 
@@
 The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013; @nichols_ecological_2008].
-[**EB edit:** I added the following to justify dung beetles more, and then made the questions a second paragraph] Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits [e.g., size, foraging style, resource-use, [@decastro-arrazola_trait-based_2023; @ospina-garces_dung_2018]].
+Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits [e.g., size, foraging style, resource-use, [@decastro-arrazola_trait-based_2023; @ospina-garces_dung_2018]].
 Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001].
@@
 \begin{flushleft} 
-\textbf{NEED}
-\end{flushleft}
-
-1. rarefaction curves  
-1. add biomass to table 1  
-1. add totals of habitat in table 2 to table 1.
-
-\begin{flushleft}  
-\textbf{OUTLINE}
-\end{flushleft}
-
-\begin{flushleft} 
 \textbf{Paragraph 1: }Overview: While there may indeed be effects of patch area and connectivity on dung beetle abundance on community composition, other (landscape-scale) drivers appear to be more important.
@@
 Another potential explanation is that species in the matrix are acting as a source population which feeds into patches with edge acting as a drift fence for directing movement.
-The latter example seems more likely because flies, which have similar resource dependencies were also found to have a similar interaction ** fried **, and also studies on dung beetles in the tropics found more dramatic differences in dung beetle populations in fragments and matrix [@bustamante-sanchez_dung_2004; @WOS:000180797500010; @WOS:000288810500026].
+The latter example seems more likely because flies, which have similar resource dependencies were also found to have a similar interaction [@fried_habitat_2005], and also studies on dung beetles in the tropics found more dramatic differences in dung beetle populations in fragments and matrix [@bustamante-sanchez_dung_2004; @WOS:000180797500010; @WOS:000288810500026].
 
@@
 While the design of the experimental blocks control for as much as possible, landscape features across the sight are different enough to influence species composition.
-Several factors may explain this pattern, including soil quality which influences where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019] as well as matrix permeability, canopy cover, and different land-use types (e.g., roadways, water sources)** harvey **[@WOS:000651446900001].
+Several factors may explain this pattern, including soil quality which influences where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019] as well as matrix permeability, canopy cover, and different land-use types (e.g., roadways, water sources)[@WOS:000241362400032; @WOS:000651446900001].
 Not only do these impact dung beetle movement, but also these landscape features can affect mammal movement, which in turn limits dung availability for beetles.
@@
 
-Despite dung beetles being a well studied model system, most of the current work in fragmented habitats come from tropical climates[@nichols_global_2007], with any work in temperate locations coming from Europe where _Aphodius_ beetles dominate[@roslin_distribution_2001].
-Many studies in the tropics have found patterns of habitat preference between open fields and continuous forest similar to what we did in this study[@arellano_response_2008; @WOS:000349242000004].
+Despite ample work documenting their ecological importance, there is a surprising lack of research on dung beetle diversity and corridors outside of the tropics[@nichols_global_2007], with any work in temperate locations coming from Europe where _Aphodius_ beetles dominate[@roslin_distribution_2001].
+Many studies in the tropics have found patterns of habitat preference between open fields and continuous forest similar to what we did in this study[@WOS:000349242000004], and some work that found that forest species were using living fences as corridors[@arellano_response_2008].
 However, differences in composition were much more distinctly different with some species being exclusive to habitat types unlike we see in this paper.
 Of the limited work out of the south eastern United States we did recover assemblages similar to other work [@conover_phenology_2019; @nealis_habitat_1977; @young_pines_nodate], and even returned results to another study where differences between treatment types were non-significant while variation on a larger scale was more apparent[@young_dung_2023].
-In comparison to other studies from SRS our results were similar to thise  of small mammals which were more likely to be using matrix land *mabry*.
+In comparison to other studies from SRS our results were similar to those of small mammals which were more likely to be using matrix land *mabry*.
 We did expect some species to be more dominant in patches with higher edge ratios as was found with ants at SRS [@resasco_landscape_2014], but species were evenly present in all patches.
@@
 
-- potential effects of seasonality
-
-- Sampling wasn't designed to look for potential effects of edge proximity, and so we couldn't test for previously observed pattern of generalist/opportunistic species more common near edges.
-
-- used great bait, but possible some species (eg specialists) not attracted
+There were some limitations of this study which should be addressed in any future work to capture a clearer picture of how populations are being altered by habitat connectivity.
+First, sampling for this study was conducted over a two month period in the summer of 2024, sufficient data was collected but due to a lack of available resources temporal patterns were obscured due to inconsistent sampling periods.
+This is highly important since dung beetles exhibit consistent patterns of seasonality [@conover_phenology_2019; @davis_feeding_1966].
+Another potential avenue for improvement is lowering the grain size of sampling and changing trap placement to better understand the effects of edge proximity and connectivity.
+We also used one of the best bait types for collecting dung beetles in pig dung [@marsh_optimising_2013], but a mix of differently sourced baits may have been more optimal as more diverse baits attract more diverse species [@frank_nutrient_2017; @frank_search_2018; @gimenez_gomez_dung_2021].
 
@@
 
-- tough to compare when most of the work has been done in the tropics; surprisingly few studies of this species-rich and ecolgically vital guild in the SE US.
-"Despite ample work documenting their ecological importance, there is a suprising lack of research on dung beetle diversity and corridors outside of the tropics.
-That said, tropical studies find stark contrast between fragments/pasture.
-They have been found to use corridors, but it is...."
-
-- experimental approaches to track movement and distinguish beetles already in pacthes vs. those coming from outside. Our preliminary data with MRR indicates individuals do move through corridors.
+Moving forward, there is a bounty of knowledge yet to be collected for dung beetles in the south eastern United States.
+Relative to what is already known in tropical ecosystems, this study provides a glimpse on how dung beetles are responding to fragmentation and connectivity in subtropic pine habitats, but much can still be learned about dung beetles in these spaces and the role they play in ecosystems as well as how they interact with landscapes.
+The unique role of dung beetles as waste removers is fascinating enough and of great importance for maintaining landscapes so dominated by pastures like Southern US [@cheng_dweller_2022; @WOS:000368243300001].
+Not only waste removal but also secondary effect such as seed dispersal and parasitic reduction are important to understand in the context of fragmentation [@fincher_effects_1975; @vulinec_dung_2002].
+SRS provides an excellent experimental design for direct comparison of landscapes but another main goal of the site is to study dispersal, dung beetles would be an excellent system for studying movement.
+They already show promise, in preliminary trials we released beetles from both the connected and rectangular patches with baited traps in the central patch.
+We recovered one recapture which originated from the connected patch, indicating that individuals do move through corridors.
 
@@
 
--   dung beetles are robust
-
--   implications for ecosystem processes.
-
--   what do think about corridor design and considerations for fragmented landscapes
-
--   practical applications think about the beetles
-
-\begin{flushleft} 
-\textbf{A few things to consider incorporating}
-\end{flushleft}
-
-- ellen wind paper/ connectivity
-
-This study advances our understanding of the factors shaping dung beetle community composition in temperate regions of the southeastern United States.
-In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches.
-Our main findings emphasized: (1) Habitat type and patch shape were the main driving factors for dung beetle species abundances were composed, however effects were species specific.
-(2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level.
-(3) Species diversity metrics were also relatively even across patch types however varied widely by sampling blocks.
-These results suggest that dung beetle species are fully capable of permeating fragmented landscapes and that habitat type and connectivity shape community compositions, but landscape effects of a larger scale are driving changes in biodiversity.
-
-### Abundance
-
-The total number of dung beetle species captured N = `r total_spp` and the total individuals N = `r total_N` are similar to those in studies conducted in similar habitats and regions of the US [@conover_phenology_2019; @nealis_habitat_1977; @stanbrook_dung_2022].
-
-Dung beetle abundances were significantly lowest in rectangular patches.
-Matrix patches had the most total beetles collected and consistently had the highest counts for the most dominant species.
-However abundances were not significantly different between matrix, connected, and winged patches.
-This may suggest that isolation is leading to a decrease in abundances, but it is more likely that populations in the matrix are acting as a source, and since connected and winged patches have higher edge to area ratios dung beetles were more likely to move into those patches.
-Past studies have shown trends where habitat type and forest regeneration stages are key in the partitioning of dung beetle assemblages[@conover_phenology_2019; @WOS:000386714700011; @arellano_response_2008] which supports why we might see these differences in local populations.
-In addition distance from habitat edge has been attributed to differences in spillover from source populations[@WOS:000707546700001], emphasizing the importance of habitat edge in the context of our study.
-
-We also observed that patch effects were not equally proportional for all species.
-The abundances of all dominant species were positively influenced by matrix habitat, but *Aphoidius alloblackburneus* responded more positively than any other species.
-*Phanaeus igneus* and *Ateuchus lecontei* also had higher positive effects in rectangle patches.
-
-### Richness
-
-While abundances were different between patch types, we did not detect any patterns of species richness in our modeling.
-Total species counts were very even across patch types and sampling blocks.
-This difference indicates that even though habitat type or landscape features are informing occupational preferences, land use is more or less the same between patches and matrix.
-This differs from past work where species richness is lower in forest fragments [@WOS:000178929600002]
-
-### Diversity
-
-Analysis of Hill numbers indicated that species compositions were even between patch types.
-Comparing values of Simpson's Diversity between patches determined that dominant species were not prevalent in any one treatment type.
-Likewise, the Shannon index values insisted that rare species throughout the study population were even.
-However, for both metrics, values varied greatly by sampling blocks.
-Similar to our richness results this suggests that variation between patch and matrix is not distinct enough to limit the land use of dung beetles within our study site.
-
-In contrast to what we observed with species richness it appears that assemblages larger landscape patterns are effecting the composition of assemblages across the study site.
-
-*here write about low sample sizes and reasons for why dung beetle communities would be so similar across treatment types.
-Dispersal, scale (roslin), maybe again a good reason to expect matrix as a source pop.*
-
-notes that are important for discussion
-
--   habitat preference and seasonality (conover 2019, nealis 1977)
-> -   cvig - open sand chaparral, open grassland preference nealis 1977
-> -   mbis - open sand chapparal preference nealis 1977
-> -   open - slight preference towards open chap and shaded chap, little open grassland nealis 1977
-> -   pign - sharing forest habitat but need more info conover 2019
-> -   pvin - open habitat but share forest with pign conover 2019
-> -   alec - forest preference conover 2019
-
--   habitat again but only part of variation form habitat specifically aphodius also phenology (roslin 2001)
--   spatial scales (roslin 2000)
--   morphology
-
-> -   Ospina 2018: wing shape py habitat preference and differences in wing shape determined by species groupings.
-Beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies).
-Large body beetles tending towards lower energy flight strategies.
-maybe justification for corridor preference?
-> -   stanbrook and king 2022: tunnelers preferring open habitat and tunnelers also contributing more towards dung removal.
-> -   Conover 2019: dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types
-> -   Gimenez Gomez et al 2021 -\> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used.
-However this study was done in a more tropical ecosystem so its hard to say if the same would apply to our more temperate system so more work is needed in this specific avenue.
-
--   functionality especially gas emissions (slade 2016)
--   species id and funcitonality (slade 2017)
--   faovored species dominating in fragments (resasco 2014)
--   wind direction in corridors (damschen 2014)
-
-revisit bray curtis
-
-1.  Abundances in matrix vs connected patch and why this could be happening
-
--   source pop to habiitat edge
-
-2.  species richness again supporting that matrix is more of an ideal habitat for dung beetle community
-
-3.  diversity indices community structures weren't highly different between patch types
-
-4.  bray curtis hinting at similar land uses between corridor and winged patch
-
--   like julians paper corridors benefit certain populations and more fit populations are able to make better use
+Although there is plentiful work to be done, this study paints an encouraging picture that while dung beetles face disturbances that make habitats less preferential, they are robust enough to persist throughout fragmented landscapes and the management strategies we do have to connect fragmented landscapes work to mitigate losses.
+Because assemblages remain mostly the same across these landscapes, ecosystem services should remain uninterrupted allowing continuing benefits to all community members.
+While this study of dung beetles in fragmented landscapes was conducted in a relatively protected area, in real world application we could consider the use of movement corridors within disturbed areas to help bolster effected populations, but also near areas of high importance such as pastures to provide areas of refuge for beetles.
+In this case it is important to consider the inverse of what was manipulated in this study where open field acts as matrix and forested area is focal landscape, thus it is vital to continue learning more about how landscape composition and connectivity effect dung beetle populations.
 
@@
 Might want to opt for another option, like the ones in the green notes below]
+
+This study advances our understanding of the factors shaping dung beetle community composition in subtropical regions of the southeastern United States.
+In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches.
+Our main findings emphasized: (1) Habitat type and patch shape were the main driving factors for dung beetle species abundances were composed, however effects were species specific.
+(2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level.
+(3) Species diversity metrics were also relatively even across patch types however varied widely by sampling blocks.
+These results suggest that dung beetle species are fully capable of permeating fragmented landscapes and that habitat type and connectivity shape community compositions, but landscape effects of a larger scale are driving changes in biodiversity.
+
+### Abundance
+
+The total number of dung beetle species captured N = `r total_spp` and the total individuals N = `r total_N` are similar to those in studies conducted in similar habitats and regions of the US [@conover_phenology_2019; @nealis_habitat_1977; @stanbrook_dung_2022].
+
+Dung beetle abundances were significantly lowest in rectangular patches.
+Matrix patches had the most total beetles collected and consistently had the highest counts for the most dominant species.
+However abundances were not significantly different between matrix, connected, and winged patches.
+This may suggest that isolation is leading to a decrease in abundances, but it is more likely that populations in the matrix are acting as a source, and since connected and winged patches have higher edge to area ratios dung beetles were more likely to move into those patches.
+Past studies have shown trends where habitat type and forest regeneration stages are key in the partitioning of dung beetle assemblages[@conover_phenology_2019; @WOS:000386714700011; @arellano_response_2008] which supports why we might see these differences in local populations.
+In addition distance from habitat edge has been attributed to differences in spillover from source populations[@WOS:000707546700001], emphasizing the importance of habitat edge in the context of our study.
+
+We also observed that patch effects were not equally proportional for all species.
+The abundances of all dominant species were positively influenced by matrix habitat, but *Aphoidius alloblackburneus* responded more positively than any other species.
+*Phanaeus igneus* and *Ateuchus lecontei* also had higher positive effects in rectangle patches.
+
+### Richness
+
+While abundances were different between patch types, we did not detect any patterns of species richness in our modeling.
+Total species counts were very even across patch types and sampling blocks.
+This difference indicates that even though habitat type or landscape features are informing occupational preferences, land use is more or less the same between patches and matrix.
+This differs from past work where species richness is lower in forest fragments [@WOS:000178929600002]
+
+### Diversity
+
+Analysis of Hill numbers indicated that species compositions were even between patch types.
+Comparing values of Simpson's Diversity between patches determined that dominant species were not prevalent in any one treatment type.
+Likewise, the Shannon index values insisted that rare species throughout the study population were even.
+However, for both metrics, values varied greatly by sampling blocks.
+Similar to our richness results this suggests that variation between patch and matrix is not distinct enough to limit the land use of dung beetles within our study site.
+
+In contrast to what we observed with species richness it appears that assemblages larger landscape patterns are effecting the composition of assemblages across the study site.
+
+*here write about low sample sizes and reasons for why dung beetle communities would be so similar across treatment types.
+Dispersal, scale (roslin), maybe again a good reason to expect matrix as a source pop.*
+
+notes that are important for discussion
+
+-   habitat preference and seasonality (conover 2019, nealis 1977)
+> -   cvig - open sand chaparral, open grassland preference nealis 1977
+> -   mbis - open sand chapparal preference nealis 1977
+> -   open - slight preference towards open chap and shaded chap, little open grassland nealis 1977
+> -   pign - sharing forest habitat but need more info conover 2019
+> -   pvin - open habitat but share forest with pign conover 2019
+> -   alec - forest preference conover 2019
+
+-   habitat again but only part of variation form habitat specifically aphodius also phenology (roslin 2001)
+-   spatial scales (roslin 2000)
+-   morphology
+
+> -   Ospina 2018: wing shape py habitat preference and differences in wing shape determined by species groupings.
+Beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies).
+Large body beetles tending towards lower energy flight strategies.
+maybe justification for corridor preference?
+> -   stanbrook and king 2022: tunnelers preferring open habitat and tunnelers also contributing more towards dung removal.
+> -   Conover 2019: dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types
+> -   Gimenez Gomez et al 2021 -\> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used.
+However this study was done in a more tropical ecosystem so its hard to say if the same would apply to our more temperate system so more work is needed in this specific avenue.
+
+-   functionality especially gas emissions (slade 2016)
+-   species id and funcitonality (slade 2017)
+-   faovored species dominating in fragments (resasco 2014)
+-   wind direction in corridors (damschen 2014)
+
+revisit bray curtis
+
+1.  Abundances in matrix vs connected patch and why this could be happening
+
+-   source pop to habiitat edge
+
+2.  species richness again supporting that matrix is more of an ideal habitat for dung beetle community
+
+3.  diversity indices community structures weren't highly different between patch types
+
+4.  bray curtis hinting at similar land uses between corridor and winged patch
+
+-   like julians paper corridors benefit certain populations and more fit populations are able to make better use
+
+\begin{flushleft}  
+\textbf{NEED}
+\end{flushleft}
+
+1. rarefaction curves  
+1. add biomass to table 1  
+1. add totals of habitat in table 2 to table 1.

### 2025-07-07 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
yepperioni

@@
 
-## Methods
+## METHODS
 
@@
 However, differences in composition were much more distinctly different with some species being exclusive to habitat types unlike we see in this paper.
-Of the limited work out of the south eastern United States we did recover assemblages similar to other work [@conover_phenology_2019; @nealis_habitat_1977; @young_pines_nodate], and even returned results to another study where differences between treatment types were non-significant while variation on a larger scale was more apparent[@young_dung_2023].
-In comparison to other studies from SRS our results were similar to those of small mammals which were more likely to be using matrix land *mabry*.
+Of the limited work out of the south eastern United States we did recover assemblages similar to other work [@conover_phenology_2019; @nealis_habitat_1977], and even returned results to another study where differences between treatment types were non-significant while variation on a larger scale was more apparent[@young_dung_2023].
+In comparison to other studies from SRS our results were similar to those of small mammals which were more likely to be using matrix land [@mabry_influence_2003].
 We did expect some species to be more dominant in patches with higher edge ratios as was found with ants at SRS [@resasco_landscape_2014], but species were evenly present in all patches.
@@
 
-```{r echo=FALSE, fig.align='center', fig.cap="Experimental block with the different patches and distance between them.", message=FALSE, warning=FALSE, label=srs-map, out.width="40%"}
+```{r echo=FALSE, fig.align='center', fig.cap="All experimental blocks at SRS. Orange indicates blocks sampled.", message=FALSE, warning=FALSE, label=srs-map, out.width="40%"}
 
@@
 
-```{r, label = pitfalls, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Experimental block with the different patches and distance between them.", out.width="50%", fig.align='center'}
+```{r, label = pitfalls, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Placement of traps within patch types and in the matrix as well as their orientation.", out.width="50%", fig.align='center'}
 
@@
 \newpage
-```{r, label = patch, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle abundance in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+```{r, label = patch, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle abundance in three different treatments and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
 
@@
 
-```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle abundance in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle abundance of the top 6 most abundant species in three different treatments and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
 
@@
 
-```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle secies richness in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle secies richness in three different treatments and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
 
@@
 
-```{r, label = shannon, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Shannon diversity in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+```{r, label = shannon, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Shannon diversity in three different treatments and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
 
@@
 
-```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index in three different --- and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index in three different treatments and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
 
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "AOV TABLE CAPTION") %>%
+      caption = "Model of species abundance by patch type.") %>%
   kable_styling(bootstrap_options = c("striped"),
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "aovM3 patch type effects on...") %>%
+      caption = "AOV patch type effects on species abundance and interavtive term.") %>%
   kable_styling(bootstrap_options = c("striped"),
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "M3 patch type effects on...") %>%
+      caption = "M3 patch type effects on species abundance of top 6 most abundant speceies and interactive term for patch type and species ID.") %>%
   kable_styling(bootstrap_options = c("striped"),

### 2025-07-08 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
update for sharing with cmmte

@@
 bibliography: ["eric-thesis.bib"]
-output: bookdown::pdf_document2
+output:
+  bookdown::pdf_document2:
+    toc: false
+# output: bookdown::pdf_document2
   # pdf_document:
@@
 ```
+\newpage
 
@@
 
-We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site.
+I thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site.
 I also wanted to specifically thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry.
@@
 
+\begin{flushleft}
+
 SRS: Savannah River Site.
-GLMM: Generalized Linear Mixed Model
+
+GLMM: Generalized Linear Mixed Model.
+
+\end{flushleft}
 
@@
 Corridors have become an important tool in order to combat the negative effects of fragmentation, however they are difficult to study in natural systems without incurring confounding effects.
-To observe changes in insect community composition as an effect of landscape features we sampled dung beetles in a landscape scale experiment.
-We did not see a difference in species richness or diversity, but dung beetle abundances were higher in continuous forest habitat and open habitat patches connected by a corridors than in isolated patches.
+To observe changes in insect community composition as an effect of landscape features I sampled dung beetles in a landscape scale experiment.
+I did not see a difference in species richness or diversity, but dung beetle abundances were higher in continuous forest habitat and open habitat patches connected by a corridors than in isolated patches.
 
@@
 Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003; @fletcher_is_2018].
-However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012; @haddad_corridor_2003; @jennings_composition_2006; [@harrison_habitat_1999]].
+However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012; @haddad_corridor_2003; @jennings_composition_2006; @harrison_habitat_1999].
 This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions  [@haddad_habitat_2015].
@@
 
-By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes [@fletcher_jr_landscape_2023; @resasco_contribution_2017].
+By measuring changes in biodiversity and species richness within experimental designs I am able to isolate factors might be contributing to ecological patterns and processes [@fletcher_jr_landscape_2023; @resasco_contribution_2017].
 Past studies have endeavored to experimentally measure changes in community compositions as a result of connecting habitats with corridors [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002]  Yet very few have directly compared matrix and patch populations.
@@
 
-Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources [@hanski_dung_1991], interact with corridors in their landscapes.
-We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch area to edge ratio and distance to edge [@tewksbury_corridors_2002].
+Here, I aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources [@hanski_dung_1991], interact with corridors in their landscapes.
+I sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch area to edge ratio and distance to edge [@tewksbury_corridors_2002].
 To ask the questions: (1) How are species abundances of dung beetles distributed within isolated and connected patches and what are their relationships with assemblages in matrix habitat?
@@
 
-In the months of July and August 2024 dung beetles were sampled in 4 blocks spread across SRS, baited pitfall traps were placed in one of each patch type and in one matrix plot per block (Figure \@ref(fig:pitfalls)).
-Traps were placed in groups of 3 in the centers of each patch approximately 250 meters from the midpoint of the central patch 40 m from patch edge.
+Dung beetles were sampled in July and August 2024 in four of the SRS blocks (8, 52, 53n, and 54; Figure \@ref(fig:srs-map)).
+In each block sampling was conducted using baited pitfall traps placed in each patch type as well as the matrix surrounding the patches (Figure \@ref(fig:pitfalls)).
+Traps were placed in groups of 3 in the centers of each patch, approximately 250 meters from the midpoint of the central patch 40 m from patch edge.
 Pitfalls were oriented in a triangular pattern with the bottom two traps positioned towards the center patch, each trap 20 m apart.
@@
 Individual pitfall traps consisted of two components, a 10cm tall by 8 cm wide cylinder base topped with a funnel with a 10cm wide rim.
-We sourced pig feces from the University of Florida Swine Barn Unit.
+I sourced pig feces from the University of Florida Swine Barn Unit.
 Bait was processed into 5cm wide balls and wrapped in a layer of coffee filter material.
@@
 
-All dung beetles were counted and identified to species as described in @nemes_illustrated_2015-1 and @edmonds_taxonomic_2023.
+All dung beetles were counted and identified to species using @nemes_illustrated_2015 and @edmonds_taxonomic_2023.
 Fifteen individuals of each species with adequate captures were dried to equilibrium and weighed for biomass measurements.
-Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods.
+Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods upon completion of all analyses.
 
@@
 Second, by manipulating *q* we can gain an understanding of compositional shifts otherwise obscured while using species richness [@chao_rarefaction_2014].
-We compared community composition by increasing magnitudes of diversity components (i.e., *qD*) of *0D* (i.e., species richness), *1D* (i.e., Shannon entropy), and *2D* (i.e, Simpson Diversity).
+I compared community composition by increasing magnitudes of diversity components (i.e., *qD*) of *0D* (i.e., species richness), *1D* (i.e., Shannon entropy), and *2D* (i.e, Simpson Diversity).
 Diversity numbers and species richness were calculated using the package `hill` [@li_hillr_2018] for the R statistical programming language [@posit_team_rstudio_2025].
@@
 
-To test for the effects of connectivity on abundance, species richness, and species diversity we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix.
-For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution [@bolker_generalized_2009].
+To test for the effects of connectivity on abundance, species richness, and species diversity I compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix.
+For abundance and richness I used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution [@bolker_generalized_2009].
 Compared (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type.
-We included the identity of the sampling block as a random effects.
-To model our diversity metrics we took a similar approach, but this time using linear mixed effects models with a Gaussian distribution [@chao_rarefaction_2014].
+I included the identity of the sampling block as a random effects.
+To model my diversity metrics I took a similar approach, but this time using linear mixed effects models with a Gaussian distribution [@chao_rarefaction_2014].
 In all models, the reference level for patch type was Matrix.
 All models were fit using `lme4` package [@bates_fitting_2015].
-Prior to conducting our modeling we evaluated the the suitability of our data with *qqplots* generated with the `DARMa` package [@hartig_dharma_2024].
-
-We used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation.
+Prior to conducting my modeling I evaluated the the suitability of my data with *qqplots* generated with the `DARMa` package [@hartig_dharma_2024].
+
+I used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation.
 Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.
@@
   summarise(n_spp=n_distinct(species))
+
+## Abund by patch
+
+btl_sums_patch<-btl_data %>%
+  select(-sum) %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
+  group_by(patch) %>%
+  summarize(n=sum(n, na.rm=TRUE))
+
+n_corr<-btl_sums_patch %>% filter(patch=="c") %>% select(n) %>% as.numeric()
+n_mat<-btl_sums_patch %>% filter(patch=="m") %>% select(n) %>% as.numeric()
+n_rect<-btl_sums_patch %>% filter(patch=="r") %>% select(n) %>% as.numeric()
+n_wing<-btl_sums_patch %>% filter(patch=="w") %>% select(n) %>% as.numeric()
 
@@
 
-Overall, I collected N = `r total_N` dung beetles belonging to N = `r total_spp` species.
-The N = `r how_many_common` most dominant species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1).
+Overall, I collected N = `r total_N` dung beetles (N = `r n_corr` in Connected patches, N = `r n_wing` in Winged patches, N = `r n_rect` in Rectangle patches, N = `r n_mat` in the Matrix).
+These beetles belonged to N = `r total_spp` species; the N = `r how_many_common` most dominant species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1).
 All but four species were captured in every patch type.
@@
 
-\singlespacing
-
-\begin{flushleft} 
-\textbf{Paragraph 1: }Overview: While there may indeed be effects of patch area and connectivity on dung beetle abundance on community composition, other (landscape-scale) drivers appear to be more important.
-\end{flushleft}
+<!-- Paragraph 1: Overview: While there may indeed be effects of patch area and connectivity on dung beetle abundance on community composition, other (landscape-scale) drivers appear to be more important.
+-->
 
@@
 In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches.
-Our main findings emphasized: (1) Habitat type and patch shape were the main driving factors for determining how dung beetle species abundances were composed, however effects were species specific.
+My main findings emphasized: (1) Habitat type and patch shape were the main driving factors for determining how dung beetle species abundances were composed, however effects were species specific.
 (2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level.
@@
 
-\begin{flushleft} 
-\textbf{Paragraph 2: }Effects of Edge vs. Connectivity
-\end{flushleft}
-
-In our comparison of dung beetle compositions, connectivity and habitat edge accounted for differences in dung beetle abundances, yet species richness and diversity were even across patch types and in  the matrix.
+<!-- Paragraph 2: Effects of Edge vs. Connectivity -->
+
+In my comparison of dung beetle compositions, connectivity and habitat edge accounted for differences in dung beetle abundances, yet species richness and diversity were even across patch types and in  the matrix.
 Total beetle counts were consistently highest in the matrix, in comparison abundances is connected patches were not significantly different while those of the winged and rectangular patches were lower.
-Additionally, we also observed that patch effects were not equally proportional for all species.
+Additionally, I also observed that patch effects were not equally proportional for all species.
 This suggests that the species captured could be exhibiting habitat preferences between the open patches versus forested matrix.
@@
 The latter example seems more likely because flies, which have similar resource dependencies were also found to have a similar interaction [@fried_habitat_2005], and also studies on dung beetles in the tropics found more dramatic differences in dung beetle populations in fragments and matrix [@bustamante-sanchez_dung_2004; @WOS:000180797500010; @WOS:000288810500026].
-
-\begin{flushleft} 
-\textbf{Paragraph 3: }What then are the drivers of dung beetle diversity and abundance?
-\end{flushleft}
+<!-- Paragraph 3: }What then are the drivers of dung beetle diversity and abundance? -->
 
@@
 
-\begin{flushleft} 
-\textbf{Paragraph 4: }That said, there were some interspecific differences in abundance. 
-\end{flushleft}
+<!-- Paragraph 4: That said, there were some interspecific differences in abundance.  -->
 
@@
 Beetles from the genus _Aphodius_ do tend to show patterns of habitat specificity[@roslin_distribution_2001], so it is not unexpected that they might show a preference towards either forested matrix or open patches.
-We did not detect that any species was more positively associated with open patches despite expectations that some species ( eg.
+I did not detect that any species was more positively associated with open patches despite expectations that some species ( eg.
 _Canthon vigilans_, _Melanocanthon bispinatus_) would prefer open habitat [@conover_phenology_2019; @nealis_habitat_1977].
-Altogether this could be another sign pointing towards matrix acting as a source population, and since open habitat was much less dominant in our experimental system beetles could be moving into patches from habitat edge.
-
-\begin{flushleft} 
-\textbf{Paragraph 5:} Our results largely echo/constrast with previous ones from the SRS and elsewhere 
-\end{flushleft}
+Altogether this could be another sign pointing towards matrix acting as a source population, and since open habitat was much less dominant in my experimental system beetles could be moving into patches from habitat edge.
+
+<!-- Paragraph 5: Our results largely echo/constrast with previous ones from the SRS and elsewhere  -->
 
 Despite ample work documenting their ecological importance, there is a surprising lack of research on dung beetle diversity and corridors outside of the tropics[@nichols_global_2007], with any work in temperate locations coming from Europe where _Aphodius_ beetles dominate[@roslin_distribution_2001].
-Many studies in the tropics have found patterns of habitat preference between open fields and continuous forest similar to what we did in this study[@WOS:000349242000004], and some work that found that forest species were using living fences as corridors[@arellano_response_2008].
-However, differences in composition were much more distinctly different with some species being exclusive to habitat types unlike we see in this paper.
-Of the limited work out of the south eastern United States we did recover assemblages similar to other work [@conover_phenology_2019; @nealis_habitat_1977], and even returned results to another study where differences between treatment types were non-significant while variation on a larger scale was more apparent[@young_dung_2023].
-In comparison to other studies from SRS our results were similar to those of small mammals which were more likely to be using matrix land [@mabry_influence_2003].
-We did expect some species to be more dominant in patches with higher edge ratios as was found with ants at SRS [@resasco_landscape_2014], but species were evenly present in all patches.
-
-\begin{flushleft} 
-\textbf{Paragraph 6: }Caveats
-\end{flushleft}
+Many studies in the tropics have found patterns of habitat preference between open fields and continuous forest similar to what I did in this study[@WOS:000349242000004], and some work that found that forest species were using living fences as corridors[@arellano_response_2008].
+However, differences in composition were much more distinctly different with some species being exclusive to habitat types unlike I see in this paper.
+Of the limited work out of the south eastern United States I did recover assemblages similar to other work [@conover_phenology_2019; @nealis_habitat_1977], and even returned results to another study where differences between treatment types were non-significant while variation on a larger scale was more apparent[@young_dung_2023].
+In comparison to other studies from SRS my results were similar to those of small mammals which were more likely to be using matrix land [@mabry_influence_2003].
+I did expect some species to be more dominant in patches with higher edge ratios as was found with ants at SRS [@resasco_landscape_2014], but species were evenly present in all patches.
+
+<!-- Paragraph 6: Caveats -->
 
@@
 Another potential avenue for improvement is lowering the grain size of sampling and changing trap placement to better understand the effects of edge proximity and connectivity.
-We also used one of the best bait types for collecting dung beetles in pig dung [@marsh_optimising_2013], but a mix of differently sourced baits may have been more optimal as more diverse baits attract more diverse species [@frank_nutrient_2017; @frank_search_2018; @gimenez_gomez_dung_2021].
-
-\begin{flushleft} 
-\textbf{Paragraph 7: }Moving forward
-\end{flushleft}
+I also used one of the best bait types for collecting dung beetles in pig dung [@marsh_optimising_2013], but a mix of differently sourced baits may have been more optimal as more diverse baits attract more diverse species [@frank_nutrient_2017; @frank_search_2018; @gimenez_gomez_dung_2021].
+
+<!-- Paragraph 7: Moving forward -->
 
@@
 SRS provides an excellent experimental design for direct comparison of landscapes but another main goal of the site is to study dispersal, dung beetles would be an excellent system for studying movement.
-They already show promise, in preliminary trials we released beetles from both the connected and rectangular patches with baited traps in the central patch.
-We recovered one recapture which originated from the connected patch, indicating that individuals do move through corridors.
-
-\begin{flushleft} 
-\textbf{Paragraph 8: } Close with takeaways for conservation and management
-\end{flushleft}
+They already show promise, in preliminary trials I released beetles from both the connected and rectangular patches with baited traps in the central patch.
+I recovered one recapture which originated from the connected patch, indicating that individuals do move through corridors.
+
+<!-- Paragraph 8: Close with takeaways for conservation and management -->
 
@@
 
-```{r abundance-patch, echo=FALSE,message = FALSE,warning=FALSE}
-# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
-
-btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
-spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
-
-btl_sums_patch<-btl_data %>%
-  select(-sum) %>%
-  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
-  group_by(patch) %>%
-  summarize(n=sum(n, na.rm=TRUE)) %>%
-  arrange(patch,desc(n)) %>%
-  mutate(patch=case_when(
-    patch == "c" ~ "Corridor",
-    patch == "m" ~ "Matrix",
-    patch == "w" ~ "Winged",
-    patch == "r" ~ "Rectangle",
-    .default = as.character(patch)))
-
-abund_patch <- knitr::kable(btl_sums_patch,
-      digits = 2,
-      align="lr",
-      format="latex",
-      row.names = FALSE,
-      booktabs=T,
-      linesep = "", #removes the blank line after every 5 lines
-      caption = "Total dung beetles captured in all replicates of a patch type.") %>%
-  kable_styling(bootstrap_options = c("striped"),
-                full_width = F,
-                latex_options=c("scale_down","hold_position"),
-                font_size = 12,
-                position = "center") %>%
-  # column_spec(1, italic = T) %>%
-  row_spec(0, bold = T)
-  # row_spec(1:15, hline_after = T)
-abund_patch
-
-```
+<!-- ```{r abundance-patch, echo=FALSE,message = FALSE,warning=FALSE} -->
+<!-- # https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html -->
+
+<!-- btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv")) -->
+<!-- spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv")) -->
+
+<!-- btl_sums_patch<-btl_data %>% -->
+<!--   select(-sum) %>% -->
+<!--   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% -->
+<!--   group_by(patch) %>% -->
+<!--   summarize(n=sum(n, na.rm=TRUE)) %>% -->
+<!--   arrange(patch,desc(n)) %>% -->
+<!--   mutate(patch=case_when( -->
+<!--     patch == "c" ~ "Corridor", -->
+<!--     patch == "m" ~ "Matrix", -->
+<!--     patch == "w" ~ "Winged", -->
+<!--     patch == "r" ~ "Rectangle", -->
+<!--     .default = as.character(patch))) -->
+
+<!-- # if wanted to make a bar graph -->
+<!-- # ggplot(btl_sums_patch, aes(x=patch, y=n)) +  -->
+<!-- #   geom_bar(stat = "identity") -->
+<!-- #  -->
+<!-- abund_patch <- knitr::kable(btl_sums_patch, -->
+<!--       digits = 2, -->
+<!--       align="lr", -->
+<!--       format="latex", -->
+<!--       row.names = FALSE, -->
+<!--       booktabs=T, -->
+<!--       linesep = "", #removes the blank line after every 5 lines -->
+<!--       caption = "Total dung beetles captured in all replicates of a patch type.") %>% -->
+<!--   kable_styling(bootstrap_options = c("striped"), -->
+<!--                 full_width = F, -->
+<!--                 latex_options=c("scale_down","hold_position"), -->
+<!--                 font_size = 12, -->
+<!--                 position = "center") %>% -->
+<!--   # column_spec(1, italic = T) %>% -->
+<!--   row_spec(0, bold = T) -->
+<!--   # row_spec(1:15, hline_after = T) -->
+<!-- abund_patch -->
+
+<!-- ``` -->
 
@@
 taxa_bray <- vegan::vegdist(btl_spec_patch, method = "bray", binary = FALSE)
+
 # taxa_bray
-
+tb<-as.matrix(taxa_bray)
+tb[lower.tri(tb)] <- 0
+tb<-as.data.frame(tb)
+tb<-tb %>% 
+  mutate_if(is.numeric, round, digits=3) %>% 
+  select(-Corridor)
+
+row.names.remove <- c("Winged")
+tb<-tb[!(row.names(tb) %in% row.names.remove), ]
+tb[tb == 0] <- ""
+col1<-row.names(tb)
+col1<-as.data.frame(col1)
+
+tb<-bind_cols(col1,tb) %>% 
+  rename(" "=col1)
 # bray<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/bray_patches.csv") 
-bray<-read_csv(here("corridor_docs","eric_ms_thesis","tables","bray_patches.csv")) %>% kable(
+# bray<-read_csv(here("corridor_docs","eric_ms_thesis","tables","bray_patches.csv")) %>% 
+  bray<-tb %>% 
+  kable(
       digits = 3,
@@
   # column_spec(1, italic = T) %>%
-  row_spec(0, bold = T)
+  row_spec(0, bold = T) %>% 
+  column_spec(1, bold = T)
   # row_spec(1:15, hline_after = T)
@@
   replace_na(list(group=" ")) %>% 
+  rename(SE = std.error,
+         Estimate=estimate,
+         Term=term,
+         Group=group,
+         Effect=effect,
+         Statistic=statistic,
+         P=p.value) %>% 
+  mutate(Effect=if_else(Effect=="ran_pars","Random",Effect)) %>% 
+  mutate(Effect=if_else(Effect=="fixed","Fixed",Effect)) %>% 
+  mutate(Group=if_else(Group=="block","Block",Group)) %>% 
+  mutate(Term=gsub("patch_type","patch type - ",Term)) %>% 
+  mutate(Term=gsub("cor__","correlation: ",Term)) %>% 
+  mutate(Term=gsub("sd__","SD: ",Term)) %>% 
+  mutate(Term=gsub("[.]"," x ",Term)) %>% 
+  mutate(SE=round(SE,3)) %>% 
+  mutate(Estimate=round(Estimate,3)) %>% 
+  mutate(Statistic=round(Statistic,3)) %>% 
+  mutate(P=round(P,3)) %>% 
+  mutate_all(as.character) %>% 
+  mutate(P=if_else(as.numeric(P)<0.001,"< 0.001",P)) %>% 
+  replace_na(list(SE="",Statistic="",P="")) %>%
+
   # need to round numbers above to 4, convert to character, then na_replace
@@
     .default = as.character(term))) |>
+  mutate(p.value=as.character(p.value)) %>% 
+  mutate(p.value=if_else(as.numeric(p.value)<0.001,"< 0.001",p.value)) %>% 
+  rename(P=p.value,
+         Term=term,
+         Statistic=statistic) %>% 
 # aovM3<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM3.csv") %>% 
@@
   replace_na(list(group=" ")) %>% 
+  # mutate(p.value=as.character(p.value)) %>% 
+  # mutate(p.value=if_else(as.numeric(p.value)<0.001,"< 0.001",p.value)) %>% 
+  rename(P=p.value,
+         SE = std.error,
+         Estimate=estimate,
+         Term=term,
+         Group=group,
+         Effect=effect,
+         Statistic=statistic) %>%
+
+  mutate(Effect=if_else(Effect=="ran_pars","Random",Effect)) %>% 
+  mutate(Effect=if_else(Effect=="fixed","Fixed",Effect)) %>% 
+  mutate(Group=if_else(Group=="block","Block",Group)) %>% 
+  mutate(Term=gsub("patch_type","patch type - ",Term)) %>% 
+  mutate(Term=gsub("cor__","correlation: ",Term)) %>% 
+  mutate(Term=gsub("sd__","SD: ",Term)) %>% 
+  mutate(Term=gsub("sp_code","Species - ",Term)) %>% 
+  mutate(Term=gsub("[.]"," x ",Term)) %>% 
+  mutate(Term=gsub("[:]"," : ",Term)) %>% 
+  mutate(SE=round(SE,3)) %>% 
+  mutate(Estimate=round(Estimate,3)) %>% 
+  mutate(Statistic=round(Statistic,3)) %>% 
+  mutate(P=round(P,3)) %>% 
+  mutate_all(as.character) %>% 
+  mutate(P=if_else(as.numeric(P)<0.001,"< 0.001",P)) %>% 
+  replace_na(list(SE="",Statistic="",P="")) %>%
+
   # need to round numbers above to 4, convert to character, then na_replace
@@
    mutate(term=case_when(
-    term == "patch_typeMatrix" ~ "Patch Type: Matrix",
-    term == "patch_typeRectangle" ~ "Patch Type: Rectangle",
-    term == "patch_typeWinged" ~ "Patch Type: Winged",
+    term == "patch_typeMatrix" ~ "Patch Type - Matrix",
+    term == "patch_typeRectangle" ~ "Patch Type - Rectangle",
+    term == "patch_typeWinged" ~ "Patch Type - Winged",
+    term == "patch_typeConnected" ~ "Patch Type - Connected",
     .default = as.character(term))) |>
+  rename(P=p.value,
+         SE = std.error,
+         Estimate=estimate,
+         Term=term,
+         Effect=effect,
+         Statistic=statistic) %>%
+
+  mutate(Effect=if_else(Effect=="ran_pars","Random",Effect)) %>% 
+  mutate(Effect=if_else(Effect=="fixed","Fixed",Effect)) %>%
+
+  mutate(SE=round(SE,3)) %>% 
+  mutate(Estimate=round(Estimate,3)) %>% 
+  mutate(Statistic=round(Statistic,3)) %>% 
+  mutate(P=round(P,3)) %>% 
+  mutate_all(as.character) %>% 
+  mutate(P=if_else(as.numeric(P)<0.001,"< 0.001",P)) %>% 
+  replace_na(list(SE="",Statistic="",P="")) %>%
+
 # M_rich<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_rich.csv") %>% 
@@
   replace_na(list(group=" ")) %>%
+
+  rename(SE = std.error,
+         Estimate=estimate,
+         Term=term,
+         Group=group,
+         Effect=effect,
+         Statistic=statistic) %>%
+
+  mutate(Effect=if_else(Effect=="ran_pars","Random",Effect)) %>% 
+  mutate(Effect=if_else(Effect=="fixed","Fixed",Effect)) %>% 
+  mutate(Group=if_else(Group=="block","Block",Group)) %>%
+
+  mutate(Term=gsub("patch_type","patch type - ",Term)) %>% 
+  mutate(Term=gsub("cor__","correlation: ",Term)) %>% 
+  mutate(Term=gsub("sd__","SD: ",Term)) %>% 
+  mutate(SE=round(SE,3)) %>% 
+  mutate(Estimate=round(Estimate,3)) %>% 
+  mutate(Statistic=round(Statistic,3)) %>%
+
+  mutate_all(as.character) %>%
+
+  replace_na(list(SE="",Statistic="")) %>%
+
   # need to round numbers above to 4, convert to character, then na_replace
@@
   replace_na(list(group=" ")) %>% 
+  rename(SE = std.error,
+         Estimate=estimate,
+         Term=term,
+         Group=group,
+         Effect=effect,
+         Statistic=statistic) %>%
+
+  mutate(Effect=if_else(Effect=="ran_pars","Random",Effect)) %>% 
+  mutate(Effect=if_else(Effect=="fixed","Fixed",Effect)) %>% 
+  mutate(Group=if_else(Group=="block","Block",Group)) %>%
+
+  mutate(Term=gsub("patch_type","patch type - ",Term)) %>% 
+  mutate(Term=gsub("cor__","correlation: ",Term)) %>% 
+  mutate(Term=gsub("sd__","SD: ",Term)) %>% 
+  mutate(SE=round(SE,3)) %>% 
+  mutate(Estimate=round(Estimate,3)) %>% 
+  mutate(Statistic=round(Statistic,3)) %>%
+
+  mutate_all(as.character) %>%
+
+  replace_na(list(SE="",Statistic="")) %>%
+
   # need to round numbers above to 4, convert to character, then na_replace
@@
 He later began to explore this curiosity deeper in beginning his graduate education at the University of Florida as a Master's Student under the supervision of Emilio Bruna.
-
-\newpage
-
-## Text Sandbox
-
-As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented.
-This degradation lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].
-Effects from isolation can vary, however as habitats are broken down community structures are significantly altered [@WOS:000419965700012].
-Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003].
-As disturbance continues to intensify, it is becoming increasingly more important to understand how different taxonomic groups.
-Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes.
-
-Here, we aim to determine how connectivity and fragmentation affect Species Richness and Diversity, Abundance, and functional diversity.
-We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge.
-To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors.
-Additionally we asked (2) Are corridors benefiting any one functional trait over another?
-Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling.
-
-Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services [@hevia_trait-based_2017].
-[**EB edit:** these last sentences don't really make the case for experiments, whcih were the initial emphasis of the paragraph.
-Might want to opt for another option, like the ones in the green notes below]
-
-This study advances our understanding of the factors shaping dung beetle community composition in subtropical regions of the southeastern United States.
-In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches.
-Our main findings emphasized: (1) Habitat type and patch shape were the main driving factors for dung beetle species abundances were composed, however effects were species specific.
-(2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level.
-(3) Species diversity metrics were also relatively even across patch types however varied widely by sampling blocks.
-These results suggest that dung beetle species are fully capable of permeating fragmented landscapes and that habitat type and connectivity shape community compositions, but landscape effects of a larger scale are driving changes in biodiversity.
-
-### Abundance
-
-The total number of dung beetle species captured N = `r total_spp` and the total individuals N = `r total_N` are similar to those in studies conducted in similar habitats and regions of the US [@conover_phenology_2019; @nealis_habitat_1977; @stanbrook_dung_2022].
-
-Dung beetle abundances were significantly lowest in rectangular patches.
-Matrix patches had the most total beetles collected and consistently had the highest counts for the most dominant species.
-However abundances were not significantly different between matrix, connected, and winged patches.
-This may suggest that isolation is leading to a decrease in abundances, but it is more likely that populations in the matrix are acting as a source, and since connected and winged patches have higher edge to area ratios dung beetles were more likely to move into those patches.
-Past studies have shown trends where habitat type and forest regeneration stages are key in the partitioning of dung beetle assemblages[@conover_phenology_2019; @WOS:000386714700011; @arellano_response_2008] which supports why we might see these differences in local populations.
-In addition distance from habitat edge has been attributed to differences in spillover from source populations[@WOS:000707546700001], emphasizing the importance of habitat edge in the context of our study.
-
-We also observed that patch effects were not equally proportional for all species.
-The abundances of all dominant species were positively influenced by matrix habitat, but *Aphoidius alloblackburneus* responded more positively than any other species.
-*Phanaeus igneus* and *Ateuchus lecontei* also had higher positive effects in rectangle patches.
-
-### Richness
-
-While abundances were different between patch types, we did not detect any patterns of species richness in our modeling.
-Total species counts were very even across patch types and sampling blocks.
-This difference indicates that even though habitat type or landscape features are informing occupational preferences, land use is more or less the same between patches and matrix.
-This differs from past work where species richness is lower in forest fragments [@WOS:000178929600002]
-
-### Diversity
-
-Analysis of Hill numbers indicated that species compositions were even between patch types.
-Comparing values of Simpson's Diversity between patches determined that dominant species were not prevalent in any one treatment type.
-Likewise, the Shannon index values insisted that rare species throughout the study population were even.
-However, for both metrics, values varied greatly by sampling blocks.
-Similar to our richness results this suggests that variation between patch and matrix is not distinct enough to limit the land use of dung beetles within our study site.
-
-In contrast to what we observed with species richness it appears that assemblages larger landscape patterns are effecting the composition of assemblages across the study site.
-
-*here write about low sample sizes and reasons for why dung beetle communities would be so similar across treatment types.
-Dispersal, scale (roslin), maybe again a good reason to expect matrix as a source pop.*
-
-notes that are important for discussion
-
--   habitat preference and seasonality (conover 2019, nealis 1977)
-> -   cvig - open sand chaparral, open grassland preference nealis 1977
-> -   mbis - open sand chapparal preference nealis 1977
-> -   open - slight preference towards open chap and shaded chap, little open grassland nealis 1977
-> -   pign - sharing forest habitat but need more info conover 2019
-> -   pvin - open habitat but share forest with pign conover 2019
-> -   alec - forest preference conover 2019
-
--   habitat again but only part of variation form habitat specifically aphodius also phenology (roslin 2001)
--   spatial scales (roslin 2000)
--   morphology
-
-> -   Ospina 2018: wing shape py habitat preference and differences in wing shape determined by species groupings.
-Beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies).
-Large body beetles tending towards lower energy flight strategies.
-maybe justification for corridor preference?
-> -   stanbrook and king 2022: tunnelers preferring open habitat and tunnelers also contributing more towards dung removal.
-> -   Conover 2019: dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types
-> -   Gimenez Gomez et al 2021 -\> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used.
-However this study was done in a more tropical ecosystem so its hard to say if the same would apply to our more temperate system so more work is needed in this specific avenue.
-
--   functionality especially gas emissions (slade 2016)
--   species id and funcitonality (slade 2017)
--   faovored species dominating in fragments (resasco 2014)
--   wind direction in corridors (damschen 2014)
-
-revisit bray curtis
-
-1.  Abundances in matrix vs connected patch and why this could be happening
-
--   source pop to habiitat edge
-
-2.  species richness again supporting that matrix is more of an ideal habitat for dung beetle community
-
-3.  diversity indices community structures weren't highly different between patch types
-
-4.  bray curtis hinting at similar land uses between corridor and winged patch
-
--   like julians paper corridors benefit certain populations and more fit populations are able to make better use
-
-\begin{flushleft}  
-\textbf{NEED}
-\end{flushleft}
-
-1. rarefaction curves  
-1. add biomass to table 1  
-1. add totals of habitat in table 2 to table 1.

### 2025-07-08 corridor_docs/eric_ms_thesis/text_sandbox.Rmd
update for sharing with cmmte

---
title: "Text Sandbox"
author: "ERIC ESCOBAR-CHENA"
# date: "\\vspace{-1em}"
# header-includes:
 # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
#  - \usepackage{lineno}
#  - \linenumbers
#  - \usepackage[default]{sourcesanspro}
#  - \usepackage{fancyhdr} # modify header into left, center, right
#  - \pagestyle{fancy} # cues the use of fancy headers
#  - \fancyfoot{} # cues the use of fancy footer
#  - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
#  - \fancyhead[R]{p. \thepage} # name and page on header right
#  - \usepackage{subcaption}
#  - \usepackage{float}
# # # - \fancyfoot[R]{ }
#  - \usepackage{setspace}

header-includes:
  - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
  - \usepackage{float} #to keep figure in the location placed instead of end
  - \floatplacement{figure}{H}
  - \usepackage{booktabs}
  - \raggedbottom
  - \usepackage{fancyhdr}
  - \pagestyle{plain}
  - \fancyfoot[C]{\thepage}
  - \fancyhead[L]{}
  - \fancyhead[R]{\small{EEC MS Thesis Draft}}
  - \thispagestyle{fancy}
  - \usepackage{tabu}
  - \usepackage{makecell}
  - \usepackage{pdflscape}
  - \usepackage{longtable}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \setlength\parindent{22pt}
  - \usepackage{indentfirst}
  - \renewenvironment{abstract}{}{} #suppress the abstract
  - \usepackage{setspace}
  - \setcounter{secnumdepth}{0} # removes numbering from sections and subsections
  - \usepackage{sectsty}\allsectionsfont{\raggedright}
  - \usepackage{sectsty}\sectionfont{\centering}{\normalfont} # center # headings
  - \usepackage{sectsty}\subsectionfont{\centering}{\normalfont} # center ## headings
  - \usepackage{sectsty}\subsubsectionfont{\bfseries} # center ## headings
  # - \usepackage{sectsty}\subsubsectionfont{\itshape} # no centering but italic subsu
  # - \usepackage{sectsty}\subsubsectionfont{\centering}{ital} # center ### headings
  # - \usepackage{sectsty}\sectionfont{\centering\color{darkmidnightblue}}

 # - doublespacing # manipulate line spacing 
# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
# - \usepackage{xcolor} # changes colors of text
# # http://latexcolor.com/ to choose colors
# # define the colors used above with rgb codes
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# # to center to top level (#) headings (my name) and change color of font
# # more info # https://www.ctan.org/pkg/sectsty
# - \usepackage{sectsty}
# - \sectionfont{\centering\color{darkmidnightblue}}
# # to center to sub-sub-headings (###) and change color of font
# - \usepackage{sectsty} 
# - \subsectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} 
# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
# - \usepackage{wrapfig} 
# - \newcommand{\blandscape}{\begin{landscape}}
# - \newcommand{\elandscape}{\end{landscape}}
# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
# - \usepackage{indentfirst}
# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
bibliography: ["eric-thesis.bib"]
output: bookdown::pdf_document2
  # pdf_document:
  #   includes:
  #     # in_header: style/preamble.tex
  #     template: null
  #     keep_tex: no
  #     extra_dependencies: ["float"]
      # number_sections: true
# documentclass: style/nsf2
biblio-style: apalike
csl: ecology.csl
---

## Text Sandbox

As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented. This degradation lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015]. Effects from isolation can vary, however as habitats are broken down community structures are significantly altered [@WOS:000419965700012]. Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes with the goal of minimizing negative consequences of fragmentation[@haddad_corridor_2003]. As disturbance continues to intensify, it is becoming increasingly more important to understand how different taxonomic groups. Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes.

Here, we aim to determine how connectivity and fragmentation affect Species Richness and Diversity, Abundance, and functional diversity. We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge. To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors. Additionally we asked (2) Are corridors benefiting any one functional trait over another? Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling.

Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services [@hevia_trait-based_2017]. [**EB edit:** these last sentences don't really make the case for experiments, whcih were the initial emphasis of the paragraph. Might want to opt for another option, like the ones in the green notes below]

This study advances our understanding of the factors shaping dung beetle community composition in subtropical regions of the southeastern United States. In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches. Our main findings emphasized: (1) Habitat type and patch shape were the main driving factors for dung beetle species abundances were composed, however effects were species specific. (2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level. (3) Species diversity metrics were also relatively even across patch types however varied widely by sampling blocks. These results suggest that dung beetle species are fully capable of permeating fragmented landscapes and that habitat type and connectivity shape community compositions, but landscape effects of a larger scale are driving changes in biodiversity.

### Abundance

Dung beetle abundances were significantly lowest in rectangular patches. Matrix patches had the most total beetles collected and consistently had the highest counts for the most dominant species. However abundances were not significantly different between matrix, connected, and winged patches. This may suggest that isolation is leading to a decrease in abundances, but it is more likely that populations in the matrix are acting as a source, and since connected and winged patches have higher edge to area ratios dung beetles were more likely to move into those patches. Past studies have shown trends where habitat type and forest regeneration stages are key in the partitioning of dung beetle assemblages[@conover_phenology_2019; @WOS:000386714700011; @arellano_response_2008] which supports why we might see these differences in local populations. In addition distance from habitat edge has been attributed to differences in spillover from source populations[@WOS:000707546700001], emphasizing the importance of habitat edge in the context of our study.

We also observed that patch effects were not equally proportional for all species. The abundances of all dominant species were positively influenced by matrix habitat, but *Aphoidius alloblackburneus* responded more positively than any other species. *Phanaeus igneus* and *Ateuchus lecontei* also had higher positive effects in rectangle patches.

### Richness

While abundances were different between patch types, we did not detect any patterns of species richness in our modeling. Total species counts were very even across patch types and sampling blocks. This difference indicates that even though habitat type or landscape features are informing occupational preferences, land use is more or less the same between patches and matrix. This differs from past work where species richness is lower in forest fragments [@WOS:000178929600002]

### Diversity

Analysis of Hill numbers indicated that species compositions were even between patch types. Comparing values of Simpson's Diversity between patches determined that dominant species were not prevalent in any one treatment type. Likewise, the Shannon index values insisted that rare species throughout the study population were even. However, for both metrics, values varied greatly by sampling blocks. Similar to our richness results this suggests that variation between patch and matrix is not distinct enough to limit the land use of dung beetles within our study site.

In contrast to what we observed with species richness it appears that assemblages larger landscape patterns are effecting the composition of assemblages across the study site.

*here write about low sample sizes and reasons for why dung beetle communities would be so similar across treatment types. Dispersal, scale (roslin), maybe again a good reason to expect matrix as a source pop.*

notes that are important for discussion

-   habitat preference and seasonality (conover 2019, nealis 1977)
> -   cvig - open sand chaparral, open grassland preference nealis 1977
> -   mbis - open sand chapparal preference nealis 1977
> -   open - slight preference towards open chap and shaded chap, little open grassland nealis 1977
> -   pign - sharing forest habitat but need more info conover 2019
> -   pvin - open habitat but share forest with pign conover 2019
> -   alec - forest preference conover 2019

-   habitat again but only part of variation form habitat specifically aphodius also phenology (roslin 2001)
-   spatial scales (roslin 2000)
-   morphology

> -   Ospina 2018: wing shape py habitat preference and differences in wing shape determined by species groupings. Beetles with wider wings showed more preference towards open habitat, contrary to what is already known in literature on other taxa (butterflies). Large body beetles tending towards lower energy flight strategies. maybe justification for corridor preference?
> -   stanbrook and king 2022: tunnelers preferring open habitat and tunnelers also contributing more towards dung removal.
> -   Conover 2019: dung source being less of an issue than habitat type but some species were more responsive to specific bait types so future studies should use a mix of multiple bait types
> -   Gimenez Gomez et al 2021 -\> similar outcome some beetles were extra sensitive to specififc baits so a mix of baits should be used. However this study was done in a more tropical ecosystem so its hard to say if the same would apply to our more temperate system so more work is needed in this specific avenue.

-   functionality especially gas emissions (slade 2016)
-   species id and funcitonality (slade 2017)
-   faovored species dominating in fragments (resasco 2014)
-   wind direction in corridors (damschen 2014)

revisit bray curtis

1.  Abundances in matrix vs connected patch and why this could be happening

-   source pop to habiitat edge

2.  species richness again supporting that matrix is more of an ideal habitat for dung beetle community

3.  diversity indices community structures weren't highly different between patch types

4.  bray curtis hinting at similar land uses between corridor and winged patch

-   like julians paper corridors benefit certain populations and more fit populations are able to make better use

\begin{flushleft}
\textbf{NEED}
\end{flushleft}

1. rarefaction curves
1. add biomass to table 1
1. add totals of habitat in table 2 to table 1.

### 2025-07-09 corridor_docs/eric_ms_thesis/03_biomass_analysis.R
updates, corrected aov0 for total n

@@ -18,9 +18,11 @@ spp_bmass<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","
             n_btls=n_distinct(weight)) %>% 
   arrange(desc(avg_bmass))
 
-biomass<-btl_data %>%
+
+
+biomass_per_spp<-btl_data %>%
   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
-  group_by(species,patch,block) %>%
+  group_by(species) %>%
   mutate(patch=case_when(
     patch == "c" ~ "Connected",
     patch == "m" ~ "Matrix",
@@ -28,11 +30,80 @@ biomass<-btl_data %>%
     patch == "r" ~ "Rectangle",
     .default = as.character(patch))) %>%
   summarize(n=sum(n, na.rm=TRUE)) %>%
-  arrange(patch,desc(n)) %>%
+  arrange(desc(n)) %>%
   rename(sp_code=species) %>% 
   left_join(spp_bmass,by="sp_code") %>% 
   select(-sd, -n_btls) %>% 
   mutate(spp_biomass=n*avg_bmass)
+
+biomass_per_spp<-biomass_per_spp %>% 
+  drop_na() %>% 
+  ungroup() %>% 
+  rename(species=sp_code)
+
+ggplot(biomass_per_spp, 
+       aes(
+         x=reorder(species,-spp_biomass), 
+         y=spp_biomass
+         )
+       ) + 
+  geom_bar(stat = "identity")
+
+
+biomass_2<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
+  group_by(species, patch,block) %>%
+  # group_by(block) %>%
+  # group_by(patch,block) %>%
+  mutate(patch=case_when(
+    patch == "c" ~ "Connected",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  rename(sp_code=species) %>% 
+  left_join(spp_bmass,by="sp_code") %>%
+  # select(-n.y) %>% 
+  # rename(n=n.x) %>% 
+  # select(-sd, -n_btls) %>% 
[…]

### 2025-07-09 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -197,10 +197,35 @@ ggplot(spp_abund_top6,
 
 
 
+# comparing total abundance by patches ------------------------------------
+
+total_abund_each_patch<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
+  mutate(patch=case_when(
+    patch == "c" ~ "Connected",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) %>%
+  group_by(patch,block) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(patch,desc(n)) %>%
+  # rename(sp_code=species) %>%
+  mutate(patch_block=paste(patch,block,sep="_")) %>% 
+  ungroup() %>% 
+  select(-patch,-block) %>% 
+  separate(patch_block,c("patch_type","block"),sep="_",remove=TRUE) 
+
+#releveling
+total_abund_each_patch<-total_abund_each_patch %>% 
+  mutate(patch_type=as.factor(patch_type)) 
+total_abund_each_patch$patch_type<-relevel(total_abund_each_patch$patch_type,"Matrix")
+
+
 
 # Global model: Take 1 abundance, indep of species ID
 M0 <- glmer(n ~ patch_type + (1 + patch_type | block), 
-            data = spp_abund, 
+            data = total_abund_each_patch, 
             family = poisson)
 
 summary(M0)
@@ -256,6 +281,9 @@ plot_model(M2, type = "pred")
 
 
 M2 %>% broom::tidy()
+
+
+# species by patch type ---------------------------------------------------
 
 
 # Model 3

### 2025-07-09 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
updates

@@
   - \usepackage{indentfirst}
+  # - \usepackage{caption}
+  # - \captionsetup[figure]{name={Figure 1-},labelsep=period}
+  # - \captionsetup[table]{name={Table 1-},labelsep=period}
   - \renewenvironment{abstract}{}{} #suppress the abstract
@@
   - \usepackage{sectsty}\allsectionsfont{\raggedright}
-  - \usepackage{sectsty}\sectionfont{\centering}{\normalfont} # center # headings
-  - \usepackage{sectsty}\subsectionfont{\centering}{\normalfont} # center ## headings
+  # - \usepackage{sectsty}\sectionfont{\centering}{\normalfont} # center # headings
+  # - \usepackage{sectsty}\subsectionfont{\centering}{\normalfont} # center ## headings
   - \usepackage{sectsty}\subsubsectionfont{\bfseries} # center ## headings
@@
   bookdown::pdf_document2:
-    toc: false
+    # toc: false
+    number_sections: true
+    fig_caption: true
+
+# output: bookdown::word_document2
+# output: word_document
+# output: html_document
+
 # output: bookdown::pdf_document2
@@
 
-## INTRODUCTION
-
 As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented [@diaz_pervasive_2019; @ma_global_2023; @taubert_global_2018].
@@
 
-## METHODS
-
+\begin{center} 
+\textbf{Methods}
+\end{center}
 ### Study site
@@
 
-## RESULTS
+\begin{center} 
+\textbf{Results}
+\end{center}
 
@@
 
-## DISCUSSION
-
+\begin{center} 
+\textbf{Discussion}
+\end{center}
 <!-- Paragraph 1: Overview: While there may indeed be effects of patch area and connectivity on dung beetle abundance on community composition, other (landscape-scale) drivers appear to be more important.
@@
 
-## REFERENCES
+\begin{center} 
+\textbf{References}
+\end{center}
 
@@
 
-## FIGURES & TABLES
-
 <!-- (text - Nathan Fig of plots) -->
@@
 
-knitr::include_graphics("images/sp_abund.png")
+knitr::include_graphics("images/total_abund.png")
 

### 2025-07-09 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -235,7 +235,7 @@ aovM0<-Anova(M0)
 plot_model(M0, type = "pred")
 
 
-
+ggsave("./corridor_docs/eric_ms_thesis/images/total_abund.png",width = 10, height = 10, units = "cm")
 # EMILIO ADDITION - EXPORTING TABLE RESULTS TO UPLOAD TO THESIS DOC
 
 # REQUIRES CLEANUP, BUT HERE IS HOW TO GET THE STATS TABLES IN KABLEEXTRA FOR YOUR THESIS

### 2025-07-10 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
some more fleshing out

@@
 Another potential explanation is that species in the matrix are acting as a source population which feeds into patches with edge acting as a drift fence for directing movement.
-The latter example seems more likely because flies, which have similar resource dependencies were also found to have a similar interaction [@fried_habitat_2005], and also studies on dung beetles in the tropics found more dramatic differences in dung beetle populations in fragments and matrix [@bustamante-sanchez_dung_2004; @WOS:000180797500010; @WOS:000288810500026].
+The latter example seems more likely because flies, which have similar resource dependencies were also found to have a similar interaction [@fried_habitat_2005], and also studies on dung beetles in the tropics found more dramatic differences in dung beetle populations in fragments and matrix [@WOS:000288810500026].
 <!-- Paragraph 3: }What then are the drivers of dung beetle diversity and abundance? -->
@@
 While the design of the experimental blocks control for as much as possible, landscape features across the sight are different enough to influence species composition.
-Several factors may explain this pattern, including soil quality which influences where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019] as well as matrix permeability, canopy cover, and different land-use types (e.g., roadways, water sources)[@WOS:000241362400032; @WOS:000651446900001].
+Several factors may explain this pattern, at the time this study was conducted block 8 had the densest matrix of any of the other sampling blocks which could have hindered the scent permeability of our bait.
+Other habitat characteristics might have been influential as well, including soil quality and forest cover which can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019]
 Not only do these impact dung beetle movement, but also these landscape features can affect mammal movement, which in turn limits dung availability for beetles.
+In this regard it is also important to consider other geographical differences, block 8 was nearest to roadways used by all employees at SRS and the eastern side of the sight had large bodies of water.
+Both of which would effect the land-use by larger mammals[@WOS:000241362400032; Dechen Quinn et al 2013; @WOS:000651446900001].
+Additionally, deer and feral hogs are the most dominant mammals at the site, two species that do well in patchy landscapes(Castillo-Contreras et al.
+2018, Fraser et al.
+2019).
+This might explain the lack of variation in richness or diversity in patches while differences are more apparent at greater distances across the site.
 
@@
 _Aphodius alloblackburneus_ had a disproportionate positive effect to being captured in matrix patches as compared to other species.
-Beetles from the genus _Aphodius_ do tend to show patterns of habitat specificity[@roslin_distribution_2001], so it is not unexpected that they might show a preference towards either forested matrix or open patches.
+Beetles from the genus _Aphodius_ do tend to show patterns of habitat specificity[@roslin_distribution_2001], and many preferring forested habitat*frank*,  so it is not unexpected that they might show a preference towards forested matrix.
 I did not detect that any species was more positively associated with open patches despite expectations that some species ( eg.
@@
 \newpage
-```{r, label = patch, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle abundance in three different treatments and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+```{r, label = patch, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average total dung beetle abundance by patch type with standard deviations as error bars.", out.width="50%", fig.align='center'}
 
@@
 
-```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle abundance of the top 6 most abundant species in three different treatments and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average abundance of the top 6 most abundant species by patch type.", out.width="50%", fig.align='center'}
 
@@
 
-```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle secies richness in three different treatments and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle secies richness by patch type.", out.width="50%", fig.align='center'}
 
@@
 
-```{r, label = shannon, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Shannon diversity in three different treatments and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+```{r, label = shannon, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Shannon diversity by patch type.", out.width="50%", fig.align='center'}
 
@@
 
-```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index in three different treatments and the forest matrix surrounding patches.", out.width="50%", fig.align='center'}
+```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index y patch type.", out.width="50%", fig.align='center'}
 

### 2025-07-10 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
update and formatting edits

@@
   - \usepackage{longtable}
+  - \usepackage{lscape}
   - \newcommand{\blandscape}{\begin{landscape}}
@@
   bookdown::pdf_document2:
-    # toc: false
+    toc: false
     number_sections: true
     fig_caption: true
-
+classoption: 12pt
 # output: bookdown::word_document2
@@
 I sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch area to edge ratio and distance to edge [@tewksbury_corridors_2002].
-To ask the questions: (1) How are species abundances of dung beetles distributed within isolated and connected patches and what are their relationships with assemblages in matrix habitat?
-(2) Does species richness vary when movement between patches is facilitated by movement corridors?
-(3) Are there shifts in species diversity/composition and what are the implications for ecosystem service functionality?
+I do so I address these  questions:
+
+(1) How does the abundance of dung beetles differ among and between isolated and connected patches and how does this compare with abundance in matrix habitat?
+(2) Does species richness and diversity differ between among connected patches, unconnected patches, and the matrix habitat?
+(3) What are the implications of changes in community composition for ecosystem services and function?
 
@@
 
-Our study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, USA (33.208$^\circ$ N, 81.408$^\circ$ W, Figure \@ref(fig:srs-map)).
+My study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, USA (33.208$^\circ$ N, 81.408$^\circ$ W, Figure \@ref(fig:srs-map)).
 in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals [@tewksbury_corridors_2002].
@@
 While the design of the experimental blocks control for as much as possible, landscape features across the sight are different enough to influence species composition.
-Several factors may explain this pattern, at the time this study was conducted block 8 had the densest matrix of any of the other sampling blocks which could have hindered the scent permeability of our bait.
+Several factors may explain this pattern, at the time this study was conducted block 8 had the densest matrix of any of the other sampling blocks which could have hindered the scent permeability of the bait.
 Other habitat characteristics might have been influential as well, including soil quality and forest cover which can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019]
@@
 In this regard it is also important to consider other geographical differences, block 8 was nearest to roadways used by all employees at SRS and the eastern side of the sight had large bodies of water.
-Both of which would effect the land-use by larger mammals[@WOS:000241362400032; Dechen Quinn et al 2013; @WOS:000651446900001].
-Additionally, deer and feral hogs are the most dominant mammals at the site, two species that do well in patchy landscapes(Castillo-Contreras et al.
-2018, Fraser et al.
-2019).
+Both of which would effect the land-use by larger mammals [@WOS:000241362400032; @dechen_quinn_landscape_2013; @WOS:000651446900001].
+Additionally, deer and feral hogs are the most dominant mammals at the site, two species that do well in patchy landscapes [@castillo-contreras_urban_2018; @fraser_connectivity_2019].
 This might explain the lack of variation in richness or diversity in patches while differences are more apparent at greater distances across the site.
@@
 _Aphodius alloblackburneus_ had a disproportionate positive effect to being captured in matrix patches as compared to other species.
-Beetles from the genus _Aphodius_ do tend to show patterns of habitat specificity[@roslin_distribution_2001], and many preferring forested habitat*frank*,  so it is not unexpected that they might show a preference towards forested matrix.
-I did not detect that any species was more positively associated with open patches despite expectations that some species ( eg.
-_Canthon vigilans_, _Melanocanthon bispinatus_) would prefer open habitat [@conover_phenology_2019; @nealis_habitat_1977].
+Beetles from the genus _Aphodius_ do tend to show patterns of habitat specificity [@roslin_distribution_2001], and many preferring forested habitat*frank*,  so it is not unexpected that they might show a preference towards forested matrix.
+I did not detect that any species was more positively associated with open patches despite expectations that some species (e.g., _Canthon vigilans_, _Melanocanthon bispinatus_) would prefer open habitat [@conover_phenology_2019; @nealis_habitat_1977].
 Altogether this could be another sign pointing towards matrix acting as a source population, and since open habitat was much less dominant in my experimental system beetles could be moving into patches from habitat edge.
@@
 
-Despite ample work documenting their ecological importance, there is a surprising lack of research on dung beetle diversity and corridors outside of the tropics[@nichols_global_2007], with any work in temperate locations coming from Europe where _Aphodius_ beetles dominate[@roslin_distribution_2001].
-Many studies in the tropics have found patterns of habitat preference between open fields and continuous forest similar to what I did in this study[@WOS:000349242000004], and some work that found that forest species were using living fences as corridors[@arellano_response_2008].
+Despite ample work documenting their ecological importance, there is a surprising lack of research on dung beetle diversity and corridors outside of the tropics [@nichols_global_2007], with any work in temperate locations coming from Europe where _Aphodius_ beetles dominate [@roslin_distribution_2001].
+Many studies in the tropics have found patterns of habitat preference between open fields and continuous forest similar to what I did in this study [@WOS:000349242000004], and some work that found that forest species were using living fences as corridors [@arellano_response_2008].
 However, differences in composition were much more distinctly different with some species being exclusive to habitat types unlike I see in this paper.
-Of the limited work out of the south eastern United States I did recover assemblages similar to other work [@conover_phenology_2019; @nealis_habitat_1977], and even returned results to another study where differences between treatment types were non-significant while variation on a larger scale was more apparent[@young_dung_2023].
+Of the limited work out of the south eastern United States I did recover assemblages similar to other work [@conover_phenology_2019; @nealis_habitat_1977], and even returned results to another study where differences between treatment types were non-significant while variation on a larger scale was more apparent [@young_dung_2023].
 In comparison to other studies from SRS my results were similar to those of small mammals which were more likely to be using matrix land [@mabry_influence_2003].
@@
 
-```{r echo=FALSE, fig.align='center', fig.cap="All experimental blocks at SRS. Orange indicates blocks sampled.", message=FALSE, warning=FALSE, label=srs-map, out.width="40%"}
+```{r echo=FALSE, fig.align='center', fig.cap="All experimental blocks at SRS.
+Filled squares indicate the blocks in which dung beetles were sampled.
+Each block has a unique ID number.", message=FALSE, warning=FALSE, label=srs-map, out.width="60%"}
 
@@
 
-```{r, label = patches, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Experimental block with the different patches and distance between them.", out.width="50%", fig.align='center'}
+```{r, label = patches, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Overhead view of one of the experimental blocks indicating the different types of patches and the distance between them (from Nathan et al.
+2018 and reused with permission).", out.width="70%", fig.align='center'}
 
@@
 
-```{r, label = pitfalls, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Placement of traps within patch types and in the matrix as well as their orientation.", out.width="50%", fig.align='center'}
+```{r, label = pitfalls, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Diagram indicating the placement pitfall traps in each patch and in the matrix.
+Traps in patches were approximately 40 m from the patch edge.", out.width="70%", fig.align='center'}
 
@@
 \newpage
-```{r, label = patch, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average total dung beetle abundance by patch type with standard deviations as error bars.", out.width="50%", fig.align='center'}
+```{r, label = patch, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average total dung beetle abundance by patch type with standard deviations as error bars.", out.width="70%", fig.align='center'}
 
@@
 
-```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average abundance of the top 6 most abundant species by patch type.", out.width="50%", fig.align='center'}
+```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average abundance of the top 6 most abundant species by patch type.", out.width="70%", fig.align='center'}
 
@@
 
-```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle secies richness by patch type.", out.width="50%", fig.align='center'}
+```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle species richness in each patch type.
+The point shapes indicates the block in which each patch was located.", out.width="70%", fig.align='center'}
 
@@
 
-```{r, label = shannon, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Shannon diversity by patch type.", out.width="50%", fig.align='center'}
+```{r, label = shannon, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Shannon diversity by patch type.", out.width="70%", fig.align='center'}
 
@@
 
-```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index y patch type.", out.width="50%", fig.align='center'}
+```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index y patch type.", out.width="70%", fig.align='center'}
 
@@
 \newpage
-
+\blandscape
 ```{r Table-MO, echo=FALSE,message = FALSE,warning=FALSE}
@@
 ```
-
+\elandscape
 \newpage
@@
 
+\blandscape
 ```{r Table-M3, echo=FALSE,message = FALSE,warning=FALSE}

### 2025-07-10 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
added date

@@
 author: "ERIC ESCOBAR-CHENA"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 # date: "\\vspace{-1em}"
@@
   - \fancyhead[L]{}
-  - \fancyhead[R]{\small{EEC MS Thesis Draft}}
+  - \fancyhead[R]{\tiny{EEC MS Thesis Draft (`r format(Sys.time(), '%d %B %Y')`)}}
   - \thispagestyle{fancy}

### 2025-07-14 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
add ridiculous grad school required ch1 and ch3

@@
 CHAPTER 1\\  
+INTRODUCTION\\  
+\end{center}
+
+\doublespacing
+
+Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes [@haddad_corridor_2003].
+These movements are hypothesized to prevent species diversity from declining in fragments, as well as help maintain the ecosystem services provided by these species (at both the patch- and landscape-level)[@burt_ants_2022].
+Although there is some evidence that animals disperse between patches via corridors, and that connected patches have higher species diversity than unconnected ones, little work to date has investigated the consequences of these corridor-driven patterns for ecosystem services.
+
+\newpage
+
+\begin{center}
+CHAPTER 2\\  
 THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES\\  
@@
 
+\newpage
+
+\begin{center}
+CHAPTER 3\\  
+CONCLUSIONS AND FUTURE DIRECTIONS\\  
+\end{center}
+
+\doublespacing 
 Moving forward, there is a bounty of knowledge yet to be collected for dung beetles in the south eastern United States.

### 2025-07-15 corridor_docs/eric_ms_thesis/03_biomass_analysis.R
trying out some biomass models

@@ -5,6 +5,8 @@
 
 library(tidyverse)
 library(here)
+library(lme4)
+library(sjPlot)
 # load data ---------------------------------------------------------------
 
 
@@ -41,6 +43,7 @@ biomass_per_spp<-biomass_per_spp %>%
   ungroup() %>% 
   rename(species=sp_code)
 
+
 ggplot(biomass_per_spp, 
        aes(
          x=reorder(species,-spp_biomass), 
@@ -48,6 +51,15 @@ ggplot(biomass_per_spp,
          )
        ) + 
   geom_bar(stat = "identity")
+
+biomass_top_6 <- biomass_per_spp %>% 
+  filter(species=="cvig"|
+           species=="alec"|
+           species=="pign"|
+           species=="dcar"|
+           species=="open"|
+           species=="aaeg")
+
 
 
 biomass_2<-btl_data %>%
@@ -87,6 +99,9 @@ biomass_2 %>%
   group_by(patch) %>% 
   summarize(tot_bm=sum(spp_biomass))
 
+biomass_2<-biomass_2 %>% 
+  mutate(patch=as.factor(patch)) 
+biomass_2$patch<-relevel(biomass_2$patch,"Matrix")
 
 patch_type_mean_bm<-
 biomass_2 %>% 
@@ -104,6 +119,29 @@ ggplot(patch_type_mean_bm,
        )
 ) + 
   geom_bar(stat = "identity")
+
+M5 <- lmer(spp_biomass ~ patch + (1 + patch | block), 
+            data = biomass_2)
+
+plot_model(M5, type = "pred")
+
+summary(M5)
+
+
+biomass_2_top_6 <- biomass_2 %>% 
+  filter(sp_code=="cvig"|
+           sp_code=="alec"|
+           sp_code=="pign"|
+           sp_code=="dcar"|
+           sp_code=="open"|
+           sp_code=="aaeg")
+
+M6 <- lmer(spp_biomass ~ patch + (1 + patch | block), 
+           data = biomass_2_top_6)
+
+summary(M6)
+
+plot_model(M6, type = "pred")
 
 b2<-biomass %>%
   ungroup() %>% 

### 2025-07-17 corridor_docs/eric_ms_thesis/03_biomass_analysis.R
more models!

@@ -143,6 +143,26 @@ summary(M6)
 
 plot_model(M6, type = "pred")
 
+M7 <- lmer(spp_biomass ~ sp_code + patch + (1 | block), 
+     data = biomass_2_top_6)
+
+summary(M7)
+
+plot_model(M7, type = "pred", terms = "sp_code")
+
+plot_model(M7, type = "pred", terms = c("sp_code", "patch"))
+
+M8 <- lmer(spp_biomass ~ n + sp_code + (1| block),
+           data = biomass_2_top_6)
+
+summary(M8)
+
+plot_model(M8, type = "pred")
+
+plot_model(M8, type = "pred", terms = c("n", "sp_code"))
+ 
+plot_model(M8, type = "pred", terms = c("sp_code", "n"))
+
 b2<-biomass %>%
   ungroup() %>% 
   select(-c(sp_code,n,avg_bmass)) %>% 

### 2025-07-17 corridor_docs/eric_ms_thesis/03_biomass_analysis.R
sp biomass v abund

@@ -185,3 +185,111 @@ ggplot(b2, aes(
   pivot_wider(
     names_from = sp_code,
     values_from = n_patch) 
+
+  
+
+# ind biomass v abundance -------------------------------------------------
+
+  # https://www.sciencedirect.com/science/article/pii/S0169534707000985?via%3Dihub
+  
+  body_mass_v_abund<-btl_data %>%
+    pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
+    group_by(species, patch,block) %>%
+    # group_by(block) %>%
+    # group_by(patch,block) %>%
+    mutate(patch=case_when(
+      patch == "c" ~ "Connected",
+      patch == "m" ~ "Matrix",
+      patch == "w" ~ "Winged",
+      patch == "r" ~ "Rectangle",
+      .default = as.character(patch))) %>%
+    summarize(n=sum(n, na.rm=TRUE)) %>%
+    arrange(desc(n)) %>%
+    rename(sp_code=species) %>% 
+    left_join(spp_bmass,by="sp_code") %>%
+    # select(-n.y) %>% 
+    # rename(n=n.x) %>% 
+    # select(-sd, -n_btls) %>% 
+    mutate(spp_biomass=n*avg_bmass) %>% 
+    ungroup() %>% 
+    select(-sd) %>% 
+    # select(-sp_code,-avg_bmass) %>% 
+    drop_na() 
+  
+  indiv_bmass<-body_mass_v_abund %>% 
+    select(sp_code,avg_bmass) %>% 
+    distinct() %>% 
+    arrange(desc(avg_bmass)) %>% 
+    mutate(sp_code = fct_inorder(factor(sp_code))) %>% 
+    group_by(sp_code) %>% 
+    mutate(shape=cur_group_id()) %>% 
+    mutate(shape=as.factor(shape))
+
+
+# overall species mass v abundance ----------------------------------------
+
+    library(RColorBrewer)
+  library(geomtextpath)
+  
+  
+  
+  total_n<- body_mass_v_abund %>% group_by(sp_code) %>% 
+    summarize(total_n=sum(n))
+  total_n<-full_join(indiv_bmass,total_n)
+  total_n %>% 
+    ggplot(aes(x=avg_bmass, y=total_n, color=sp_code)) + 
+    # scale_shape_manual(values = total_n$shape)+
+    # ggplot(aes(x=log(avg_bmass), y=log(total_n), color=sp_code,shape=sp_code)) + 
+    geom_point()+
[…]

### 2025-07-17 corridor_docs/eric_ms_thesis/03_biomass_analysis.R
fixed biomass

@@ -13,14 +13,25 @@ library(sjPlot)
 btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
 spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
 
-spp_bmass<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","spp_weights.csv")) %>% 
+# need to add how many beetles were included in each raw weight (some too small)
+btls_per_tray<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","spp_weights.csv")) %>%
   group_by(sp_code) %>% 
-  summarize(avg_bmass=mean(weight),
-            sd=sd(weight),
-            n_btls=n_distinct(weight)) %>% 
-  arrange(desc(avg_bmass))
+  summarize(n_tray=n()) %>% 
+  mutate(btls_per_tray=case_when(
+    n_tray==15~1,
+    n_tray==5~3,
+    n_tray==3~5,
+    n_tray==1~15,
+    .default = as.integer(n_tray)
+  ))
 
 
+spp_bmass<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","spp_weights.csv")) %>% 
+  left_join(btls_per_tray,by="sp_code")  %>% 
+  mutate(indiv_bmass=weight/btls_per_tray) %>% 
+  group_by(sp_code) %>% 
+  summarize(avg_ind_bmass=mean(indiv_bmass),
+            sd = sd(indiv_bmass))
 
 biomass_per_spp<-btl_data %>%
   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
@@ -35,8 +46,8 @@ biomass_per_spp<-btl_data %>%
   arrange(desc(n)) %>%
   rename(sp_code=species) %>% 
   left_join(spp_bmass,by="sp_code") %>% 
-  select(-sd, -n_btls) %>% 
-  mutate(spp_biomass=n*avg_bmass)
+  # select(-sd, -n_btls) %>% 
+  mutate(spp_biomass=n*avg_ind_bmass)
 
 biomass_per_spp<-biomass_per_spp %>% 
   drop_na() %>% 
@@ -80,7 +91,7 @@ biomass_2<-btl_data %>%
[…]

### 2025-07-18 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
post-defense edits

@@
 Fifteen individuals of each species with adequate captures were dried to equilibrium and weighed for biomass measurements.
-Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods upon completion of all analyses.
+**[EDIT SENTENCE ABOVE AS FOLLOWS:** Fifteen individuals of each species with adequate captures (was there a numeric cutoff?) were dried at XXX degrees until they reached a stable mass.
+We weighed each beetle using a BRAND, MODEL microbalance, except for species whose individuals were extremely small.
+Individual biomass for these species was estimated by weighing batches of N = 5  (i.e., ---------) or N = 3  (i.e., ---------) beetles and then calculating the average biomass per beetle.
+These values of individual biomass were then used to estimate the total dung beetle biomass -- all species combined and for each individual species - in each block and patch.
+**]** Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods upon completion of all analyses.
 
@@
 Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.
+**[need to add to the last paragraph:** that you did these analyses for the dung beetle biomass as well.
+And it would be cool for guild...**]**
 
@@
 
-### Beetle Abundance
+### Dung Beetle Abundance
 
@@
 
-### Beetle Richness and Diversity
+### Dung Beetle Richness and Diversity
 
@@
 Again, none of the other patch types had significant effects: matrix ($\beta$ = -0.107, SE = 0.411, t = -0.259), rectangle ($\beta$ = 0.059, SE = 0.411, t = 0.143), and winged ($\beta$ = –0.480, SE = 0.411, t = –1.169).
+
+### Dung Beetle Biomass
+
+**ADD RESULTS FOR DB BIOMASS ANALYSES**
 
@@
 This might explain the lack of variation in richness or diversity in patches while differences are more apparent at greater distances across the site.
+**[need to add and discuss soil compaction at 8 here in this paragraph: 1-2 sentences]**
 
@@
 
-<!-- Paragraph 5: Our results largely echo/constrast with previous ones from the SRS and elsewhere  -->
+<!-- Paragraph 5: Beetle Biomass  -->
+
+**[Paragraph 6: biomass]** The patterns in dung beetle biomass largeley echo the ones for richness and abundnace,.
+these have potential implications for ecosystem functioning.
+PS might want to add someting to the Future directions about "Interestingly, dung beetles appear to be an exception to the globally and taxonomically robust rule that abundance  is negatively correlated with individual biomass....
+
+<!-- Paragraph 6: Our results largely echo/constrast with previous ones from the SRS and elsewhere  -->
 
@@
 
-<!-- Paragraph 6: Caveats -->
+<!-- Paragraph 7: spatial scale of experimental plots vs. dung detectability  -->
+
+**[Paragraph 7: spatial scale of experimental plots vs.
+dung detectability]** Finally, it is important to emphasize that while dung beetles are capable of long distance flight and detecting dung at distances of over X meters, the results I observed are lijkely not the result of a mismatch between the spatial scale of the experimental replicates and dung beetle movement.
+Put another way, some beetles were almost certainly drawn by the dung used in baits from the matrix into the plots.
+If this were a big effect we'd expect all plots to be similar to the matrix, or at least to each other.
+the fact that rectangular are lower suggests that...]
+
+<!-- Paragraph 7: Caveats -->
 

### 2025-07-23 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
beetle weights

@@
 All dung beetles were counted and identified to species using @nemes_illustrated_2015 and @edmonds_taxonomic_2023.
-Fifteen individuals of each species with adequate captures were dried to equilibrium and weighed for biomass measurements.
-**[EDIT SENTENCE ABOVE AS FOLLOWS:** Fifteen individuals of each species with adequate captures (was there a numeric cutoff?) were dried at XXX degrees until they reached a stable mass.
-We weighed each beetle using a BRAND, MODEL microbalance, except for species whose individuals were extremely small.
-Individual biomass for these species was estimated by weighing batches of N = 5  (i.e., ---------) or N = 3  (i.e., ---------) beetles and then calculating the average biomass per beetle.
-These values of individual biomass were then used to estimate the total dung beetle biomass -- all species combined and for each individual species - in each block and patch.
-**]** Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods upon completion of all analyses.
+Fifteen individuals of each species with adequate captures were dried and weighed for biomass measurements.
+Ten species had the required counts for weighing which I dried in drying oven until all specimen reached a stable mass.
+I weighed each beetle using an Ohaus Adventurer Pro AV53 microbalance, with the exception of individuals whose weights were extremely small.
+Individual biomass for these species was estimated by weighing in batches of N = 3 (i.e., *Ateuchus lecontei*), N = 5 (i.e., *Onthophagus pennsylvanicus*), or N = 15 (i.e., *Aphodius alloblackburneus*) and then calculated by average biomass per beetle.
+Values for individual biomass were then used to estimate the totall dung beetle biomass for each -- all species combined and for each individual species - in each block and patch.
+Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods upon completion of all analyses.
 

### 2025-07-24 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
anoother sentance
added soil compaction and agriculture sentance

@@
 Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.
-**[need to add to the last paragraph:** that you did these analyses for the dung beetle biomass as well.
-And it would be cool for guild...**]**
+Likewise, linear mixed-effects models were used to compare the average biomass of individual dung beetles in each patch type, once again including block as a random effect.
+
+**[need to add to the last paragraph:** that you did these analyses for the dung beetle biomass as well. And it would be cool for guild...**]**
 
@@
 
-**ADD RESULTS FOR DB BIOMASS ANALYSES**
-
 \begin{center} 
@@
 Several factors may explain this pattern, at the time this study was conducted block 8 had the densest matrix of any of the other sampling blocks which could have hindered the scent permeability of the bait.
-Other habitat characteristics might have been influential as well, including soil quality and forest cover which can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019]
-Not only do these impact dung beetle movement, but also these landscape features can affect mammal movement, which in turn limits dung availability for beetles.
+Other habitat characteristics might have been influential as well, including soil quality and forest cover which can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019].
+Furthermore in the context of my study site, land-use history could be an important driver to consider as much of the land would have previously been used for agriculture and during development of experimental units cleared with heavy machinery might experience heavy soil compaction *brudvig*.
+Another important consideration to take is that these landscape features can affect mammal movement, which in turn limits dung availability for beetles.
 In this regard it is also important to consider other geographical differences, block 8 was nearest to roadways used by all employees at SRS and the eastern side of the sight had large bodies of water.
@@
 However, differences in composition were much more distinctly different with some species being exclusive to habitat types unlike I see in this paper.
-Of the limited work out of the south eastern United States I did recover assemblages similar to other work [@conover_phenology_2019; @nealis_habitat_1977], and even returned results to another study where differences between treatment types were non-significant while variation on a larger scale was more apparent [@young_dung_2023].
+Of the limited work out of the south eastern United States I did recover assemblages similar to other work [@conover_phenology_2019; @nealis_habitat_1977], and even returned results similar to another study where differences between treatment types were non-significant while variation on a larger scale was more apparent [@young_dung_2023].
 In comparison to other studies from SRS my results were similar to those of small mammals which were more likely to be using matrix land [@mabry_influence_2003].
@@
 He later began to explore this curiosity deeper in beginning his graduate education at the University of Florida as a Master's Student under the supervision of Emilio Bruna.
+He received his Master of Science from the University of Florida in 2025.

### 2025-07-24 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
some more

@@
 
+Dung beetle biomass follows a pattern very similar to abundance where biomass is the highest in matrix plots, followed by connected, then winged, with a large drop in rectangle patches (Figure X).
+However, variation between individual beetle biomass
+is to large to detect a pattern in statistical analysis (Table X).
+
 \begin{center} 
@@
 
-**[Paragraph 7: spatial scale of experimental plots vs.
-dung detectability]** Finally, it is important to emphasize that while dung beetles are capable of long distance flight and detecting dung at distances of over X meters, the results I observed are lijkely not the result of a mismatch between the spatial scale of the experimental replicates and dung beetle movement.
+**[Paragraph 7: spatial scale of experimental plots vs. dung detectability]**
+
+Finally, it is important to emphasize that while dung beetles are capable of long distance flight and detecting dung at distances of over 50 meters *gray*, the results I observed are likely the result of a mismatch between the spatial scale of the experimental replicates and dung beetle movement.
 Put another way, some beetles were almost certainly drawn by the dung used in baits from the matrix into the plots.
 If this were a big effect we'd expect all plots to be similar to the matrix, or at least to each other.
-the fact that rectangular are lower suggests that...]
+The fact that rectangular are lower suggests that some sort of landscape effect is apparent, likely related to habitat edge since connected and winged patches were the most similar.
 

### 2025-07-24 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
updates for final submission

@@
 # library(grid)
-# library(gridExtra)
-# library(png)
+library(gridExtra)
+library(png)
 # library(egg)
-# library(magick)
+library(magick)
 # library(citr)
@@
 spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+bmass<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","biomass_per_spp.csv"))
 
@@
   summarise(n_spp=n_distinct(species))
+
+# biomass
+bmass<-bmass %>%
+  rename(sp_code=species) %>% 
+  left_join(spp_codes) %>% 
+  mutate(species=tolower(species)) %>% 
+  mutate(species=paste(genus, species, sep=" ")) %>% 
+  select(species,
+         `Individual Biomass (mean)`=avg_ind_bmass,
+         SD=sd) %>% 
+  mutate(SD=round(SD,3)) %>% 
+  mutate(SD=as.character(SD)) %>% 
+  replace_na(list(SD=""))
 
@@
 Individual biomass for these species was estimated by weighing in batches of N = 3 (i.e., *Ateuchus lecontei*), N = 5 (i.e., *Onthophagus pennsylvanicus*), or N = 15 (i.e., *Aphodius alloblackburneus*) and then calculated by average biomass per beetle.
-Values for individual biomass were then used to estimate the totall dung beetle biomass for each -- all species combined and for each individual species - in each block and patch.
+Values for individual biomass were then used to estimate the total biomass of each species in each patch, as well as the total beetle biomass (i.e., all species combined) in each patch.
 Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods upon completion of all analyses.
@@
 Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.
-Likewise, linear mixed-effects models were used to compare the average biomass of individual dung beetles in each patch type, once again including block as a random effect.
-
-**[need to add to the last paragraph:** that you did these analyses for the dung beetle biomass as well. And it would be cool for guild...**]**
+**Likewise, linear mixed-effects models were used to compare for differences between patch types in (1) total dung beetle biomass, and (2) the total biomass of each of the ten species, again including block as a random effect.
+**
 
@@
 
-```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average abundance of the top 6 most abundant species by patch type.", out.width="70%", fig.align='center'}
+```{r spcodes}
+
+sp_code_table<-data.frame(sp_code=c("aaeg", "cvig", "dcar", "open", "pign", "alec"))
+species_names<-spp_codes %>% 
+  filter(sp_code%in%sp_code_table$sp_code) %>% 
+  mutate(species=tolower(species)) %>% 
+  mutate(species=paste(species=genus,species,sep=" ")) %>% 
+  select(sp_code, species) %>% 
+mutate(species=paste(sp_code,species,sep=": ")) 
+species_names<-paste(species_names$species,collapse=", ")
+
+```
+
+```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = paste("Average abundance of the top 6 most abundant species by patch type.
+Species codes: ", species_names,".",sep=""), out.width="70%", fig.align='center'}
 
@@
 
+```
+
+```{r, label = sp-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "LEGEND", out.width="70%", fig.align='center'}
+
+knitr::include_graphics("images/sp_biomass_plot.png")
+# knitr::include_graphics("images/sp_abundance_plot.png")
+# grid.arrange(p1[1], p2[1], nrow = 1)
+
+```
+
+```{r, label = patch-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "LEGEND", out.width="70%", fig.align='center'}
+
+knitr::include_graphics("images/avg_bmass_patch_plot.png")
+```
+
+```{r, label = block-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "LEGEND", out.width="70%", fig.align='center'}
+
+knitr::include_graphics("images/avg_bmass_block_plot.png")
 ```
@@
   mutate(Group=if_else(Group=="block","Block",Group)) %>% 
-  mutate(Term=gsub("patch_type","patch type - ",Term)) %>% 
+  mutate(Term=gsub("patch_type","Patch type: ",Term)) %>% 
   mutate(Term=gsub("cor__","correlation: ",Term)) %>% 
-  mutate(Term=gsub("sd__","SD: ",Term)) %>% 
+  mutate(Term=gsub("sd__","SD ",Term)) %>% 
   mutate(Term=gsub("[.]"," x ",Term)) %>% 
@@
 \elandscape
+
+\newpage
+
+```{r Table-aovM51, echo=FALSE,message = FALSE,warning=FALSE}
+lmerM51<-read_csv(here("corridor_docs","eric_ms_thesis","tables","lmerM5_1.csv")) %>%
+    mutate(term=case_when(
+    term == "patchConnected" ~ "Patch: Connected",
+    term == "patchRectangle" ~ "Patch: Rectangle",
+    term == "patchWinged" ~ "Patch: Winged",
+    term == "sd__(Intercept)" ~ "SD (Intercept)",
+    term == "sd__Observation" ~ "SD (Observation)",
+    .default = as.character(term))) |>
+mutate(effect=if_else(effect=="ran_pars","Random",effect)) %>% 
+  mutate(effect=if_else(effect=="fixed","Fixed",effect)) %>% 
+  mutate(group=if_else(effect=="block","Block",group)) %>% 
+  mutate(std.error=round(std.error,3)) %>% 
+  mutate(statistic=round(statistic,3)) %>% 
+  mutate(std.error=(as.character(std.error))) %>% 
+  mutate(statistic=(as.character(statistic))) %>% 
+  replace_na(list(group="",std.error="",statistic="")) %>% 
+  rename(SE = std.error,
+         Estimate=estimate,
+         Term=term,
+         Group=group,
+         Effect=effect,
+         Statistic=statistic) %>% 
+  kable(
+      digits = 3,
+      align="lllccc",
+      format="latex",
+      row.names = FALSE,
+      booktabs=T,
+      linesep = "", #removes the blank line after every 5 lines
+      caption = "FILL IN CAPTION TOTAL BIOMASS BY PATCH ALL POOLED.") %>%
+  kable_styling(bootstrap_options = c("striped"),
+                full_width = F,
+                latex_options=c("scale_down","hold_position"),
+                font_size = 12,
+                position = "center") %>%
+  # column_spec(1, italic = T) %>%
+  row_spec(0, bold = T)
+  # row_spec(1:15, hline_after = T)
+lmerM51
+```
+\newpage
+
+```{r Table-aovM52, echo=FALSE,message = FALSE,warning=FALSE}
+
+sp_code_table<-data.frame(sp_code=c("bpro", "cvig", "dcar", "mbis", "ocon", "open", "pign", "pvin"))
+species_names<-spp_codes %>% 
+  filter(sp_code%in%sp_code_table$sp_code) %>% 
+  mutate(species=tolower(species)) %>% 
+  mutate(species=paste(species=genus,species,sep=" ")) %>% 
+  select(sp_code, species) %>% 
+mutate(species=paste(sp_code,species,sep=": ")) 
+species_names<-paste(species_names$species,collapse=", ")
+
+lmerM52<-read_csv(here("corridor_docs","eric_ms_thesis","tables","lmerM5_2.csv")) %>%
+    mutate(term=case_when(
+    term == "patchConnected" ~ "Patch: Connected",
+    term == "patchRectangle" ~ "Patch: Rectangle",
+    term == "patchWinged" ~ "Patch: Winged",
+    term == "sd__(Intercept)" ~ "SD (Intercept)",
+    term == "sd__Observation" ~ "SD (Observation)",
+    .default = as.character(term))) |>
+  mutate(term=case_when(
+term == "sp_codealec" ~ "Species: alec",
+    term == "sp_codebpro" ~ "Species: bpro",
+term == "sp_codecvig" ~ "Species: bvig",
+term == "sp_codedcar" ~ "Species: dcar",
+term == "sp_codembis" ~ "Species: mbis",
+term == "sp_codeocon" ~ "Species: ocon",
+term == "sp_codeopen" ~ "Species: open",
+term == "sp_codepign" ~ "Species: pign",
+term == "sp_codepvin" ~ "Species: pvin",
+  .default = as.character(term))) |>
+
+mutate(effect=if_else(effect=="ran_pars","Random",effect)) %>% 
+  mutate(effect=if_else(effect=="fixed","Fixed",effect)) %>% 
+  mutate(group=if_else(effect=="block","Block",group)) %>% 
+  mutate(std.error=round(std.error,3)) %>% 
+  mutate(statistic=round(statistic,3)) %>% 
+  mutate(std.error=(as.character(std.error))) %>% 
+  mutate(statistic=(as.character(statistic))) %>% 
+  replace_na(list(group="",std.error="",statistic="")) %>% 
+  rename(SE = std.error,
+         Estimate=estimate,
+         Term=term,
+         Group=group,
+         Effect=effect,
+         Statistic=statistic) %>% 
+  kable(
+      digits = 3,
+      align="lllccc",
+      format="latex",
+      row.names = FALSE,
+      booktabs=T,
+      linesep = "", #removes the blank line after every 5 lines
+      caption = paste("FILL IN CAPTION BIOMASS by SP AND PATCH. Species codes = ",species_names, sep="")
+      ) %>%
+  kable_styling(bootstrap_options = c("striped"),
+                full_width = F,
+                latex_options=c("scale_down","hold_position"),
+                font_size = 12,
+                position = "center") %>%
+  # column_spec(1, italic = T) %>%
+  row_spec(0, bold = T)
+  # row_spec(1:15, hline_after = T)
+
+lmerM52
+```
+
 \newpage
@@
 ```{r Table-M3, echo=FALSE,message = FALSE,warning=FALSE}
+
+sp_code_table<-data.frame(sp_code=c("alec", "cvig", "dcar", "open", "pign"))
+species_names<-spp_codes %>% 
+  filter(sp_code%in%sp_code_table$sp_code) %>% 
+  mutate(species=tolower(species)) %>% 
+  mutate(species=paste(species=genus,species,sep=" ")) %>% 
+  select(sp_code, species) %>% 
+mutate(species=paste(sp_code,species,sep=": ")) 
+species_names<-paste(species_names$species,collapse=", ")
+
 M3<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M3.csv")) %>%
@@
   mutate(Group=if_else(Group=="block","Block",Group)) %>% 
-  mutate(Term=gsub("patch_type","patch type - ",Term)) %>% 
-  mutate(Term=gsub("cor__","correlation: ",Term)) %>% 
-  mutate(Term=gsub("sd__","SD: ",Term)) %>% 
-  mutate(Term=gsub("sp_code","Species - ",Term)) %>% 
-  mutate(Term=gsub("[.]"," x ",Term)) %>% 
-  mutate(Term=gsub("[:]"," : ",Term)) %>% 
+  mutate(Term=gsub("patch_type","Patch type: ",Term)) %>% 
+  mutate(Term=gsub("cor__","Correlation: ",Term)) %>% 
+  mutate(Term=gsub("sd__","SD ",Term)) %>% 
+  mutate(Term=gsub("sp_code","Species: ",Term)) %>% 
+  # mutate(Term=gsub("[.]"," x ",Term)) %>% 
+  mutate(Term=gsub(":P"," x P",Term)) %>% 
   mutate(SE=round(SE,3)) %>% 
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "M3 patch type effects on species abundance of top 6 most abundant speceies and interactive term for patch type and species ID.") %>%
+      caption = "Effects of patch type on the abundance of the 6 most abundant species and interaction term for patch type and species.") %>%
   kable_styling(bootstrap_options = c("striped"),
@@
    mutate(term=case_when(
-    term == "patch_typeMatrix" ~ "Patch Type - Matrix",
-    term == "patch_typeRectangle" ~ "Patch Type - Rectangle",
-    term == "patch_typeWinged" ~ "Patch Type - Winged",
-    term == "patch_typeConnected" ~ "Patch Type - Connected",
+    term == "patch_typeMatrix" ~ "Patch Type: Matrix",
+    term == "patch_typeRectangle" ~ "Patch Type: Rectangle",
+    term == "patch_typeWinged" ~ "Patch Type: Winged",
+    term == "patch_typeConnected" ~ "Patch Type: Connected",
     .default = as.character(term))) |>
@@
 
-  mutate(Term=gsub("patch_type","patch type - ",Term)) %>% 
+  mutate(Term=gsub("patch_type","Patch type: ",Term)) %>% 
   mutate(Term=gsub("cor__","correlation: ",Term)) %>% 
-  mutate(Term=gsub("sd__","SD: ",Term)) %>% 
+  mutate(Term=gsub("sd__Observation","SD (Observation)",Term)) %>% 
+  mutate(Term=if_else(str_detect(Term,"sd__(Intercept)"),"SD (Intercept)",Term)) %>% 
   mutate(SE=round(SE,3)) %>% 
@@
 
-  mutate(Term=gsub("patch_type","patch type - ",Term)) %>% 
+  mutate(Term=gsub("patch_type","Patch type: ",Term)) %>% 
   mutate(Term=gsub("cor__","correlation: ",Term)) %>% 
-  mutate(Term=gsub("sd__","SD: ",Term)) %>% 
+  mutate(Term=gsub("sd__(Intercept)","SD (Intercept)",Term)) %>% 
   mutate(SE=round(SE,3)) %>% 

### 2025-07-24 corridor_docs/eric_ms_thesis/03_biomass_analysis.R

@@ -50,18 +50,55 @@ biomass_per_spp<-btl_data %>%
   mutate(spp_biomass=n*avg_ind_bmass)
 
 biomass_per_spp<-biomass_per_spp %>% 
-  drop_na() %>% 
+  drop_na(avg_ind_bmass) %>% 
   ungroup() %>% 
   rename(species=sp_code)
 
 
+
+write_csv(biomass_per_spp,here("corridor_docs","eric_ms_thesis","ms_data","data_raw","biomass_per_spp.csv"))
+
+biomass_per_spp<-
+biomass_per_spp %>% 
+rename(sp_code=species) %>% 
+  left_join(spp_codes) %>% 
+  mutate(species=tolower(species)) %>% 
+  mutate(species=paste(genus, species, sep=" ")) 
+
+sp_biomass_plot<-
 ggplot(biomass_per_spp, 
        aes(
          x=reorder(species,-spp_biomass), 
          y=spp_biomass
          )
        ) + 
-  geom_bar(stat = "identity")
+  labs(x="Species",y="Total biomass collected")+
+  geom_bar(stat = "identity")+
+  theme_classic()+
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
+  theme(axis.text.x = element_text(face = 'italic'))
+
+ggsave("corridor_docs/eric_ms_thesis/images/sp_biomass_plot.png", width = 4, height = 4, units = "in")
+
+
+
+sp_abundance_plot<-
+ggplot(biomass_per_spp, 
+       aes(
+         x=reorder(species,-spp_biomass), 
+         y=n
+       )
+) + 
+  labs(x="Species",y="Total No. collected")+
+  geom_bar(stat = "identity")+
+  theme_classic()+
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
+  theme(axis.text.x = element_text(face = 'italic'))
+
+ggsave("corridor_docs/eric_ms_thesis/images/sp_abundance_plot.png", width = 4, height = 4, units = "in")
+
+
+
 
 biomass_top_6 <- biomass_per_spp %>% 
   filter(species=="cvig"|
@@ -123,20 +160,85 @@ biomass_2 %>%
             sd_bm=sd(tot_bm))
 
 
+avg_bmass_patch_plot<-
 ggplot(patch_type_mean_bm, 
[…]

### 2025-07-25 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
push writing

@@
 
-Dung beetle biomass follows a pattern very similar to abundance where biomass is the highest in matrix plots, followed by connected, then winged, with a large drop in rectangle patches (Figure X).
-However, variation between individual beetle biomass
-is to large to detect a pattern in statistical analysis (Table X).
+Pooling dung beetle biomass by patch type follows a pattern very similar to abundance where biomass is the highest in matrix plots, followed by connected, then winged, with a large drop in rectangle patches (Figure 7).
+When comparing to biomass collected in matrix plots, statistical analysis showed a non-significant relationship with connected ($\beta$ = -7.337, SE = 8.659, t = -0.847; Table 4) and winged patches ($\beta$ = -9.608, SE = 8.659, t = -1.11; Table 4), while significant results were returned for rectangle patches ($\beta$ = -23.296, SE = 8.659, t = -2.69; Table 4).
+Block effects continue to be important for explaining high variation as the total biomass collected in block 52 nearly quadrupled that of block 8 (Table 8).
+
+Parsing biomass by species emphasizes the morphological variability between our study species as some of the most common contribute very little to total biomass (i.e.
+*Ateuchus lecontei*, *Aphodius allobloackburneus*; Figure 6).
+Linear mixed effects models further showcases these species specific differences (Table 5).
 
@@
 _Aphodius alloblackburneus_ had a disproportionate positive effect to being captured in matrix patches as compared to other species.
-Beetles from the genus _Aphodius_ do tend to show patterns of habitat specificity [@roslin_distribution_2001], and many preferring forested habitat*frank*,  so it is not unexpected that they might show a preference towards forested matrix.
+Beetles from the genus _Aphodius_ do tend to show patterns of habitat specificity [@roslin_distribution_2001], and many preferring forested habitat[@frank_land_2017],  so it is not unexpected that they might show a preference towards forested matrix.
 I did not detect that any species was more positively associated with open patches despite expectations that some species (e.g., _Canthon vigilans_, _Melanocanthon bispinatus_) would prefer open habitat [@conover_phenology_2019; @nealis_habitat_1977].
@@
 
-**[Paragraph 6: biomass]** The patterns in dung beetle biomass largeley echo the ones for richness and abundnace,.
+**[Paragraph 6: biomass]** The patterns in dung beetle biomass largeley echo the ones for richness and abundance,.
 these have potential implications for ecosystem functioning.
@@
 
-Finally, it is important to emphasize that while dung beetles are capable of long distance flight and detecting dung at distances of over 50 meters *gray*, the results I observed are likely the result of a mismatch between the spatial scale of the experimental replicates and dung beetle movement.
+Finally, it is important to emphasize that while dung beetles are capable of long distance flight and detecting dung at distances of over 50 meters [@WOS:000707546700001], the results I observed are likely the result of a mismatch between the spatial scale of the experimental replicates and dung beetle movement.
 Put another way, some beetles were almost certainly drawn by the dung used in baits from the matrix into the plots.

### 2025-07-25 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
biomass discussion

@@
 
-**[Paragraph 6: biomass]** The patterns in dung beetle biomass largeley echo the ones for richness and abundance,.
-these have potential implications for ecosystem functioning.
-PS might want to add someting to the Future directions about "Interestingly, dung beetles appear to be an exception to the globally and taxonomically robust rule that abundance  is negatively correlated with individual biomass....
+The patterns in dung beetle biomass largely echo what we observed for abundance, however this distinction is important to make since dung beetles vary greatly in terms of morphology and functionality [@ospina-garces_dung_2018].
+Understanding these patterns is of particular interest since biomass is understood to be positively associated with dung removal [@WOS:000287168100017].
+Because of this we can expect that the magnitude of removal might be greater in areas of larger biomass (i.e.
+matrix or connected patches), however future work should aim to directly measure this potential pattern.
+In addition to this, dung beetles appear to be an exception to the globally and taxonomically robust rule that abundance is negatively correlated with individual biomass [@white_relationships_2007].
+More intensive sampling could determine whether this trend is truly apparent.
 

### 2025-07-25 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
scale discussion

@@
 Put another way, some beetles were almost certainly drawn by the dung used in baits from the matrix into the plots.
-If this were a big effect we'd expect all plots to be similar to the matrix, or at least to each other.
-The fact that rectangular are lower suggests that some sort of landscape effect is apparent, likely related to habitat edge since connected and winged patches were the most similar.
+If this were an overarching effect we'd expect all plots to be similar to the matrix, or at least to each other.
+The fact that rectangular are have lower abundance suggests that some sort of landscape effect is apparent, likely related to habitat edge since connected and winged patches were the most similar.
+To remedy this issue any future work in this site should either focus solely on dung beetles with less dispersal ability or consider mark release recapture sampling in order to better isolate movements within and between patches.
 

### 2025-07-25 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
lastish edits and bib

@@
 
-When comparing the overall abundance of beetles (all species combined) across patch types, matrix plots had the highest captures, followed by connected patches, then winged, with the fewest in rectangular patches.
-Abundances from connected patches were not significantly different from those in the matrix while rectangle and winged patches had significantly fewer than the matrix ($\beta$ = -0.528 and -0.303 respectively, P \< 0.001, Table 4).
+When comparing the overall abundance of beetles (all species combined) across patch types, matrix plots had the highest captures, followed by connected patches, then winged, with the fewest in rectangular patches (Figure 4).
+Abundances from connected patches were not significantly different from those in the matrix while rectangle and winged patches had significantly fewer than the matrix ($\beta$ = -0.528 and -0.303 respectively, P \< 0.001, Table 2).
 
 Statistical analysis of abundance focused on the six most abundant species.
-A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 7).
+A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 6).
 The baseline abundance corresponds to the abundance of *Aphodius alloblackburneus* in matrix patches.
@@
 
-Plotting species richness by patch type reveals consistent richness across patch types with some variation between sampling blocks (Figure 5).The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N).
+Plotting species richness by patch type reveals consistent richness across patch types with some variation between sampling blocks (Figure 9).The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N).
 Modeling the effect of patch type on species richness with block as a random effect determined there was no significant differences among patch types.
@@
 
-Biodiversity was also even among patch types, however metric values were varied by sampling blocks.
-For Shannon Diversity, the estimated mean in matrix patches was 5.202 (SE = 0.665, t = 7.825).
-None of the alternative patch types showed statistically significant differences compared to matrix: connected ($\beta$ = –0.085, SE = 0.495, t = 0.17), rectangle ($\beta$ = 0.234, SE = 0.495, t = 0.472), or winged ($\beta$ = –0.251, SE = 0.495, t = –0.506).
-Similarly, for Simpson's Diversity, the average value in matrix patches was 4.181 (SE = 0.640, t = 6.535).
-Again, none of the other patch types had significant effects: matrix ($\beta$ = -0.107, SE = 0.411, t = -0.259), rectangle ($\beta$ = 0.059, SE = 0.411, t = 0.143), and winged ($\beta$ = –0.480, SE = 0.411, t = –1.169).
+Biodiversity was also even among patch types, however metric values were varied by sampling blocks (Figure 10).
+For Shannon Diversity, the estimated mean in matrix patches was 5.202 (SE = 0.665, t = 7.825; Table 8).
+None of the alternative patch types showed statistically significant differences compared to matrix: connected ($\beta$ = –0.085, SE = 0.495, t = 0.17; Table 8), rectangle ($\beta$ = 0.234, SE = 0.495, t = 0.472; Table 8), or winged ($\beta$ = –0.251, SE = 0.495, t = –0.506; Table 8).
+Similarly, Simpson's Diversity values were fairly even (Figure 11), the average value in matrix patches was 4.181 (SE = 0.640, t = 6.535; Table 9).
+Again, none of the other patch types had significant effects: matrix ($\beta$ = -0.107, SE = 0.411, t = -0.259; Table 9), rectangle ($\beta$ = 0.059, SE = 0.411, t = 0.143; Table 9), and winged ($\beta$ = –0.480, SE = 0.411, t = –1.169; Table 9).
 
@@
 
-```{r, label = sp-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "LEGEND", out.width="70%", fig.align='center'}
+```{r, label = sp-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Total dung beetle biomass collected for each species with sufficient weights.", out.width="70%", fig.align='center'}
 
@@
 
-```{r, label = patch-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "LEGEND", out.width="70%", fig.align='center'}
+```{r, label = patch-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average total biomass by patch type with standard deviation as error bars.", out.width="70%", fig.align='center'}
 
@@
 
-```{r, label = block-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "LEGEND", out.width="70%", fig.align='center'}
+```{r, label = block-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average total biomass by sampling block with standard deviation as error bars.", out.width="70%", fig.align='center'}
 
@@
 <!-- ``` -->
-
-\newpage
-
-```{r bray-curtis, echo=FALSE,message = FALSE,warning=FALSE}
-btl_spec_patch<- btl_data |>
-  pivot_longer(pvin:osyl, names_to = "species", values_to = "n") |>
-  group_by(species,patch) |>
-  summarise(n=sum(n, na.rm=TRUE)) |>
-  arrange(patch,desc(n)) |>
-  pivot_wider(values_from=n,names_from=species) |>
-  mutate(patch=case_when(
-    patch == "c" ~ "Corridor",
-    patch == "m" ~ "Matrix",
-    patch == "w" ~ "Winged",
-    patch == "r" ~ "Rectangle",
-    .default = as.character(patch))) |>
-  column_to_rownames("patch")
-
-taxa_bray <- vegan::vegdist(btl_spec_patch, method = "bray", binary = FALSE)
-
-# taxa_bray
-tb<-as.matrix(taxa_bray)
-tb[lower.tri(tb)] <- 0
-tb<-as.data.frame(tb)
-tb<-tb %>% 
-  mutate_if(is.numeric, round, digits=3) %>% 
-  select(-Corridor)
-
-row.names.remove <- c("Winged")
-tb<-tb[!(row.names(tb) %in% row.names.remove), ]
-tb[tb == 0] <- ""
-col1<-row.names(tb)
-col1<-as.data.frame(col1)
-
-tb<-bind_cols(col1,tb) %>% 
-  rename(" "=col1)
-# bray<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/bray_patches.csv") 
-# bray<-read_csv(here("corridor_docs","eric_ms_thesis","tables","bray_patches.csv")) %>% 
-  bray<-tb %>% 
-  kable(
-      digits = 3,
-      align="lcc",
-      format="latex",
-      row.names = FALSE,
-      booktabs=T,
-      linesep = "", #removes the blank line after every 5 lines
-      caption = "Bray-Curtis dissimilarity comparing composition of dung beetles between patches.") %>%
-  kable_styling(bootstrap_options = c("striped"),
-                full_width = F,
-                latex_options=c("scale_down","hold_position"),
-                font_size = 12,
-                position = "center") %>%
-  # column_spec(1, italic = T) %>%
-  row_spec(0, bold = T) %>% 
-  column_spec(1, bold = T)
-  # row_spec(1:15, hline_after = T)
-bray
-```
 
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "FILL IN CAPTION TOTAL BIOMASS BY PATCH ALL POOLED.") %>%
+      caption = "Linear mixed effects model of pooled total biomass by patch type including sampling block as a random effect.") %>%
   kable_styling(bootstrap_options = c("striped"),
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = paste("FILL IN CAPTION BIOMASS by SP AND PATCH. Species codes = ",species_names, sep="")
+      caption = paste("Linear mixed effects model of biomass by species and patch type including sampling block as a random effect. Species codes = ",species_names, sep="")
       ) %>%
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "AOV patch type effects on species abundance and interavtive term.") %>%
+      caption = "AOV patch type effects on species abundance and interactive term.") %>%
   kable_styling(bootstrap_options = c("striped"),
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "Effects of patch type on the abundance of the 6 most abundant species and interaction term for patch type and species.") %>%
+      caption = "Generalized linear mixed model of abundance by patch type of the 6 most abundant species and interaction term for patch type and species including sampling block as a random effect.") %>%
   kable_styling(bootstrap_options = c("striped"),
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "Richness Model") %>%
+      caption = "Linear mixed effects model of species richness by patch type.") %>%
   kable_styling(bootstrap_options = c("striped"),
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "Shannon Diversity Model") %>%
+      caption = "Linear mixed effects model of Shannon Diversity by patch type.") %>%
   kable_styling(bootstrap_options = c("striped"),
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "Simpson's Index Model") %>%
+      caption = "Linear mixed effects model of Simpson's Diversity by patch type.") %>%
   kable_styling(bootstrap_options = c("striped"),

### 2025-07-25 corridor_docs/eric_ms_thesis/03_biomass_analysis.R
correct axis label

@@ -65,6 +65,21 @@ rename(sp_code=species) %>%
   mutate(species=tolower(species)) %>% 
   mutate(species=paste(genus, species, sep=" ")) 
 
+
+
+# figure of indiv biomass vs abund ----------------------------------------
+
+# are the largest species the least common?
+
+
+
+# Basic scatter plot.
+p1 <- ggplot(biomass_per_spp, aes(x=avg_ind_bmass, y=n)) + 
+  geom_point( color="navyblue") +
+  geom_smooth(method=lm , color="black", se=TRUE) +
+  theme_classic()
+labs(x="Species",y="Total biomass collected")+
+
 sp_biomass_plot<-
 ggplot(biomass_per_spp, 
        aes(
@@ -198,7 +213,7 @@ avg_bmass_block_plot<-
   ) + 
   geom_errorbar(aes(x=block, ymin=avg_bm-sd_bm, ymax=avg_bm+sd_bm), 
                 width=0.2, colour="black", alpha=0.9, size=0.5) +
-  labs(x="Patch Type",y= "Biomass (Mean ± SD)")+
+  labs(x="Block",y= "Biomass (Mean ± SD)")+
   geom_bar(stat = "identity")+
   theme_classic() 
 

### 2025-07-25 corridor_docs/eric_ms_thesis/MS_thesis.Rmd

@@
 Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.
-**Likewise, linear mixed-effects models were used to compare for differences between patch types in (1) total dung beetle biomass, and (2) the total biomass of each of the ten species, again including block as a random effect.
-**
+Likewise, I used linear mixed-effects models to compare for differences between patch types in (1) total dung beetle biomass, and (2) the total biomass of each of the ten species, again including block as a random effect.
 
@@
 
-When comparing the overall abundance of beetles (all species combined) across patch types, matrix plots had the highest captures, followed by connected patches, then winged, with the fewest in rectangular patches (Figure 4).
-Abundances from connected patches were not significantly different from those in the matrix while rectangle and winged patches had significantly fewer than the matrix ($\beta$ = -0.528 and -0.303 respectively, P \< 0.001, Table 2).
+When comparing the overall abundance of beetles (all species combined) across patch types, matrix plots had the highest captures, followed by connected patches, then winged, with the fewest in rectangular patches (Figure \@ref(fig:patch)).
+Abundances from connected patches were not significantly different from those in the matrix while rectangle and winged patches had significantly fewer than the matrix ($\beta$ = -0.528 and -0.303 respectively, P \< 0.001, Table \@ref(tab:Table-MO)).
 
 Statistical analysis of abundance focused on the six most abundant species.
-A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 6).
+A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table \@ref(tab:Table-M3)).
 The baseline abundance corresponds to the abundance of *Aphodius alloblackburneus* in matrix patches.
 Compared to this baseline, results were highly variable, emphasizing species specific responses to patch type.
-Similarly, species showed to have disproportionate responses to patch since interaction terms varied widely (Table 6).
+Similarly, species showed to have disproportionate responses to patch since interaction terms varied widely (Table \@ref(tab:Table-M3)).
 
@@
 
-Plotting species richness by patch type reveals consistent richness across patch types with some variation between sampling blocks (Figure 9).The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N).
+Plotting species richness by patch type reveals consistent richness across patch types with some variation between sampling blocks (Figure \@ref(fig:richness)).The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N).
 Modeling the effect of patch type on species richness with block as a random effect determined there was no significant differences among patch types.
-Comparing treatments using matrix patches as a baseline resulted in no significant differences in connected ($\beta$ = 0.216, p = 0.839; Table 7), rectangle ($\beta$ = -0.548, p = 0.583; Table 7), and winged patches ($\beta$ = -0.095, p = 0.663; Table 7).
-
-Biodiversity was also even among patch types, however metric values were varied by sampling blocks (Figure 10).
-For Shannon Diversity, the estimated mean in matrix patches was 5.202 (SE = 0.665, t = 7.825; Table 8).
-None of the alternative patch types showed statistically significant differences compared to matrix: connected ($\beta$ = –0.085, SE = 0.495, t = 0.17; Table 8), rectangle ($\beta$ = 0.234, SE = 0.495, t = 0.472; Table 8), or winged ($\beta$ = –0.251, SE = 0.495, t = –0.506; Table 8).
-Similarly, Simpson's Diversity values were fairly even (Figure 11), the average value in matrix patches was 4.181 (SE = 0.640, t = 6.535; Table 9).
-Again, none of the other patch types had significant effects: matrix ($\beta$ = -0.107, SE = 0.411, t = -0.259; Table 9), rectangle ($\beta$ = 0.059, SE = 0.411, t = 0.143; Table 9), and winged ($\beta$ = –0.480, SE = 0.411, t = –1.169; Table 9).
+Comparing treatments using matrix patches as a baseline resulted in no significant differences in connected ($\beta$ = 0.216, p = 0.839; Table \@ref(tab:Table-M-rich)), rectangle ($\beta$ = -0.548, p = 0.583; Table \@ref(tab:Table-M-rich)), and winged patches ($\beta$ = -0.095, p = 0.663; Table \@ref(tab:Table-M-rich)).
+
+Biodiversity was also even not significantly different between patch types, however the values of the different metrics varied significantly across sampling blocks (Figures \@ref(fig:shannon)-\@ref(fig:simpson)).
+For Shannon Diversity, the estimated mean in matrix patches was 5.202 (SE = 0.665, t = 7.825; Table \@ref(tab:Table-M-shan)).
+None of the alternative patch types showed statistically significant differences compared to matrix: connected ($\beta$ = –0.085, SE = 0.495, t = 0.17), rectangle ($\beta$ = 0.234, SE = 0.495, t = 0.472), or winged ($\beta$ = –0.251, SE = 0.495, t = –0.506; Table \@ref(tab:Table-M-shan)).
+Results for Simpson's Diversity were similar (Table \@ref(tab:Table-M-simp)), with the average value in matrix patches was 4.181 (SE = 0.640, t = 6.535).
+Again, none of the other patch types were significantly different from the others: matrix ($\beta$ = -0.107, SE = 0.411, t = -0.259), rectangle ($\beta$ = 0.059, SE = 0.411, t = 0.143), and winged ($\beta$ = –0.480, SE = 0.411, t = –1.169; Table \@ref(tab:Table-M-simp)).
 
 ### Dung Beetle Biomass
-
-Pooling dung beetle biomass by patch type follows a pattern very similar to abundance where biomass is the highest in matrix plots, followed by connected, then winged, with a large drop in rectangle patches (Figure 7).
-When comparing to biomass collected in matrix plots, statistical analysis showed a non-significant relationship with connected ($\beta$ = -7.337, SE = 8.659, t = -0.847; Table 4) and winged patches ($\beta$ = -9.608, SE = 8.659, t = -1.11; Table 4), while significant results were returned for rectangle patches ($\beta$ = -23.296, SE = 8.659, t = -2.69; Table 4).
-Block effects continue to be important for explaining high variation as the total biomass collected in block 52 nearly quadrupled that of block 8 (Table 8).
-
-Parsing biomass by species emphasizes the morphological variability between our study species as some of the most common contribute very little to total biomass (i.e.
-*Ateuchus lecontei*, *Aphodius allobloackburneus*; Figure 6).
-Linear mixed effects models further showcases these species specific differences (Table 5).
+Patterns of biomass by patch type were similar to those for abundance: biomass was highest in matrix plots, followed by connected and winged, with biomass in rectangle patches being far lower than in the other locations (Figure \@ref(fig:patch-bmass)).
+There was no significant difference in the total beetle biomass of matrix plots when compared to connected ($\beta$ = -7.337, SE = 8.659, t = -0.847; Table \@ref(tab:Table-aovM51)) and winged patches ($\beta$ = -9.608, SE = 8.659, t = -1.11; Table \@ref(tab:Table-aovM51)), but biomass in rectangle patches was significantly different ($\beta$ = -23.296, SE = 8.659, t = -2.69; Table \@ref(tab:Table-aovM51)).
+Block effects continue to be important for explaining high variation as the total biomass collected in block 52 was nearly quadruple that collected in Block 8 (Figure \@ref(fig:block-bmass)).
+Linear mixed effects models also indicate significant differences between the biomass of different species  (Table \@ref(tab:Table-aovM52), Figure \@ref(fig:sp-bmass)).
 
@@
 Another potential explanation is that species in the matrix are acting as a source population which feeds into patches with edge acting as a drift fence for directing movement.
-The latter example seems more likely because flies, which have similar resource dependencies were also found to have a similar interaction [@fried_habitat_2005], and also studies on dung beetles in the tropics found more dramatic differences in dung beetle populations in fragments and matrix [@WOS:000288810500026].
+The latter example seems more likely - small mammals at SRS, which are a potential source of dung, appear to preferentially use matrix habitat [@mabry_influence_2003] and flies, which have similar resource dependencies were also found to have a similar interaction [@fried_habitat_2005].
+Moreover studies on dung beetles in the tropics found more dramatic differences in dung beetle populations in fragments and matrix [@WOS:000288810500026].
 <!-- Paragraph 3: }What then are the drivers of dung beetle diversity and abundance? -->
@@
 Block 8 generally had the lowest species richness and biodiversity while 53n had the highest.
-While the design of the experimental blocks control for as much as possible, landscape features across the sight are different enough to influence species composition.
-Several factors may explain this pattern, at the time this study was conducted block 8 had the densest matrix of any of the other sampling blocks which could have hindered the scent permeability of the bait.
-Other habitat characteristics might have been influential as well, including soil quality and forest cover which can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019].
-Furthermore in the context of my study site, land-use history could be an important driver to consider as much of the land would have previously been used for agriculture and during development of experimental units cleared with heavy machinery might experience heavy soil compaction *brudvig*.
-Another important consideration to take is that these landscape features can affect mammal movement, which in turn limits dung availability for beetles.
-In this regard it is also important to consider other geographical differences, block 8 was nearest to roadways used by all employees at SRS and the eastern side of the sight had large bodies of water.
-Both of which would effect the land-use by larger mammals [@WOS:000241362400032; @dechen_quinn_landscape_2013; @WOS:000651446900001].
+While the experimental design attempted to control for the effects of patch size and edge, there could be large (and potential unknown) environmental gradients across the SRS landscape that could influence the observed patterns in diversity and abundance.
+For instance, at the time of my study Block 8 had the densest matrix of any of the blocks.
+This could have hindered the diffusion of bait scent, leading to lower capture rates in this block.
+Other habitat characteristics that might dsiffer among blocks could have been influential as well -- for example, soil quality and forest cover can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019].
+The same is true for land-use history; much of the SES land would have previously been used for agriculture, and during development of experimental units cleared with heavy machinery might experience heavy soil compaction *brudvig*.
+Landscape features could also affect mammal movement, which might in turn limit dung availability - Block 8 was nearest to roadways used by all employees at SRS, and the eastern side of the site had large bodies of water upon which mammals are dependent [@WOS:000241362400032; @dechen_quinn_landscape_2013; @WOS:000651446900001].
 Additionally, deer and feral hogs are the most dominant mammals at the site, two species that do well in patchy landscapes [@castillo-contreras_urban_2018; @fraser_connectivity_2019].
-This might explain the lack of variation in richness or diversity in patches while differences are more apparent at greater distances across the site.
-**[need to add and discuss soil compaction at 8 here in this paragraph: 1-2 sentences]**
-
-<!-- Paragraph 4: That said, there were some interspecific differences in abundance.  -->
-
-Species specific differences are apparent but do not follow any particular trend.
+The locally variable patterns of abundance in the landscape of these generalist large mammals, coupled with ability to easily move through modified landscapes, might help explain the limited variation in richness or diversity within blocks (i.e., between patch types) but large block effects.
+Deer also commonly forage along the edges of habitat patches and forest ecotones.
+If they are spending more time in these locations, the higher abundance and diversity in connected and winged patches might in part be due to  'drift-fence' effects **citation of levey drift fence**.
+
+Species-specific differences are apparent but do not follow any particular trend.
 _Aphodius alloblackburneus_ had a disproportionate positive effect to being captured in matrix patches as compared to other species.
-Beetles from the genus _Aphodius_ do tend to show patterns of habitat specificity [@roslin_distribution_2001], and many preferring forested habitat[@frank_land_2017],  so it is not unexpected that they might show a preference towards forested matrix.
-I did not detect that any species was more positively associated with open patches despite expectations that some species (e.g., _Canthon vigilans_, _Melanocanthon bispinatus_) would prefer open habitat [@conover_phenology_2019; @nealis_habitat_1977].
-Altogether this could be another sign pointing towards matrix acting as a source population, and since open habitat was much less dominant in my experimental system beetles could be moving into patches from habitat edge.
+Beetles from the genus _Aphodius_ do tend to show patterns of habitat specificity [@roslin_distribution_2001], and many preferring forested habitat [@frank_land_2017],  so it is not unexpected that they might show a preference towards forested matrix.
+I did not detect that any species was more positively associated with open patches despite suggestions that some species (e.g., _Canthon vigilans_, _Melanocanthon bispinatus_) prefered open habitat [@conover_phenology_2019; @nealis_habitat_1977].
+This could be another sign pointing towards matrix acting as a source population, and since open habitat was much less dominant in my experimental system beetles could be moving into patches from habitat edge.
 
@@
 
-The patterns in dung beetle biomass largely echo what we observed for abundance, however this distinction is important to make since dung beetles vary greatly in terms of morphology and functionality [@ospina-garces_dung_2018].
-Understanding these patterns is of particular interest since biomass is understood to be positively associated with dung removal [@WOS:000287168100017].
+The patterns in dung beetle biomass largely echo what we observed for abundance.
+Although biomass is understood to be positively associated with dung removal [@WOS:000287168100017], dung beetle species vary greatly in terms of morphology and functionality [@ospina-garces_dung_2018] so evaluating species-specific patterns is particularly important.
 Because of this we can expect that the magnitude of removal might be greater in areas of larger biomass (i.e.
 matrix or connected patches), however future work should aim to directly measure this potential pattern.
+Indeed comparing the biomass of different species emphasizes the morphological variability between our study species - some of the most common species (e.g., *Ateuchus lecontei*, *Aphodius allobloackburneus*) contribute very little to total biomass (Figure \@ref(fig:sp-bmass)).
 In addition to this, dung beetles appear to be an exception to the globally and taxonomically robust rule that abundance is negatively correlated with individual biomass [@white_relationships_2007].
@@
 
-Despite ample work documenting their ecological importance, there is a surprising lack of research on dung beetle diversity and corridors outside of the tropics [@nichols_global_2007], with any work in temperate locations coming from Europe where _Aphodius_ beetles dominate [@roslin_distribution_2001].
-Many studies in the tropics have found patterns of habitat preference between open fields and continuous forest similar to what I did in this study [@WOS:000349242000004], and some work that found that forest species were using living fences as corridors [@arellano_response_2008].
-However, differences in composition were much more distinctly different with some species being exclusive to habitat types unlike I see in this paper.
-Of the limited work out of the south eastern United States I did recover assemblages similar to other work [@conover_phenology_2019; @nealis_habitat_1977], and even returned results similar to another study where differences between treatment types were non-significant while variation on a larger scale was more apparent [@young_dung_2023].
-In comparison to other studies from SRS my results were similar to those of small mammals which were more likely to be using matrix land [@mabry_influence_2003].
+Despite ample work documenting their ecological importance, there is a surprising lack of research on dung beetle diversity and corridors outside of the tropics [@nichols_global_2007], with work in temperate locations coming primarily from Europe where _Aphodius_ beetles dominate [@roslin_distribution_2001].
+Many studies conducted in the tropics found strong preference by some species for open fields vs.
+continuous forest [@WOS:000349242000004], while othersfound that forest species were using living fences as corridors [@arellano_response_2008].
+I did not observe such intense habitat specificity, which may not be especially common outside of the tropics.
+I found similar patterns of community composition as did the small number of prior dung beetle biodiversity surveys conducted in the southeastern United States [@conover_phenology_2019; @nealis_habitat_1977], including at least one more where differences between treatment types were non-significant while variation on a larger scale was more apparent [@young_dung_2023].
 I did expect some species to be more dominant in patches with higher edge ratios as was found with ants at SRS [@resasco_landscape_2014], but species were evenly present in all patches.
@@
 The fact that rectangular are have lower abundance suggests that some sort of landscape effect is apparent, likely related to habitat edge since connected and winged patches were the most similar.
-To remedy this issue any future work in this site should either focus solely on dung beetles with less dispersal ability or consider mark release recapture sampling in order to better isolate movements within and between patches.
+To remedy this issue any future work in this site should either focus solely on dung beetles with more limited dispersal ability or consider conducting mark-release-recapture experiments in an attempt to document movements within and between patches.
 
@@
 
-```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index y patch type.", out.width="70%", fig.align='center'}
+```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index by patch type.", out.width="70%", fig.align='center'}
 
@@
       linesep = "", #removes the blank line after every 5 lines
-      caption = "AOV patch type effects on species abundance and interactive term.") %>%
+      caption = "Results of an analysis of assesing the effects of species identity, patch type, and their interaction on the abundance of dung beetle species.") %>%
   kable_styling(bootstrap_options = c("striped"),

### 2025-07-25 corridor_docs/eric_ms_thesis/03_biomass_analysis.R
bmass_v_abund

@@ -74,11 +74,23 @@ rename(sp_code=species) %>%
 
 
 # Basic scatter plot.
-p1 <- ggplot(biomass_per_spp, aes(x=avg_ind_bmass, y=n)) + 
+bmass_v_abund <- ggplot(biomass_per_spp, aes(x=avg_ind_bmass, y=n)) + 
   geom_point( color="navyblue") +
-  geom_smooth(method=lm , color="black", se=TRUE) +
-  theme_classic()
-labs(x="Species",y="Total biomass collected")+
+  geom_smooth(method=lm , color="black", se=FALSE) +
+  geom_text(
+    label=biomass_per_spp$sp_code, 
+    nudge_x = 0.025, nudge_y = 0.025, 
+    check_overlap = T
+  )+
+  theme_classic()+
+labs(x="Avg. Individual Biomass",y="Total Number Captured")
+  
+
+ggsave("corridor_docs/eric_ms_thesis/images/bmass_v_abund.png", width = 4, height = 4, units = "in")
+
+# total biomass per species plot ------------------------------------------
+
+
 
 sp_biomass_plot<-
 ggplot(biomass_per_spp, 

### 2025-07-25 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
doc formating

@@
 bibliography: ["eric-thesis.bib"]
-output:
-  bookdown::pdf_document2:
-    toc: false
-    number_sections: true
-    fig_caption: true
-classoption: 12pt
-# output: bookdown::word_document2
+#output:
+#  bookdown::pdf_document2:
+#    toc: false
+#    number_sections: true
+#    fig_caption: true
+#classoption: 12pt
+output: bookdown::word_document2
 # output: word_document
@@
 This could have hindered the diffusion of bait scent, leading to lower capture rates in this block.
-Other habitat characteristics that might dsiffer among blocks could have been influential as well -- for example, soil quality and forest cover can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019].
+Other habitat characteristics that might differ among blocks could have been influential as well -- for example, soil quality and forest cover can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019].
 The same is true for land-use history; much of the SES land would have previously been used for agriculture, and during development of experimental units cleared with heavy machinery might experience heavy soil compaction *brudvig*.

### 2025-07-25 corridor_docs/eric_ms_thesis/MS_thesis.Rmd

@@
 #classoption: 12pt
-output: bookdown::word_document2
+# output: bookdown::word_document2
 # output: word_document
@@
 
-# output: bookdown::pdf_document2
-  # pdf_document:
-  #   includes:
-  #     # in_header: style/preamble.tex
-  #     template: null
-  #     keep_tex: no
-  #     extra_dependencies: ["float"]
-      # number_sections: true
+output: bookdown::pdf_document2
+pdf_document:
+  includes:
+    # in_header: style/preamble.tex
+    template: null
+    keep_tex: no
+    extra_dependencies: ["float"]
+    number_sections: true
 # documentclass: style/nsf2

### 2025-08-01 corridor_docs/eric_ms_thesis/EscobarChena_etal_ms.Rmd
reorg of repo for MS

--- 2025-07-25 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
@@
 ---
-title: "The Influence of Connectivity on Dung Beetle Communities"
-author: "ERIC ESCOBAR-CHENA"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-# date: "\\vspace{-1em}"
-# header-includes:
- # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
-#  - \usepackage{lineno}
-#  - \linenumbers
-#  - \usepackage[default]{sourcesanspro}
-#  - \usepackage{fancyhdr} # modify header into left, center, right
-#  - \pagestyle{fancy} # cues the use of fancy headers
-#  - \fancyfoot{} # cues the use of fancy footer
-#  - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
-#  - \fancyhead[R]{p. \thepage} # name and page on header right
-#  - \usepackage{subcaption}
-#  - \usepackage{float}
-# # # - \fancyfoot[R]{ }
-#  - \usepackage{setspace}
+title             : "Effect of experimental habitat connectivity on dung beetle communities"
+shorttitle        : "dung beetles and habitat corridors"
+
+author: 
+  - name          : "Eric Escobar-Chena"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    address       : "University of Florida"
+    email         : "eescobarchena@ufl.edu"
+    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
+      - Conceptualization
+      - Methodology
+      - Investigation
+      - Formal analysis
+      - Data curation
+      - Software
+      - Visualization
+      - Writing - Original Draft Preparation
+      - Writing - Review & Editing
+  - name          : "Julian Resasco"
+    affiliation   : "2"
+
+    role:
+      - Conceptualization
+      - Methodology
+      - Writing - Original Draft Preparation
+      - Writing - Review & Editing
+
+  - name          : "Emilio M. Bruna"
+    affiliation   : "1,3"
+    role:
+      - Conceptualization
+      - Methodology
+      - Investigation
+      - Formal analysis
+      - Data curation
+      - Visualization
+      - Writing - Original Draft Preparation
+      - Writing - Review & Editing
+      - Project administration
+      - Supervision
+
+affiliation:
+  - id            : "1"
+    institution   : "Department of Wildlife Ecology & Conservation, University of Florida, PO Box 110530, Gainesville, Florida, 32611-0430 USA"
+  - id            : "2"
+    institution   : "University of Colorado, Boulder"
+  - id            : "3"
+    institution   : "Center for Latin American Studies, University of Florida, PO Box 110530, Gainesville, Florida, 32611-0530 USA"
+
+note:       | 
+  A preprint of this article has been posted at ----- Preprints (link).
+
+  ORCID:
+    EEC: ---- |
+    EB: 0000-0003-3381-8477 |
+    JR: ----- |
+
+    \newpage
+
+authornote: |
+  <!-- EEC: Department of Wildlife Ecology & Conservation, University of Florida, PO Box 110530, Gainesville, Florida, 32611-0430 USA -->
+  <!-- JR: University of Colorado, Boulder -->
+  <!-- EB: Department of Wildlife Ecology & Conservation, University of Florida, PO Box 110530, Gainesville, Florida, 32611-0430 USA and Center for Latin American Studies, University of Florida, PO Box 110530, Gainesville, Florida, 32611-0530 USA -->
+
+    We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site.
+We also wanted to specifically thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry.
+Funding was provided by -----.
+
+abstract: |
+
+  Habitat fragmentation threatens biodiversity across the globe as habitat loss, isolation, and edge effects become increasingly prevalent.
+Corridors have become an important tool in order to combat the negative effects of fragmentation, however they are difficult to study in natural systems without incurring confounding effects.
+To observe changes in insect community composition as an effect of landscape features we sampled dung beetles in a landscape scale experiment.
+We did not see a difference in species richness or diversity, but dung beetle abundances were higher in continuous forest habitat and open habitat patches connected by a corridors than in isolated patches.
+
+  <!-- https://tinyurl.com/ybremelq -->
+
+keywords          : "keywords"
+wordcount         : "4437"
+
+bibliography      : "eric-thesis.bib"
+csl               : "ecology.csl"
+floatsintext      : no
+linenumbers       : yes
+draft             : no
+mask              : no
+
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+replace_ampersands: yes
+
+classoption       : "man, donotrepeattitle"
+output            : papaja::apa6_pdf
+
+fig_caption: yes #yes will automatically add "Figure X" to the start 
+# of the caption and reorders automatically if the figure position changes
+always_allow_html: yes
 
 header-includes:
-  - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
-  - \usepackage{float} #to keep figure in the location placed instead of end
-  - \floatplacement{figure}{H}
-  - \usepackage{booktabs}
   - \raggedbottom
-  - \usepackage{fancyhdr}
-  - \pagestyle{plain}
-  - \fancyfoot[C]{\thepage}
-  - \fancyhead[L]{}
-  - \fancyhead[R]{\tiny{EEC MS Thesis Draft (`r format(Sys.time(), '%d %B %Y')`)}}
-  - \thispagestyle{fancy}
+  - \usepackage{endfloat} #[nomarkers] excludes the {insert figure x around here] from main text.
+The others exclude the list of tables and figures.
+https://cs.brown.edu/about/system/managed/latex/doc/endfloat.pdf
+  - \usepackage{setspace}\doublespacing
+  - \usepackage{lineno}
+  - \linenumbers
   - \usepackage{tabu}
-  - \usepackage{makecell}
   - \usepackage{pdflscape}
-  - \usepackage{longtable}
-  - \usepackage{lscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
-  - \setlength\parindent{22pt}
-  - \usepackage{indentfirst}
-  # - \usepackage{caption}
-  # - \captionsetup[figure]{name={Figure 1-},labelsep=period}
-  # - \captionsetup[table]{name={Table 1-},labelsep=period}
-  - \renewenvironment{abstract}{}{} #suppress the abstract
-  - \usepackage{setspace}
-  - \setcounter{secnumdepth}{0} # removes numbering from sections and subsections
   - \usepackage{sectsty}\allsectionsfont{\raggedright}
-  # - \usepackage{sectsty}\sectionfont{\centering}{\normalfont} # center # headings
-  # - \usepackage{sectsty}\subsectionfont{\centering}{\normalfont} # center ## headings
-  - \usepackage{sectsty}\subsubsectionfont{\bfseries} # center ## headings
-  # - \usepackage{sectsty}\subsubsectionfont{\itshape} # no centering but italic subsu
-  # - \usepackage{sectsty}\subsubsectionfont{\centering}{ital} # center ### headings
-  # - \usepackage{sectsty}\sectionfont{\centering\color{darkmidnightblue}}
-
- # - doublespacing # manipulate line spacing 
-# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
-# - \usepackage{xcolor} # changes colors of text
-# # http://latexcolor.com/ to choose colors
-# # define the colors used above with rgb codes
-# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
-# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
-# # to center to top level (#) headings (my name) and change color of font
-# # more info # https://www.ctan.org/pkg/sectsty
-# - \usepackage{sectsty}
-# - \sectionfont{\centering\color{darkmidnightblue}}
-# # to center to sub-sub-headings (###) and change color of font
-# - \usepackage{sectsty} 
-# - \subsectionfont{\centering\color{darkmidnightblue}}
-# - \usepackage{sectsty} 
-# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
-# - \usepackage{wrapfig} 
-# - \newcommand{\blandscape}{\begin{landscape}}
-# - \newcommand{\elandscape}{\end{landscape}}
-# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
-# - \usepackage{indentfirst}
-# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
-bibliography: ["eric-thesis.bib"]
-#output:
-#  bookdown::pdf_document2:
-#    toc: false
-#    number_sections: true
-#    fig_caption: true
-#classoption: 12pt
-# output: bookdown::word_document2
-# output: word_document
-# output: html_document
-
-output: bookdown::pdf_document2
-pdf_document:
-  includes:
-    # in_header: style/preamble.tex
-    template: null
-    keep_tex: no
-    extra_dependencies: ["float"]
-    number_sections: true
-# documentclass: style/nsf2
-biblio-style: apalike
-csl: ecology.csl
+
 ---
 
-```{r GlobalOptions, include = FALSE}
-options(knitr.duplicate.label = 'allow')
-options(knitr.graphics.error = FALSE)  # This allows for the include_graphics to not use the path from root 
-knitr::opts_chunk$set(echo = TRUE, fig.align="center")
-knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
-knitr::opts_chunk$set(tab.pos = "H", out.extra = "")
-```
-
 ```{r setup, include = FALSE}
+library("papaja")
+
+# r_refs("
+# here("corridor_docs","eric_ms_thesis","eric-thesis.bib")
+# ")
 
@@
 
+```{r GlobalOptions, include = FALSE}
+options(knitr.duplicate.label = 'allow')
+options(knitr.graphics.error = FALSE)  # This allows for the include_graphics to not use the path from root 
+knitr::opts_chunk$set(echo = TRUE, fig.align="center")
+knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
+knitr::opts_chunk$set(tab.pos = "H", out.extra = "")
+```
+
 ```{r load-data, cached=TRUE, echo = FALSE, warning=FALSE, message = FALSE}
@@
 ```
-\newpage
-
-\vspace*{5cm}
-
-\begin{center}
-THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES
-\end{center}
-
-\vspace{5cm}
-
-\begin{center}
-By
-\end{center}
-
-\vspace{1cm}
-
-\begin{center}
-ERIC ESCOBAR-CHENA
-\end{center}
-
-\vspace{5cm}
-
-\begin{center}
-A THESIS PRESENTED TO THE GRADUATE SCHOOL\\  
-OF THE UNIVERSITY OF FLORIDA IN PARTIAL FULFILLMENT\\ 
-OF THE REQUIREMENTS FOR THE DEGREE OF\\  
-MASTER OF SCIENCE\\
-\end{center}
-
-\begin{center}
-UNIVERSITY OF FLORIDA
-\end{center}
-
-\begin{center}
-2025
-\end{center}
-
-\newpage
-
-\vspace*{11cm}
-
-\begin{center}
-© 2025 Eric Escobar-Chena
-\end{center}
-
-\newpage
-
-\vspace*{10cm} 
-<!-- _Your dedication is typed here.
-It should begin with the word “To.” (To my Mom is a typical dedication) If your dedication is longer than a single line it should be single-spaced, and centered vertically and horizontally_ -->
-
-\begin{center}
-To my mom, dad, and two sisters. This thesis is a chapter in our immigrant story. 
-\end{center}
-
-\newpage
-
-\begin{center}
-
-\textbf{ACKNOWLEDGMENTS}
-
-\end{center}
-<!-- _Acknowledgments must be written in complete sentences. Do not use direct address. For example, instead of Thanks, Mom and Dad!, you should say I thank my parents._  -->
-\doublespacing
-
-I thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site.
-I also wanted to specifically thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry.
-
-\newpage
-
-\begin{center}  
-\textbf{LIST OF ABBREVIATIONS}
-\end{center}
-
-\vspace{1cm}
-
-\begin{flushleft}
-
-SRS: Savannah River Site.
-
-GLMM: Generalized Linear Mixed Model.
-
-\end{flushleft}
-
-\newpage
-
-\vspace*{5cm}
-
-\singlespacing
-
-\begin{center}  
-Abstract of Thesis Presented to the Graduate School\\  
-of the University of Florida in Partial Fulfillment of the\\  
-Requirements for the Degree of Master of Science.\\   
-\end{center}
-
-\vspace*{1cm}
-
-\begin{center}
-THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES
-\end{center}
-
-\begin{center}
-By
-\end{center}
-
-\begin{center}
-Eric Escobar-Chena   
-\end{center}
-
-\begin{center}
-August 2025   
-\end{center}
-
-\begin{flushleft}
-Chair: Emilio Bruna\\  
-Major: Wildlife Ecology and Conservation\\   
-\end{flushleft}
-
-\doublespacing
-
-Habitat fragmentation threatens biodiversity across the globe as habitat loss, isolation, and edge effects become increasingly prevalent.
-Corridors have become an important tool in order to combat the negative effects of fragmentation, however they are difficult to study in natural systems without incurring confounding effects.
-To observe changes in insect community composition as an effect of landscape features I sampled dung beetles in a landscape scale experiment.
-I did not see a difference in species richness or diversity, but dung beetle abundances were higher in continuous forest habitat and open habitat patches connected by a corridors than in isolated patches.
-
-\newpage
-
-\begin{center}
-CHAPTER 1\\  
-INTRODUCTION\\  
-\end{center}
-
-\doublespacing
-
-Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes [@haddad_corridor_2003].
-These movements are hypothesized to prevent species diversity from declining in fragments, as well as help maintain the ecosystem services provided by these species (at both the patch- and landscape-level)[@burt_ants_2022].
-Although there is some evidence that animals disperse between patches via corridors, and that connected patches have higher species diversity than unconnected ones, little work to date has investigated the consequences of these corridor-driven patterns for ecosystem services.
-
-\newpage
-
-\begin{center}
-CHAPTER 2\\  
-THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES\\  
-\end{center}
-
-\doublespacing
+
+## INTRODUCTION
 
@@
 
-\begin{center} 
-\textbf{Methods}
-\end{center}
+## METHODS
+
 ### Study site
@@
 
-\begin{center} 
-\textbf{Results}
-\end{center}
+## RESULTS
 
@@
 
-\begin{center} 
-\textbf{Discussion}
-\end{center}
+## DISCUSSION
+
 <!-- Paragraph 1: Overview: While there may indeed be effects of patch area and connectivity on dung beetle abundance on community composition, other (landscape-scale) drivers appear to be more important.
@@
 
-**[Paragraph 7: spatial scale of experimental plots vs. dung detectability]**
-
 Finally, it is important to emphasize that while dung beetles are capable of long distance flight and detecting dung at distances of over 50 meters [@WOS:000707546700001], the results I observed are likely the result of a mismatch between the spatial scale of the experimental replicates and dung beetle movement.
@@
 
-\newpage
-
-\begin{center}
-CHAPTER 3\\  
-CONCLUSIONS AND FUTURE DIRECTIONS\\  
-\end{center}
-
-\doublespacing 
 Moving forward, there is a bounty of knowledge yet to be collected for dung beetles in the south eastern United States.
@@
 
-\newpage
-
-\begin{center} 
-\textbf{References}
-\end{center}
-
-\smallskip
+## LTERATURE CITED
 
@@
 :::
-
-\newpage
-
-<!-- (text - Nathan Fig of plots) -->
-
-\newpage
-
-<!-- ```{r, label = pitfalls, echo = FALSE, message = FALSE, warning = FALSE,fig.cap = 'Experimental landscapes indicating location of the pitfall traps.',  out.width="25%", fig.align='center'} -->
-
-<!-- a<-readPNG("images/pitfalls.png") -->
-
-<!-- b<-readPNG("images/srs-map.png") -->
-
-<!-- grid.arrange(rasterGrob(a),rasterGrob(b),ncol=1) -->
-
-<!-- ``` -->
-
-```{r echo=FALSE, fig.align='center', fig.cap="All experimental blocks at SRS.
-Filled squares indicate the blocks in which dung beetles were sampled.
-Each block has a unique ID number.", message=FALSE, warning=FALSE, label=srs-map, out.width="60%"}
-
-knitr::include_graphics("images/srs-map.png")
-
-```
-
-```{r, label = patches, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Overhead view of one of the experimental blocks indicating the different types of patches and the distance between them (from Nathan et al.
-2018 and reused with permission).", out.width="70%", fig.align='center'}
-
-knitr::include_graphics("images/patches-pic.png")
-
-```
-
-```{r, label = pitfalls, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Diagram indicating the placement pitfall traps in each patch and in the matrix.
-Traps in patches were approximately 40 m from the patch edge.", out.width="70%", fig.align='center'}
-
-knitr::include_graphics("images/pitfalls.png")
-
-```
-
-<!-- ```{r Figure_hill, echo=FALSE,message = FALSE,warning=FALSE} -->
-
-<!-- library(iNEXT) -->
-
-<!-- # -->
-
-<!-- # btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv")) -->
-
-<!-- # spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv")) -->
-
-<!-- btl_sums_patch<-btl_data %>% -->
-
-<!--   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% -->
-
-<!--   group_by(species,patch,block) %>% -->
-
-<!--   mutate(patch=case_when( -->
-
-<!--     patch == "c" ~ "Corridor", -->
-
-<!--     patch == "m" ~ "Matrix", -->
-
-<!--     patch == "w" ~ "Winged", -->
-
-<!--     patch == "r" ~ "Rectangle", -->
-
-<!--     .default = as.character(patch))) %>% -->
-
-<!--   summarize(n=sum(n, na.rm=TRUE)) %>% -->
-
-<!--   arrange(patch,desc(n)) %>% -->
-
-<!--   rename(sp_code=species, -->
-
-<!--          n_patch=n) %>% -->
-
-<!--   mutate(pb=paste(patch,block,sep="-")) %>%  -->
-
-<!--   ungroup() %>%  -->
-
-<!--   select(-patch,-block) %>%  -->
-
-<!--   pivot_wider( -->
-
-<!--     names_from = pb, -->
-
-<!--     values_from = c(n_patch))  -->
-
-<!-- div.data <- btl_sums_patch |> -->
-
-<!--   column_to_rownames(var = "sp_code") -->
-
-<!-- div.btl <- iNEXT(div.data, q = c(0,1,2), datatype = "abundance") -->
-
-<!-- z <- iNEXT(div.data, q=c(0,1,2), datatype="abundance") -->
-
-<!-- ggiNEXT(z, facet.var="Order.q", color.var="Assemblage") + -->
-
-<!--   theme_bw(base_size=18) -->
-
-<!-- ``` -->
-
-\newpage
-```{r, label = patch, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average total dung beetle abundance by patch type with standard deviations as error bars.", out.width="70%", fig.align='center'}
-
-knitr::include_graphics("images/total_abund.png")
-
-```
-
-```{r spcodes}
-
-sp_code_table<-data.frame(sp_code=c("aaeg", "cvig", "dcar", "open", "pign", "alec"))
-species_names<-spp_codes %>% 
-  filter(sp_code%in%sp_code_table$sp_code) %>% 
-  mutate(species=tolower(species)) %>% 
-  mutate(species=paste(species=genus,species,sep=" ")) %>% 
-  select(sp_code, species) %>% 
-mutate(species=paste(sp_code,species,sep=": ")) 
-species_names<-paste(species_names$species,collapse=", ")
-
-```
-
-```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = paste("Average abundance of the top 6 most abundant species by patch type.
-Species codes: ", species_names,".",sep=""), out.width="70%", fig.align='center'}
-
-knitr::include_graphics("images/sp_abund_patch.png")
-
-```
-
-```{r, label = sp-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Total dung beetle biomass collected for each species with sufficient weights.", out.width="70%", fig.align='center'}
-
-knitr::include_graphics("images/sp_biomass_plot.png")
-# knitr::include_graphics("images/sp_abundance_plot.png")
-# grid.arrange(p1[1], p2[1], nrow = 1)
-
-```
-
-```{r, label = patch-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average total biomass by patch type with standard deviation as error bars.", out.width="70%", fig.align='center'}
-
-knitr::include_graphics("images/avg_bmass_patch_plot.png")
-```
-
-```{r, label = block-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average total biomass by sampling block with standard deviation as error bars.", out.width="70%", fig.align='center'}
-
-knitr::include_graphics("images/avg_bmass_block_plot.png")
-```
-
-```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle species richness in each patch type.
-The point shapes indicates the block in which each patch was located.", out.width="70%", fig.align='center'}
-
-knitr::include_graphics("images/sp_richness_patches.png")
-```
-
-```{r, label = shannon, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Shannon diversity by patch type.", out.width="70%", fig.align='center'}
-
-knitr::include_graphics("images/sp_shannon_patches.png")
-
-```
-
-```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index by patch type.", out.width="70%", fig.align='center'}
-
-knitr::include_graphics("images/sp_simpson_patches.png")
-
-```
-
-<!-- ```{r sp_richness_patches, echo = FALSE, fig.cap = "Dung beetle secies richness in three different --- and the forest matrix surrounding patches.", out.width = '100%'} -->
-
-<!-- knitr::include_graphics("./corridor_beetles/corridor_docs/eric_ms_thesis/images/sp_richness_patches.png") -->
-
-<!-- ``` -->
 
@@
 \newpage
-
-<!-- ```{r abundance-patch, echo=FALSE,message = FALSE,warning=FALSE} -->
-<!-- # https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html -->
-
-<!-- btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv")) -->
-<!-- spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv")) -->
-
-<!-- btl_sums_patch<-btl_data %>% -->
-<!--   select(-sum) %>% -->
-<!--   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% -->
-<!--   group_by(patch) %>% -->
-<!--   summarize(n=sum(n, na.rm=TRUE)) %>% -->
-<!--   arrange(patch,desc(n)) %>% -->
-<!--   mutate(patch=case_when( -->
-<!--     patch == "c" ~ "Corridor", -->
-<!--     patch == "m" ~ "Matrix", -->
-<!--     patch == "w" ~ "Winged", -->
-<!--     patch == "r" ~ "Rectangle", -->
-<!--     .default = as.character(patch))) -->
-
-<!-- # if wanted to make a bar graph -->
-<!-- # ggplot(btl_sums_patch, aes(x=patch, y=n)) +  -->
-<!-- #   geom_bar(stat = "identity") -->
-<!-- #  -->
-<!-- abund_patch <- knitr::kable(btl_sums_patch, -->
-<!--       digits = 2, -->
-<!--       align="lr", -->
-<!--       format="latex", -->
-<!--       row.names = FALSE, -->
-<!--       booktabs=T, -->
-<!--       linesep = "", #removes the blank line after every 5 lines -->
-<!--       caption = "Total dung beetles captured in all replicates of a patch type.") %>% -->
-<!--   kable_styling(bootstrap_options = c("striped"), -->
-<!--                 full_width = F, -->
-<!--                 latex_options=c("scale_down","hold_position"), -->
-<!--                 font_size = 12, -->
-<!--                 position = "center") %>% -->
-<!--   # column_spec(1, italic = T) %>% -->
-<!--   row_spec(0, bold = T) -->
-<!--   # row_spec(1:15, hline_after = T) -->
-<!-- abund_patch -->
-
-<!-- ``` -->
-
-\newpage
-
-<!-- ```{r Table-aovMO, echo=FALSE,message = FALSE,warning=FALSE} -->
-<!-- aovM0<-read_csv(here("corridor_docs","eric_ms_thesis","tables","aovM0.csv")) %>% -->
-<!--   mutate(term=case_when( -->
-<!--     term == "patch_type" ~ "Patch type", -->
-<!--     .default = as.character(term))) |> -->
-<!-- # aovM0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM0.csv") %>%  -->
-<!--   kable( -->
-<!--       digits = 3, -->
-<!--       align="lcc", -->
-<!--       format="latex", -->
-<!--       row.names = FALSE, -->
-<!--       booktabs=T, -->
-<!--       linesep = "", #removes the blank line after every 5 lines -->
-<!--       caption = "AOV TABLE CAPTION") %>% -->
-<!--   kable_styling(bootstrap_options = c("striped"), -->
-<!--                 full_width = F, -->
-<!--                 latex_options=c("scale_down","hold_position"), -->
-<!--                 font_size = 12, -->
-<!--                 position = "center") %>% -->
-<!--   # column_spec(1, italic = T) %>% -->
-<!--   row_spec(0, bold = T) -->
-<!--   # row_spec(1:15, hline_after = T) -->
-<!-- aovM0 -->
-<!-- ``` -->
-
-\newpage
 \blandscape
+
 ```{r Table-MO, echo=FALSE,message = FALSE,warning=FALSE}
@@
 ```
+
 \elandscape
-
 \newpage
@@
 ```
+
 \newpage
@@
 \newpage
-
 \blandscape
+
 ```{r Table-M3, echo=FALSE,message = FALSE,warning=FALSE}
@@
 ```
+
 \elandscape
@@
 
-\vspace*{1cm}
+```{r, label = patch, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average total dung beetle abundance by patch type with standard deviations as error bars.", out.width="70%", fig.align='center'}
+
+knitr::include_graphics("images/total_abund.png")
+
+```
+
+\newpage
+
+```{r spcodes}
+
+sp_code_table<-data.frame(sp_code=c("aaeg", "cvig", "dcar", "open", "pign", "alec"))
+species_names<-spp_codes %>% 
+  filter(sp_code%in%sp_code_table$sp_code) %>% 
+  mutate(species=tolower(species)) %>% 
+  mutate(species=paste(species=genus,species,sep=" ")) %>% 
+  select(sp_code, species) %>% 
+mutate(species=paste(sp_code,species,sep=": ")) 
+species_names<-paste(species_names$species,collapse=", ")
+
+```
+
+\newpage
+
+```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = paste("Average abundance of the top 6 most abundant species by patch type.
+Species codes: ", species_names,".",sep=""), out.width="70%", fig.align='center'}
+
+knitr::include_graphics("images/sp_abund_patch.png")
+
+```
+
+\newpage
+
+```{r, label = sp-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Total dung beetle biomass collected for each species with sufficient weights.", out.width="70%", fig.align='center'}
+
+knitr::include_graphics("images/sp_biomass_plot.png")
+# knitr::include_graphics("images/sp_abundance_plot.png")
+# grid.arrange(p1[1], p2[1], nrow = 1)
+
+```
+
+\newpage
+
+```{r, label = patch-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average total biomass by patch type with standard deviation as error bars.", out.width="70%", fig.align='center'}
+
+knitr::include_graphics("images/avg_bmass_patch_plot.png")
+```
+
+\newpage
+
+```{r, label = block-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average total biomass by sampling block with standard deviation as error bars.", out.width="70%", fig.align='center'}
+
+knitr::include_graphics("images/avg_bmass_block_plot.png")
+```
+
+\newpage
+
+```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle species richness in each patch type.
+The point shapes indicates the block in which each patch was located.", out.width="70%", fig.align='center'}
+
+knitr::include_graphics("images/sp_richness_patches.png")
+```
+
+\newpage
+
+```{r, label = shannon, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Shannon diversity by patch type.", out.width="70%", fig.align='center'}
+
+knitr::include_graphics("images/sp_shannon_patches.png")
+
+```
+
+\newpage
+
+```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index by patch type.", out.width="70%", fig.align='center'}
+
+knitr::include_graphics("images/sp_simpson_patches.png")
+
+```
+
+\newpage
 
 \begin{center}
-\textbf{BIOGRAPHICAL SKETCH}
 \end{center}
-
-<!-- _A biographical sketch is required of all candidates.
-The biographical sketch should be in narrative form.
-Third person, past tense, it typically includes the educational background of the candidate.
-The author should have replaced this paragraph with their own._ -->
-
-\doublespacing
-
-Eric Escobar-Chena completed his Bachelors education at Virginia Commonwealth University in 2023.
-During his time there he developed a fondness for insects which grew into a curiosity of the natural world.
-He later began to explore this curiosity deeper in beginning his graduate education at the University of Florida as a Master's Student under the supervision of Emilio Bruna.
-He received his Master of Science from the University of Florida in 2025.

### 2025-08-01 corridor_docs/eric_ms_thesis/EscobarChena_etal_ms.tex
reorg of repo for MS

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  man, donotrepeattitle]{apa6}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\ifLuaTeX
  \usepackage[english]{selnolig} % disable illegal ligatures
\fi
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\makeatletter
\usepackage{etoolbox}
\patchcmd{\maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\makeatother
\keywords{keywords\newline\indent Word count: 4437}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\raggedbottom
\usepackage{endfloat}
\usepackage{setspace}\doublespacing
\usepackage{lineno}
\linenumbers
\usepackage{tabu}
\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{sectsty}\allsectionsfont{\raggedright}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Effect of experimental habitat connectivity on dung beetle communities},
  pdfauthor={Eric Escobar-Chena1, Julian Resasco2, \& Emilio M. Bruna1,3},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Effect of experimental habitat connectivity on dung beetle communities}
\author{Eric Escobar-Chena\textsuperscript{1}, Julian Resasco\textsuperscript{2}, \& Emilio M. Bruna\textsuperscript{1,3}}
\date{}

\shorttitle{dung beetles and habitat corridors}

\authornote{

We thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site. We also wanted to specifically thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry. Funding was provided by -----.

The authors made the following contributions. Eric Escobar-Chena: Conceptualization, Methodology, Investigation, Formal analysis, Data curation, Software, Visualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Julian Resasco: Conceptualization, Methodology, Writing - Original Draft Preparation, Writing - Review \& Editing; Emilio M. Bruna: Conceptualization, Methodology, Investigation, Formal analysis, Data curation, Visualization, Writing - Original Draft Preparation, Writing - Review \& Editing, Project administration, Supervision.

Correspondence concerning this article should be addressed to Eric Escobar-Chena, University of Florida. E-mail: \href{mailto:eescobarchena@ufl.edu}{\nolinkurl{eescobarchena@ufl.edu}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Wildlife Ecology \& Conservation, University of Florida, PO Box 110530, Gainesville, Florida, 32611-0430 USA\\\textsuperscript{2} University of Colorado, Boulder\\\textsuperscript{3} Center for Latin American Studies, University of Florida, PO Box 110530, Gainesville, Florida, 32611-0530 USA}

\note{

A preprint of this article has been posted at ----- Preprints (link).

ORCID:
EEC: ---- \textbar{}
EB: 0000-0003-3381-8477 \textbar{}
JR: ----- \textbar{}

\newpage

}

\abstract{%
Habitat fragmentation threatens biodiversity across the globe as habitat loss, isolation, and edge effects become increasingly prevalent. Corridors have become an important tool in order to combat the negative effects of fragmentation, however they are difficult to study in natural systems without incurring confounding effects. To observe changes in insect community composition as an effect of landscape features we sampled dung beetles in a landscape scale experiment. We did not see a difference in species richness or diversity, but dung beetle abundances were higher in continuous forest habitat and open habitat patches connected by a corridors than in isolated patches.
}

\begin{document}
\maketitle

\subsection{INTRODUCTION}\label{introduction}

As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented (Taubert et al. 2018, Díaz et al. 2019, Ma et al. 2023). Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems (Fahrig 2003, Fletcher et al. 2018). However, as habitats are broken down community structures are significantly altered (Harrison and Bruna 1999, Haddad et al. 2003, Jennings and Tallamy 2006, Laurance et al. 2018). This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions (Haddad 2015).

Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation (Haddad et al. 2003). By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy (Forman 1995). The resulting changes in species composition are important to identify because any species impacted would have corresponding effects depending on how they interact with the ecosystem (Zhou et al. 2023). Any gain or loss in key members of an community could disrupt processes which on their own could shape ecosystems (Cuke and Srivastava 2016), or effect other organisms which rely on said interaction (Wu et al. 2011). Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity (Debinski and Holt 2000).

By measuring changes in biodiversity and species richness within experimental designs I am able to isolate factors might be contributing to ecological patterns and processes (Resasco et al. 2017, Fletcher Jr. et al. 2023). Past studies have endeavored to experimentally measure changes in community compositions as a result of connecting habitats with corridors (Tewksbury et al. 2002, Collins et al. 2017, Graham et al. 2022) Yet very few have directly compared matrix and patch populations. Land use is different from one species to another so it is vital to understand where compositions are distributed and what processes might be driving population differences(Haddad 1999).

Dung beetles have emerged as a model system with which to test spatial ecology hypotheses (Roslin 2000, Rös et al. 2012).They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al. 2024). The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung (Nichols et al. 2008, Iwasa et al. 2015, Slade et al. 2016b). Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits {[}e.g., size, foraging style, resource-use, (Ospina-Garcés et al. 2018, deCastro-Arrazola et al. 2023){]}. Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors (Gray et al. 2022). Past studies have also focused on how landscape structure alters the community compositions of dung beetles(Costa et al. 2017), yet large landscape scale experimental studies with carefully controlled and replicated treatments are non-existent for this model species.

Here, I aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources (Hanski and Cambefort 1991), interact with corridors in their landscapes. I sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch area to edge ratio and distance to edge (Tewksbury et al. 2002). I do so I address these questions:

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\tightlist
\item
  How does the abundance of dung beetles differ among and between isolated and connected patches and how does this compare with abundance in matrix habitat?
\item
  Does species richness and diversity differ between among connected patches, unconnected patches, and the matrix habitat?
\item
  What are the implications of changes in community composition for ecosystem services and function?
\end{enumerate}

\subsection{METHODS}\label{methods}

\subsubsection{Study site}\label{study-site}

My study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, USA (33.208\(^\circ\) N, 81.408\(^\circ\) W, Figure \ref{fig:srs-map}). in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals (Tewksbury et al. 2002). Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna (Figure \ref{fig:patches}). In each block the central patch (100 \(\times\) 100 m) is always connected to one peripheral patch with identical dimensions by a 150 \(\times\) 25 m corridor, this will hereafter be referred to as the connected patch. The remaining patches are either ``winged'' or ``rectangular''. The winged patch is also 100 \(\times\) 100 m, however they exhibit their characteristic wings in the form of two 75 \(\times\) 25 m offshoots meant to account for the extra area and edge space the corridor provides. The rectangular patch is 100 \(\times\) 137.5 m also the same area as the space of the connected patch plus the corridor. Each block has a duplicate of either the winged or rectangle patch, all peripheral patches being 150 m from the center patch. For this study sampling was done in one of each patch type and in one matrix plot per block, all matrix blocks were set up 150 m away from the center as well.

\subsubsection{Dung beetle sampling}\label{dung-beetle-sampling}

Dung beetles were sampled in July and August 2024 in four of the SRS blocks (8, 52, 53n, and 54; Figure \ref{fig:srs-map}). In each block sampling was conducted using baited pitfall traps placed in each patch type as well as the matrix surrounding the patches (Figure \ref{fig:pitfalls}). Traps were placed in groups of 3 in the centers of each patch, approximately 250 meters from the midpoint of the central patch 40 m from patch edge. Pitfalls were oriented in a triangular pattern with the bottom two traps positioned towards the center patch, each trap 20 m apart. Plots in the matrix were set up in a similar fashion with the center point 250 m from the center placed equidistant between adjacent patches. Individual pitfall traps consisted of two components, a 10cm tall by 8 cm wide cylinder base topped with a funnel with a 10cm wide rim. I sourced pig feces from the University of Florida Swine Barn Unit. Bait was processed into 5cm wide balls and wrapped in a layer of coffee filter material. For each sample period, traps were buried flush with the ground and baited with pig dung between 8-9 pm and picked up 12 hours later, all beetles captured were stored in ethanol for further processing. In total 16 sampling rounds were carried out with 4 rounds per block, 196 samples were collected.

All dung beetles were counted and identified to species using Nemes and Price (2015) and Edmonds (2023). Fifteen individuals of each species with adequate captures were dried and weighed for biomass measurements. Ten species had the required counts for weighing which I dried in drying oven until all specimen reached a stable mass. I weighed each beetle using an Ohaus Adventurer Pro AV53 microbalance, with the exception of individuals whose weights were extremely small. Individual biomass for these species was estimated by weighing in batches of N = 3 (i.e., \emph{Ateuchus lecontei}), N = 5 (i.e., \emph{Onthophagus pennsylvanicus}), or N = 15 (i.e., \emph{Aphodius alloblackburneus}) and then calculated by average biomass per beetle. Values for individual biomass were then used to estimate the total biomass of each species in each patch, as well as the total beetle biomass (i.e., all species combined) in each patch. Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods upon completion of all analyses.

\subsubsection{Analyses}\label{analyses}

Biodiversity between patch types was compared using Hill numbers, a set of indexes developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity (Jost 2006). They are an alternative to more specialized metrics such as alpha, beta, and gamma diversities while being more standardized than other indexes such as Renyi or HCDT entropies, of which both groups of metrics are less intuitive for interpretation. Hill numbers are now the preferred metric for describing community dynamics for two reasons. First, they are extrapolated from the same equation, manipulating a single parameter (i.e., \emph{q}) to arrive at estimates of richness and diversity. Second, by manipulating \emph{q} we can gain an understanding of compositional shifts otherwise obscured while using species richness (Chao et al. 2014). I compared community composition by increasing magnitudes of diversity components (i.e., \emph{qD}) of \emph{0D} (i.e., species richness), \emph{1D} (i.e., Shannon entropy), and \emph{2D} (i.e, Simpson Diversity). Diversity numbers and species richness were calculated using the package \texttt{hill} (Li 2018) for the R statistical programming language (Posit team 2025). Diversity numbers were calculated using package \texttt{iNEXT} (Hsieh et al. 2016). Bray-Curtis dissimilarity values were calculated using package \texttt{Vegan} (Oksanen et al. 2025). Dung beetles were assigned traits by waste removal guild and habitat preference.

To test for the effects of connectivity on abundance, species richness, and species diversity I compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix. For abundance and richness I used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution (Bolker et al. 2009). Compared (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type. I included the identity of the sampling block as a random effects. To model my diversity metrics I took a similar approach, but this time using linear mixed effects models with a Gaussian distribution (Chao et al. 2014). In all models, the reference level for patch type was Matrix. All models were fit using \texttt{lme4} package (Bates et al. 2015). Prior to conducting my modeling I evaluated the the suitability of my data with \emph{qqplots} generated with the \texttt{DARMa} package (Hartig 2024).

I used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation. Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability. Likewise, I used linear mixed-effects models to compare for differences between patch types in (1) total dung beetle biomass, and (2) the total biomass of each of the ten species, again including block as a random effect.

\subsection{RESULTS}\label{results}

Overall, I collected N = 5213 dung beetles (N = 1359 in Connected patches, N = 1199 in Winged patches, N = 942 in Rectangle patches, N = 1713 in the Matrix). These beetles belonged to N = 16 species; the N = 6 most dominant species comprised of 93.9\% of all captures: \textit{Canthon vigilans} (N = 1473), \textit{Ateuchus lecontei} (N = 1115), \textit{Phanaeus igneus} (N = 958), \textit{Aphodius alloblackburneus} (N = 585), \textit{Dichotomius carolinus} (N = 556), and \textit{Onthophagus pennsylvanicus} (N = 207; Table 1). All but four species were captured in every patch type. \emph{Onthophagus concinnus} was only found in the matrix and winged patches, while \emph{Onthophagus striatulus} was only captured in matrix habitat and rectangular patches. \emph{Geotrupes blackburnii} and \emph{Onthophagus tuberculifrons} were the only species restricted to one patch type (winged and matrix, respectively). All species were within their native ranges.

\subsubsection{Dung Beetle Abundance}\label{dung-beetle-abundance}

When comparing the overall abundance of beetles (all species combined) across patch types, matrix plots had the highest captures, followed by connected patches, then winged, with the fewest in rectangular patches (Figure \ref{fig:patch}). Abundances from connected patches were not significantly different from those in the matrix while rectangle and winged patches had significantly fewer than the matrix (\(\beta\) = -0.528 and -0.303 respectively, P \textless{} 0.001, Table \ref{tab:Table-MO}).

Statistical analysis of abundance focused on the six most abundant species. A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table \ref{tab:Table-M3}). The baseline abundance corresponds to the abundance of \emph{Aphodius alloblackburneus} in matrix patches. Compared to this baseline, results were highly variable, emphasizing species specific responses to patch type. Similarly, species showed to have disproportionate responses to patch since interaction terms varied widely (Table \ref{tab:Table-M3}).

\subsubsection{Dung Beetle Richness and Diversity}\label{dung-beetle-richness-and-diversity}

Plotting species richness by patch type reveals consistent richness across patch types with some variation between sampling blocks (Figure \ref{fig:richness}).The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N). Modeling the effect of patch type on species richness with block as a random effect determined there was no significant differences among patch types. Comparing treatments using matrix patches as a baseline resulted in no significant differences in connected (\(\beta\) = 0.216, p = 0.839; Table \ref{tab:Table-M-rich}), rectangle (\(\beta\) = -0.548, p = 0.583; Table \ref{tab:Table-M-rich}), and winged patches (\(\beta\) = -0.095, p = 0.663; Table \ref{tab:Table-M-rich}).

Biodiversity was also even not significantly different between patch types, however the values of the different metrics varied significantly across sampling blocks (Figures \ref{fig:shannon}-\ref{fig:simpson}). For Shannon Diversity, the estimated mean in matrix patches was 5.202 (SE = 0.665, t = 7.825; Table \ref{tab:Table-M-shan}). None of the alternative patch types showed statistically significant differences compared to matrix: connected (\(\beta\) = --0.085, SE = 0.495, t = 0.17), rectangle (\(\beta\) = 0.234, SE = 0.495, t = 0.472), or winged (\(\beta\) = --0.251, SE = 0.495, t = --0.506; Table \ref{tab:Table-M-shan}). Results for Simpson's Diversity were similar (Table \ref{tab:Table-M-simp}), with the average value in matrix patches was 4.181 (SE = 0.640, t = 6.535). Again, none of the other patch types were significantly different from the others: matrix (\(\beta\) = -0.107, SE = 0.411, t = -0.259), rectangle (\(\beta\) = 0.059, SE = 0.411, t = 0.143), and winged (\(\beta\) = --0.480, SE = 0.411, t = --1.169; Table \ref{tab:Table-M-simp}).

\subsubsection{Dung Beetle Biomass}\label{dung-beetle-biomass}

Patterns of biomass by patch type were similar to those for abundance: biomass was highest in matrix plots, followed by connected and winged, with biomass in rectangle patches being far lower than in the other locations (Figure \ref{fig:patch-bmass}). There was no significant difference in the total beetle biomass of matrix plots when compared to connected (\(\beta\) = -7.337, SE = 8.659, t = -0.847; Table \ref{tab:Table-aovM51}) and winged patches (\(\beta\) = -9.608, SE = 8.659, t = -1.11; Table \ref{tab:Table-aovM51}), but biomass in rectangle patches was significantly different (\(\beta\) = -23.296, SE = 8.659, t = -2.69; Table \ref{tab:Table-aovM51}). Block effects continue to be important for explaining high variation as the total biomass collected in block 52 was nearly quadruple that collected in Block 8 (Figure \ref{fig:block-bmass}). Linear mixed effects models also indicate significant differences between the biomass of different species (Table \ref{tab:Table-aovM52}, Figure \ref{fig:sp-bmass}).

\subsection{DISCUSSION}\label{discussion}

This study advances our understanding of the factors shaping dung beetle community composition in temperate regions of the southeastern United States. In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches. My main findings emphasized: (1) Habitat type and patch shape were the main driving factors for determining how dung beetle species abundances were composed, however effects were species specific. (2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level. (3) Species diversity metrics were also relatively even across patch types however varied widely by sampling blocks. These results suggest that while there may indeed be effects of patch structure and connectivity on dung beetle abundances and community composition, other landscape scale drivers appear to be more prominent for species richness and diversity.

In my comparison of dung beetle compositions, connectivity and habitat edge accounted for differences in dung beetle abundances, yet species richness and diversity were even across patch types and in the matrix. Total beetle counts were consistently highest in the matrix, in comparison abundances is connected patches were not significantly different while those of the winged and rectangular patches were lower. Additionally, I also observed that patch effects were not equally proportional for all species. This suggests that the species captured could be exhibiting habitat preferences between the open patches versus forested matrix. Another potential explanation is that species in the matrix are acting as a source population which feeds into patches with edge acting as a drift fence for directing movement. The latter example seems more likely - small mammals at SRS, which are a potential source of dung, appear to preferentially use matrix habitat (Mabry et al. 2003) and flies, which have similar resource dependencies were also found to have a similar interaction (Fried et al. 2005). Moreover studies on dung beetles in the tropics found more dramatic differences in dung beetle populations in fragments and matrix (Barragan et al. 2011).

Although richness and diversity were the same among treatments, there was notable variability between sampling blocks. Block 8 generally had the lowest species richness and biodiversity while 53n had the highest. While the experimental design attempted to control for the effects of patch size and edge, there could be large (and potential unknown) environmental gradients across the SRS landscape that could influence the observed patterns in diversity and abundance. For instance, at the time of my study Block 8 had the densest matrix of any of the blocks. This could have hindered the diffusion of bait scent, leading to lower capture rates in this block. Other habitat characteristics that might differ among blocks could have been influential as well -- for example, soil quality and forest cover can determine where beetles can reproduce (Arellano et al. 2008, Conover et al. 2019). The same is true for land-use history; much of the SES land would have previously been used for agriculture, and during development of experimental units cleared with heavy machinery might experience heavy soil compaction \emph{brudvig}. Landscape features could also affect mammal movement, which might in turn limit dung availability - Block 8 was nearest to roadways used by all employees at SRS, and the eastern side of the site had large bodies of water upon which mammals are dependent (Harvey et al. 2006, Dechen Quinn et al. 2013, Barahona-Segovia 2021). Additionally, deer and feral hogs are the most dominant mammals at the site, two species that do well in patchy landscapes (Castillo-Contreras et al. 2018, Fraser et al. 2019). The locally variable patterns of abundance in the landscape of these generalist large mammals, coupled with ability to easily move through modified landscapes, might help explain the limited variation in richness or diversity within blocks (i.e., between patch types) but large block effects. Deer also commonly forage along the edges of habitat patches and forest ecotones. If they are spending more time in these locations, the higher abundance and diversity in connected and winged patches might in part be due to `drift-fence' effects \textbf{citation of levey drift fence}.

Species-specific differences are apparent but do not follow any particular trend. \emph{Aphodius alloblackburneus} had a disproportionate positive effect to being captured in matrix patches as compared to other species. Beetles from the genus \emph{Aphodius} do tend to show patterns of habitat specificity (Roslin and Koivunen 2001), and many preferring forested habitat (Frank et al. 2017b), so it is not unexpected that they might show a preference towards forested matrix. I did not detect that any species was more positively associated with open patches despite suggestions that some species (e.g., \emph{Canthon vigilans}, \emph{Melanocanthon bispinatus}) prefered open habitat (Nealis 1977, Conover et al. 2019). This could be another sign pointing towards matrix acting as a source population, and since open habitat was much less dominant in my experimental system beetles could be moving into patches from habitat edge.

The patterns in dung beetle biomass largely echo what we observed for abundance. Although biomass is understood to be positively associated with dung removal (Slade et al. 2011), dung beetle species vary greatly in terms of morphology and functionality (Ospina-Garcés et al. 2018) so evaluating species-specific patterns is particularly important. Because of this we can expect that the magnitude of removal might be greater in areas of larger biomass (i.e.~matrix or connected patches), however future work should aim to directly measure this potential pattern. Indeed comparing the biomass of different species emphasizes the morphological variability between our study species - some of the most common species (e.g., \emph{Ateuchus lecontei}, \emph{Aphodius allobloackburneus}) contribute very little to total biomass (Figure \ref{fig:sp-bmass}). In addition to this, dung beetles appear to be an exception to the globally and taxonomically robust rule that abundance is negatively correlated with individual biomass (White et al. 2007). More intensive sampling could determine whether this trend is truly apparent.

Despite ample work documenting their ecological importance, there is a surprising lack of research on dung beetle diversity and corridors outside of the tropics (Nichols et al. 2007), with work in temperate locations coming primarily from Europe where \emph{Aphodius} beetles dominate (Roslin and Koivunen 2001). Many studies conducted in the tropics found strong preference by some species for open fields vs.~continuous forest (Damborsky et al. 2015), while othersfound that forest species were using living fences as corridors (Arellano et al. 2008). I did not observe such intense habitat specificity, which may not be especially common outside of the tropics. I found similar patterns of community composition as did the small number of prior dung beetle biodiversity surveys conducted in the southeastern United States (Nealis 1977, Conover et al. 2019), including at least one more where differences between treatment types were non-significant while variation on a larger scale was more apparent (Young et al. 2023). I did expect some species to be more dominant in patches with higher edge ratios as was found with ants at SRS (Resasco et al. 2014), but species were evenly present in all patches.

Finally, it is important to emphasize that while dung beetles are capable of long distance flight and detecting dung at distances of over 50 meters (Gray et al. 2022), the results I observed are likely the result of a mismatch between the spatial scale of the experimental replicates and dung beetle movement. Put another way, some beetles were almost certainly drawn by the dung used in baits from the matrix into the plots. If this were an overarching effect we'd expect all plots to be similar to the matrix, or at least to each other. The fact that rectangular are have lower abundance suggests that some sort of landscape effect is apparent, likely related to habitat edge since connected and winged patches were the most similar. To remedy this issue any future work in this site should either focus solely on dung beetles with more limited dispersal ability or consider conducting mark-release-recapture experiments in an attempt to document movements within and between patches.

There were some limitations of this study which should be addressed in any future work to capture a clearer picture of how populations are being altered by habitat connectivity. First, sampling for this study was conducted over a two month period in the summer of 2024, sufficient data was collected but due to a lack of available resources temporal patterns were obscured due to inconsistent sampling periods. This is highly important since dung beetles exhibit consistent patterns of seasonality (Davis 1966, Conover et al. 2019). Another potential avenue for improvement is lowering the grain size of sampling and changing trap placement to better understand the effects of edge proximity and connectivity. I also used one of the best bait types for collecting dung beetles in pig dung (Marsh et al. 2013), but a mix of differently sourced baits may have been more optimal as more diverse baits attract more diverse species (Frank et al. 2017a, 2018, Giménez Gómez et al. 2021).

Moving forward, there is a bounty of knowledge yet to be collected for dung beetles in the south eastern United States. Relative to what is already known in tropical ecosystems, this study provides a glimpse on how dung beetles are responding to fragmentation and connectivity in subtropic pine habitats, but much can still be learned about dung beetles in these spaces and the role they play in ecosystems as well as how they interact with landscapes. The unique role of dung beetles as waste removers is fascinating enough and of great importance for maintaining landscapes so dominated by pastures like Southern US (Slade et al. 2016a, Cheng et al. 2022). Not only waste removal but also secondary effect such as seed dispersal and parasitic reduction are important to understand in the context of fragmentation (Fincher 1975, Vulinec 2002). SRS provides an excellent experimental design for direct comparison of landscapes but another main goal of the site is to study dispersal, dung beetles would be an excellent system for studying movement. They already show promise, in preliminary trials I released beetles from both the connected and rectangular patches with baited traps in the central patch. I recovered one recapture which originated from the connected patch, indicating that individuals do move through corridors.

Although there is plentiful work to be done, this study paints an encouraging picture that while dung beetles face disturbances that make habitats less preferential, they are robust enough to persist throughout fragmented landscapes and the management strategies we do have to connect fragmented landscapes work to mitigate losses. Because assemblages remain mostly the same across these landscapes, ecosystem services should remain uninterrupted allowing continuing benefits to all community members. While this study of dung beetles in fragmented landscapes was conducted in a relatively protected area, in real world application we could consider the use of movement corridors within disturbed areas to help bolster effected populations, but also near areas of high importance such as pastures to provide areas of refuge for beetles. In this case it is important to consider the inverse of what was manipulated in this study where open field acts as matrix and forested area is focal landscape, thus it is vital to continue learning more about how landscape composition and connectivity effect dung beetle populations.

\subsection{LTERATURE CITED}\label{lterature-cited}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-arellano_response_2008}
Arellano, L., J. L. Leon-Cortes, and G. Halffter. 2008. \href{https://doi.org/10.1111/j.1752-4598.2008.00033.x}{Response of dung beetle assemblages to landscape structure in remnant natural and modified habitats in southern {Mexico}}. Insect Conservation and Diversity 1:253--262.

\bibitem[\citeproctext]{ref-WOS:000651446900001}
Barahona-Segovia, R. M. 2021. \href{https://doi.org/10.1007/s10841-021-00320-z}{Until death do us part: Abundance and survival of necrophagous beetle species associated with fox scats in fragmented landscapes}. Journal of Insect Conservation 25:521--530.

\bibitem[\citeproctext]{ref-WOS:000288810500026}
Barragan, F., C. E. Moreno, F. Escobar, G. Halffter, and D. Navarrete. 2011. \href{https://doi.org/10.1371/journal.pone.0017976}{Negative impacts of human land use on dung beetle functional diversity}. Plos One 6.

\bibitem[\citeproctext]{ref-bates_fitting_2015}
Bates, D., M. Mächler, B. Bolker, and S. Walker. 2015. \href{https://doi.org/10.18637/jss.v067.i01}{Fitting {Linear} {Mixed}-{Effects} {Models} {Using} lme4}. Journal of Statistical Software 67:1--48.

\bibitem[\citeproctext]{ref-bolker_generalized_2009}
Bolker, B. M., M. E. Brooks, C. J. Clark, S. W. Geange, J. R. Poulsen, M. H. H. Stevens, and J.-S. S. White. 2009. \href{https://doi.org/10.1016/j.tree.2008.10.008}{Generalized linear mixed models: A practical guide for ecology and evolution}. Trends in Ecology \& Evolution 24:127--135.

\bibitem[\citeproctext]{ref-castillo-contreras_urban_2018}
Castillo-Contreras, R., J. Carvalho, E. Serrano, G. Mentaberre, X. Fernández-Aguilar, A. Colom, C. González-Crespo, S. Lavín, and J. R. López-Olvera. 2018. \href{https://doi.org/10.1016/j.scitotenv.2017.09.277}{Urban wild boars prefer fragmented areas with food resources near natural corridors}. Science of The Total Environment 615:282--288.

\bibitem[\citeproctext]{ref-chao_rarefaction_2014}
Chao, A., N. J. Gotelli, T. C. Hsieh, E. L. Sander, K. H. Ma, R. K. Colwell, and A. M. Ellison. 2014. \href{https://doi.org/10.1890/13-0133.1}{Rarefaction and extrapolation with {Hill} numbers: A framework for sampling and estimation in species diversity studies}. Ecological Monographs 84:45--67.

\bibitem[\citeproctext]{ref-cheng_dweller_2022}
Cheng, J., F. Y. Li, Y. Wang, Y. Wang, X. Liu, J. Zhang, Z. Wang, Y. Li, H. Wang, Z. Yang, and M. A. Potter. 2022. \href{https://doi.org/10.1016/j.agee.2022.107873}{Dweller and tunneler dung beetles synergistically accelerate decomposition of cattle and horse dung in a semi-arid steppe}. Agriculture, Ecosystems \& Environment 329:107873.

\bibitem[\citeproctext]{ref-collins_fragmentation_2017}
Collins, C. D., C. Banks‐Leite, L. A. Brudvig, B. L. Foster, W. M. Cook, E. I. Damschen, A. Andrade, M. Austin, J. L. Camargo, D. A. Driscoll, R. D. Holt, W. F. Laurance, A. O. Nicholls, and J. L. Orrock. 2017. \href{https://doi.org/10.1111/ecog.02607}{Fragmentation affects plant community composition over time}. Ecography 40:119--130.

\bibitem[\citeproctext]{ref-conover_phenology_2019}
Conover, D., J. Dubeux, and X. Martini. 2019. \href{https://doi.org/10.1093/ee/nvz068}{Phenology, distribution, and diversity of dung beetles ({Coleoptera}: {Scarabaeidae}) in north {Florida}'s pastures and forests}. Environmental Entomology 48:847--855.

\bibitem[\citeproctext]{ref-WOS:000398662200001}
Costa, C., V. H. F. Oliveira, R. Maciel, W. Beiroz, V. Korasaki, and J. Louzada. 2017. \href{https://doi.org/10.7717/peerj.3125}{Variegated tropical landscapes conserve diverse dung beetle communities}. Peerj 5.

\bibitem[\citeproctext]{ref-WOS:000374962000009}
Cuke, M., and D. S. Srivastava. 2016. \href{https://doi.org/10.1007/s10980-015-0316-z}{Divergent effects of tropical forest fragmentation and conversion on leaf litter decomposition}. Landscape Ecology 31:1037--1050.

\bibitem[\citeproctext]{ref-WOS:000349242000004}
Damborsky, M. P., M. C. Alvarez Bohle, M. G. Ibarra Polesel, E. A. Porcel, and J. L. Fontana. 2015. \href{https://doi.org/10.1007/s13744-014-0257-2}{Spatial and temporal variation of dung beetle assemblages in a fragmented landscape at eastern humid chaco}. Neotropical Entomology 44:30--39.

\bibitem[\citeproctext]{ref-davis_feeding_1966}
Davis, L. V. 1966. \href{https://www.jstor.org/stable/24333363}{Feeding habits and seasonal distribution of scarab beetles in the {North} {Carolina} piedmont}. Journal of the Elisha Mitchell Scientific Society 82:212--220.

\bibitem[\citeproctext]{ref-debinski_survey_2000}
Debinski, D. M., and R. D. Holt. 2000. \href{https://doi.org/10.1046/j.1523-1739.2000.98081.x}{A survey and overview of habitat fragmentation experiments}. Conservation Biology 14:342--355.

\bibitem[\citeproctext]{ref-decastro-arrazola_trait-based_2023}
deCastro-Arrazola, I., N. R. Andrew, M. P. Berg, A. Curtsdotter, J.-P. Lumaret, R. Menéndez, M. Moretti, B. Nervo, E. S. Nichols, F. Sánchez-Piñero, A. M. C. Santos, K. S. Sheldon, E. M. Slade, and J. Hortal. 2023. \href{https://doi.org/10.1111/1365-2656.13829}{A trait-based framework for dung beetle functional ecology}. Journal of Animal Ecology 92:44--65.

\bibitem[\citeproctext]{ref-dechen_quinn_landscape_2013}
Dechen Quinn, A. C., D. M. Williams, and W. F. Porter. 2013. \href{https://doi.org/10.1644/11-MAMM-A-221.1}{Landscape structure influences space use by white-tailed deer}. Journal of Mammalogy 94:398--407.

\bibitem[\citeproctext]{ref-diaz_pervasive_2019}
Díaz, S., J. Settele, E. S. Brondízio, H. T. Ngo, J. Agard, A. Arneth, P. Balvanera, K. A. Brauman, S. H. M. Butchart, K. M. A. Chan, L. A. Garibaldi, K. Ichii, J. Liu, S. M. Subramanian, G. F. Midgley, P. Miloslavich, Z. Molnár, D. Obura, A. Pfaff, S. Polasky, A. Purvis, J. Razzaque, B. Reyers, R. R. Chowdhury, Y.-J. Shin, I. Visseren-Hamakers, K. J. Willis, and C. N. Zayas. 2019. \href{https://doi.org/10.1126/science.aax3100}{Pervasive human-driven decline of life on {Earth} points to the need for transformative change}. Science 366:eaax3100.

\bibitem[\citeproctext]{ref-edmonds_taxonomic_2023}
Edmonds, W. D. 2023. \href{https://journals.flvc.org/mundi/article/view/134908}{Taxonomic review of the {North} {American} dung beetle genus \emph{{Melanocanthon} {Halffter}}, 1958 ({Coleoptera}: {Scarabaeidae}: {Scarabaeinae}: {Deltochilini})}. Insecta Mundi.

\bibitem[\citeproctext]{ref-fahrig_effects_2003}
Fahrig, L. 2003. \href{https://doi.org/10.1146/annurev.ecolsys.34.011802.132419}{Effects of habitat fragmentation on biodiversity}. Annual Review of Ecology, Evolution, and Systematics 34:487--515.

\bibitem[\citeproctext]{ref-fincher_effects_1975}
Fincher, G. T. 1975. \href{https://doi.org/10.2307/3279480}{Effects of dung beetle activity on the number of nematode parasites acquired by grazing cattle}. The Journal of Parasitology 61:759.

\bibitem[\citeproctext]{ref-fletcher_jr_landscape_2023}
Fletcher Jr., R. J., T. A. H. Smith, N. Kortessis, E. M. Bruna, and R. D. Holt. 2023. \href{https://doi.org/10.1002/ecy.4037}{Landscape experiments unlock relationships among habitat loss, fragmentation, and patch-size effects}. Ecology 104:e4037.

\bibitem[\citeproctext]{ref-fletcher_is_2018}
Fletcher, R. J., R. K. Didham, C. Banks-Leite, J. Barlow, R. M. Ewers, J. Rosindell, R. D. Holt, A. Gonzalez, R. Pardini, E. I. Damschen, F. P. L. Melo, L. Ries, J. A. Prevedello, T. Tscharntke, W. F. Laurance, T. Lovejoy, and N. M. Haddad. 2018. \href{https://doi.org/10.1016/j.biocon.2018.07.022}{Is habitat fragmentation good for biodiversity?} Biological Conservation 226:9--15.

\bibitem[\citeproctext]{ref-forman_general_1995}
Forman, R. T. T. 1995. \href{https://doi.org/10.1007/BF00133027}{Some general principles of landscape and regional ecology}. Landscape Ecology 10:133--142.

\bibitem[\citeproctext]{ref-frank_search_2018}
Frank, K., A. Brückner, N. Blüthgen, and T. Schmitt. 2018. \href{https://doi.org/10.1007/s00049-018-0266-4}{In search of cues: Dung beetle attraction and the significance of volatile composition of dung}. Chemoecology 28:145--152.

\bibitem[\citeproctext]{ref-frank_nutrient_2017}
Frank, K., A. Brückner, A. Hilpert, M. Heethoff, and N. Blüthgen. 2017a. \href{https://doi.org/10.1038/s41598-017-12265-y}{Nutrient quality of vertebrate dung as a diet for dung beetles}. Scientific Reports 7:12141.

\bibitem[\citeproctext]{ref-frank_land_2017}
Frank, K., M. Hülsmann, T. Assmann, T. Schmitt, and N. Blüthgen. 2017b. \href{https://doi.org/10.1016/j.agee.2017.04.010}{Land use affects dung beetle communities and their ecosystem service in forests and grasslands}. Agriculture, Ecosystems \& Environment 243:114--122.

\bibitem[\citeproctext]{ref-fraser_connectivity_2019}
Fraser, D. L., K. Ironside, R. K. Wayne, and E. E. Boydston. 2019. \href{https://doi.org/10.1007/s10980-019-00824-9}{Connectivity of mule deer ({Odocoileus} hemionus) populations in a highly fragmented urban landscape}. Landscape Ecology 34:1097--1115.

\bibitem[\citeproctext]{ref-fried_habitat_2005}
Fried, J. H., D. J. Levey, and J. A. Hogsette. 2005. \href{https://doi.org/10.1007/s00442-005-0023-6}{Habitat corridors function as both drift fences and movement conduits for dispersing flies}. Oecologia 143:645--651.

\bibitem[\citeproctext]{ref-gimenez_gomez_dung_2021}
Giménez Gómez, V. C., J. R. Verdú, S. J. E. Velazco, and G. A. Zurita. 2021. \href{https://doi.org/10.1111/een.13001}{Dung beetle trophic ecology: Are we misunderstanding resources attraction?} Ecological Entomology 46:552--561.

\bibitem[\citeproctext]{ref-graham_impact_2022}
Graham, C. D. K., C. R. Warneke, M. Weber, and L. A. Brudvig. 2022. \href{https://doi.org/10.1007/s10980-022-01529-2}{The impact of habitat fragmentation on domatia-dwelling mites and a mite-plant-fungus tritrophic interaction}. Landscape Ecology 37:3029--3041.

\bibitem[\citeproctext]{ref-WOS:000707546700001}
Gray, R. E. J., L. F. Rodriguez, O. T. Lewis, A. Y. C. Chung, O. Ovaskainen, and E. M. Slade. 2022. \href{https://doi.org/10.1111/1365-2664.14049}{Movement of forest-dependent dung beetles through riparian buffers in {Bornean} oil palm plantations}. Journal of Applied Ecology 59:238--250.

\bibitem[\citeproctext]{ref-haddad_corridor_1999}
Haddad, N. M. 1999. \href{https://doi.org/10.1086/303163}{Corridor use predicted from behaviors at habitat boundaries.} The American Naturalist 153:215--227.

\bibitem[\citeproctext]{ref-haddad_habitat_2015}
Haddad, N. M. 2015, March. \href{https://www-science-org.lp.hscl.ufl.edu/doi/10.1126/sciadv.1500052}{Habitat fragmentation and its lasting impact on {Earth}'s ecosystems {\textbar} {Science} {Advances}}.

\bibitem[\citeproctext]{ref-haddad_corridor_2003}
Haddad, N. M., D. R. Bowne, A. Cunningham, B. J. Danielson, D. J. Levey, S. Sargent, and T. Spira. 2003. \href{https://doi.org/10.1890/0012-9658(2003)084\%5B0609:CUBDT\%5D2.0.CO;2}{Corridor use by diverse taxa}. Ecology 84:609--615.

\bibitem[\citeproctext]{ref-hanski_dung_1991}
Hanski, I., and Y. Cambefort. 1991. Dung {Beetle} {Ecology}. Princeton University Press.

\bibitem[\citeproctext]{ref-harrison_habitat_1999}
Harrison, S., and E. Bruna. 1999. \href{https://doi.org/10.1111/j.1600-0587.1999.tb00496.x}{Habitat fragmentation and large-scale conservation: What do we know for sure?} Ecography 22:225--232.

\bibitem[\citeproctext]{ref-hartig_dharma_2024}
Hartig, F. 2024. \href{https://CRAN.R-project.org/package=DHARMa}{{DHARMa}: {Residual} {Diagnostics} for {Hierarchical} ({Multi}-{Level} / {Mixed}) {Regression} {Models}}.

\bibitem[\citeproctext]{ref-WOS:000241362400032}
Harvey, C. A., A. Medina, D. Merlo Sanchez, S. Vilchez, B. Hernandez, J. C. Saenz, J. M. Maes, F. Casanoves, and F. L. Sinclair. 2006. \href{https://doi.org/10.1890/1051-0761(2006)016\%7B\%5B\%7D1986:POADID\%5D2.0.CO;2}{Patterns of animal diversity in different forms of tree cover in agricultural landscapes}. Ecological Applications 16:1986--1999.

\bibitem[\citeproctext]{ref-hasan_dung_2024}
Hasan, F., K. J. Wallace, S. V. Fowler, L. A. Schipper, Z. Hemmings, J. D. Berson, and A. D. Barnes. 2024. \href{https://doi.org/10.1111/1365-2435.14607}{Dung beetles drive direct and indirect changes in ecosystem multifunctionality}. Functional Ecology 38:1971--1983.

\bibitem[\citeproctext]{ref-hsieh_inext_2016}
Hsieh, T. C., K. H. Ma, and A. Chao. 2016. \href{https://doi.org/10.1111/2041-210X.12613}{{iNEXT}: An {R} package for rarefaction and extrapolation of species diversity ({Hill} numbers)}. Methods in Ecology and Evolution 7:1451--1456.

\bibitem[\citeproctext]{ref-WOS:000350444300013}
Iwasa, M., Y. Moki, and J. Takahashi. 2015. \href{https://doi.org/10.1093/ee/nvu023}{Effects of the activity of coprophagous insects on greenhouse gas emissions from cattle dung pats and changes in amounts of nitrogen, carbon, and energy}. Environmental Entomology 44:106--113.

\bibitem[\citeproctext]{ref-jennings_composition_2006}
Jennings, V. H., and D. W. Tallamy. 2006. \href{https://doi.org/10.1093/ee/35.6.1550}{Composition and abundance of ground-dwelling {Coleoptera} in a fragmented and continuous forest}. Environmental Entomology 35:1550--1560.

\bibitem[\citeproctext]{ref-jost_entropy_2006}
Jost, L. 2006. \href{https://doi.org/10.1111/j.2006.0030-1299.14714.x}{Entropy and diversity}. Oikos 113:363--375.

\bibitem[\citeproctext]{ref-WOS:000419965700012}
Laurance, W. F., J. L. C. Camargo, P. M. Fearnside, T. E. Lovejoy, G. B. Williamson, R. C. G. Mesquita, C. F. J. Meyer, P. E. D. Bobrowiec, and S. G. W. Laurance. 2018. \href{https://doi.org/10.1111/brv.12343}{An {Amazonian} rainforest and its fragments as a laboratory of global change}. Biological Reviews 93:223--247.

\bibitem[\citeproctext]{ref-li_hillr_2018}
Li, D. 2018. \href{https://doi.org/10.21105/joss.01041}{{hillR}: Taxonomic, functional, and phylogenetic diversity and similarity through {Hill} {Numbers}}. Journal of Open Source Software 3:1041.

\bibitem[\citeproctext]{ref-ma_global_2023}
Ma, J., J. Li, W. Wu, and J. Liu. 2023. \href{https://doi.org/10.1038/s41467-023-39221-x}{Global forest fragmentation change from 2000 to 2020}. Nature Communications 14:3752.

\bibitem[\citeproctext]{ref-mabry_influence_2003}
Mabry, K. E., E. A. Dreelin, and G. W. Barrett. 2003. \href{https://www.jstor.org/stable/1383631}{Influence of landscape elements on population densities and habitat use of three small-mammal species}. Journal of Mammalogy 84:20--25.

\bibitem[\citeproctext]{ref-marsh_optimising_2013}
Marsh, C. J., J. Louzada, W. Beiroz, and R. M. Ewers. 2013. \href{https://doi.org/10.1371/journal.pone.0073147}{Optimising bait for pitfall trapping of {Amazonian} dung beetles ({Coleoptera}: {Scarabaeinae})}. Plos One 8:e73147.

\bibitem[\citeproctext]{ref-nealis_habitat_1977}
Nealis, V. G. 1977. \href{https://doi.org/10.1139/z77-017}{Habitat associations and community analysis of {South} {Texas} dung beetles ({Coleoptera}: {Scarabaeinae})}. Canadian Journal of Zoology 55:138--147.

\bibitem[\citeproctext]{ref-nemes_illustrated_2015}
Nemes, S. N., and D. L. Price. 2015. \href{https://doi.org/10.1656/045.022.0208}{Illustrated keys to the {Scarabaeinae} ({Coleoptera}: {Scarabaeidae}) of {Maryland}}. Northeastern Naturalist 22:318--344.

\bibitem[\citeproctext]{ref-nichols_global_2007}
Nichols, E., T. Larsen, S. Spector, A. L. Davis, F. Escobar, M. Favila, K. Vuline, and S. R. Network. 2007. \href{https://doi.org/10.1016/j.biocon.2007.01.023}{Global dung beetle response to tropical forest modification and fragmentation: {A} quantitative literature review and meta-analysis}. Biological Conservation 137:1--19.

\bibitem[\citeproctext]{ref-nichols_ecological_2008}
Nichols, E., S. Spector, J. Louzada, T. Larsen, S. Amezquita, and M. E. Favila. 2008. \href{https://doi.org/10.1016/j.biocon.2008.04.011}{Ecological functions and ecosystem services provided by {Scarabaeinae} dung beetles}. Biological Conservation 141:1461--1474.

\bibitem[\citeproctext]{ref-oksanen_vegan_2025}
Oksanen, J., G. L. Simpson, F. G. Blanchet, R. Kindt, P. Legendre, P. R. Minchin, R. B. O'Hara, P. Solymos, M. H. H. Stevens, E. Szoecs, H. Wagner, M. Barbour, M. Bedward, B. Bolker, D. Borcard, G. Carvalho, M. Chirico, M. D. Caceres, S. Durand, H. B. A. Evangelista, R. FitzJohn, M. Friendly, B. Furneaux, G. Hannigan, M. O. Hill, L. Lahti, D. McGlinn, M.-H. Ouellette, E. R. Cunha, T. Smith, A. Stier, C. J. F. T. Braak, J. Weedon, and T. Borman. 2025. \href{https://CRAN.R-project.org/package=vegan}{Vegan: {Community} {Ecology} {Package}}.

\bibitem[\citeproctext]{ref-ospina-garces_dung_2018}
Ospina-Garcés, S. M., F. Escobar, M. L. Baena, A. L. V. Davis, and C. H. Scholtz. 2018. \href{https://doi.org/10.1007/s10682-018-9958-z}{Do dung beetles show interrelated evolutionary trends in wing morphology, flight biomechanics and habitat preference?} Evolutionary Ecology 32:663--682.

\bibitem[\citeproctext]{ref-posit_team_rstudio_2025}
Posit team. 2025. \href{http://www.posit.co/}{{RStudio}: {Integrated} {Development} {Environment} for {R}}. Posit Software, PBC, Boston, MA.

\bibitem[\citeproctext]{ref-resasco_contribution_2017}
Resasco, J., E. M. Bruna, N. M. Haddad, C. Banks-Leite, and C. R. Margules. 2017. \href{https://doi.org/10.1111/ecog.02546}{The contribution of theory and experiments to conservation in fragmented landscapes}. Ecography 40:109--118.

\bibitem[\citeproctext]{ref-resasco_landscape_2014}
Resasco, J., N. M. Haddad, J. L. Orrock, D. Shoemaker, L. A. Brudvig, E. I. Damschen, J. J. Tewksbury, and D. J. Levey. 2014. \href{https://doi.org/10.1890/14-0169.1}{Landscape corridors can increase invasion by an exotic species and reduce diversity of native species}. Ecology 95:2033--2039.

\bibitem[\citeproctext]{ref-ros_how_2012}
Rös, M., F. Escobar, and G. Halffter. 2012. \href{https://doi.org/10.1111/j.1472-4642.2011.00834.x}{How dung beetles respond to a human-modified variegated landscape in {Mexican} cloud forest: A study of biodiversity integrating ecological and biogeographical perspectives}. Diversity and Distributions 18:377--389.

\bibitem[\citeproctext]{ref-roslin_dung_2000}
Roslin, T. 2000. \href{https://doi.org/10.1034/j.1600-0706.2000.910213.x}{Dung beetle movements at two spatial scales}. Oikos 91:323--335.

\bibitem[\citeproctext]{ref-roslin_distribution_2001}
Roslin, T., and A. Koivunen. 2001. \href{https://doi.org/10.1007/s004420000565}{Distribution and abundance of dung beetles in fragmented landscapes}. Oecologia 127:69--77.

\bibitem[\citeproctext]{ref-WOS:000287168100017}
Slade, E. M., D. J. Mann, and O. T. Lewis. 2011. \href{https://doi.org/10.1016/j.biocon.2010.08.011}{Biodiversity and ecosystem function of tropical forest dung beetles under contrasting logging regimes}. BIOLOGICAL CONSERVATION 144:166--174.

\bibitem[\citeproctext]{ref-WOS:000368243300001}
Slade, E. M., T. Riutta, T. Roslin, and H. L. Tuomisto. 2016a. \href{https://doi.org/10.1038/srep18140}{The role of dung beetles in reducing greenhouse gas emissions from cattle farming}. Scientific Reports 6.

\bibitem[\citeproctext]{ref-slade_disentangling_2016}
Slade, E. M., T. Roslin, M. Santalahti, and T. Bell. 2016b. \href{https://doi.org/10.1111/oik.02640}{Disentangling the ``brown world' faecal-detritus interaction web: Dung beetle effects on soil microbial properties}. Oikos 125:629--635.

\bibitem[\citeproctext]{ref-taubert_global_2018}
Taubert, F., R. Fischer, J. Groeneveld, S. Lehmann, M. S. Müller, E. Rödig, T. Wiegand, and A. Huth. 2018. \href{https://doi.org/10.1038/nature25508}{Global patterns of tropical forest fragmentation}. Nature 554:519--522.

\bibitem[\citeproctext]{ref-tewksbury_corridors_2002}
Tewksbury, J. J., D. J. Levey, N. M. Haddad, S. Sargent, J. L. Orrock, A. Weldon, B. J. Danielson, J. Brinkerhoff, E. I. Damschen, and P. Townsend. 2002. \href{https://doi.org/10.1073/pnas.202242699}{Corridors affect plants, animals, and their interactions in fragmented landscapes}. Proceedings of the National Academy of Sciences 99:12923--12926.

\bibitem[\citeproctext]{ref-vulinec_dung_2002}
Vulinec, K. 2002. \href{https://doi.org/10.1111/j.1744-7429.2002.tb00541.x}{Dung beetle communities and seed dispersal in primary forest and disturbed land in amazonia}. Biotropica 34:297--309.

\bibitem[\citeproctext]{ref-white_relationships_2007}
White, E. P., S. K. M. Ernest, A. J. Kerkhoff, and B. J. Enquist. 2007. \href{https://doi.org/10.1016/j.tree.2007.03.007}{Relationships between body size and abundance in ecology}. Trends in Ecology \& Evolution 22:323--330.

\bibitem[\citeproctext]{ref-wu_brown-world_2011}
Wu, X., J. E. Duffy, P. B. Reich, and S. Sun. 2011. \href{https://doi.org/10.1890/10-0808.1}{A brown-world cascade in the dung decomposer food web of an alpine meadow: Effects of predator interactions and warming}. Ecological Monographs 81:313--328.

\bibitem[\citeproctext]{ref-young_dung_2023}
Young, K., A. Tsalickis, T. N. Sheehan, K. D. Klepzig, M. S. Caterino, and J. A. Hartshorn. 2023. Dung beetle community composition in the presence and absence of mesopredators in a longleaf pine forest. Southeastern Naturalist 22:78--90.

\bibitem[\citeproctext]{ref-zhou_linking_2023}
Zhou, G., Y. Huan, L. Wang, Y. Lan, T. Liang, B. Shi, and Q. Zhang. 2023. \href{https://doi.org/10.1016/j.scitotenv.2023.162261}{Linking ecosystem services and circuit theory to identify priority conservation and restoration areas from an ecological network perspective}. Science of The Total Environment 873:162261.

\end{CSLReferences}

\newpage

\begin{table}[!h]
\centering
\caption{\label{tab:summary-table}Dung beetle species sampled in the SRS site and their total abundance over the course of the study.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{>{}lcrcccc}
\toprule
\textbf{Species} & \textbf{Guild} & \textbf{N} & \textbf{Matrix} & \textbf{Corridor} & \textbf{Winged} & \textbf{Rectangular}\\
\midrule
\em{Canthon vigilans} & roll & 1473 & x & x & x & x\\
\em{Ateuchus lecontei} & tunnell & 1115 & x & x & x & x\\
\em{Phanaeus igneus} & tunnell & 958 & x & x & x & x\\
\em{Dichotomius carolinus} & tunnell & 556 & x & x & x & x\\
\em{Aphodius alloblackburneus} & dwell & 585 & x & x & x & x\\
\em{Onthophagus pennsylvanicus} & tunnell & 207 & x & x & x & x\\
\em{Melanocanthon bispinatus} & roll & 83 & x & x & x & x\\
\em{Phanaeus vindex} & tunnell & 133 & x & x & x & x\\
\em{Boreocanthon probus} & roll & 47 & x & x & x & x\\
\em{Copris minutus} & tunnell & 24 & x & x & x & x\\
\em{Deltochilum gibbosum} & roll & 14 & x & x & x & x\\
\em{Aphodius oximus} & dwell & 11 & x & x & x & x\\
\em{Geotrupes blackburnii} & tunnell & 1 &  &  & x & \\
\em{Onthophagus concinnus} & tunnell & 2 & x &  & x & \\
\em{Onthophagus striatulus} & tunnell & 3 & x &  &  & x\\
\em{Onthophagus tuberculifrons} & tunnell & 1 & x &  &  & \\
\bottomrule
\end{tabular}}
\end{table}

\newpage
\begin{landscape}

\begin{table}[!h]
\centering
\caption{\label{tab:Table-MO}Model of species abundance by patch type.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{lcclccl}
\toprule
\textbf{Effect} & \textbf{Group} & \textbf{Term} & \textbf{Estimate} & \textbf{SE} & \textbf{Statistic} & \textbf{P}\\
\midrule
Fixed &  & (Intercept) & 5.896 & 0.294 & 20.062 & < 0.001\\
Fixed &  & Patch type: Connected & -0.125 & 0.139 & -0.9 & 0.368\\
Fixed &  & Patch type: Rectangle & -0.528 & 0.107 & -4.919 & < 0.001\\
Fixed &  & Patch type: Winged & -0.303 & 0.072 & -4.188 & < 0.001\\
Random & Block & SD (Intercept) & 0.585 &  &  & \\
Random & Block & correlation: (Intercept) x Patch type: Connected & -0.932 &  &  & \\
Random & Block & correlation: (Intercept) x Patch type: Rectangle & -0.744 &  &  & \\
Random & Block & correlation: (Intercept) x Patch type: Winged & -0.967 &  &  & \\
Random & Block & SD Patch type: Connected & 0.266 &  &  & \\
Random & Block & correlation: Patch type: Connected x Patch type: Rectangle & 0.865 &  &  & \\
Random & Block & correlation: Patch type: Connected x Patch type: Winged & 0.816 &  &  & \\
Random & Block & SD Patch type: Rectangle & 0.194 &  &  & \\
Random & Block & correlation: Patch type: Rectangle x Patch type: Winged & 0.653 &  &  & \\
Random & Block & SD Patch type: Winged & 0.116 &  &  & \\
\bottomrule
\end{tabular}}
\end{table}

\end{landscape}
\newpage

\begin{table}[!h]
\centering
\caption{\label{tab:Table-aovM51}Linear mixed effects model of pooled total biomass by patch type including sampling block as a random effect.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{lllccc}
\toprule
\textbf{Effect} & \textbf{Group} & \textbf{Term} & \textbf{Estimate} & \textbf{SE} & \textbf{Statistic}\\
\midrule
Fixed &  & (Intercept) & 61.593 & 17.199 & 3.581\\
Fixed &  & Patch: Connected & -7.337 & 8.659 & -0.847\\
Fixed &  & Patch: Rectangle & -23.296 & 8.659 & -2.69\\
Fixed &  & Patch: Winged & -9.608 & 8.659 & -1.11\\
Random & block & SD (Intercept) & 32.145 &  & \\
Random & Residual & SD (Observation) & 12.246 &  & \\
\bottomrule
\end{tabular}}
\end{table}

\newpage

\begin{table}[!h]
\centering
\caption{\label{tab:Table-aovM52}Linear mixed effects model of biomass by species and patch type including sampling block as a random effect. Species codes = bpro: Boreocanthon probus, cvig: Canthon vigilans, dcar: Dichotomius carolinus, mbis: Melanocanthon bispinatus, ocon: Onthophagus concinnus, open: Onthophagus pennsylvanicus, pign: Phanaeus igneus, pvin: Phanaeus vindex}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{lllccc}
\toprule
\textbf{Effect} & \textbf{Group} & \textbf{Term} & \textbf{Estimate} & \textbf{SE} & \textbf{Statistic}\\
\midrule
Fixed &  & (Intercept) & 1.047 & 2.226 & 0.471\\
Fixed &  & Species: alec & 0.572 & 1.949 & 0.293\\
Fixed &  & Species: bpro & 0.078 & 1.949 & 0.04\\
Fixed &  & Species: bvig & 13.639 & 1.949 & 7\\
Fixed &  & Species: dcar & 20.454 & 1.949 & 10.497\\
Fixed &  & Species: mbis & 0.090 & 1.949 & 0.046\\
Fixed &  & Species: ocon & -0.040 & 1.949 & -0.02\\
Fixed &  & Species: open & -0.012 & 1.949 & -0.006\\
Fixed &  & Species: pign & 14.484 & 1.949 & 7.433\\
Fixed &  & Species: pvin & 1.853 & 1.949 & 0.951\\
Fixed &  & Patch: Connected & -0.734 & 1.232 & -0.595\\
Fixed &  & Patch: Rectangle & -2.330 & 1.232 & -1.89\\
Fixed &  & Patch: Winged & -0.961 & 1.232 & -0.78\\
Random & block & SD (Intercept) & 3.154 &  & \\
Random & Residual & SD (Observation) & 5.511 &  & \\
\bottomrule
\end{tabular}}
\end{table}

\newpage

\begin{table}[!h]
\centering
\caption{\label{tab:Table-aovMO3}Results of an analysis of assesing the effects of species identity, patch type, and their interaction on the abundance of dung beetle species.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{lccl}
\toprule
\textbf{Term} & \textbf{Statistic} & \textbf{df} & \textbf{P}\\
\midrule
Species & 1044.680 & 5 & < 0.001\\
Patch type & 254.638 & 3 & < 0.001\\
Species x Patch type & 110.372 & 15 & < 0.001\\
\bottomrule
\end{tabular}}
\end{table}

\newpage
\begin{landscape}

\begin{table}[!h]
\centering
\caption{\label{tab:Table-M3}Generalized linear mixed model of abundance by patch type of the 6 most abundant species and interaction term for patch type and species including sampling block as a random effect.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{lcclccl}
\toprule
\textbf{Effect} & \textbf{Group} & \textbf{Term} & \textbf{Estimate} & \textbf{SE} & \textbf{Statistic} & \textbf{P}\\
\midrule
Fixed &  & (Intercept) & 4.17 & 0.261 & 15.948 & < 0.001\\
Fixed &  & Species: alec & 0.234 & 0.078 & 2.989 & 0.003\\
Fixed &  & Species: cvig & 0.457 & 0.075 & 6.107 & < 0.001\\
Fixed &  & Species: dcar & -0.518 & 0.096 & -5.407 & < 0.001\\
Fixed &  & Species: open & -1.373 & 0.13 & -10.549 & < 0.001\\
Fixed &  & Species: pign & -0.105 & 0.085 & -1.231 & 0.218\\
Fixed &  & Patch type: Connected & -0.749 & 0.103 & -7.257 & < 0.001\\
Fixed &  & Patch type: Rectangle & -1.487 & 0.136 & -10.913 & < 0.001\\
Fixed &  & Patch type: Winged & -1.188 & 0.121 & -9.815 & < 0.001\\
Fixed &  & Species: alec x Patch type: Connected & 0.495 & 0.13 & 3.809 & < 0.001\\
Fixed &  & Species: cvig x Patch type: Connected & 0.595 & 0.124 & 4.8 & < 0.001\\
Fixed &  & Species: dcar x Patch type: Connected & 0.567 & 0.153 & 3.715 & < 0.001\\
Fixed &  & Species: open x Patch type: Connected & 0.573 & 0.201 & 2.852 & 0.004\\
Fixed &  & Species: pign x Patch type: Connected & 0.761 & 0.135 & 5.636 & < 0.001\\
Fixed &  & Species: alec x Patch type: Rectangle & 0.899 & 0.162 & 5.561 & < 0.001\\
Fixed &  & Species: cvig x Patch type: Rectangle & 0.834 & 0.158 & 5.287 & < 0.001\\
Fixed &  & Species: dcar x Patch type: Rectangle & 0.828 & 0.188 & 4.399 & < 0.001\\
Fixed &  & Species: open x Patch type: Rectangle & 0.794 & 0.243 & 3.266 & 0.001\\
Fixed &  & Species: pign x Patch type: Rectangle & 1.228 & 0.165 & 7.434 & < 0.001\\
Fixed &  & Species: alec x Patch type: Winged & 0.819 & 0.146 & 5.611 & < 0.001\\
Fixed &  & Species: cvig x Patch type: Winged & 0.987 & 0.14 & 7.072 & < 0.001\\
Fixed &  & Species: dcar x Patch type: Winged & 1.02 & 0.165 & 6.182 & < 0.001\\
Fixed &  & Species: open x Patch type: Winged & 0.41 & 0.24 & 1.711 & 0.087\\
Fixed &  & Species: pign x Patch type: Winged & 1.037 & 0.151 & 6.853 & < 0.001\\
Random & Block & SD (Intercept) & 0.509 &  &  & \\
\bottomrule
\end{tabular}}
\end{table}

\end{landscape}
\newpage

\begin{table}[!h]
\centering
\caption{\label{tab:Table-M-rich}Linear mixed effects model of species richness by patch type.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{lcclcc}
\toprule
\textbf{Effect} & \textbf{Term} & \textbf{Estimate} & \textbf{SE} & \textbf{Statistic} & \textbf{P}\\
\midrule
Fixed & (Intercept) & 2.398 & 0.151 & 15.906 & < 0.001\\
Fixed & Patch Type: Connected & -0.047 & 0.216 & -0.216 & 0.829\\
Fixed & Patch Type: Rectangle & -0.121 & 0.22 & -0.548 & 0.583\\
Fixed & Patch Type: Winged & -0.095 & 0.218 & -0.436 & 0.663\\
\bottomrule
\end{tabular}}
\end{table}

\newpage

\begin{table}[!h]
\centering
\caption{\label{tab:Table-M-shan}Linear mixed effects model of Shannon Diversity by patch type.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{lcclcc}
\toprule
\textbf{Effect} & \textbf{Group} & \textbf{Term} & \textbf{Estimate} & \textbf{SE} & \textbf{Statistic}\\
\midrule
Fixed &  & (Intercept) & 5.202 & 0.665 & 7.825\\
Fixed &  & Patch type: Connected & 0.085 & 0.495 & 0.171\\
Fixed &  & Patch type: Rectangle & 0.234 & 0.495 & 0.472\\
Fixed &  & Patch type: Winged & -0.251 & 0.495 & -0.506\\
Random & Block & sd\_\_(Intercept) & 1.13 &  & \\
Random & Residual & SD (Observation) & 0.701 &  & \\
\bottomrule
\end{tabular}}
\end{table}

\newpage

\begin{table}[!h]
\centering
\caption{\label{tab:Table-M-simp}Linear mixed effects model of Simpson's Diversity by patch type.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{lcclcc}
\toprule
\textbf{Effect} & \textbf{Group} & \textbf{Term} & \textbf{Estimate} & \textbf{SE} & \textbf{Statistic}\\
\midrule
Fixed &  & (Intercept) & 4.181 & 0.64 & 6.535\\
Fixed &  & Patch type: Connected & -0.107 & 0.411 & -0.259\\
Fixed &  & Patch type: Rectangle & 0.059 & 0.411 & 0.143\\
Fixed &  & Patch type: Winged & -0.48 & 0.411 & -1.169\\
Random & Block & sd\_\_(Intercept) & 1.14 &  & \\
Random & Residual & sd\_\_Observation & 0.581 &  & \\
\bottomrule
\end{tabular}}
\end{table}

\newpage

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{images/total_abund}

}

\caption{Average total dung beetle abundance by patch type with standard deviations as error bars.}\label{fig:patch}
\end{figure}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sp\_code\_table}\OtherTok{\textless{}{-}}\FunctionTok{data.frame}\NormalTok{(}\AttributeTok{sp\_code=}\FunctionTok{c}\NormalTok{(}\StringTok{"aaeg"}\NormalTok{, }\StringTok{"cvig"}\NormalTok{, }\StringTok{"dcar"}\NormalTok{, }\StringTok{"open"}\NormalTok{, }\StringTok{"pign"}\NormalTok{, }\StringTok{"alec"}\NormalTok{))}
\NormalTok{species\_names}\OtherTok{\textless{}{-}}\NormalTok{spp\_codes }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(sp\_code}\SpecialCharTok{\%in\%}\NormalTok{sp\_code\_table}\SpecialCharTok{$}\NormalTok{sp\_code) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{species=}\FunctionTok{tolower}\NormalTok{(species)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{species=}\FunctionTok{paste}\NormalTok{(}\AttributeTok{species=}\NormalTok{genus,species,}\AttributeTok{sep=}\StringTok{" "}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{select}\NormalTok{(sp\_code, species) }\SpecialCharTok{\%\textgreater{}\%} 
\FunctionTok{mutate}\NormalTok{(}\AttributeTok{species=}\FunctionTok{paste}\NormalTok{(sp\_code,species,}\AttributeTok{sep=}\StringTok{": "}\NormalTok{)) }
\NormalTok{species\_names}\OtherTok{\textless{}{-}}\FunctionTok{paste}\NormalTok{(species\_names}\SpecialCharTok{$}\NormalTok{species,}\AttributeTok{collapse=}\StringTok{", "}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\newpage

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{images/sp_abund_patch}

}

\caption{Average abundance of the top 6 most abundant species by patch type. Species codes: alec: Ateuchus lecontei, cvig: Canthon vigilans, dcar: Dichotomius carolinus, open: Onthophagus pennsylvanicus, pign: Phanaeus igneus, aaeg: Aphodius alloblackburneus.}\label{fig:abundance}
\end{figure}

\newpage

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{images/sp_biomass_plot}

}

\caption{Total dung beetle biomass collected for each species with sufficient weights.}\label{fig:sp-bmass}
\end{figure}

\newpage

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{images/avg_bmass_patch_plot}

}

\caption{Average total biomass by patch type with standard deviation as error bars.}\label{fig:patch-bmass}
\end{figure}

\newpage

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{images/avg_bmass_block_plot}

}

\caption{Average total biomass by sampling block with standard deviation as error bars.}\label{fig:block-bmass}
\end{figure}

\newpage

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{images/sp_richness_patches}

}

\caption{Dung beetle species richness in each patch type. The point shapes indicates the block in which each patch was located.}\label{fig:richness}
\end{figure}

\newpage

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{images/sp_shannon_patches}

}

\caption{Dung beetle Shannon diversity by patch type.}\label{fig:shannon}
\end{figure}

\newpage

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{images/sp_simpson_patches}

}

\caption{Dung beetle Simpson's index by patch type.}\label{fig:simpson}
\end{figure}

\newpage

\begin{center}
\end{center}

\end{document}

### 2025-08-01 corridor_docs/eric_ms_thesis/ms_forms_files/eec_ms_proposal.Rmd
reorg of repo for MS

--- 2024-05-02 corridor_docs/masters_cmtee_proposal/eec_ms_proposal.Rmd
@@
 date: 'updated: `r Sys.Date()`'
-bibliography: "`r here::here('corridor_docs', 'eric-thesis.bib')`"
-csl: "`r here::here('corridor_docs', 'ecology.csl')`"
+bibliography: "`r here::here('corridor_docs', 'eric_ms_thesis','eric-thesis.bib')`"
+csl: "`r here::here('corridor_docs', 'eric_ms_thesis', 'ecology.csl')`"
 output: pdf_document
@@
 
-<!-- knitr::include_graphics(here("corridor_docs","proposal","images", "plots1.png")) -->
+<!-- knitr::include_graphics(here("corridor_docs",, 'eric_ms_theis',"ms_proposal","images", "plots1.png")) -->
 <!-- ``` -->

### 2025-08-01 corridor_docs/radar/radar_team_info.Rmd
reorg of repo for MS

---
title: "radar team ufl"
author: "Emilio M. Bruna"
date: "2024-04-18"
output: html_document
bibliography: "`r here::here('corridor_docs', 'radar','radar_refs','insect_radar.bib')`"
csl: "`r here::here('corridor_docs', 'eric_ms_thesis', 'ecology.csl')`"
---

---
nocite: |
  @*
---

---
# Personnel

## Entomology
[Norman C. Leppla](https://entnemdept.ufl.edu/people-directory/norm-leppla/)

[Isaac Esquivel](https://nfrec.ifas.ufl.edu/faculty-directory/isaac-esquivel/)
William Piwowarek (MS Student)

## Agronomy
[Marcelo Wallau](https://agronomy.ifas.ufl.edu/people/marcelo-wallau/)

## Department of Agricultural and Biological Engineering
[Jasmeet Judge](https://abe.ufl.edu/people/faculty/jasmeet-judge/)

## Electrical and Computer Engineering
[Tan Wong](https://wong.ece.ufl.edu/)
David Greene (Ph.D. student)

# Meeting 30 September

technically a bis-static radar because there are 2 antennas. Will be using a digital system to send digitally modified waverform received by a digital

Called harmonic because transmitted at 9 gigahertz and retransmitted at quite that.

Need a power receiver because signal is weak. 5 m tower height, 0.5 km range

How plants and animals move is at the core of ecological theory and central to conservation -  I cannot overemphasize how important it is - and yet for the most of the planets species it is a total mystery because they're just too small for the existing methods and technologies. You are working on one of the holy grails of ecology.

Ability to add lidar is incredibly exciting because we could relate habitat structure to animal movements. This would be a great add-in to a potential NSF grant

# Potential NSF applications

NSF 23-578: Research Resources Cluster (Innovation: Instrumentation)

_"The Instrumentation Programmatic Area supports the design of novel and innovative instrumentation and associated methods that address a clearly defined gap in biologists’ ability to capture observations of biological phenomena and that have the potential to be broadly applicable in biology. Proposed projects may include instrumentation for observing any level of biological phenomena (e.g., molecular, cellular, organismal, ecosystem, biome), and may propose either new and innovative instrumentation; instrumentation that significantly improves the accuracy, resolution, or throughput of data capture; or advancements that reduce costs of instrument construction or operation."_

Recent Awards from this solicitation:  [link](https://www.nsf.gov/awardsearch/advancedSearchResult?ProgEleCode=165Y&BooleanElement=Any&BooleanRef=Any&ActiveAwards=true#results)

## Grant meeting 7 october 2024

1. IFAS grant to visit agency contacts

1. 1 page summary to NSF PO

Questions for PO:

- better as an EAGER?

main difference - working on an elecrtronically sterable array onstead of mechanical array (italian just keeops spinning lokoing for signal - this is not a bvery durable or desirable feature). An eletronically receivable array doesnt require movement. Electronically receivable at this frequency are not uncommon, but they are used to track bigger things and are more powerful (signal). Possibility is also that could be able to track multiple simultanously. (whether practicval or doable is a sepoerate question).

EB question how would this be better for me as a biologist: smaller? cheaper?

A from Tan Wong: what others have built is very specialized, not an "easy-to-follow" recepie

Our group's goal is to make it more portable, more user-friendly, easier to set-up and maintain.

Bees carry 25% of weight in pollen, so think of

processing of high volume data stream, models

main probelm - vegetation penetration, At this frequeny,. radar signal could not penetrate. If you want to be use. a frequency that is more powerfiul. not just power *(spatial respolution isw based on bandwidth, how far you can go, propagatilon environment of radar signal). regulatory problem (FCC) too - need to be in differnt bands  only allowed to transmit in certain bands.Certain bands we dont need lisence to operate on (wifi, blue tooth) but these are hevily congested. The amount of power we expect to be reflected from transponder. Probably no hope to use these bands due to how polluted they are. Bands chosen are based on italian paper, You can aappl;y for the bands (like we did) at certain sites. Can emphasize this is something you can do at fuled stations that will become center for research in the topic.

Doing this in forest is a bigger

### 2025-08-01 corridor_docs/eric_ms_thesis/04_GLMM.R

@@ -1,668 +1,0 @@
-library(tidyverse)
-library(here)
-library(kableExtra)
-library(sjPlot)
-library(lme4)
-# READ THIS ABOUT GLMM REFERENCE LEVELS
-# https://stats.stackexchange.com/questions/628348/when-i-change-my-reference-level-on-my-glmer-in-r-why-do-the-p-values-change-an
-
-# load and prep data ------------------------------------------------------
-
-
-btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
-spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
-
-
-# calclulate Hill Nos. ----------------------------------------------------
-
-
-# using hillR
-# https://github.com/daijiang/hillR/blob/master/README.md
-# https://daijiang.r-universe.dev/hillR
-library(hillR)
-
-
-hill_data<-btl_data %>%
-  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
-  group_by(species,patch,block) %>%
-  mutate(patch=case_when(
-    patch == "c" ~ "Connected",
-    patch == "m" ~ "Matrix",
-    patch == "w" ~ "Winged",
-    patch == "r" ~ "Rectangle",
-    .default = as.character(patch))) %>%
-  summarize(n=sum(n, na.rm=TRUE)) %>%
-  arrange(patch,desc(n)) %>%
-  rename(sp_code=species,
-         n_patch=n) %>%
-  mutate(patch_block=paste(patch,block,sep="_")) %>% 
-  ungroup() %>% 
-  select(-patch,-block) %>% 
-  pivot_wider(
-    names_from = sp_code,
-    values_from = n_patch) 
-
-
-
-spp_abund<-btl_data %>%
-  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
-  mutate(patch=case_when(
-    patch == "c" ~ "Connected",
-    patch == "m" ~ "Matrix",
-    patch == "w" ~ "Winged",
-    patch == "r" ~ "Rectangle",
-    .default = as.character(patch))) %>%
-  group_by(species,patch,block) %>%
-  summarize(n=sum(n, na.rm=TRUE)) %>%
-  arrange(patch,desc(n)) %>%
-  rename(sp_code=species) %>%
[…]

### 2025-08-01 corridor_docs/eric_ms_thesis/code/02_beetle_biomass_cleanup.R

@@ -1,0 +1,56 @@
+# Script to determine if weights of btls in drying oven have stabilized  
+# date: 19 March 2025
+
+# load libraries ----------------------------------------------------------
+
+library(tidyverse)
+library(here)
+
+
+# load data ---------------------------------------------------------------
+
+btl_wt<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","btl_weight_equilibrium.csv"))
+
+btl_wt<-btl_wt %>% 
+  pivot_longer(wet_weight:weight_5,names_to = "interval") %>% 
+  mutate(interval=if_else(interval=="wet_weight","weight_0",interval)) %>% 
+  arrange(sample_id,sp_code,interval) %>% 
+  group_by(sample_id,sp_code) %>% 
+  mutate(perc_loss=((lag(value)-(value))/lag(value)*100)) %>% 
+  mutate(sample=paste(sp_code,sample_id,sep="-"))
+  
+btl_wt
+
+
+btl_wt %>% 
+  # filter(interval!="weight_1") %>%
+  # filter(interval!="weight_0") %>%
+  ggplot( aes(x=interval, y=value, group=sample, color=sample)) +
+  geom_line()
+
+
+
+#  btl weights ------------------------------------------------------------
+
+
+plate_weight<-btl_wt %>% 
+  ungroup() %>% 
+  filter(interval=="weight_5") %>% 
+  filter(sp_code=="plate") %>% 
+  select(value)  
+
+
+
+corrected_btl_wt<-
+btl_wt %>% 
+  filter(interval=="weight_5") %>% 
+  mutate(plate_weight=plate_weight$value) %>% 
+  filter(sp_code!="plate") %>% 
+  group_by(sp_code) %>% 
+  mutate(corrected_wt=value-plate_weight) %>% 
+  summarize(mean_wt=mean(corrected_wt))
+
+
+btl_wt_clean<-write_csv(corrected_btl_wt,here("corridor_docs","eric_ms_thesis","ms_data","data_clean","corrected_btl_wt.csv"))
+
+

### 2025-08-01 corridor_docs/eric_ms_thesis/code/03_biodiversity_metrics.R

@@ -1,0 +1,1188 @@
+# Script to import, clean, and organize EEC's Beetle data
+# date: 25 september 2024
+
+# load libraries ----------------------------------------------------------
+
+library(tidyverse)
+library(here)
+library(lubridate) # to deal with dates https://lubridate.tidyverse.org/
+library(vegan)
+library(permute)
+library(lattice)
+library(ggmosaic)
+library(iNEXT)
+# load data ---------------------------------------------------------------
+
+btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
+
+
+# SUMMARY STATS -----------------------------------------------------------
+
+# total beetles captured --------------------------------------------------
+total_N<-btl_data %>% 
+  summarise(n=sum(sum, na.rm=TRUE))
+total_N
+# How many species total?
+total_spp<-btl_data %>% 
+  select(-sum) %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n))) %>% 
+  summarise(n_spp=n_distinct(species))
+total_spp
+
+# Number captured per species (all data pooled) ---------------------------
+btl_order<-btl_data %>%
+  select(-sum) %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(species) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  mutate(species = reorder(species, desc(n)))
+btl_order
+
+
+# Number captured per block -----------------------------------------------
+N_block<-btl_data %>%
+  select(-sum) %>% 
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
+  group_by(block) %>%
+  summarize(n=sum(n, na.rm=TRUE))
+N_block
+
+# number species per block ------------------------------------------------
+
+N_spp_block<-btl_data %>%
[…]

### 2025-08-01 corridor_docs/eric_ms_thesis/code/05_biomass_analysis.R

@@ -1,0 +1,438 @@
+# Script to import, clean, and organize EEC's Beetle data
+# date: 25 september 2024
+
+# load libraries ----------------------------------------------------------
+
+library(tidyverse)
+library(here)
+library(lme4)
+library(sjPlot)
+# load data ---------------------------------------------------------------
+
+
+btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
+spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
+
+# need to add how many beetles were included in each raw weight (some too small)
+btls_per_tray<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","spp_weights.csv")) %>%
+  group_by(sp_code) %>% 
+  summarize(n_tray=n()) %>% 
+  mutate(btls_per_tray=case_when(
+    n_tray==15~1,
+    n_tray==5~3,
+    n_tray==3~5,
+    n_tray==1~15,
+    .default = as.integer(n_tray)
+  ))
+
+
+spp_bmass<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","spp_weights.csv")) %>% 
+  left_join(btls_per_tray,by="sp_code")  %>% 
+  mutate(indiv_bmass=weight/btls_per_tray) %>% 
+  group_by(sp_code) %>% 
+  summarize(avg_ind_bmass=mean(indiv_bmass),
+            sd = sd(indiv_bmass))
+
+biomass_per_spp<-btl_data %>%
+  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
+  group_by(species) %>%
+  mutate(patch=case_when(
+    patch == "c" ~ "Connected",
+    patch == "m" ~ "Matrix",
+    patch == "w" ~ "Winged",
+    patch == "r" ~ "Rectangle",
+    .default = as.character(patch))) %>%
+  summarize(n=sum(n, na.rm=TRUE)) %>%
+  arrange(desc(n)) %>%
+  rename(sp_code=species) %>% 
+  left_join(spp_bmass,by="sp_code") %>% 
+  # select(-sd, -n_btls) %>% 
+  mutate(spp_biomass=n*avg_ind_bmass)
+
[…]

### 2025-08-01 corridor_docs/eric_ms_thesis/EscobarChena_etal_ms.tex
i/my to we/our

@@
 
-By measuring changes in biodiversity and species richness within experimental designs I am able to isolate factors might be contributing to ecological patterns and processes (Resasco et al.
+By measuring changes in biodiversity and species richness within experimental designs we were able to isolate factors might be contributing to ecological patterns and processes (Resasco et al.
 2017, Fletcher Jr.
@@
 
-Here, I aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources (Hanski and Cambefort 1991), interact with corridors in their landscapes.
-I sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch area to edge ratio and distance to edge (Tewksbury et al.
+Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources (Hanski and Cambefort 1991), interact with corridors in their landscapes.
+We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch area to edge ratio and distance to edge (Tewksbury et al.
 2002).
-I do so I address these questions:
+We do so to address these questions:
 
@@
 
-My study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, USA (33.208\(^\circ\) N, 81.408\(^\circ\) W, Figure \ref{fig:srs-map}).
+Our study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, USA (33.208\(^\circ\) N, 81.408\(^\circ\) W, Figure \ref{fig:srs-map}).
 in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals (Tewksbury et al.
@@
 Individual pitfall traps consisted of two components, a 10cm tall by 8 cm wide cylinder base topped with a funnel with a 10cm wide rim.
-I sourced pig feces from the University of Florida Swine Barn Unit.
+We sourced pig feces from the University of Florida Swine Barn Unit.
 Bait was processed into 5cm wide balls and wrapped in a layer of coffee filter material.
@@
 Fifteen individuals of each species with adequate captures were dried and weighed for biomass measurements.
-Ten species had the required counts for weighing which I dried in drying oven until all specimen reached a stable mass.
-I weighed each beetle using an Ohaus Adventurer Pro AV53 microbalance, with the exception of individuals whose weights were extremely small.
+Ten species had the required counts for weighing which we dried in drying oven until all specimen reached a stable mass.
+We weighed each beetle using an Ohaus Adventurer Pro AV53 microbalance, with the exception of individuals whose weights were extremely small.
 Individual biomass for these species was estimated by weighing in batches of N = 3 (i.e., \emph{Ateuchus lecontei}), N = 5 (i.e., \emph{Onthophagus pennsylvanicus}), or N = 15 (i.e., \emph{Aphodius alloblackburneus}) and then calculated by average biomass per beetle.
@@
 2014).
-I compared community composition by increasing magnitudes of diversity components (i.e., \emph{qD}) of \emph{0D} (i.e., species richness), \emph{1D} (i.e., Shannon entropy), and \emph{2D} (i.e, Simpson Diversity).
+We compared community composition by increasing magnitudes of diversity components (i.e., \emph{qD}) of \emph{0D} (i.e., species richness), \emph{1D} (i.e., Shannon entropy), and \emph{2D} (i.e, Simpson Diversity).
 Diversity numbers and species richness were calculated using the package \texttt{hill} (Li 2018) for the R statistical programming language (Posit team 2025).
@@
 
-To test for the effects of connectivity on abundance, species richness, and species diversity I compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix.
-For abundance and richness I used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution (Bolker et al.
+To test for the effects of connectivity on abundance, species richness, and species diversity we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix.
+For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution (Bolker et al.
 2009).
 Compared (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type.
-I included the identity of the sampling block as a random effects.
-To model my diversity metrics I took a similar approach, but this time using linear mixed effects models with a Gaussian distribution (Chao et al.
+We included the identity of the sampling block as a random effects.
+To model diversity metrics we took a similar approach, but this time using linear mixed effects models with a Gaussian distribution (Chao et al.
 2014).
@@
 2015).
-Prior to conducting my modeling I evaluated the the suitability of my data with \emph{qqplots} generated with the \texttt{DARMa} package (Hartig 2024).
-
-I used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation.
+Prior to conducting our modeling we evaluated the the suitability of our data with \emph{qqplots} generated with the \texttt{DARMa} package (Hartig 2024).
+
+We used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation.
 Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.
-Likewise, I used linear mixed-effects models to compare for differences between patch types in (1) total dung beetle biomass, and (2) the total biomass of each of the ten species, again including block as a random effect.
+Likewise, we used linear mixed-effects models to compare for differences between patch types in (1) total dung beetle biomass, and (2) the total biomass of each of the ten species, again including block as a random effect.
 
@@
 
-Overall, I collected N = 5213 dung beetles (N = 1359 in Connected patches, N = 1199 in Winged patches, N = 942 in Rectangle patches, N = 1713 in the Matrix).
+Overall, we collected N = 5213 dung beetles (N = 1359 in Connected patches, N = 1199 in Winged patches, N = 942 in Rectangle patches, N = 1713 in the Matrix).
 These beetles belonged to N = 16 species; the N = 6 most dominant species comprised of 93.9\% of all captures: \textit{Canthon vigilans} (N = 1473), \textit{Ateuchus lecontei} (N = 1115), \textit{Phanaeus igneus} (N = 958), \textit{Aphodius alloblackburneus} (N = 585), \textit{Dichotomius carolinus} (N = 556), and \textit{Onthophagus pennsylvanicus} (N = 207; Table 1).
@@
 In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches.
-My main findings emphasized: (1) Habitat type and patch shape were the main driving factors for determining how dung beetle species abundances were composed, however effects were species specific.
+Our main findings emphasized: (1) Habitat type and patch shape were the main driving factors for determining how dung beetle species abundances were composed, however effects were species specific.
 (2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level.
@@
 
-In my comparison of dung beetle compositions, connectivity and habitat edge accounted for differences in dung beetle abundances, yet species richness and diversity were even across patch types and in the matrix.
+In our comparison of dung beetle compositions, connectivity and habitat edge accounted for differences in dung beetle abundances, yet species richness and diversity were even across patch types and in the matrix.
 Total beetle counts were consistently highest in the matrix, in comparison abundances is connected patches were not significantly different while those of the winged and rectangular patches were lower.
-Additionally, I also observed that patch effects were not equally proportional for all species.
+Additionally, we also observed that patch effects were not equally proportional for all species.
 This suggests that the species captured could be exhibiting habitat preferences between the open patches versus forested matrix.
@@
 While the experimental design attempted to control for the effects of patch size and edge, there could be large (and potential unknown) environmental gradients across the SRS landscape that could influence the observed patterns in diversity and abundance.
-For instance, at the time of my study Block 8 had the densest matrix of any of the blocks.
+For instance, at the time of our study Block 8 had the densest matrix of any of the blocks.
 This could have hindered the diffusion of bait scent, leading to lower capture rates in this block.
@@
 2017b), so it is not unexpected that they might show a preference towards forested matrix.
-I did not detect that any species was more positively associated with open patches despite suggestions that some species (e.g., \emph{Canthon vigilans}, \emph{Melanocanthon bispinatus}) prefered open habitat (Nealis 1977, Conover et al.
+We did not detect that any species was more positively associated with open patches despite suggestions that some species (e.g., \emph{Canthon vigilans}, \emph{Melanocanthon bispinatus}) prefered open habitat (Nealis 1977, Conover et al.
 2019).
-This could be another sign pointing towards matrix acting as a source population, and since open habitat was much less dominant in my experimental system beetles could be moving into patches from habitat edge.
+This could be another sign pointing towards matrix acting as a source population, and since open habitat was much less dominant in our experimental system beetles could be moving into patches from habitat edge.
 
@@
 2008).
-I did not observe such intense habitat specificity, which may not be especially common outside of the tropics.
-I found similar patterns of community composition as did the small number of prior dung beetle biodiversity surveys conducted in the southeastern United States (Nealis 1977, Conover et al.
+We did not observe such intense habitat specificity, which may not be especially common outside of the tropics.
+We found similar patterns of community composition as did the small number of prior dung beetle biodiversity surveys conducted in the southeastern United States (Nealis 1977, Conover et al.
 2019), including at least one more where differences between treatment types were non-significant while variation on a larger scale was more apparent (Young et al.
 2023).
-I did expect some species to be more dominant in patches with higher edge ratios as was found with ants at SRS (Resasco et al.
+We did expect some species to be more dominant in patches with higher edge ratios as was found with ants at SRS (Resasco et al.
 2014), but species were evenly present in all patches.
@@
 Finally, it is important to emphasize that while dung beetles are capable of long distance flight and detecting dung at distances of over 50 meters (Gray et al.
-2022), the results I observed are likely the result of a mismatch between the spatial scale of the experimental replicates and dung beetle movement.
+2022), the results we observed are likely the result of a mismatch between the spatial scale of the experimental replicates and dung beetle movement.
 Put another way, some beetles were almost certainly drawn by the dung used in baits from the matrix into the plots.
@@
 Another potential avenue for improvement is lowering the grain size of sampling and changing trap placement to better understand the effects of edge proximity and connectivity.
-I also used one of the best bait types for collecting dung beetles in pig dung (Marsh et al.
+We also used one of the best bait types for collecting dung beetles in pig dung (Marsh et al.
 2013), but a mix of differently sourced baits may have been more optimal as more diverse baits attract more diverse species (Frank et al.
@@
 SRS provides an excellent experimental design for direct comparison of landscapes but another main goal of the site is to study dispersal, dung beetles would be an excellent system for studying movement.
-They already show promise, in preliminary trials I released beetles from both the connected and rectangular patches with baited traps in the central patch.
-I recovered one recapture which originated from the connected patch, indicating that individuals do move through corridors.
+They already show promise, in preliminary trials we released beetles from both the connected and rectangular patches with baited traps in the central patch.
+We recovered one recapture which originated from the connected patch, indicating that individuals do move through corridors.
 

### 2025-08-02 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
im sleepy

--- 2025-07-25 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
@@
 
-# output: bookdown::pdf_document2
-  # pdf_document:
+#output: bookdown::pdf_document2
+#pdf_document:
 #  includes:
-  #     # in_header: style/preamble.tex
+    # in_header: style/preamble.tex
 #    template: null
@@
 knitr::include_graphics("images/avg_bmass_block_plot.png")
+```
+
+```{r, label = abund-bmass, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Trendline of average dung beetle biomass by total individuals captured.", out.width="70%", fig.align='center'}
+
+knitr::include_graphics("images/bmass_v_abund.png")
 ```

### 2025-08-02 corridor_docs/2024_corridor_mtg_notes.Rmd

---
title: "Corridor Meeting"
author: "Emilio M. Bruna"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# INTRODUCTORY OVERVIEW

## Nick Haddad
### Overview of the Corridor Expeoriment

**Capacity Building**

-   79 undergrads
-   39 postbac scientists
-   27 grad students
-   9 postdocs

**Overarching Question**

Do corridors work for dispersal (phase 1) - Builds on a priort project
investigating the effects of corridor *length* - Corridor project
originally designed around 1 sp of butterfly (spicebush swallowtail) -
corridors and wind - maybe use this to ask questions about volatiles and
detection? - spicebush swallowtail only species that *didn't* use
corridors

Do corridors increase diversity (phase 2) - 25 years of plant diversity
surveys - effect of fire ants on other ant species - spiders

Plant population dynamics and species interactions - restoration plants
and interactions (with Lauren Sullivan MSU) - observed depredation of
butterflies, maybe by prayinf mantis?

Outreach - Conservation Corridor - research digests - 30 x 30
initiatives

EB notes: reminds me very much of fragmentation work

-   interactions & community composition, jump over population dynamics
-   limited matrix comparisons

## Lars Brudvig

### The Remnant Project

**Overarching question** How best to restore longleaf pine remnants?

-   At SRS in 1951 most of LLP forest was cleared for ag, but there were
    remnants, perhaps in woodlots
-   this was followed by the largest planting effort in the world, with
    planted forest and interspersed patches of llp
-   remnant project: a remnant of forest adjacent to an ag field that
    was abandoned and replanted (post ag).
-   each of remnant and post-ag site had two plots, with in the post-ag
    one being thinned and the other being unthinned
-   Questions:
    -   What are consequences of past ag land use?
    -   what are consequences of restoration thinning treatment
    -   can thinning ameliiorate legacies of ag land use?
    -   to what extent can remnants serve as sources for restoration and
        post-ag land areas?

-   Lesson 1: effects of ag history are long-lived
    -   ag history alters plant
        composition.
    -   about 60 spp of plants only foundin remnant fragments even
        though there have been decades for them to move only meters into
        adjacent restoration areas
    -   ag legacy on soils
-   Lesson 2
    - (missed)

-   Lesson 3: what limits plant colonization post ag
    - limited seed production? In absence of thinning low flowering and seed production, much higher seed production
    - limited seed dispersal? most of occurrence in post-ag is in seed additiona plots, with more establishment in thinning. def some limitation.
    - establishment / persistence
    - many plant species will manually resestablish when manually added

## Ellen Damschen
### SERDP Project: restoring the ground layer community

- Brett Mattingly, John Orrick, Ellen, and Lars
- thinking about gradients in fire history 
- study also at ft stuart and fort liberty (formerly ft bragg)
- how to most effectively do restoration?
- asking about role of adding seeds relative to a bunch of other processes? 
    - consumer exclusion
    - elimination of competition by spraying with herbicide
- initial: 108 sites, 30 million seeds of 25 species, veg data 2010-2013

- seed addition consistently increases diversity over time
- other ket factords seed coat, liiter depth (fire burns off), temp 
- once established, recovered populations persist and spread
- management can affect recovery and spread
    - prescribed fire and thinning
    - adding seeds

# STUDENT PRESENTATIONS

## Melissa Burt (did PhD with Susan Whitehead at VTech)
### Ants

- >144K ants identified, 60+ species
- 45% fire ants
- spp richness increasing over time everywhere, at higher rate in connected (effect of fire ants)
- patches with corridors have fewwr species than isolatoned and high edge patches
- fewer species 
EB note - has dung beetle bycatch in pitfalls

## Katherine
### PLant arthropod interactions

- how do different factors influence multiple plant-arhropod interactions that can influence reproductive output: pollination, florivory, and depredation of pollinators by spiders
- how do these factors influence pollination rate?

used a founder species (_Carphephorus bellidifolius_)
founder species = species inplanted in 2007(?) experiment where introduced into different patches in edge-interior grid pattern

- observes pollinator visits for 10 min
- collect folivores and spiders _after_ the 10 min visit
- surveyed floral abundnace within 5 m of plant
- collected seeds ar end of season to measure pollination rate

## Thomas
### Edge Effects, matrix, and spillover

- edge effects and spilloiver
- population density leads to spillover
- cons of spillover - disease, predators, etc.
- 2009 (PNAS) study by lars et al - spillover out of connected patches into matrix greater, edge dependent decreases
- spillover driver by diversity at the patch level
- shortly after - spillover effect
- 2011 matrix thinned for red cockaded woodpecker

## Angelica Bautista
### Seed emergence and establishment in LLP

### 2025-08-02 corridor_docs/eric_ms_thesis/EscobarChena_etal_ms.Rmd

@@
 
-By measuring changes in biodiversity and species richness within experimental designs I am able to isolate factors might be contributing to ecological patterns and processes [@fletcher_jr_landscape_2023; @resasco_contribution_2017].
+By measuring changes in biodiversity and species richness within experimental designs we were able to isolate factors might be contributing to ecological patterns and processes [@fletcher_jr_landscape_2023; @resasco_contribution_2017].
 Past studies have endeavored to experimentally measure changes in community compositions as a result of connecting habitats with corridors [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002]  Yet very few have directly compared matrix and patch populations.
@@
 
-Here, I aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources [@hanski_dung_1991], interact with corridors in their landscapes.
-I sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch area to edge ratio and distance to edge [@tewksbury_corridors_2002].
-I do so I address these  questions:
+Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources [@hanski_dung_1991], interact with corridors in their landscapes.
+We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch area to edge ratio and distance to edge [@tewksbury_corridors_2002].
+We do so to address these questions:
 
@@
 
-My study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, USA (33.208$^\circ$ N, 81.408$^\circ$ W, Figure \@ref(fig:srs-map)).
+Our study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, USA (33.208$^\circ$ N, 81.408$^\circ$ W, Figure \@ref(fig:srs-map)).
 in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals [@tewksbury_corridors_2002].
@@
 Fifteen individuals of each species with adequate captures were dried and weighed for biomass measurements.
-Ten species had the required counts for weighing which I dried in drying oven until all specimen reached a stable mass.
-I weighed each beetle using an Ohaus Adventurer Pro AV53 microbalance, with the exception of individuals whose weights were extremely small.
+Ten species had the required counts for weighing which we dried in drying oven until all specimen reached a stable mass.
+We weighed each beetle using an Ohaus Adventurer Pro AV53 microbalance, with the exception of individuals whose weights were extremely small.
 Individual biomass for these species was estimated by weighing in batches of N = 3 (i.e., *Ateuchus lecontei*), N = 5 (i.e., *Onthophagus pennsylvanicus*), or N = 15 (i.e., *Aphodius alloblackburneus*) and then calculated by average biomass per beetle.
@@
 Second, by manipulating *q* we can gain an understanding of compositional shifts otherwise obscured while using species richness [@chao_rarefaction_2014].
-I compared community composition by increasing magnitudes of diversity components (i.e., *qD*) of *0D* (i.e., species richness), *1D* (i.e., Shannon entropy), and *2D* (i.e, Simpson Diversity).
+We compared community composition by increasing magnitudes of diversity components (i.e., *qD*) of *0D* (i.e., species richness), *1D* (i.e., Shannon entropy), and *2D* (i.e, Simpson Diversity).
 Diversity numbers and species richness were calculated using the package `hill` [@li_hillr_2018] for the R statistical programming language [@posit_team_rstudio_2025].
@@
 
-To test for the effects of connectivity on abundance, species richness, and species diversity I compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix.
-For abundance and richness I used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution [@bolker_generalized_2009].
+To test for the effects of connectivity on abundance, species richness, and species diversity we compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix.
+For abundance and richness we used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution [@bolker_generalized_2009].
 Compared (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type.
-I included the identity of the sampling block as a random effects.
-To model my diversity metrics I took a similar approach, but this time using linear mixed effects models with a Gaussian distribution [@chao_rarefaction_2014].
+We included the identity of the sampling block as a random effects.
+To model diversity metrics we took a similar approach, but this time using linear mixed effects models with a Gaussian distribution [@chao_rarefaction_2014].
 In all models, the reference level for patch type was Matrix.
 All models were fit using `lme4` package [@bates_fitting_2015].
-Prior to conducting my modeling I evaluated the the suitability of my data with *qqplots* generated with the `DARMa` package [@hartig_dharma_2024].
-
-I used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation.
+Prior to conducting our modeling we evaluated the the suitability of our data with *qqplots* generated with the `DARMa` package [@hartig_dharma_2024].
+
+We used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation.
 Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.
-Likewise, I used linear mixed-effects models to compare for differences between patch types in (1) total dung beetle biomass, and (2) the total biomass of each of the ten species, again including block as a random effect.
+Likewise, we used linear mixed-effects models to compare for differences between patch types in (1) total dung beetle biomass, and (2) the total biomass of each of the ten species, again including block as a random effect.
 
@@
 
-Overall, I collected N = `r total_N` dung beetles (N = `r n_corr` in Connected patches, N = `r n_wing` in Winged patches, N = `r n_rect` in Rectangle patches, N = `r n_mat` in the Matrix).
+Overall, we collected N = `r total_N` dung beetles (N = `r n_corr` in Connected patches, N = `r n_wing` in Winged patches, N = `r n_rect` in Rectangle patches, N = `r n_mat` in the Matrix).
 These beetles belonged to N = `r total_spp` species; the N = `r how_many_common` most dominant species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1).
@@
 
-In my comparison of dung beetle compositions, connectivity and habitat edge accounted for differences in dung beetle abundances, yet species richness and diversity were even across patch types and in  the matrix.
+In our comparison of dung beetle compositions, connectivity and habitat edge accounted for differences in dung beetle abundances, yet species richness and diversity were even across patch types and in  the matrix.
 Total beetle counts were consistently highest in the matrix, in comparison abundances is connected patches were not significantly different while those of the winged and rectangular patches were lower.
-Additionally, I also observed that patch effects were not equally proportional for all species.
+Additionally, we also observed that patch effects were not equally proportional for all species.
 This suggests that the species captured could be exhibiting habitat preferences between the open patches versus forested matrix.
@@
 Beetles from the genus _Aphodius_ do tend to show patterns of habitat specificity [@roslin_distribution_2001], and many preferring forested habitat [@frank_land_2017],  so it is not unexpected that they might show a preference towards forested matrix.
-I did not detect that any species was more positively associated with open patches despite suggestions that some species (e.g., _Canthon vigilans_, _Melanocanthon bispinatus_) prefered open habitat [@conover_phenology_2019; @nealis_habitat_1977].
-This could be another sign pointing towards matrix acting as a source population, and since open habitat was much less dominant in my experimental system beetles could be moving into patches from habitat edge.
+We did not detect that any species was more positively associated with open patches despite suggestions that some species (e.g., _Canthon vigilans_, _Melanocanthon bispinatus_) prefered open habitat [@conover_phenology_2019; @nealis_habitat_1977].
+This could be another sign pointing towards matrix acting as a source population, and since open habitat was much less dominant in our experimental system beetles could be moving into patches from habitat edge.
 
@@
 continuous forest [@WOS:000349242000004], while othersfound that forest species were using living fences as corridors [@arellano_response_2008].
-I did not observe such intense habitat specificity, which may not be especially common outside of the tropics.
-I found similar patterns of community composition as did the small number of prior dung beetle biodiversity surveys conducted in the southeastern United States [@conover_phenology_2019; @nealis_habitat_1977], including at least one more where differences between treatment types were non-significant while variation on a larger scale was more apparent [@young_dung_2023].
-I did expect some species to be more dominant in patches with higher edge ratios as was found with ants at SRS [@resasco_landscape_2014], but species were evenly present in all patches.
+We did not observe such intense habitat specificity, which may not be especially common outside of the tropics.
+We found similar patterns of community composition as did the small number of prior dung beetle biodiversity surveys conducted in the southeastern United States [@conover_phenology_2019; @nealis_habitat_1977], including at least one more where differences between treatment types were non-significant while variation on a larger scale was more apparent [@young_dung_2023].
+We did expect some species to be more dominant in patches with higher edge ratios as was found with ants at SRS [@resasco_landscape_2014], but species were evenly present in all patches.
 
@@
 
-Finally, it is important to emphasize that while dung beetles are capable of long distance flight and detecting dung at distances of over 50 meters [@WOS:000707546700001], the results I observed are likely the result of a mismatch between the spatial scale of the experimental replicates and dung beetle movement.
+Finally, it is important to emphasize that while dung beetles are capable of long distance flight and detecting dung at distances of over 50 meters [@WOS:000707546700001], the results we observed are likely the result of a mismatch between the spatial scale of the experimental replicates and dung beetle movement.
 Put another way, some beetles were almost certainly drawn by the dung used in baits from the matrix into the plots.
@@
 Another potential avenue for improvement is lowering the grain size of sampling and changing trap placement to better understand the effects of edge proximity and connectivity.
-I also used one of the best bait types for collecting dung beetles in pig dung [@marsh_optimising_2013], but a mix of differently sourced baits may have been more optimal as more diverse baits attract more diverse species [@frank_nutrient_2017; @frank_search_2018; @gimenez_gomez_dung_2021].
+We also used one of the best bait types for collecting dung beetles in pig dung [@marsh_optimising_2013], but a mix of differently sourced baits may have been more optimal as more diverse baits attract more diverse species [@frank_nutrient_2017; @frank_search_2018; @gimenez_gomez_dung_2021].
 
